### 2018-01-16 MSF_Paper_FA.tex
MSF Paper

\documentclass[twocolumn]{article}
\usepackage{authblk}
\usepackage{blindtext}
\usepackage{multicol}% http://ctan.org/pkg/multicols
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{cite}
\usepackage[table,xcdraw]{xcolor}
\newcommand{\FA}[1]{\begingroup\color{magenta}#1\endgroup}
\newcommand{\TODO}[1]{\begingroup\color{red}#1\endgroup}
%opening
\begin{document}

\title{\textit{MSF: Modulated Sub-path Finder} }

\author[1]{Mariam R Farman}
\author[1,2]{Fabian Amman}
\author[1]{Ivo L Hofacker}
\affil[1]{Institute for Theoretical Chemistry,Theoretical Biochemistry Group,University of Vienna,Austria}
\affil[2]{Department of Chromosome Biology, Max F. Perutz Laboratories,University of Vienna,Austria}

\maketitle

\begin{abstract}
Differentially expressed genes play an important role in giving
insights to a phenotype. These DEGs are then used to identify the
pathways altered during changing biological conditions. We developed a
novel approach to find the maximal significantly dis-regulated
sub-paths or cluster of genes from the cell signaling network, giving
these sub-paths an overall significance of modulation by combining the
individual p-values of the genes.
\end{abstract}

\section*{Keywords}

Differentially expressed genes, Pathway analysis, Combining \textit{p}-value, Cell signaling network, Modulated Sub-paths

\section*{Background}

High throughput sequencing techniques have been widely used to yield
the differentially expressed genes (DEG)~\cite{DEG}. To this end,
changes in transcript abundance is measured, e.g.~by next generation
sequencing techniques, and interpreted as an indicator of differential
expression of genes. Differentially expressed genes can be used to get
insights into the mechanism underlying differences conditions, such as
healthy versus diseased. In the first instance, the differential gene
expression analysis informs about the magnitude of expression changes
between the conditions which is the log fold change, a sign of the log fold change and the confidence level
of observing a authentic change, often expressed as
\textit{p}-value. These lists of differentially expressed genes are
then used to extract meaningful insights, for example the genes that
could be important for a particular condition or maybe the target gene
of any infection or disease. Out of all high throughput analysis
methods currently in use, pathway-based analysis has become an
important tool to further interpret the results of a DGEA and to
acquire understandings of the perturbations in a biological
system. Biological pathways are set of genes contained in a functional
unit. They help to identify pathways or networks that may be altered
during a change of condition providing important information about
diseases and its treatment process~\cite{Khatri2012}. Pathway-based
methods use the predefined pathways or networks such as
KEGG~\cite{Kegg} and Reactome~\cite{Reactome}, the expression
measurements of the genes obtained from differential gene expression
analysis (DGEA) in combination with statistical methods and algorithms to
identify specifically modulated pathways and processes~\cite{Campos}.

The existing pathway-based analysis approaches use different research
designs, which can be categorized into ORA (Over-representation
analysis), FCS (Functional class scoring) and pathway topology based
methods. ORA is the first and the most basic method of pathway
analysis. It uses a user defined cut-off for the log-fold change and
p-value from the DGEA (most commonly using absolute log-fold change
$\geq$ 2 and \textit{p}-value $\leq$ 0.05) to define a list of
differentially expressed genes. Subsequently, sets of genes associated
with annotated pathways are tested for being overrepresented in the
set of differentially expressed genes. To this end, hypergeometric
distribution, chi-square tests, binomial probability or the Fisher’s
exact test are used. Thereby the information of the topology of the
pathways are ignored~\cite{Bayer}. Futhermore, ORA assumes the
pathways are independent of each other and ignores the fact that
biological pathways cross-talk and overlap~\cite{Khatri2012,Campos}.

Unlike ORA, FCS has no artificial cut-off to define DEG list and does
not assume genes to be independent of each other. FCS works in three
step, first it calculates the gene-level statistics including correlation of molecular measurements, ANOVA, Q-statistic, signal-to-noise ratio, t-test and Z-score. In the second step the statistics of individual genes are transformed to a individual pathway-level statistic and finally the
pathway-level statistics are accessed. Although FCs covers some of the
limitations from ORA, it still lacks the topology, cross-talk and
overlap of the pathways~\cite{Khatri2012,Campos}. Pathway topology
based methods are similar to FCS except that they consider the
topology of each gene during the gene-level
statistics but still are unable to consider a link between different pathways~\cite{Khatri2012}.

On these grounds we propose a novel approach to make use of the rich
pathway annotation resources available to gain additional functional
insights from basic DGEA. To this, we
start with the presupposition that expression of neighboring genes
within a functional pathway are not independent from each
other. Rather, they are often regulating each others expression or are
part of the same regulon~\cite{Michalak}. Second, we
understand that the categorization of links between genes into labeled
pathways is often an arbitrary one, given the extensive cross talk
between different pathways. Although this categories have proven to be
useful in many situations, they force a certain perspective onto the
interpretation of novel data. Based on this two principles, we aim to
find sub-modules within predefined networks which exhibits as a whole
significant differential expression changes. As input information on
functional links between genes, provided by e.g. KEGG or Reactome, and
information on the differential expression status of single genes,
resulting from a DGEA, are required. As a result the analysis returns
sub-modules and their joint confidence scores, reflecting how the
perturbation is migrated through the network. Furthermore, the entry
points of perturbation in the networks and overlap with conventional
pathway categories are returned. All of this can be helpful to
understand cause end effect of a stimulus and might inform about
potential points of intervention. The proposed algorithm was
implemented as a java program, which was named Modulated SubPath
Finder (\texttt{MSF}).

\section*{Methods}
\texttt{MSF} is developed as a novel heuristic approach in Java to
find the modulated sub-paths of gene interactions from the cell
signaling network. The network has nodes corresponding to genes and
edges representing their relationship. The pipeline of identification
of modulated sub-paths from a network by \texttt{MSF} is depicted in
Fig (1).

\texttt{MSF} uses the individual gene's \textit{p}-values generated
from the DGEA. The \textit{p}-value
expresses the probability that the hypothesis of unmodified gene
expression can be rejected for a given statistical model. To find
significantly modulated sub-paths individual \textit{p}-values of the
vicinal genes in the global network are combined into a single
combined \textit{p}-value, using a statistical method for combining
dependent \textit{p}-values described by~\cite{Hartung}. Using the
inverse normal method, individual gene \textit{p}-values $p_i$ are
first transformed to its corresponding normal score $t_i$,

\[ t_i = \Phi^{-1}(p_i) \]

Then using these normal scores, the correlation \textit{Co} between genes are calculated,

\[ Co (t_i, t_j) =\varrho \]

followed by correction of the correlation \textbf{k}. The normal
scores, correlations and the correction of correlation are applied to
the inverse normal function to calculate the individual
\textit{p}-value for a sub-path \textit{Cp}. \newline

$  {\scriptstyle Cp(\varrho,K) = \frac{\sum_{i=1}^{n} \lambda_i t_i}{\sqrt{\sum_{i=1}^{n} \lambda_i^2 +[(\sum_{i=1}^{n} \lambda_i)^2- \sum_{i=1}^{n} \lambda_i^2]\{\varrho+k\cdot\sqrt{\frac{2}{n+1}} (1-\varrho)}}} $\\ \newline

Lambda $\lambda$ are the weights for each gene, For the moment equal
weights are given to all genes and \textbf{k} used is 0.2 as used by authors. An
overall \textit{p}-value of a sub-path will express the significance
of all genes in the sub-path being modulated together.

To reduce the complexity to score all possible connected sub-modules
\texttt{MSF} applies a three steps heuristic as described in the
following.\newline

\subsection*{Overview of our method}

\textbf{Initial Modulated Sub-paths}

\texttt{MSF} constructs the first sub-path starting with the genes
associated with the lowest (most significant) \textit{p}-value from
the network. It tries to extend the
sub-path to the most significant neighboring gene. A single combined
\textit{p}-value is calculated for the two genes. If the combined
\textit{p}-value is smaller than the minimal individual gene's
\textit{p}-values, the extended sub-path is accepted. Then the next most significant neighboring gene is added to the sub-path and the combined \textit{p}-value is calculated, if the new combined p-value of three genes is smaller than the combined p-value of the first two genes,the extended sub-path is accepted. This step is iteratively repeated until no further extension is accepted. In this
case the process starts over with all remaining genes not yet in a
significant modulated sub-path. This step identifies all the trivial
sub-paths that are modulated in the whole network~(Fig.~1b).\newline

\textbf{Extending Modulated Sub-paths}

In the next step the initial modulated sub-paths are used to check if
they could further be extended beyond the immediate neighborhood.
This is done by testing all possible extension up to N genes for
all genes in the sub-path. Again, this step is iteratively repeated until no
further genes are added to the significant differentially expressed
sub-paths. This steps bridges small gaps of genes without a clear
differential signal in the DGEA~(Fig.~1c).\newline

\textbf{Merging Modulated Sub-paths}

After detection and extension of the modulated sub-paths, they are
tested if combined sub-path score is better than on their own. The most significant gene interaction of the first gene of a sub-path is taken and checked if it interacts with a gene in the second sub-path. If the two
paths merge with the connector of at most N genes and the
combined \textit{p}-value of the merged sub-path including the
bridging genes in between is less than the individual
\textit{p}-values of the two sub-paths, the two sub-paths are merged
together to one big modulated sub-module~(Fig.~1d). This step is repeated iteratively until no sub-paths could be merged to the sub-module.\newline

\textbf{Finding Sources \& Sinks}

In a last post processing step \texttt{MSF} identifies the trigger
points of the modulated sub-path. These trigger genes are the sources
of the sub-modules with only outgoing edges. These genes can be
interpreted as the entry points of perturbation from where the
stimulus causes downstream effects. In the same spirit the most
downstream genes of the modulated sub-path are identified and defined
as sinks. Sinks can be interpreted as the effectors where the
integrated information within the signal transduction network is set
to action.

\section*{Results}

\subsection*{Case Study}

To demonstrate the usefulness \texttt{MSF} is applied to a RNAseq data
set of primary human monocyte-derived macrophages (MDMs) infected with
Ebola virus~\cite{Olejnik} (GSE84188). Ebola Virus (EBOV)
belongs to the Filoviridea family; filamentous, enveloped and single
stranded RNA viruses. EBOV causes hemorrhagic fever in humans,
inducing the host innate and adaptive immune response to be unable to
control virus infection~\cite{Prins}. Until now there are no approved
antiviral drugs for the treatment of Ebola virus infection~\cite{Konde,Rhein}. 
The initial targets of EBOV are the macrophages and
dendritic immune cells~\cite{Falasca,Rhein}. Ebola Virus inhibits the critical
innate immune response of the host, which includes the activation of
alpha/beta interferon (IFN-$\alpha /
\beta$)~\cite{Prins,Konde,Cardenas}. The EBOV viral protein VP35
targets the host type~I interferon IFN-$\alpha / \beta$ to block the
early antiviral immune response
~\cite{Prins,Konde,Falasca,Cardenas,Olejnik}. It has been proposed
that IFN-$\alpha / \beta$ could be used through clinical trials to
design antiviral drug against Ebola. The aim of testing the EBOV infection data with \texttt{MSF} was to  identify the modulated sub-paths and check if \texttt{MSF} is able to identify the IFN-$\alpha / \beta$ gene as one of the sources.

The EBOV infection data has three time-points six hour (6hpi), one day (1dpi) and two day (2dpi) post infection. At 6hpi, five modulated sub-modules were identified. Most of the genes part of the sub-moldules were chemokines (CXCL10, CCL8) and Interleukin genes (IL6, IL27, IL23) which serve as a subset of CD4+ T
cells. IFNB1  and IFNA1 were both identified as two of the sources. Most of the sources identified by \texttt{MSF} were type~I interferon induced genes. At 1dpi nine modulated sub-modules were identified, IFNA1 was identified as one of the sources. For the last time-point again IFNB1 and IFNA1 were identified as the two sources out of the possible sources from seven sub-modules identified.

As stated earlier IFN-$\alpha / \beta$ could be one of the target
genes of Ebola infection. We were able to successfully identify IFNA1
as a source in all three time-points and INFB1 in two of the time-points. Although IFNA1 and IFNB1 were two of the most
significant gene in the later time points, \texttt{MSF} was able to detect them as
a source in the early time-point when the genes were not significant
based on the individual DGEA alone. These sources can help the biologist as the starting point of clinical testing for drugs and vaccines against an infection.

\subsection*{Benchmark}

To better understand the functional role of the sub-modules identified by  \texttt{MSF}, we linked the  sub-modules to the predefined KEGG pathways and Reactome
pathways through gene enrichment analysis. We used Cytoscape~\cite{Cyto} plugins to identify the enriched patwhays, CytoKegg~\cite{Cytokegg} for KEGG and Reactome FI~\cite{Reactome} for Reactome. We compared the pathways enriched in \texttt{MSF} identified sub-modules to SPIA~\cite{Tarca} for KEGG and Reactome pathway enrichment analysis.

At 6hpi Cytokegg identified the sub-module genes to be enriched in toll-like receptor signaling, TNF signaling, IL-17 signaling and NF-kappa signaling pathways, as also stated by~\cite{Olejnik}. The earliest response EBOV
induces on  cytokine is via TLR4-mediated signaling~\cite{Olejnik}, Gene enrichment analysis of sub-module genes at 6hpi showed toll-like receptor signaling being most significantly dis-regulated pathway, this important signaling pathway was not identified by SPIA along with TNF signaling and IL-17 signaling (Table 1). On the later time-points \texttt{MSF} and SPIA showed consensus on the important pathways e.g Chemokine signaling, Cytokine-Cytokine receptor interaction, NF-kappa B signaling ,Cytosolic DNA-sensing  and RIG-I-like receptor signaling as being dis-regulated supplement (Table 3,4). SPIA identifies Cytokine-Cytokine receptor interaction and Chemokine signaling on the top for all the time
points and also shows apoptosis at the earliest time-point, where infection
just started.

Likewise we compared gene enrichment results of sub-module genes identified by \texttt{MSF} using Reactome FI to the Reactome pathway enrichment analysis results. Although MSF and Reactome show good agreement on the dis-regulated pathways e.g Signaling by Interleukins, Interleukin-10 signaling, Cytokine Signaling in Immune system, Chemokine receptors bind chemokines, RIG-I/MDA5 mediated induction of IFN-alpha/beta pathways, NF-kB activation, Reactome as well fails to identify the Toll-like recpetor signaling. On the other hand, \texttt{MSF} could not identify the dis-regulation of phases of cell cycle. This shows that \texttt{MSF} is able to identify the modulated genes in the sub-modules at the earliest time-point of infection even though when the signal transduction was weak.

\subsection*{Robustness}

To show the robustness and stability of our method, Poisson
distributed noise was added to the sample read counts. 
Then DGEA was carried out on the disturbed data with the
same parameters as for the native data, followed by analysis with
\texttt{MSF}. Using this criteria, the goal was to spot almost
identical modulated sub-paths in the real and noisy data.

Simulation was also conducted on \texttt{MSF} to test the results
obtained from real data, showing a chance of less than 10 percent to
get more sources than observed.

\section*{Discussion}

The discussion should include the implications of the article results
in view of prior work in this field.

\section*{Conclusions}

Please state what you think are the main conclusions that can be
realistically drawn from the findings in the paper, taking care not to
make claims that cannot be supported.

\subsection*{Availability of data and material}

The Ebola infection dataset analyzed during the current study are
available in the GEO repository, {\scriptsize
  https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE84188.}  The
network file used is from Reactome Functional interactions (FIs)
Version 2016

\subsection*{Author contributions}
In order to give appropriate credit to each author of an article, the individual
contributions of each author to the manuscript should be detailed in this section. We
recommend using author initials and then stating briefly how they contributed.

\subsection*{Competing interests}

The authors declare that they have no competing interests.

\subsection*{Grant information}
Please state who funded the work discussed in this article, whether it is your employer,
a grant funder etc. Please do not list funding that you have that is not relevant to this
specific piece of research. For each funder, please state the funder’s name, the grant
number where applicable, and the individual to whom the grant was assigned.
If your work was not funded by any grants, please include the line: ‘The author(s)
declared that no grants were involved in supporting this work.’

\subsection*{Acknowledgements}
This section should acknowledge anyone who contributed to the research or the
article but who does not qualify as an author based on the criteria provided earlier
(e.g. someone or an organisation that provided writing assistance). Please state how
they contributed; authors should obtain permission to acknowledge from all those
mentioned in the Acknowledgements section.

\begin{thebibliography}{9}

	\bibitem{DEG}
	Malone, J. H., \& Oliver, B. (2011). Microarrays, deep sequencing and the true measure of the transcriptome. BMC Biology, 9(1), 34. doi:10.1186/1741-7007-9-34
	\bibitem{Khatri2012}
	Khatri, P., Sirota, M., \& Butte, A. J. (2012). Ten Years of Pathway Analysis: Current Approaches and Outstanding Challenges. PLoS Computational Biology, 8(2). doi:10.1371/journal.pcbi.1002375
	\bibitem{Kegg}
	Kanehisa, M. (2000). KEGG: Kyoto Encyclopedia of Genes and Genomes. Nucleic Acids Research, 28(1), 27-30. doi:10.1093/nar/28.1.27
	\bibitem{Reactome}
	Croft, D., \& Croft, D. (2010). Reactome: A database of biological pathways. Nature Precedings. doi:10.1038/npre.2010.5025
	\bibitem{Campos}
	García-Campos, M. A., Espinal-Enríquez, J., \& Hernández-Lemus, E. (2015). Pathway Analysis: State of the Art. Frontiers in Physiology, 6. doi:10.3389/fphys.2015.00383 
	\bibitem{Bayer}
	Bayerlová, M., Jung, K., Kramer, F., Klemm, F., Bleckmann, A., \& Beißbarth, T. (2015). Comparative study on gene set and pathway topology-based enrichment methods. BMC Bioinformatics, 16(1). doi:10.1186/s12859-015-0751-5   
	\bibitem{Hartung}
	Hartung, J. (1999). A Note on Combining Dependent Tests of Significance. Biometrical Journal, 41(7), 849-855. doi:10.1002/(sici)1521-4036(199911)41:7<849::aid-bimj849>3.0.co;2-t 5

	\bibitem {Prins}
	Prins, K., Cardenas, W. and Basler, C. (2009). Ebola Virus Protein VP35 Impairs the Function of Interferon Regulatory Factor-Activating Kinases IKK and TBK-1. Journal of Virology, 83(7), pp.3069-3077.

	\bibitem {Konde}
	Konde, M., Baker, D., Traore, F., Sow, M., Camara, A., Barry, A., Mara, D., Barry, A., Cone, M., Kaba, I., Richard, A., Beavogui, A., Günther, S., Pintilie, M. and Fish, E. (2017). Interferon β-1a for the treatment of Ebola virus disease: A historically controlled, single-arm proof-of-concept trial. PLOS ONE, 12(2).

	\bibitem {Rhein}
	Rhein, B., Powers, L., Rogers, K., Anantpadma, M., Singh, B., Sakurai, Y., Bair, T., Miller-Hunt, C., Sinn, P., Davey, R., Monick, M. and Maury, W. (2015). Interferon-γ Inhibits Ebola Virus Infection. PLOS Pathogens, 11(11), p.e1005263.

	\bibitem {Falasca}
	Falasca, L., Agrati, C., Petrosillo, N., Di Caro, A., Capobianchi, M., Ippolito, G. and Piacentini, M. (2015).
	 Molecular mechanisms of Ebola virus pathogenesis: focus on cell death. Cell Death and Differentiation, 22(8), pp.1250-1259.

	\bibitem {Cardenas}
	Cardenas, W., Loo, Y., Gale, M., Hartman, A., Kimberlin, C., Martinez-Sobrido, L., Saphire, E. and Basler, C. (2006). Ebola Virus VP35 Protein Binds Double-Stranded RNA and Inhibits Alpha/Beta Interferon Production Induced by RIG-I Signaling. Journal of Virology, 80(11), pp.5168-5178.

	\bibitem {Olejnik}
	Olejnik, J., Forero, A., Deflubé, L., Hume, A., Manhart, W., Nishida, A., Marzi, A., Katze, M., Ebihara, H., Rasmussen, A. and Mühlberger, E. (2017). Ebolaviruses Associated with Differential Pathogenicity Induce Distinct Host Responses in Human Macrophages. Journal of Virology, 91(11), pp.e00179-17.

	\bibitem {Dilley}
	Dilley, K., Voorhies, A., Luthra, P., Puri, V., Stockwell, T., Lorenzi, H., Basler, C. and Shabman, R. (2017). The Ebola virus VP35 protein binds viral immunostimulatory and host RNAs identified through deep sequencing. PLOS ONE, 12(6), p.e0178717.

	\bibitem{Michalak}
	Michalak, P. (2008). Coexpression, coregulation, and cofunctionality of neighboring genes in eukaryotic genomes. Genomics, 91(3), 243-248. doi:10.1016/j.ygeno.2007.11.002

	\bibitem{Cyto}
	http://www.cytoscape.org/

	\bibitem{Cytokegg}
	http://apps.cytoscape.org/apps/cytokegg

	\bibitem{Reactome}
	https://wiki.reactome.org/index.php?title=ReactomeFIViz

	\bibitem{Tarca}
	Tarca, A. L., Draghici, S., Khatri, P., Hassan, S. S., Mittal, P., Kim, J., … Romero, R. (2009). A novel signaling pathway impact analysis. Bioinformatics, 25(1), 75–82. http://doi.org/10.1093/bioinformatics/btn577

\end{thebibliography}

\begin{figure*}[ht!]

	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/Tgfb1} 
		\caption{}
	\end{subfigure}
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep1}
		\caption{}
	\end{subfigure}
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep2} 
		\caption{}
	\end{subfigure}
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep3}
		\caption{}
	\end{subfigure}
	\caption{Overview of \texttt{MSF} process showing the steps to identify the modulated sub-paths. (a) showing a network of genes with green nodes as genes with \textit{p}-values $< $ 0.05, and blue nodes are genes with \textit{p}-values $>$ 0.05, (b) the genes circled red show the two initial modulated sub-paths, sub-path1 found one with M,K,C and the second modulated sub-path2 S,V,T, (c) shows the modulated sub-path1 being extended to M,K,C,A,D, (d) shows both the modulated sub=paths merged with the addition of 3 genes Z,W,P.}
\end{figure*}

\begin{table*}[]
	\centering
	\caption{Comparison of gene enrichment analysis of MSF 6hpi sub-modules with SPIA pathway analysis}
	\label{Comparison of gene enrichment analysis of MSF 6hpi sub-modules with SPIA pathway analysis}
	\begin{tabular}{|l|l|}
		\hline
		\textbf{MSF Sub-module Gene Enrichment}              & \textbf{SPIA Pathway Analysis}            \\ \hline
		Toll-like receptor signaling pathway                 & Cytokine-cytokine receptor interaction    \\ \hline
		TNF signaling pathway                                & Chemokine signaling pathway               \\ \hline
		Prion diseases                                       & Apoptosis                                 \\ \hline
		Pertussis                                            & Taste transduction                        \\ \hline
		Type II diabetes mellitus                            & Influenza A                               \\ \hline
		Thyroid cancer                                       & NF-kappa B signaling pathway              \\ \hline
		Toxoplasmosis                                        & Type II diabetes mellitus                 \\ \hline
		Prolactin signaling pathway                          & Osteoclast differentiation                \\ \hline
		Cytosolic DNA-sensing pathway                        & Cytosolic DNA-sensing pathway             \\ \hline
		Chagas disease (American trypanosomiasis)            & Natural killer cell mediated cytotoxicity \\ \hline
		RIG-I-like receptor signaling pathway                & Measles                                   \\ \hline
		Th1 and Th2 cell differentiation                     & Amyotrophic lateral sclerosis (ALS)       \\ \hline
		Chemokine signaling pathway                          & Salivary secretion                        \\ \hline
		IL-17 signaling pathway                              & HTLV-I infection                          \\ \hline
		AGE-RAGE signaling pathway in diabetic complications & MAPK signaling pathway                    \\ \hline
		Measles                                              & Calcium signaling pathway                 \\ \hline
		Jak-STAT signaling pathway                           & Notch signaling pathway                   \\ \hline
		Leishmaniasis                                        & Graft-versus-host disease                 \\ \hline
		Endometrial cancer                                   & Maturity onset diabetes of the young      \\ \hline
		ECM-receptor interaction                             & Herpes simplex infection                  \\ \hline

	\end{tabular}
\end{table*}

\begin{table*}[]
	\centering
	\caption{Comparison of gene enrichment analysis of MSF 6hpi sub-modules with Reactome pathway enrichment analysis}
	\label{my-label}
	\begin{tabular}{|l|l|}
		\hline
		\textbf{MSF Sub-module Gene Enrichment}          & \textbf{Reactome Pathway Enrichment Analysis}   \\ \hline
		Signaling by Interleukins                        & Interferon alpha/beta signaling                 \\ \hline
		Interleukin-10 signaling                         & Interferon Signaling                            \\ \hline
		Cytokine Signaling in Immune system              & Interleukin-10 signaling                        \\ \hline
		Interleukin-6 family signaling                   & Cytokine Signaling in Immune system             \\ \hline
		Growth hormone receptor signaling                & Signaling by Interleukins                       \\ \hline
		Chemokine receptors bind chemokines              & Interferon gamma signaling                      \\ \hline
		Interleukin-4 and 13 signaling                   & Interleukin-4 and 13 signaling                  \\ \hline
		Signalling by NGF                                & Chemokine receptors bind chemokines             \\ \hline
		IL-6-type cytokine receptor ligand interactions  & RIG-I/MDA5 mediated induction of IFN-alpha/beta \\ \hline
		GPVI-mediated activation cascade                 & Negative regulators of RIG-I/MDA5 signaling     \\ \hline
		MyD88 dependent cascade initiated on endosome    & Nucleotide-binding domain                       \\ \hline
		Toll Like Receptor 7/8 (TLR7/8) Cascade          & NF-kB activation through FADD/RIP-1 pathway     \\ \hline
		Activated TLR4 signalling                        & Ovarian tumor domain proteases                  \\ \hline
		CD28 dependent PI3K/Akt signaling                & GPCR ligand binding                             \\ \hline
		Toll Like Receptor 9 (TLR9) Cascade              & Interleukin-1 processing                        \\ \hline
		CD28 co-stimulation                              & TRAF3-dependent IRF activation pathway          \\ \hline
		NGF signalling via TRKA from the plasma membrane & Interleukin-6 family signaling                  \\ \hline
		Toll Like Receptor 3 (TLR3) Cascade              & Class A/1 (Rhodopsin-like receptors)            \\ \hline
		TRIF-mediated TLR3/TLR4 signaling                & Inflammasomes                                   \\ \hline
		MyD88-independent TLR3/TLR4 cascade              & Interleukin-1 signaling                         \\ \hline
	\end{tabular}
\end{table*}

\begin{figure*}[th]
	\includegraphics[width=16cm]{AllDEGSources}
	\caption{}
	\label{fig:alldegsources}
\end{figure*}

\begin{figure*}[p]
	\centering
	\includegraphics[width=14cm]{"SubPath Robustness 6H"}
\caption{}
\label{fig:subpath-robustness-6h}
\includegraphics[width=14cm]{Betweeness}
\caption{}
\label{fig:betweeness}
\end{figure*}

\end{document}

### 2018-01-16 MSF_Paper_FA.tex
changing path in includefig to make it pdflatex-ing
Update MSF_Paper_FA.tex

@@
 Likewise we compared gene enrichment results of sub-module genes identified by \texttt{MSF} using Reactome FI to the Reactome pathway enrichment analysis results.
-Although MSF and Reactome show good agreement on the dis-regulated pathways e.g Signaling by Interleukins, Interleukin-10 signaling, Cytokine Signaling in Immune system, Chemokine receptors bind chemokines, RIG-I/MDA5 mediated induction of IFN-alpha/beta pathways, NF-kB activation, Reactome as well fails to identify the Toll-like recpetor signaling.
+Although MSF and Reactome show good agreement on the dis-regulated pathways e.g Signaling by Interleukins, Interleukin-10 signaling, Cytokine Signaling in Immune system, Chemokine receptors bind chemokines, RIG-I/MDA5 mediated induction of IFN-alpha/beta pathways, NF-kB activation, Reactome as well fails to identify the Toll-like receptor signaling.
 On the other hand, \texttt{MSF} could not identify the dis-regulation of phases of cell cycle.
@@
 identical modulated sub-paths in the real and noisy data.
+This procedure was carried out 100 times and every time the genes from the sub-modules identified from noisy data were compared to the sub-modules identified from the native data.
+The robustness of MSF for the three time-points is shown in~(Fig.~2).
+For the last time-point (2dpi) 84\% genes of the sub-modules from native data were retrieved, for 1dpi the robustness was about 79\%.
+The robustness was the lowest at 6hpi nearly 71\%, this could be due to that the infection merely started and the signal transdcution was weak as compared to the later time-points and over that Poisson distributed noise was added to the data.
 
@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/Tgfb1} 
+		\includegraphics[width=10.0cm]{Tgfb1.png} 
 		\caption{}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep1}
+		\includegraphics[width=10.0cm]{TGFBstep1.png}
 		\caption{}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep2} 
+		\includegraphics[width=10.0cm]{TGFBstep2.png} 
 		\caption{}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep3}
+		\includegraphics[width=10.0cm]{TGFBstep3.png}
 		\caption{}

### 2018-02-20 MSF_Paper_FA.tex
nothing
Added 6hpi network

--- 2018-01-16 MSF_Paper_FA.tex
@@
 
-\title{\textit{MSF: Modulated Sub-path Finder} }
+\title{\textit{MSF: Modulated Sub-graph Finder} }
 
@@
 novel approach to find the maximal significantly dis-regulated
-sub-paths or cluster of genes from the cell signaling network, giving
-these sub-paths an overall significance of modulation by combining the
-individual p-values of the genes.
+sub-graphs or cluster of genes from the cell signaling network, giving
+these sub-graphs an overall significance of modulation by combining the
+individual \textit{p}-values of the genes.
 \end{abstract}
@@
 
-Differentially expressed genes, Pathway analysis, Combining \textit{p}-value, Cell signaling network, Modulated Sub-paths
+Differentially expressed genes, Pathway analysis, Combining \textit{p}-value, Cell signaling network, Modulated Sub-graphs
 
@@
 analysis. It uses a user defined cut-off for the log-fold change and
-p-value from the DGEA (most commonly using absolute log-fold change
+\textit{p}-value from the DGEA (most commonly using absolute log-fold change
 $\geq$ 2 and \textit{p}-value $\leq$ 0.05) to define a list of
@@
 interpretation of novel data. Based on this two principles, we aim to
-find sub-modules within predefined networks which exhibits as a whole
+find sub-graphs within predefined networks which exhibits as a whole
 significant differential expression changes. As input information on
@@
 resulting from a DGEA, are required. As a result the analysis returns
-sub-modules and their joint confidence scores, reflecting how the
+sub-graphs and their joint confidence scores, reflecting how the
 perturbation is migrated through the network. Furthermore, the entry
@@
 potential points of intervention. The proposed algorithm was
-implemented as a java program, which was named Modulated SubPath
+implemented as a java program, which was named Modulated Sub-graph
 Finder (\texttt{MSF}).
@@
 \texttt{MSF} is developed as a novel heuristic approach in Java to
-find the modulated sub-paths of gene interactions from the cell
-signaling network. The network has nodes corresponding to genes and
+find the modulated sub-graphs of gene interactions from the cell
+signaling network. \texttt{MSF} does not work on predefined pathways but builds up the sub-graph from scratch using the whole cell signaling network. The network has nodes corresponding to genes and
 edges representing their relationship. The pipeline of identification
-of modulated sub-paths from a network by \texttt{MSF} is depicted in
+of modulated sub-graphs from a network by \texttt{MSF} is depicted in
 Fig (1).
@@
 expression can be rejected for a given statistical model. To find
-significantly modulated sub-paths individual \textit{p}-values of the
+significantly modulated sub-graphs individual \textit{p}-values of the
 vicinal genes in the global network are combined into a single
@@
 the inverse normal function to calculate the individual
-\textit{p}-value for a sub-path \textit{Cp}. \newline
+\textit{p}-value for a sub-graph \textit{Cp}. \newline
 
@@
 weights are given to all genes and \textbf{k} used is 0.2 as used by authors. An
-overall \textit{p}-value of a sub-path will express the significance
-of all genes in the sub-path being modulated together.
+overall \textit{p}-value of a sub-graph will express the significance
+of all genes in the sub-graph being modulated together.
 
-To reduce the complexity to score all possible connected sub-modules
+To reduce the complexity to score all possible connected sub-graphs
 \texttt{MSF} applies a three steps heuristic as described in the
@@
 
-\textbf{Initial Modulated Sub-paths}
+\textbf{Initial Modulated sub-graphs}
 
-\texttt{MSF} constructs the first sub-path starting with the genes
+\texttt{MSF} constructs the first sub-graph starting with the genes
 associated with the lowest (most significant) \textit{p}-value from
 the network. It tries to extend the
-sub-path to the most significant neighboring gene. A single combined
+sub-graph to the most significant neighboring gene. A single combined
 \textit{p}-value is calculated for the two genes. If the combined
 \textit{p}-value is smaller than the minimal individual gene's
-\textit{p}-values, the extended sub-path is accepted.
-Then the next most significant neighboring gene is added to the sub-path and the combined \textit{p}-value is calculated, if the new combined p-value of three genes is smaller than the combined p-value of the first two genes,the extended sub-path is accepted.
+\textit{p}-value, the extended sub-graph is accepted.
+Then the next most significant neighboring gene is added to the sub-graph and the combined \textit{p}-value is calculated, if the new combined \textit{p}-value of three genes is smaller than the combined \textit{p}-value of the first two genes,the extended sub-graph is accepted.
 This step is iteratively repeated until no further extension is accepted.
-In this
-case the process starts over with all remaining genes not yet in a
-significant modulated sub-path. This step identifies all the trivial
-sub-paths that are modulated in the whole network~(Fig.~1b).\newline
+In this case the process starts over with all remaining genes not yet in a
+significant modulated sub-graph. This step identifies all the trivial
+sub-graphs that are modulated in the whole network~(Fig.~1b).\newline
 
-\textbf{Extending Modulated Sub-paths}
+\textbf{Extending Modulated sub-graphs}
 
-In the next step the initial modulated sub-paths are used to check if
+In the next step initial modulated sub-graphs are used to check if
 they could further be extended beyond the immediate neighborhood.
-This is done by testing all possible extension up to N genes for
-all genes in the sub-path. Again, this step is iteratively repeated until no
+This is done by testing all possible extension up to 2 genes for
+all genes in the sub-graph. Again, this step is iteratively repeated until no
 further genes are added to the significant differentially expressed
-sub-paths. This steps bridges small gaps of genes without a clear
+sub-graphs. This steps bridges small gaps of genes without a clear
 differential signal in the DGEA~(Fig.~1c).\newline
 
-\textbf{Merging Modulated Sub-paths}
+\textbf{Merging Modulated sub-graphs}
 
-After detection and extension of the modulated sub-paths, they are
-tested if combined sub-path score is better than on their own.
-The most significant gene interaction of the first gene of a sub-path is taken and checked if it interacts with a gene in the second sub-path.
-If the two
-paths merge with the connector of at most N genes and the
-combined \textit{p}-value of the merged sub-path including the
+After detection and extension of the modulated sub-graphs, they are
+tested if combined sub-graph score is better than on their own.
+The merging of the two sub-graphs is done by depth first search, If the two sub-graphs merge with the connector of at most 2 genes and the
+combined \textit{p}-value of the merged sub-graph including the
 bridging genes in between is less than the individual
-\textit{p}-values of the two sub-paths, the two sub-paths are merged
-together to one big modulated sub-module~(Fig.~1d). This step is repeated iteratively until no sub-paths could be merged to the sub-module.\newline
+\textit{p}-values of the two sub-graphs, the two sub-graphs are merged
+together to one big modulated sub-graph~(Fig.~1d). This step is repeated iteratively until no sub-graphs could be merged.\newline
 
@@
 In a last post processing step \texttt{MSF} identifies the trigger
-points of the modulated sub-path. These trigger genes are the sources
-of the sub-modules with only outgoing edges. These genes can be
-interpreted as the entry points of perturbation from where the
+points of the modulated sub-graph. These trigger genes are the sources
+of the sub-graphs with only outgoing edges. These genes can be
+interpreted as the possible entry points of perturbation from where the
 stimulus causes downstream effects. In the same spirit the most
-downstream genes of the modulated sub-path are identified and defined
+downstream genes of the modulated sub-graph are identified and defined
 as sinks. Sinks can be interpreted as the effectors where the
@@
 design antiviral drug against Ebola.
-The aim of testing the EBOV infection data with \texttt{MSF} was to  identify the modulated sub-paths and check if \texttt{MSF} is able to identify the IFN-$\alpha / \beta$ gene as one of the sources.
+The aim of testing the EBOV infection data with \texttt{MSF} was to  identify the modulated sub-graphs and check if \texttt{MSF} is able to identify the IFN-$\alpha / \beta$ gene as one of the sources.
 
 The EBOV infection data has three time-points six hour (6hpi), one day (1dpi) and two day (2dpi) post infection.
-At 6hpi, five modulated sub-modules were identified.
+At 6hpi, five modulated sub-graphs were identified.
 Most of the genes part of the sub-moldules were chemokines (CXCL10, CCL8) and Interleukin genes (IL6, IL27, IL23) which serve as a subset of CD4+ T
@@
 Most of the sources identified by \texttt{MSF} were type~I interferon induced genes.
-At 1dpi nine modulated sub-modules were identified, IFNA1 was identified as one of the sources.
-For the last time-point again IFNB1 and IFNA1 were identified as the two sources out of the possible sources from seven sub-modules identified.
+At 1dpi nine modulated sub-graphs were identified, IFNA1 was identified as one of the sources.
+For the last time-point again IFNB1 and IFNA1 were identified as the two sources out of the possible sources from seven sub-graphs identified.
 
@@
 a source in the early time-point when the genes were not significant
-based on the individual DGEA alone. These sources can help the biologist as the starting point of clinical testing for drugs and vaccines against an infection.
+based on the individual DGEA alone.
+These sources will reduce the search space for potential target gene and can help the biologist as the starting point of clinical testing for drugs and vaccines against an infection.
 
-\subsection*{Benchmark}
+\subsection*{Modulated sub-graphs at 6hpi}
 
-To better understand the functional role of the sub-modules identified by  \texttt{MSF}, we linked the  sub-modules to the predefined KEGG pathways and Reactome
+The connected modulated sub-graphs from 6hpi are shown in~(Fig.~2).
+The sub-graph shows different colors for the predefined  KEGG pathways and how they are connected to each other.
+The RIG-like receptor signaling in connected to Toll-like receptor, which intern connects to MAPK signaling, JAK-STAT signalling, CYtokine-cytokine signaling, Chemokine receptor signaling and TNF signaling.
+These pathways could be seen further leading to WNT Signaling, P13K-Akt signaling.
+
+From the other side the ECM receptor interaction is seen to lead to Focal adhesion which regulates the P13k-Akt signaling.
+Finally P13k-Akt signaling leads to FoxO signaling which eventually connects to cell cycle.
+TNF, P13k-Akt and FoxO signaling activates apoptosis and autophagy.
+
+\subsection*{Comparison}
+
+To better understand the functional role of the sub-graphs identified by \texttt{MSF}, we linked the  sub-graphs to the predefined KEGG pathways and Reactome
 pathways through gene enrichment analysis.
 We used Cytoscape~\cite{Cyto} plugins to identify the enriched patwhays, CytoKegg~\cite{Cytokegg} for KEGG and Reactome FI~\cite{Reactome} for Reactome.
-We compared the pathways enriched in \texttt{MSF} identified sub-modules to SPIA~\cite{Tarca} for KEGG and Reactome pathway enrichment analysis.
+We compared the pathways enriched in \texttt{MSF} identified sub-graphs to SPIA~\cite{Tarca} for KEGG and Reactome pathway enrichment analysis.
 
-At 6hpi Cytokegg identified the sub-module genes to be enriched in toll-like receptor signaling, TNF signaling, IL-17 signaling and NF-kappa signaling pathways, as also stated by~\cite{Olejnik}.
-The earliest response EBOV
-induces on  cytokine is via TLR4-mediated signaling~\cite{Olejnik}, Gene enrichment analysis of sub-module genes at 6hpi showed toll-like receptor signaling being most significantly dis-regulated pathway, this important signaling pathway was not identified by SPIA along with TNF signaling and IL-17 signaling (Table 1).
-On the later time-points \texttt{MSF} and SPIA showed consensus on the important pathways e.g Chemokine signaling, Cytokine-Cytokine receptor interaction, NF-kappa B signaling ,Cytosolic DNA-sensing  and RIG-I-like receptor signaling as being dis-regulated supplement (Table 3,4).
-SPIA identifies Cytokine-Cytokine receptor interaction and Chemokine signaling on the top for all the time
-points and also shows apoptosis at the earliest time-point, where infection
-just started.
+At 6hpi Cytokegg identified the sub-graph genes to be enriched in Toll-like receptor signaling, TNF signaling, IL-17 signaling and NF-kappa signaling pathways, as also stated by~\cite{Olejnik}.
+The earliest response EBOV induces on  cytokine is via TLR4-mediated signaling~\cite{Olejnik}, Gene enrichment analysis of sub-graph genes at 6hpi showed toll-like receptor signaling being most significantly dis-regulated pathway, this important signaling pathway was not identified by SPIA along with RIG-I-like receptor signaling pathway, TNF signaling pathway and IL-17 signaling (Table 1).
+On the later time-points \texttt{MSF} and SPIA showed consensus on the important pathways e.g Chemokine signaling, Cytokine-Cytokine receptor interaction, NF-kappa B signaling , Cytosolic DNA-sensing and RIG-I-like receptor signaling as being dis-regulated supplement (Table 3,4).
 
-Likewise we compared gene enrichment results of sub-module genes identified by \texttt{MSF} using Reactome FI to the Reactome pathway enrichment analysis results.
-Although MSF and Reactome show good agreement on the dis-regulated pathways e.g Signaling by Interleukins, Interleukin-10 signaling, Cytokine Signaling in Immune system, Chemokine receptors bind chemokines, RIG-I/MDA5 mediated induction of IFN-alpha/beta pathways, NF-kB activation, Reactome as well fails to identify the Toll-like recpetor signaling.
-On the other hand, \texttt{MSF} could not identify the dis-regulation of phases of cell cycle.
-This shows that \texttt{MSF} is able to identify the modulated genes in the sub-modules at the earliest time-point of infection even though when the signal transduction was weak.
+Likewise we compared gene enrichment results of sub-graph genes identified by \texttt{MSF} using Reactome FI to the Reactome predefined pathway enrichment analysis results.
+Although \texttt{MSF} and Reactome show good agreement on the dis-regulated pathways e.g Signaling by Interleukins, Interleukin-10 signaling, Cytokine Signaling in Immune system, Chemokine receptors bind chemokines, RIG-I/MDA5 mediated induction of IFN-alpha/beta pathways, NF-kB activation, Reactome as well fails to identify the Toll-like receptor signaling.
+This shows that \texttt{MSF} is able to identify the modulated genes in the sub-graphs at the earliest time-point of infection even though when the signal transduction was weak.
 
@@
 \texttt{MSF}. Using this criteria, the goal was to spot almost
-identical modulated sub-paths in the real and noisy data.
+identical modulated sub-graphs in the real and noisy data.
+This procedure was carried out 100 times and every time the genes from the sub-graphs identified from noisy data were compared to the sub-graphs identified from the native data.
+The robustness of \texttt{MSF} for the three time-points is shown in~(Fig.~2).
+For the last time-point (2dpi) 84\% genes of the sub-graphs from native data were retrieved, for 1dpi the robustness was about 79\%.
+The robustness was the lowest at 6hpi nearly 71\%, this could be due to that the infection merely started and the signal transdcution was weak as compared to the later time-points and over that Poisson distributed noise was added to the data.
 
@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/Tgfb1} 
+		\includegraphics[width=10.0cm]{Git/MSF_Paper/Tgfb1} 
 		\caption{}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep1}
+		\includegraphics[width=10.0cm]{Git/MSF_Paper/TGFBstep1}
 		\caption{}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep2} 
+		\includegraphics[width=10.0cm]{Git/MSF_Paper/TGFBstep2} 
 		\caption{}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep3}
+		\includegraphics[width=10.0cm]{Git/MSF_Paper/GFBstep3}
 		\caption{}
 	\end{subfigure}
-	\caption{Overview of \texttt{MSF} process showing the steps to identify the modulated sub-paths.
-(a) showing a network of genes with green nodes as genes with \textit{p}-values $< $ 0.05, and blue nodes are genes with \textit{p}-values $>$ 0.05, (b) the genes circled red show the two initial modulated sub-paths, sub-path1 found one with M,K,C and the second modulated sub-path2 S,V,T, (c) shows the modulated sub-path1 being extended to M,K,C,A,D, (d) shows both the modulated sub=paths merged with the addition of 3 genes Z,W,P.}
+	\caption{Overview of \texttt{MSF} process showing the steps to identify the modulated sub-graphs.
+(a) showing a network of genes with green nodes as genes with \textit{p}-values $< $ 0.05, and blue nodes are genes with \textit{p}-values $>$ 0.05, (b) the genes circled red show the two initial modulated sub-graphs, sub-graph1 found one with M,K,C and the second modulated sub-graph2 S,V,T, (c) shows the modulated sub-graph1 being extended to M,K,C,A,D, (d) shows both the modulated sub=paths merged with the addition of 3 genes Z,W,P.}
 \end{figure*}
 
-\begin{table*}[]
+\begin{figure*}
 	\centering
-	\caption{Comparison of gene enrichment analysis of MSF 6hpi sub-modules with SPIA pathway analysis}
-	\label{Comparison of gene enrichment analysis of MSF 6hpi sub-modules with SPIA pathway analysis}
-	\begin{tabular}{|l|l|}
-		\hline
-		\textbf{MSF Sub-module Gene Enrichment}              & \textbf{SPIA Pathway Analysis}            \\ \hline
-		Toll-like receptor signaling pathway                 & Cytokine-cytokine receptor interaction    \\ \hline
-		TNF signaling pathway                                & Chemokine signaling pathway               \\ \hline
-		Prion diseases                                       & Apoptosis                                 \\ \hline
-		Pertussis                                            & Taste transduction                        \\ \hline
-		Type II diabetes mellitus                            & Influenza A                               \\ \hline
-		Thyroid cancer                                       & NF-kappa B signaling pathway              \\ \hline
-		Toxoplasmosis                                        & Type II diabetes mellitus                 \\ \hline
-		Prolactin signaling pathway                          & Osteoclast differentiation                \\ \hline
-		Cytosolic DNA-sensing pathway                        & Cytosolic DNA-sensing pathway             \\ \hline
-		Chagas disease (American trypanosomiasis)            & Natural killer cell mediated cytotoxicity \\ \hline
-		RIG-I-like receptor signaling pathway                & Measles                                   \\ \hline
-		Th1 and Th2 cell differentiation                     & Amyotrophic lateral sclerosis (ALS)       \\ \hline
-		Chemokine signaling pathway                          & Salivary secretion                        \\ \hline
-		IL-17 signaling pathway                              & HTLV-I infection                          \\ \hline
-		AGE-RAGE signaling pathway in diabetic complications & MAPK signaling pathway                    \\ \hline
-		Measles                                              & Calcium signaling pathway                 \\ \hline
-		Jak-STAT signaling pathway                           & Notch signaling pathway                   \\ \hline
-		Leishmaniasis                                        & Graft-versus-host disease                 \\ \hline
-		Endometrial cancer                                   & Maturity onset diabetes of the young      \\ \hline
-		ECM-receptor interaction                             & Herpes simplex infection                  \\ \hline
-
-	\end{tabular}
-\end{table*}
-
-\begin{table*}[]
-	\centering
-	\caption{Comparison of gene enrichment analysis of MSF 6hpi sub-modules with Reactome pathway enrichment analysis}
-	\label{my-label}
-	\begin{tabular}{|l|l|}
-		\hline
-		\textbf{MSF Sub-module Gene Enrichment}          & \textbf{Reactome Pathway Enrichment Analysis}   \\ \hline
-		Signaling by Interleukins                        & Interferon alpha/beta signaling                 \\ \hline
-		Interleukin-10 signaling                         & Interferon Signaling                            \\ \hline
-		Cytokine Signaling in Immune system              & Interleukin-10 signaling                        \\ \hline
-		Interleukin-6 family signaling                   & Cytokine Signaling in Immune system             \\ \hline
-		Growth hormone receptor signaling                & Signaling by Interleukins                       \\ \hline
-		Chemokine receptors bind chemokines              & Interferon gamma signaling                      \\ \hline
-		Interleukin-4 and 13 signaling                   & Interleukin-4 and 13 signaling                  \\ \hline
-		Signalling by NGF                                & Chemokine receptors bind chemokines             \\ \hline
-		IL-6-type cytokine receptor ligand interactions  & RIG-I/MDA5 mediated induction of IFN-alpha/beta \\ \hline
-		GPVI-mediated activation cascade                 & Negative regulators of RIG-I/MDA5 signaling     \\ \hline
-		MyD88 dependent cascade initiated on endosome    & Nucleotide-binding domain                       \\ \hline
-		Toll Like Receptor 7/8 (TLR7/8) Cascade          & NF-kB activation through FADD/RIP-1 pathway     \\ \hline
-		Activated TLR4 signalling                        & Ovarian tumor domain proteases                  \\ \hline
-		CD28 dependent PI3K/Akt signaling                & GPCR ligand binding                             \\ \hline
-		Toll Like Receptor 9 (TLR9) Cascade              & Interleukin-1 processing                        \\ \hline
-		CD28 co-stimulation                              & TRAF3-dependent IRF activation pathway          \\ \hline
-		NGF signalling via TRKA from the plasma membrane & Interleukin-6 family signaling                  \\ \hline
-		Toll Like Receptor 3 (TLR3) Cascade              & Class A/1 (Rhodopsin-like receptors)            \\ \hline
-		TRIF-mediated TLR3/TLR4 signaling                & Inflammasomes                                   \\ \hline
-		MyD88-independent TLR3/TLR4 cascade              & Interleukin-1 signaling                         \\ \hline
-	\end{tabular}
-\end{table*}
-
-\begin{figure*}[th]
-	\includegraphics[width=16cm]{AllDEGSources}
-	\caption{}
-	\label{fig:alldegsources}
+	\includegraphics[width=1.7\linewidth]{Git/MSF_Paper/Merged_Network}
+	\caption{Modulated Sub-graph from 6hpi}
 \end{figure*}
@@
 \begin{figure*}[p]
-	\centering
-	\includegraphics[width=14cm]{"SubPath Robustness 6H"}
-\caption{}
-\label{fig:subpath-robustness-6h}
-\includegraphics[width=14cm]{Betweeness}
-\caption{}
-\label{fig:betweeness}
+	\includegraphics[width=16cm]{Git/MSF_Paper/3Timepoints}
+	\caption{Robustness analysis for three time-points of infection}
 \end{figure*}

### 2018-02-21 MSF_Paper_FA.tex
Updated contributions funding acknowledgments

@@
 
-The Ebola infection dataset analyzed during the current study are
+The Ebola infection RNAseq dataset analyzed during the current study are
 available in the GEO repository, {\scriptsize
-  https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE84188.}  The
+  https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE84188.}  The cell signaling
 network file used is from Reactome Functional interactions (FIs)
@@
 \subsection*{Author contributions}
-In order to give appropriate credit to each author of an article, the individual
-contributions of each author to the manuscript should be detailed in this section. We
-recommend using author initials and then stating briefly how they contributed.
+FA proposed the idea. MF implemented the method. MF designed the benchmark
+data set. MF analyzed the benchmark results. MF, FA, and IH contributed to
+the manuscript writing. MF produced all figures. All authors read and
+approved the final manuscript.
 
@@
 \subsection*{Grant information}
-Please state who funded the work discussed in this article, whether it is your employer,
-a grant funder etc. Please do not list funding that you have that is not relevant to this
-specific piece of research. For each funder, please state the funder’s name, the grant
-number where applicable, and the individual to whom the grant was assigned.
-If your work was not funded by any grants, please include the line: ‘The author(s)
-declared that no grants were involved in supporting this work.’
+
+This work was funded by the FWF (“Fonds zur Förderung der wissenschaftlichen Forschung”) within the project Internationalen Kooperationsprojektes - Intl cooperation Project (Joint Project - Lead Agency Verfahren) with the project number (I 1988-B22).
+The grant was assigned to Ivo Hofacker
 
 \subsection*{Acknowledgements}
-This section should acknowledge anyone who contributed to the research or the
-article but who does not qualify as an author based on the criteria provided earlier
-(e.g. someone or an organisation that provided writing assistance). Please state how
-they contributed; authors should obtain permission to acknowledge from all those
-mentioned in the Acknowledgements section.
+
+We thank all our colleagues who provided insight and expertise that greatly assisted the research and all reviewers for their constructive criticism.
+
+\newpage
 
@@
 		\centering
-		\includegraphics[width=10.0cm]{Git/MSF_Paper/Tgfb1} 
+		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/Tgfb1} 
 		\caption{}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Git/MSF_Paper/TGFBstep1}
+		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep1}
 		\caption{}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Git/MSF_Paper/TGFBstep2} 
+		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep2} 
 		\caption{}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Git/MSF_Paper/GFBstep3}
+		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep3}
 		\caption{}
@@
 	\caption{Overview of \texttt{MSF} process showing the steps to identify the modulated sub-graphs.
-(a) showing a network of genes with green nodes as genes with \textit{p}-values $< $ 0.05, and blue nodes are genes with \textit{p}-values $>$ 0.05, (b) the genes circled red show the two initial modulated sub-graphs, sub-graph1 found one with M,K,C and the second modulated sub-graph2 S,V,T, (c) shows the modulated sub-graph1 being extended to M,K,C,A,D, (d) shows both the modulated sub=paths merged with the addition of 3 genes Z,W,P.}
+(a) showing a network of genes with green nodes as genes with \textit{p}-values $< $ 0.05, and blue nodes are genes with \textit{p}-values $>$ 0.05, (b) the genes circled red show the two initial modulated sub-graphs, sub-graph1 found one with M,K,C and the second modulated sub-graph2 S,V,T, (c) shows the modulated sub-graph1 being extended to M,K,C,A,D, (d) shows both the modulated sub-graphs merged with the addition of 3 genes Z,W,P.}
 \end{figure*}
@@
 	\centering
-	\includegraphics[width=1.7\linewidth]{Git/MSF_Paper/Merged_Network}
+	\includegraphics[width=1.6\linewidth]{ModulatesSubPaths/MSFData/Ebola6H/Merged_Network}
 	\caption{Modulated Sub-graph from 6hpi}

### 2018-02-22 MSF_Paper_FA.tex
Conclusion updated

@@
 The sub-graph shows different colors for the predefined  KEGG pathways and how they are connected to each other.
-The RIG-like receptor signaling in connected to Toll-like receptor, which intern connects to MAPK signaling, JAK-STAT signalling, CYtokine-cytokine signaling, Chemokine receptor signaling and TNF signaling.
-These pathways could be seen further leading to WNT Signaling, P13K-Akt signaling.
+The KEGG enrichment and visualization on the modulated sub-graph was done using StringApp~\cite{StringApp} in Cytoscape.
+The sub-graph shows RIG-like receptor signaling is connected to Toll-like receptor, which intern connects to MAPK signaling, JAK-STAT signalling, Cytokine-cytokine signaling, Chemokine receptor signaling and TNF signaling.
+These pathways could be seen further leading to WNT Signaling and P13K-Akt signaling.
 
@@
 
-Please state what you think are the main conclusions that can be
-realistically drawn from the findings in the paper, taking care not to
-make claims that cannot be supported.
+\textbf{MSF} is a fast and robust tool to find modulated sub-graphs from a given network.
+So far, it is the only tool from existing approaches that builds the modulated sub-graph from scratch using the network and DEG, where the predefined pathways could be seen connected to each other.
+The Robustness analysis shows that \textbf{MSF} is robust against noise and thus finds core modulated sub-graphs even if the DGEA is not carried out properly.
+Alongside with \textbf{MSF}, we provide a good directed cell signaling network.
 
@@
 
-We thank all our colleagues who provided insight and expertise that greatly assisted the research and all reviewers for their constructive criticism.
-
-\newpage
+We thank all our colleagues who provided insight and expertise that greatly assisted the research and all reviewers for their constructive criticism. \newpage
 
@@
 
+	\bibitem{StringApp}
+	http://apps.cytoscape.org/apps/stringapp
+
 \end{thebibliography}

### 2018-02-26 MSF_Paper_FA.tex
Updated Network and descritpion

@@
 The KEGG enrichment and visualization on the modulated sub-graph was done using StringApp~\cite{StringApp} in Cytoscape.
-The sub-graph shows RIG-like receptor signaling is connected to Toll-like receptor, which intern connects to MAPK signaling, JAK-STAT signalling, Cytokine-cytokine signaling, Chemokine receptor signaling and TNF signaling.
-These pathways could be seen further leading to WNT Signaling and P13K-Akt signaling.
+The sub-graph shows activation of IRF-dependent genes, including those for type I IFNs through Toll-like receptor signaling, exhibiting strong inflammatory response as shown by elevated levels of inflammatory cytokines and chemokines.
+NF-Kappa B activation is also mediated by Toll-like receptor.
+NF-Kappa B signaling is a key regulator of cytokine and chemokine expression.
+Toll-like receptor signaling further leads to  MAPK signaling, JAK-STAT signaling and TNF signaling.
+Sensor of cytosolic DNA which activates the IRF and NF-{kappa}B transcription factors,also leads to production of type I interferon and other cytokines.
 
  From the other side the ECM receptor interaction is seen to lead to Focal adhesion which regulates the P13k-Akt signaling.
-Finally P13k-Akt signaling leads to FoxO signaling which eventually connects to cell cycle.
-TNF, P13k-Akt and FoxO signaling activates apoptosis and autophagy.
+Finally P13k-Akt signaling opens to FoxO signaling which eventually connects to cell cycle, which at later time-point is disrupted.
+TNF signaling, P13k-Akt signaling and FoxO signaling activates apoptosis and autophagy.
 
@@
 
-To show the robustness and stability of our method, Poisson
-distributed noise was added to the sample read counts. 
-Then DGEA was carried out on the disturbed data with the
-same parameters as for the native data, followed by analysis with
-\texttt{MSF}. Using this criteria, the goal was to spot almost
+To show the robustness and stability of our method, Poisson distributed noise was added to the sample read counts.
+Then DGEA was carried out on the disturbed data with the same parameters as for the native data, followed by analysis with \texttt{MSF}.
+Using this criteria, the goal was to spot almost
 identical modulated sub-graphs in the real and noisy data.
 This procedure was carried out 100 times and every time the genes from the sub-graphs identified from noisy data were compared to the sub-graphs identified from the native data.
-The robustness of \texttt{MSF} for the three time-points is shown in~(Fig.~2).
-For the last time-point (2dpi) 84\% genes of the sub-graphs from native data were retrieved, for 1dpi the robustness was about 79\%.
-The robustness was the lowest at 6hpi nearly 71\%, this could be due to that the infection merely started and the signal transdcution was weak as compared to the later time-points and over that Poisson distributed noise was added to the data.
+This was also done for the DEG obtained for each run, a cutoff was used to identify the same number of retrieved DEG as the native data.
+The robustness of \texttt{MSF} and the DEG analysis for the three time-points are shown in~(Fig.~3).
+Since we were very stringent in adding noise to the counts, which could be seen by the DEG analysis robustness.
+For the last time-point (48hpi) 84\% genes of the sub-graphs from native data were retrieved, for 24hpi robustness was about 79\%.
+Robustness was the lowest at 6hpi nearly 71\%, this could be due to that the infection merely started and the signal transdcution was weak as compared to the later time-points.
 
@@
 
-We thank all our colleagues who provided insight and expertise that greatly assisted the research and all reviewers for their constructive criticism. \newpage
+We thank all our colleagues who provided insight and expertise that greatly assisted the research and all reviewers for their constructive criticism.
+
+\subsection*{Supplementary Material}
+
+\newpage
 
@@
 	\centering
-	\includegraphics[width=1.6\linewidth]{ModulatesSubPaths/MSFData/Ebola6H/Merged_Network}
-	\caption{Modulated Sub-graph from 6hpi}
+	\includegraphics[width=1.6\linewidth]{ModulatesSubPaths/MSFData/Ebola6H/Merged6hNetwork}
+	\caption{Modulated Sub-graph from 6hpi, showing KEGG predefined pathways colored.}
+	\label{fig:merged6hnetwork}
 \end{figure*}
@@
 \begin{figure*}[p]
-	\includegraphics[width=16cm]{Git/MSF_Paper/3Timepoints}
+	\includegraphics[width=16cm]{Documents/PhDThesis/MSFPaper/DEGvsMSF2}
 	\caption{Robustness analysis for three time-points of infection}
+	\label{fig:degvsmsf2}
 \end{figure*}

### 2018-02-26 MSF_Paper_FA.tex
Update MSF_Paper_FA.tex

@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep3}
+		\includegraphics[width=10.0cm]{Git/MSF_Paper/TGFBstep3}
 		\caption{}
@@
 	\centering
-	\includegraphics[width=1.6\linewidth]{ModulatesSubPaths/MSFData/Ebola6H/Merged6hNetwork}
+	\includegraphics[width=1.6\linewidth]{Git/MSF_Paper/Merged6hNetwork}
 	\caption{Modulated Sub-graph from 6hpi, showing KEGG predefined pathways colored.}
@@
 \begin{figure*}[p]
-	\includegraphics[width=16cm]{Documents/PhDThesis/MSFPaper/DEGvsMSF2}
+	\includegraphics[width=16cm]{Git/MSF_Paper/DEGvsMSF2}
 	\caption{Robustness analysis for three time-points of infection}

### 2018-02-26 MSF_Paper_FA.tex
remove paths to figures

@@
 		\centering
-		\includegraphics[width=10.0cm]{Git/MSF_Paper/Tgfb1} 
+		\includegraphics[width=10.0cm]{Tgfb1.png} 
 		\caption{}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Git/MSF_Paper/TGFBstep1}
+		\includegraphics[width=10.0cm]{TGFBstep1.png}
 		\caption{}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Git/MSF_Paper/TGFBstep2} 
+		\includegraphics[width=10.0cm]{TGFBstep2.png} 
 		\caption{}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Git/MSF_Paper/TGFBstep3}
+		\includegraphics[width=10.0cm]{TGFBstep3.png}
 		\caption{}
@@
 	\centering
-	\includegraphics[width=1.6\linewidth]{Git/MSF_Paper/Merged6hNetwork}
+	\includegraphics[width=1.6\linewidth]{Merged6hNetwork.png}
 	\caption{Modulated Sub-graph from 6hpi, showing KEGG predefined pathways colored.}
@@
 \begin{figure*}[p]
-	\includegraphics[width=16cm]{Git/MSF_Paper/DEGvsMSF2}
+	\includegraphics[width=16cm]{DEGvsMSF2.png}
 	\caption{Robustness analysis for three time-points of infection}

### 2018-03-01 MSF_Paper_FA.tex
new version

--- 2018-02-26 MSF_Paper_FA.tex
@@
 the differentially expressed genes (DEG)~\cite{DEG}. To this end,
-changes in transcript abundance is measured, e.g.~by next generation
+changes in transcript abundance are measured, e.g.~by next generation
 sequencing techniques, and interpreted as an indicator of differential
-expression of genes. Differentially expressed genes can be used to get
-insights into the mechanism underlying differences conditions, such as
-healthy versus diseased. In the first instance, the differential gene
-expression analysis informs about the magnitude of expression changes
-between the conditions which is the log fold change, a sign of the log fold change and the confidence level
+expression of genes. DEGs can be used to get
+insights into the mechanism underlying differences between conditions of samples, such as
+healthy versus diseased. Differential gene
+expression analysis (DGEA) informs about the magnitude of expression changes
+between the conditions which are the log fold change, the sign of log fold change and the confidence level
 of observing a authentic change, often expressed as
-\textit{p}-value. These lists of differentially expressed genes are
+\textit{p}-value. These DEGs are
 then used to extract meaningful insights, for example the genes that
@@
 acquire understandings of the perturbations in a biological
-system. Biological pathways are set of genes contained in a functional
-unit. They help to identify pathways or networks that may be altered
+system. Biological pathways are sets of genes contained in a functional
+unit. DEGs help to identify pathways or networks that may be altered
 during a change of condition providing important information about
@@
 KEGG~\cite{Kegg} and Reactome~\cite{Reactome}, the expression
-measurements of the genes obtained from differential gene expression
-analysis (DGEA) in combination with statistical methods and algorithms to
+measurements of the genes obtained from DGEA in combination with statistical methods and algorithms to
 identify specifically modulated pathways and processes~\cite{Campos}.
 
-The existing pathway-based analysis approaches use different research
+KEGG (Kyoto Encyclopedia of Genes and Genomes) is a database resource for understanding the functionality and utilities of the biological system.
+KEGG pathways is a branch of KEGG database that has a collection of manually drawn pathway maps representing the molecular interaction, reaction and relation networks of cellular functions.
+Similarly Reactome is an open-source manually curated, peer-reviewed and well known  database for biological processes pathway.
+
+Existing pathway-based analysis approaches use different research
 designs, which can be categorized into ORA (Over-representation
@@
 $\geq$ 2 and \textit{p}-value $\leq$ 0.05) to define a list of
-differentially expressed genes. Subsequently, sets of genes associated
+DEGs. Subsequently, sets of genes associated
 with annotated pathways are tested for being overrepresented in the
-set of differentially expressed genes. To this end, hypergeometric
+set of DEGs. To this end, hypergeometric
 distribution, chi-square tests, binomial probability or the Fisher’s
-exact test are used. Thereby the information of the topology of the
-pathways are ignored~\cite{Bayer}. Futhermore, ORA assumes the
-pathways are independent of each other and ignores the fact that
-biological pathways cross-talk and overlap~\cite{Khatri2012,Campos}.
+exact test are used. Thereby the information of the topology of genes in the
+pathways are ignored~\cite{Bayer}. Futhermore, ORA assumes that the biological
+pathways are independent of each other and ignores the fact that they cross-talk and overlap~\cite{Khatri2012,Campos}.
 
@@
 not assume genes to be independent of each other. FCS works in three
-step, first it calculates the gene-level statistics including correlation of molecular measurements, ANOVA, Q-statistic, signal-to-noise ratio, t-test and Z-score.
-In the second step the statistics of individual genes are transformed to a individual pathway-level statistic and finally the
-pathway-level statistics are accessed. Although FCs covers some of the
+step, first it calculates the gene-level statistics including correlation of molecular measurements  using differential expression of individual genes.
+In the second step the statistics of individual genes are transformed to an individual pathway-level statistic and finally the
+pathway-level statistics are accessed. Although FCS covers some of the
 limitations from ORA, it still lacks the topology, cross-talk and
@@
 topology of each gene during the gene-level
-statistics but still are unable to consider a link between different pathways~\cite{Khatri2012}.
+statistics but still are unable to link different pathways~\cite{Khatri2012}.
 
@@
 other. Rather, they are often regulating each others expression or are
-part of the same regulon~\cite{Michalak}. Second, we
+part of the same regulon~\cite{Michalak}. We
 understand that the categorization of links between genes into labeled
 pathways is often an arbitrary one, given the extensive cross talk
-between different pathways. Although this categories have proven to be
+between different pathways. Although these categories have proven to be
 useful in many situations, they force a certain perspective onto the
 interpretation of novel data. Based on this two principles, we aim to
-find sub-graphs within predefined networks which exhibits as a whole
-significant differential expression changes. As input information on
-functional links between genes, provided by e.g. KEGG or Reactome, and
-information on the differential expression status of single genes,
+find sub-graphs within predefined networks which exhibit as a whole
+significant differential expression changes. As input, information on
+functional links between genes provided by e.g. KEGG or Reactome and
+information on the differential expression status of single genes
 resulting from a DGEA, are required. As a result the analysis returns
@@
 pathway categories are returned. All of this can be helpful to
-understand cause end effect of a stimulus and might inform about
+understand the cause end effect of a stimulus and might inform about
 potential points of intervention. The proposed algorithm was
@@
 signaling network. \texttt{MSF} does not work on predefined pathways but builds up the sub-graph from scratch using the whole cell signaling network. The network has nodes corresponding to genes and
-edges representing their relationship. The pipeline of identification
-of modulated sub-graphs from a network by \texttt{MSF} is depicted in
-Fig (1).
+edges representing their interactions.
 
@@
 combined \textit{p}-value, using a statistical method for combining
-dependent \textit{p}-values described by~\cite{Hartung}. Using the
-inverse normal method, individual gene \textit{p}-values $p_i$ are
-first transformed to its corresponding normal score $t_i$,
-
+dependent \textit{p}-values described by Hartung~\cite{Hartung}. Using the inverse of standard normal distribution function $\Phi^{-1}$, individual gene \textit{p}-values $p_i$ are
+first transformed to their corresponding normal score $t_i(i=1,...,N)$
 \[ t_i = \Phi^{-1}(p_i) \]
 
-Then using these normal scores, the correlation \textit{Co} between genes are calculated,
+Then using these normal scores, the correlation \textit{Cov} between genes is calculated,
 
-\[ Co (t_i, t_j) =\varrho \]
+\[ Cov (t_i, t_j) =\varrho\]
 
-followed by correction of the correlation \textbf{k}. The normal
-scores, correlations and the correction of correlation are applied to
+The normal scores, correlations and the correction of correlation are applied to
 the inverse normal function to calculate the individual
@@
 
-$  {\scriptstyle Cp(\varrho,K) = \frac{\sum_{i=1}^{n} \lambda_i t_i}{\sqrt{\sum_{i=1}^{n} \lambda_i^2 +[(\sum_{i=1}^{n} \lambda_i)^2- \sum_{i=1}^{n} \lambda_i^2]\{\varrho+k\cdot\sqrt{\frac{2}{n+1}} (1-\varrho)}}} $\\ \newline
+$  {\scriptstyle Cp(\varrho,k) = \frac{\sum_{i=1}^{n} \lambda_i t_i}{\sqrt{\sum_{i=1}^{n} \lambda_i^2 +[(\sum_{i=1}^{n} \lambda_i)^2- \sum_{i=1}^{n} \lambda_i^2]\{\varrho+k\cdot\sqrt{\frac{2}{n+1}} (1-\varrho)\}}}} $\\ \newline
 
-Lambda $\lambda$ are the weights for each gene, For the moment equal
-weights are given to all genes and \textbf{k} used is 0.2 as used by authors. An
+Lambda $\lambda$ are the weights for each gene and \textbf{k} is the correction of the correlation.
+For the moment, equal weights are given to all genes and \textbf{k} used is 0.2 as used by authors.
+An
 overall \textit{p}-value of a sub-graph will express the significance
@@
 \subsection*{Overview of our method}
+
+The pipeline of identification of modulated sub-graphs from a network by \texttt{MSF} is depicted in
+Fig (1).
+\newline
 
@@
 In this case the process starts over with all remaining genes not yet in a
-significant modulated sub-graph. This step identifies all the trivial
+significantly modulated sub-graph. This step identifies all the trivial
 sub-graphs that are modulated in the whole network~(Fig.~1b).\newline
@@
 they could further be extended beyond the immediate neighborhood.
-This is done by testing all possible extension up to 2 genes for
+This is done by testing all possible extension by two genes for
 all genes in the sub-graph. Again, this step is iteratively repeated until no
@@
 tested if combined sub-graph score is better than on their own.
-The merging of the two sub-graphs is done by depth first search, If the two sub-graphs merge with the connector of at most 2 genes and the
+The merging of the two sub-graphs is done by depth first search, If the two sub-graphs merge with the connector of at most two genes and the
 combined \textit{p}-value of the merged sub-graph including the
@@
 In a last post processing step \texttt{MSF} identifies the trigger
-points of the modulated sub-graph. These trigger genes are the sources
+points of the modulated sub-graphs. These trigger genes are the sources
 of the sub-graphs with only outgoing edges. These genes can be
@@
 
-To demonstrate the usefulness \texttt{MSF} is applied to a RNAseq data
+To demonstrate the usefulness \texttt{MSF} is applied to RNAseq data
 set of primary human monocyte-derived macrophages (MDMs) infected with
@@
 alpha/beta interferon (IFN-$\alpha /
-\beta$)~\cite{Prins,Konde,Cardenas}. The EBOV viral protein VP35
+\beta$)~\cite{Prins,Konde,Cardenas}. The Ebola viral protein VP35
 targets the host type~I interferon IFN-$\alpha / \beta$ to block the
@@
 
-The EBOV infection data has three time-points six hour (6hpi), one day (1dpi) and two day (2dpi) post infection.
+The EBOV infection data has three time-points 6 , 24  and 48 hour post infection (hpi).
 At 6hpi, five modulated sub-graphs were identified.
-Most of the genes part of the sub-moldules were chemokines (CXCL10, CCL8) and Interleukin genes (IL6, IL27, IL23) which serve as a subset of CD4+ T
+Most of the genes part of the sub-graphs were chemokines (CXCL10, CCL8) and Interleukin genes (IL6, IL27, IL23) which serve as a subset of CD4+ T helper
 cells.
@@
 Most of the sources identified by \texttt{MSF} were type~I interferon induced genes.
-At 1dpi nine modulated sub-graphs were identified, IFNA1 was identified as one of the sources.
+At 24hpi nine modulated sub-graphs were identified, IFNA1 was identified as one of the sources.
 For the last time-point again IFNB1 and IFNA1 were identified as the two sources out of the possible sources from seven sub-graphs identified.
@@
 based on the individual DGEA alone.
-These sources will reduce the search space for potential target gene and can help the biologist as the starting point of clinical testing for drugs and vaccines against an infection.
+Identifying these sources will reduce the search space for potential target genes and can help the biologist as the starting point of clinical testing for drugs and vaccines against an infection.
 
@@
  The connected modulated sub-graphs from 6hpi are shown in~(Fig.~2).
-The sub-graph shows different colors for the predefined KEGG pathways and how they are connected to each other.
+The graph shows different colors for predefined KEGG pathways and how they cross talk.
 The KEGG enrichment and visualization on the modulated sub-graph was done using StringApp~\cite{StringApp} in Cytoscape.
-The sub-graph shows activation of IRF-dependent genes, including those for type I IFNs through Toll-like receptor signaling, exhibiting strong inflammatory response as shown by elevated levels of inflammatory cytokines and chemokines.
-NF-Kappa B activation is also mediated by Toll-like receptor.
-NF-Kappa B signaling is a key regulator of cytokine and chemokine expression.
-Toll-like receptor signaling further leads to  MAPK signaling, JAK-STAT signaling and TNF signaling.
-Sensor of cytosolic DNA which activates the IRF and NF-{kappa}B transcription factors,also leads to production of type I interferon and other cytokines.
+The graph unveils how more than one stimulus coincide for the progression of infection.
+The activation of IRF-dependent genes (light Blue), including those for type I IFNs through Toll-like receptor signaling (yellow), exhibits strong inflammatory response as shown by elevated number of inflammatory cytokines and chemokines genes (green and pink).
+NF-Kappa B activation (off-white) is also mediated by Toll-like receptor.
+NF-Kappa B signaling is a key regulator of cytokine and chemokine expression\cite{Olejnik}.
+Toll-like receptor signaling further leads to  MAPK signaling (neon blue), JAK-STAT (red) signaling and TNF signaling (dark blue).
+Sensor of cytosolic DNA (dark grey) which activates the IRF and NF-Kappa B transcription factors, also leads to production of type I interferon and other cytokines.
 
- From the other side the ECM receptor interaction is seen to lead to Focal adhesion which regulates the P13k-Akt signaling.
-Finally P13k-Akt signaling opens to FoxO signaling which eventually connects to cell cycle, which at later time-point is disrupted.
-TNF signaling, P13k-Akt signaling and FoxO signaling activates apoptosis and autophagy.
+ Extracellular matrix (ECM) receptor interaction genes are identified by \texttt{MSF} at the other end of the graph, which are suggested to be in interaction with Ebola glycoprotein (GP)~\cite{Veljkovic}.
+ECM receptor interactions are seen to lead to Focal adhesion which regulates the P13k-Akt signaling (purple).
+P13k-Akt signaling goes to FoxO signaling which eventually connects to cell cycle, which at later time-points is disrupted.
+TNF signaling, P13k-Akt signaling and FoxO signaling activates apoptosis (black) and autophagy.
+The light gray node belong to other pathways not shown due to graph visual complexity.
 
@@
 
-To better understand the functional role of the sub-graphs identified by \texttt{MSF}, we linked the  sub-graphs to the predefined KEGG pathways and Reactome
+To further understand the functional role of the sub-graphs identified by \texttt{MSF}, we linked the  sub-graphs to the predefined KEGG pathways and Reactome
 pathways through gene enrichment analysis.
 We used Cytoscape~\cite{Cyto} plugins to identify the enriched patwhays, CytoKegg~\cite{Cytokegg} for KEGG and Reactome FI~\cite{Reactome} for Reactome.
-We compared the pathways enriched in \texttt{MSF} identified sub-graphs to SPIA~\cite{Tarca} for KEGG and Reactome pathway enrichment analysis.
+We compared the pathways enriched in \texttt{MSF} identified sub-graphs to pathway analysis by SPIA~\cite{Tarca} for KEGG and Reactome pathway enrichment analysis for Reactome.
 
 At 6hpi Cytokegg identified the sub-graph genes to be enriched in Toll-like receptor signaling, TNF signaling, IL-17 signaling and NF-kappa signaling pathways, as also stated by~\cite{Olejnik}.
-The earliest response EBOV induces on  cytokine is via TLR4-mediated signaling~\cite{Olejnik}, Gene enrichment analysis of sub-graph genes at 6hpi showed toll-like receptor signaling being most significantly dis-regulated pathway, this important signaling pathway was not identified by SPIA along with RIG-I-like receptor signaling pathway, TNF signaling pathway and IL-17 signaling (Table 1).
-On the later time-points \texttt{MSF} and SPIA showed consensus on the important pathways e.g Chemokine signaling, Cytokine-Cytokine receptor interaction, NF-kappa B signaling , Cytosolic DNA-sensing and RIG-I-like receptor signaling as being dis-regulated supplement (Table 3,4).
+The earliest response EBOV induces on cytokine is via TLR4-mediated signaling~\cite{Olejnik}, Gene enrichment analysis of sub-graph genes at 6hpi showed toll-like receptor signaling being most significantly dis-regulated pathway, this important signaling pathway was not identified by SPIA along with RIG-I-like receptor signaling pathway, TNF signaling pathway and IL-17 signaling (Table 1 in the supplement material).
+On the later time-points \texttt{MSF} and SPIA showed consensus on the important pathways e.g Chemokine signaling, Cytokine-Cytokine receptor interaction, NF-kappa B signaling, Cytosolic DNA-sensing and RIG-I-like receptor signaling as being dis-regulated (Table 2-3 in the supplement).
 
-Likewise we compared gene enrichment results of sub-graph genes identified by \texttt{MSF} using Reactome FI to the Reactome predefined pathway enrichment analysis results.
-Although \texttt{MSF} and Reactome show good agreement on the dis-regulated pathways e.g Signaling by Interleukins, Interleukin-10 signaling, Cytokine Signaling in Immune system, Chemokine receptors bind chemokines, RIG-I/MDA5 mediated induction of IFN-alpha/beta pathways, NF-kB activation, Reactome as well fails to identify the Toll-like receptor signaling.
-This shows that \texttt{MSF} is able to identify the modulated genes in the sub-graphs at the earliest time-point of infection even though when the signal transduction was weak.
+Likewise we compared gene enrichment results of sub-graph genes identified by \texttt{MSF} using Reactome FI to the Reactome predefined pathway enrichment analysis.
+Although \texttt{MSF} and Reactome showed good agreement on the dis-regulated pathways e.g Signaling by Interleukins, Interleukin-10 signaling, Cytokine Signaling in Immune system, Chemokine receptors bind chemokines, RIG-I/MDA5 mediated induction of IFN-alpha/beta pathways, NF-kB activation, Reactome as well could not identify Toll-like receptor signaling (Table 1-3 in the supplement material) .
+This shows that \texttt{MSF} is able to identify the important modulated genes in the sub-graphs at the earliest time-point of infection.
 
@@
 
-To show the robustness and stability of our method, Poisson distributed noise was added to the sample read counts.
+To assess the robustness and stability of our method, Poisson distributed noise was added to the sample read counts.
 Then DGEA was carried out on the disturbed data with the same parameters as for the native data, followed by analysis with \texttt{MSF}.
@@
 identical modulated sub-graphs in the real and noisy data.
-This procedure was carried out 100 times and every time the genes from the sub-graphs identified from noisy data were compared to the sub-graphs identified from the native data.
-This was also done for the DEG obtained for each run, a cutoff was used to identify the same number of retrieved DEG as the native data.
+This procedure was carried out 100 times and every time the genes from the sub-graphs identified from noisy data were compared to the genes of sub-graphs identified from the native data.
+This was also done for the DEG obtained for each run, a cutoff was used to compare the retrieved DEG from noisy data to the native data.
 The robustness of \texttt{MSF} and the DEG analysis for the three time-points are shown in~(Fig.~3).
-Since we were very stringent in adding noise to the counts, which could be seen by the DEG analysis robustness.
+We were very stringent in adding noise to the counts, which could be seen by the DEG analysis robustness.
 For the last time-point (48hpi) 84\% genes of the sub-graphs from native data were retrieved, for 24hpi robustness was about 79\%.
-Robustness was the lowest at 6hpi nearly 71\%, this could be due to that the infection merely started and the signal transdcution was weak as compared to the later time-points.
+Robustness was 71\% at 6hpi, this could be due to that the infection merely started and the signal transdcution was weak as compared to the later time-points.
 
-Simulation was also conducted on \texttt{MSF} to test the results
-obtained from real data, showing a chance of less than 10 percent to
-get more sources than observed.
+Simulation testing was also conducted on \texttt{MSF}, the objective was to find less sources than observed in the sub-graphs identified by \texttt{MSF} from native data.
+Simulation results showed a chance of less than 10 percent to get more sources than observed.
 
@@
 So far, it is the only tool from existing approaches that builds the modulated sub-graph from scratch using the network and DEG, where the predefined pathways could be seen connected to each other.
-The Robustness analysis shows that \textbf{MSF} is robust against noise and thus finds core modulated sub-graphs even if the DGEA is not carried out properly.
-Alongside with \textbf{MSF}, we provide a good directed cell signaling network.
+The Robustness analysis shows that \textbf{MSF} is robust against noise and thus finds core modulated sub-graphs even if the DGEA is not carried out vigilantly.
+Alongside with \textbf{MSF}, we provide a good directed cell signaling network file, which is a subset of interactions from Reactome Functional interactions (FIs)
+Version 2016 .
 
-\subsection*{Availability of data and material}
+\subsection*{Source of data}
 
 The Ebola infection RNAseq dataset analyzed during the current study are
-available in the GEO repository, {\scriptsize
-  https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE84188.}  The cell signaling
+available in the GEO repository (GSE84188). The cell signaling
 network file used is from Reactome Functional interactions (FIs)
@@
 
-This work was funded by the FWF (“Fonds zur Förderung der wissenschaftlichen Forschung”) within the project Internationalen Kooperationsprojektes - Intl cooperation Project (Joint Project - Lead Agency Verfahren) with the project number (I 1988-B22).
+This work was funded by the FWF (“Fonds zur Forderung der wissenschaftlichen Forschung”) within the project Internationalen Kooperationsprojektes - Intl cooperation Project (Joint Project - Lead Agency Verfahren) with the project number (I 1988-B22).
 The grant was assigned to Ivo Hofacker
@@
 
-\newpage
+The supplementary file contains tables of comparison between \textbf{MSF} identified sub-graph gene enrichment with SPIA and Reactome pathway analysis. 
+\clearpage
 
@@
 
+	\bibitem{Veljkovic}
+	Veljkovic, V., Glisic, S., Muller, C.
+P., Scotch, M., Branch, D.
+R., Perovic, V.
+R., … Colombatti, A.
+(2015).
+In silico analysis suggests interaction between Ebola virus and the extracellular matrix.
+Frontiers in Microbiology, 6, 135.
+http://doi.org/10.3389/fmicb.2015.00135
+
 \end{thebibliography}
@@
 \begin{figure*}[ht!]
+		\centering
+	\fbox{
+		\begin{minipage}{17 cm}
 
@@
 (a) showing a network of genes with green nodes as genes with \textit{p}-values $< $ 0.05, and blue nodes are genes with \textit{p}-values $>$ 0.05, (b) the genes circled red show the two initial modulated sub-graphs, sub-graph1 found one with M,K,C and the second modulated sub-graph2 S,V,T, (c) shows the modulated sub-graph1 being extended to M,K,C,A,D, (d) shows both the modulated sub-graphs merged with the addition of 3 genes Z,W,P.}
+			\end{minipage}
+}
 \end{figure*}
@@
 	\centering
-	\includegraphics[width=1.6\linewidth]{ModulatesSubPaths/MSFData/Ebola6H/Merged6hNetwork}
-	\caption{Modulated Sub-graph from 6hpi, showing KEGG predefined pathways colored.}
-	\label{fig:merged6hnetwork}
+	\fbox{
+		\begin{minipage}{17 cm}
+	\includegraphics[width=1.4\linewidth]{ModulatesSubPaths/MSFData/Ebola6H/newestMergedNetwork}
+	\caption{Shows connected modulated sub-graphs from 6hpi, where nodes are genes and edges are the interactions.
+The node coloring represents the predefined KEGG pathways referring to the nodes in the legend, the edges are from Reactome.
+Important genes to EBOV infection as from literature are enlarged in the graph.
+The signal flow depicted in the graph is explained in the text.}
+		\end{minipage}
+}
 \end{figure*}
@@
 \begin{figure*}[p]
+	\centering
+	\fbox{
+		\begin{minipage}{17 cm}
 			\includegraphics[width=16cm]{Documents/PhDThesis/MSFPaper/DEGvsMSF2}
-	\caption{Robustness analysis for three time-points of infection}
-	\label{fig:degvsmsf2}
+			\caption{Graph shows the percentage of DEG analysis genes and MSF identified sub-graph genes recall rates 100 times for the three different time points after the addition of Poisson distributed noise.}
+			\label{label}
+		\end{minipage}
+	}
 \end{figure*}

### 2018-03-01 MSF_Paper_FA.tex
Update MSF_Paper_FA.tex

@@
 		\begin{minipage}{17 cm}
-	\includegraphics[width=1.4\linewidth]{ModulatesSubPaths/MSFData/Ebola6H/newestMergedNetwork}
+	\includegraphics[width=1.4\linewidth]{Git/MSF_Paper/newestMergedNetwork}
 	\caption{Shows connected modulated sub-graphs from 6hpi, where nodes are genes and edges are the interactions.
@@
 		\begin{minipage}{17 cm}
-			\includegraphics[width=16cm]{Documents/PhDThesis/MSFPaper/DEGvsMSF2}
+			\includegraphics[width=16cm]{Git/MSF_Paper/DEGvsMSF2}
 			\caption{Graph shows the percentage of DEG analysis genes and MSF identified sub-graph genes recall rates 100 times for the three different time points after the addition of Poisson distributed noise.}

### 2018-03-01 MSF_Paper_FA.tex
Update MSF_Paper_FA.tex

@@
 between the conditions which are the log fold change, the sign of log fold change and the confidence level
-of observing a authentic change, often expressed as
+of observing an authentic change, often expressed as
 \textit{p}-value. These DEGs are

### 2018-03-01 MSF_Paper_FA.tex
Update MSF_Paper_FA.tex

@@
 useful in many situations, they force a certain perspective onto the
-interpretation of novel data. Based on this two principles, we aim to
+interpretation of novel data. Based on these two principles, we aim to
 find sub-graphs within predefined networks which exhibit as a whole

### 2018-03-17 MSF_Paper_FA.tex
Update MSF_Paper_FA.tex

--- 2018-03-01 MSF_Paper_FA.tex
@@
 expression analysis (DGEA) informs about the magnitude of expression changes
-between the conditions which are the log fold change, the sign of log fold change and the confidence level
+between the conditions which are often expressed as fold change, sign of fold change and the confidence level
 of observing a authentic change, often expressed as
 \textit{p}-value. These DEGs are
-then used to extract meaningful insights, for example the genes that
+then used as information converted from data, this information is further interpreted to knowledge to extract meaningful insights, for example the genes that
 could be important for a particular condition or maybe the target gene
-of any infection or disease. Out of all high throughput analysis
-methods currently in use, pathway-based analysis has become an
+of an infection or disease. To this end, pathway-based analysis has become an
 important tool to further interpret the results of a DGEA and to
 acquire understandings of the perturbations in a biological
-system. Biological pathways are sets of genes contained in a functional
+system. Biological pathways are sets of genes and their interactions forming a functional
 unit. DEGs help to identify pathways or networks that may be altered
@@
 diseases and its treatment process~\cite{Khatri2012}. Pathway-based
-methods use the predefined pathways or networks such as
+methods use predefined pathways or networks such as
 KEGG~\cite{Kegg} and Reactome~\cite{Reactome}, the expression
@@
 
-KEGG (Kyoto Encyclopedia of Genes and Genomes) is a database resource for understanding the functionality and utilities of the biological system.
-KEGG pathways is a branch of KEGG database that has a collection of manually drawn pathway maps representing the molecular interaction, reaction and relation networks of cellular functions.
-Similarly Reactome is an open-source manually curated, peer-reviewed and well known  database for biological processes pathway.
+Predefined pathways are sets of genes and their interactions categorized in a functional unit differently by different databases known as pathways.
+For example, KEGG (Kyoto Encyclopedia of Genes and Genomes) is a database resource for understanding the functionality and utilities of the biological system.
+KEGG pathways is a branch of KEGG database that has a collection of manually drawn pathway maps representing the molecular interaction, reaction and relation networks of cellular functions~\cite{Kegg}.
+Similarly, Reactome is an open-source, manually curated, peer-reviewed and well known database for signaling and metabolic molecules with their interactions formed into different  biological pathways for nineteen species~\cite{Reactome}.
 
@@
 methods. ORA is the first and the most basic method of pathway
-analysis. It uses a user defined cut-off for the log-fold change and
-\textit{p}-value from the DGEA (most commonly using absolute log-fold change
-$\geq$ 2 and \textit{p}-value $\leq$ 0.05) to define a list of
-DEGs. Subsequently, sets of genes associated
-with annotated pathways are tested for being overrepresented in the
-set of DEGs. To this end, hypergeometric
+analysis. It uses a DEG list with user defined cut-off for the log-fold change and
+\textit{p}-value (most commonly using absolute log-fold change
+$\geq$ 2 and \textit{p}-value $\leq$ 0.05). Subsequently, sets of genes associated
+with annotated pathways are tested for being over-represented in the
+set of DEGs. To this end, hyper-geometric
 distribution, chi-square tests, binomial probability or the Fisher’s
 exact test are used. Thereby the information of the topology of genes in the
-pathways are ignored~\cite{Bayer}. Futhermore, ORA assumes that the biological
+pathways are ignored~\cite{Bayer}. Furthermore, ORA assumes that the biological
 pathways are independent of each other and ignores the fact that they cross-talk and overlap~\cite{Khatri2012,Campos}.
 
-Unlike ORA, FCS has no artificial cut-off to define DEG list and does
-not assume genes to be independent of each other. FCS works in three
-step, first it calculates the gene-level statistics including correlation of molecular measurements  using differential expression of individual genes.
-In the second step the statistics of individual genes are transformed to an individual pathway-level statistic and finally the
-pathway-level statistics are accessed. Although FCS covers some of the
-limitations from ORA, it still lacks the topology, cross-talk and
+Unlike ORA, FCS has no artificial cut-off to define DEG list. FCS works in three
+step, first it calculates the gene-level statistics including correlation of molecular measurements  using differential expression of individual genes, ANOVA, t-test and  Z-score.
+In the second step the statistics of individual genes in a pathway are transformed to an individual pathway-level statistic commonly using Kolmogorov-Smirnov statistic, mean or median.
+Finally the statistical significance of the pathway-level statistics is assessed.
+Although FCS covers some of the
+limitations from ORA, it still lacks the topology of genes in a pathway, cross-talk and
 overlap of the pathways~\cite{Khatri2012,Campos}. Pathway topology
@@
 useful in many situations, they force a certain perspective onto the
-interpretation of novel data. Based on this two principles, we aim to
-find sub-graphs within predefined networks which exhibit as a whole
+interpretation of novel data. Based on these two principles, we aim to
+find sub-graphs of connected genes within cell signaling network which exhibit as a whole
 significant differential expression changes. As input, information on
@@
 implemented as a java program, which was named Modulated Sub-graph
-Finder (\texttt{MSF}).
+Finder (\texttt{MSF}). \texttt{MSF} would help transform the information obtained from DGEA into understandable knowledge of signal transduction of genes. \texttt{MSF} is freely accessible on github.
 
@@
 find the modulated sub-graphs of gene interactions from the cell
-signaling network. \texttt{MSF} does not work on predefined pathways but builds up the sub-graph from scratch using the whole cell signaling network. The network has nodes corresponding to genes and
-edges representing their interactions.
+signaling network.
+\texttt{MSF} does not use predefined gene sets in functional units, it builds up the sub-graphs from scratch using the whole cell signaling network from Reactome.
+The network has nodes corresponding to genes and edges representing their directed regulating interactions.
 
@@
 from the DGEA. The \textit{p}-value
-expresses the probability that the hypothesis of unmodified gene
+expresses the probability that the null hypothesis of unmodified gene
 expression can be rejected for a given statistical model. To find
@@
 combined \textit{p}-value, using a statistical method for combining
-dependent \textit{p}-values described by Hartung~\cite{Hartung}. Using the inverse of standard normal distribution function $\Phi^{-1}$, individual gene \textit{p}-values $p_i$ are
-first transformed to their corresponding normal score $t_i(i=1,...,N)$
-\[ t_i = \Phi^{-1}(p_i) \]
-
-Then using these normal scores, the correlation \textit{Cov} between genes is calculated,
-
-\[ Cov (t_i, t_j) =\varrho\]
-
-The normal scores, correlations and the correction of correlation are applied to
+dependent \textit{p}-values described by Hartung~\cite{Hartung}. Hartung uses the inverse of standard normal distribution function, individual gene \textit{p}-values are
+first transformed to their corresponding normal score. Then using these normal scores, the correlation between genes is calculated, the normal scores and correlation are applied to
 the inverse normal function to calculate the individual
-\textit{p}-value for a sub-graph \textit{Cp}. \newline
-
-$  {\scriptstyle Cp(\varrho,k) = \frac{\sum_{i=1}^{n} \lambda_i t_i}{\sqrt{\sum_{i=1}^{n} \lambda_i^2 +[(\sum_{i=1}^{n} \lambda_i)^2- \sum_{i=1}^{n} \lambda_i^2]\{\varrho+k\cdot\sqrt{\frac{2}{n+1}} (1-\varrho)\}}}} $\\ \newline
-
-Lambda $\lambda$ are the weights for each gene and \textbf{k} is the correction of the correlation.
-For the moment, equal weights are given to all genes and \textbf{k} used is 0.2 as used by authors.
-An
+\textit{p}-value for a sub-graph. An
 overall \textit{p}-value of a sub-graph will express the significance
@@
 
+\subsection*{Overview of our method}
+
 To reduce the complexity to score all possible connected sub-graphs
 \texttt{MSF} applies a three steps heuristic as described in the
-following.\newline
-
-\subsection*{Overview of our method}
-
-The pipeline of identification of modulated sub-graphs from a network by \texttt{MSF} is depicted in
+following.The pipeline of identification of modulated sub-graphs from a network by \texttt{MSF} is depicted in
 Fig (1).
@@
 \textit{p}-value, the extended sub-graph is accepted.
-Then the next most significant neighboring gene is added to the sub-graph and the combined \textit{p}-value is calculated, if the new combined \textit{p}-value of three genes is smaller than the combined \textit{p}-value of the first two genes,the extended sub-graph is accepted.
+Then the next most significant neighboring gene is added to the sub-graph and the combined \textit{p}-value is calculated, if the new combined \textit{p}-value of three genes is smaller than the combined \textit{p}-value of the first two genes, the extended sub-graph is accepted.
 This step is iteratively repeated until no further extension is accepted.
@@
 they could further be extended beyond the immediate neighborhood.
-This is done by testing all possible extension by two genes for
+This is done by testing all possible extension by N genes (default two genes) for
 all genes in the sub-graph. Again, this step is iteratively repeated until no
@@
 tested if combined sub-graph score is better than on their own.
-The merging of the two sub-graphs is done by depth first search, If the two sub-graphs merge with the connector of at most two genes and the
+The merging of the two sub-graphs is done by depth first search.
+If the two sub-graphs merge with the connector of at most N genes (default two genes) and the
 combined \textit{p}-value of the merged sub-graph including the
@@
 integrated information within the signal transduction network is set
-to action.
+to action. Genes in circular loops in the sub-graphs have no source or sink.
 
@@
 
+\subsection*{Robustness}
+
+To assess the robustness and stability of our method, Poisson distributed noise was added to the sample read counts.
+Then DGEA was carried out on the disturbed data with the same parameters as for the native data, followed by analysis with \texttt{MSF}.
+Using this criteria, the goal was to spot almost
+identical modulated sub-graphs in the real and noisy data.
+This procedure was carried out 100 times and every time the genes from the sub-graphs identified from noisy data were compared to the genes of sub-graphs identified from the native data.
+This was also done for the DEG obtained for each run, a cutoff was used to compare the retrieved DEG from noisy data to the native data.
+The robustness of \texttt{MSF} and the DEG analysis for the three time-points are shown in~(Fig.~2).
+We were very stringent in adding noise to the counts, which could be seen by the DEG analysis robustness.
+For the last time-point (48hpi) 84\% genes of the sub-graphs from native data were retrieved, for 24hpi robustness was about 79\%.
+Robustness was 71\% at 6hpi, this could be due to that the infection merely started and the signal transdcution was weak as compared to the later time-points.
+
+Simulation testing was also conducted on \texttt{MSF}, the objective was to find less sources than observed in the sub-graphs identified by \texttt{MSF} from native data.
+Simulation results showed a chance of less than 10 percent to get more sources than observed.
+
 \subsection*{Case Study}
 
-To demonstrate the usefulness \texttt{MSF} is applied to RNAseq data
+To demonstrate its usefulness \texttt{MSF} is applied to RNAseq data
 set of primary human monocyte-derived macrophages (MDMs) infected with
-Ebola virus~\cite{Olejnik} (GSE84188). Ebola Virus (EBOV)
+Ebola virus (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV)
 belongs to the Filoviridea family; filamentous, enveloped and single
@@
 The initial targets of EBOV are the macrophages and
-dendritic immune cells~\cite{Falasca,Rhein}. Ebola Virus inhibits the critical
+dendritic immune cells~\cite{Falasca,Rhein}. EBOV inhibits the critical
 innate immune response of the host, which includes the activation of
@@
 ~\cite{Prins,Konde,Falasca,Cardenas,Olejnik}. It has been proposed
-that IFN-$\alpha / \beta$ could be used through clinical trials to
-design antiviral drug against Ebola.
+that IFN-$\alpha / \beta$  should be
+tested against Ebola for its antiviral activity through clinical trials~\cite{Konde}.
 The aim of testing the EBOV infection data with \texttt{MSF} was to identify the modulated sub-graphs and check if \texttt{MSF} is able to identify the IFN-$\alpha / \beta$ gene as one of the sources.
 
-The EBOV infection data has three time-points 6 , 24  and 48 hour post infection (hpi).
+EBOV infection count data was downloaded from GEO (GSE84188).
+Differential gene expression analysis was performed on the count data with edgeR package (version 3.4.2)~\cite{edgeR}.
+The DEG list generated by edgeR was used as an input for \texttt{MSF}.
+The EBOV infection data has three time-points 6, 24  and 48 hour post infection (hpi).
 At 6hpi, five modulated sub-graphs were identified.
@@
 cells.
-IFNB1  and IFNA1 were both identified as two of the sources.
+IFNB1 and IFNA1 were both identified as two of the possible sources.
 Most of the sources identified by \texttt{MSF} were type~I interferon induced genes.
 At 24hpi nine modulated sub-graphs were identified, IFNA1 was identified as one of the sources.
-For the last time-point again IFNB1 and IFNA1 were identified as the two sources out of the possible sources from seven sub-graphs identified.
+For the last time-point again IFNB1 and IFNA1 were identified as the two sources out of the possible sources from seven modulated sub-graphs identified.
 
@@
 
- The connected modulated sub-graphs from 6hpi are shown in~(Fig.~2).
-The graph shows different colors for predefined KEGG pathways and how they cross talk.
-The KEGG enrichment and visualization on the modulated sub-graph was done using StringApp~\cite{StringApp} in Cytoscape.
-The graph unveils how more than one stimulus coincide for the progression of infection.
-The activation of IRF-dependent genes (light Blue), including those for type I IFNs through Toll-like receptor signaling (yellow), exhibits strong inflammatory response as shown by elevated number of inflammatory cytokines and chemokines genes (green and pink).
-NF-Kappa B activation (off-white) is also mediated by Toll-like receptor.
-NF-Kappa B signaling is a key regulator of cytokine and chemokine expression\cite{Olejnik}.
-Toll-like receptor signaling further leads to  MAPK signaling (neon blue), JAK-STAT (red) signaling and TNF signaling (dark blue).
-Sensor of cytosolic DNA (dark grey) which activates the IRF and NF-Kappa B transcription factors, also leads to production of type I interferon and other cytokines.
-
- Extracellular matrix (ECM) receptor interaction genes are identified by \texttt{MSF} at the other end of the graph, which are suggested to be in interaction with Ebola glycoprotein (GP)~\cite{Veljkovic}.
-ECM receptor interactions are seen to lead to Focal adhesion which regulates the P13k-Akt signaling (purple).
-P13k-Akt signaling goes to FoxO signaling which eventually connects to cell cycle, which at later time-points is disrupted.
-TNF signaling, P13k-Akt signaling and FoxO signaling activates apoptosis (black) and autophagy.
-The light gray node belong to other pathways not shown due to graph visual complexity.
+ Three main modulated sub-graphs identified by \texttt{MSF} at 6hpi are shown in~(Fig.~3).
+The graph shows different colored nodes as genes associated to KEGG pathways and how they cross talk.
+The KEGG association and visualization on the modulated sub-graphs was done using StringApp~\cite{StringApp} in Cytoscape~\cite{Cyto}.
+The graph unveils how more than one stimulus coincides to prevent the progression of infection.
+Sub-graph 1 shows toll-like receptor, cytokine, chemokine and jak-stat genes leading to TNF and eventually going into apoptosis.
+Sub-graph 2 shows Extra-cellular matrix (ECM) receptor interaction genes which are suggested to be in interaction with Ebola glycoprotein (GP)~\cite{Veljkovic}.
+ECM receptor, Focal adhesion and P13k-Akt signaling genes are the sources from one end of the sub-graph.
+P13k-Akt signaling goes to FoxO signaling which eventually leads to cell cycle, which at later time-points is disrupted.
+From the other end of sub-graph 2 toll-like recpetor combined with cytosolic DNA sensing and cytokine-chemokine signaling activates Jak-stat which leads to the start of apoptosis with a activation of a few apoptosis genes.
+Cytokine is also seen to sink into natural killer mediated cytotoxicity.
+Sub-graph 3 shows how some of the cytokine and chemokine genes activate the natural killer mediated cytotoxicity and apoptosis.
 
@@
 
-\subsection*{Robustness}
-
-To assess the robustness and stability of our method, Poisson distributed noise was added to the sample read counts.
-Then DGEA was carried out on the disturbed data with the same parameters as for the native data, followed by analysis with \texttt{MSF}.
-Using this criteria, the goal was to spot almost
-identical modulated sub-graphs in the real and noisy data.
-This procedure was carried out 100 times and every time the genes from the sub-graphs identified from noisy data were compared to the genes of sub-graphs identified from the native data.
-This was also done for the DEG obtained for each run, a cutoff was used to compare the retrieved DEG from noisy data to the native data.
-The robustness of \texttt{MSF} and the DEG analysis for the three time-points are shown in~(Fig.~3).
-We were very stringent in adding noise to the counts, which could be seen by the DEG analysis robustness.
-For the last time-point (48hpi) 84\% genes of the sub-graphs from native data were retrieved, for 24hpi robustness was about 79\%.
-Robustness was 71\% at 6hpi, this could be due to that the infection merely started and the signal transdcution was weak as compared to the later time-points.
-
-Simulation testing was also conducted on \texttt{MSF}, the objective was to find less sources than observed in the sub-graphs identified by \texttt{MSF} from native data.
-Simulation results showed a chance of less than 10 percent to get more sources than observed.
-
 \section*{Discussion}
@@
 So far, it is the only tool from existing approaches that builds the modulated sub-graph from scratch using the network and DEG, where the predefined pathways could be seen connected to each other.
-The Robustness analysis shows that \textbf{MSF} is robust against noise and thus finds core modulated sub-graphs even if the DGEA is not carried out vigilantly.
+The Robustness analysis shows that \textbf{MSF} is fairly robust against noise and thus finds core modulated sub-graphs even if the DGEA is not carried out vigilantly.
 Alongside with \textbf{MSF}, we provide a good directed cell signaling network file, which is a subset of interactions from Reactome Functional interactions (FIs)
@@
 
+	\bibitem{edgeR}
+	Robinson, M.
+D., Mccarthy, D.
+J., \& Smyth, G.
+K.
+(2009).
+EdgeR: a Bioconductor package for differential expression analysis of digital gene expression data.
+Bioinformatics, 26(1), 139-140.
+doi:10.1093/bioinformatics/btp616
+
 \end{thebibliography}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Git/MSF_Paper/Tgfb1} 
-		\caption{}
+		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/Tgfb1} 
+		\caption{Input}
 	\end{subfigure}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Git/MSF_Paper/TGFBstep1}
-		\caption{}
+		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep1}
+		\caption{Step 1}
 	\end{subfigure}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Git/MSF_Paper/TGFBstep2} 
-		\caption{}
+		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep2} 
+		\caption{Step 2}
 	\end{subfigure}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Git/MSF_Paper/TGFBstep3}
-		\caption{}
+		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep3}
+		\caption{Step 3}
 	\end{subfigure}
-	\caption{Overview of \texttt{MSF} process showing the steps to identify the modulated sub-graphs.
-(a) showing a network of genes with green nodes as genes with \textit{p}-values $< $ 0.05, and blue nodes are genes with \textit{p}-values $>$ 0.05, (b) the genes circled red show the two initial modulated sub-graphs, sub-graph1 found one with M,K,C and the second modulated sub-graph2 S,V,T, (c) shows the modulated sub-graph1 being extended to M,K,C,A,D, (d) shows both the modulated sub-graphs merged with the addition of 3 genes Z,W,P.}
+	\caption{Overview of \texttt{MSF} process showing 3 step toy example to identify the modulated sub-graphs.
+(a) Input: showing a network of genes with green nodes as genes with \textit{p}-values $< $ 0.05, and blue nodes are genes with \textit{p}-values $>$ 0.05, (b) Step 1: genes circled red show the two initial modulated sub-graphs identified, sub-graph 1 with genes M,K,C and sub-graph 2 with genes S,V,T, (c) Step 2: shows the modulated sub-graph 1 being extended to M,K,C,A,D while no extension in sub-graph 2 (d) Step 3: shows both the modulated sub-graphs merging together with the addition of 3 genes Z,W,P.}
+			\end{minipage}
+}
+\end{figure*}
+
+\begin{figure*}[p]
+	\centering
+	\fbox{
+		\begin{minipage}{17 cm}
+			\includegraphics[width=16cm]{Documents/PhDThesis/MSFPaper/DEGvsMSF2}
+			\caption{Graph shows the percentage of DEG analysis genes and MSF identified sub-graph genes recall rates 100 times for the three different time points after the addition of Poisson distributed noise.}
+			\label{label}
 		\end{minipage}
@@
 		\begin{minipage}{17 cm}
-	\includegraphics[width=1.4\linewidth]{Git/MSF_Paper/newestMergedNetwork}
-	\caption{Shows connected modulated sub-graphs from 6hpi, where nodes are genes and edges are the interactions.
-The node coloring represents the predefined KEGG pathways referring to the nodes in the legend, the edges are from Reactome.
+	\includegraphics[width=1.05\linewidth]{ModulatesSubPaths/MSFData/Ebola6H/SeperatedMergedNetwork}
+	\caption{Shows the three modulated sub-graphs identified by MSF at 6hpi, where nodes are genes and edges are the interactions.
+The node coloring is associated to KEGG pathways referring to the colors in the legend.
+The graph edges are from Reactome.
 Important genes to EBOV infection as from literature are enlarged in the graph.
-The signal flow depicted in the graph is explained in the text.}
-		\end{minipage}
-}
-\end{figure*}
-
-\begin{figure*}[p]
-	\centering
-	\fbox{
-		\begin{minipage}{17 cm}
-			\includegraphics[width=16cm]{Git/MSF_Paper/DEGvsMSF2}
-			\caption{Graph shows the percentage of DEG analysis genes and MSF identified sub-graph genes recall rates 100 times for the three different time points after the addition of Poisson distributed noise.}
-			\label{label}
+The signal flow corresponding to each sub-graph is depicted as simple flow chart against each sub-graph.}
 		\end{minipage}

### 2018-03-19 MSF_Paper_FA.tex
.

--- 2018-03-01 MSF_Paper_FA.tex
@@
 between the conditions which are the log fold change, the sign of log fold change and the confidence level
-of observing an authentic change, often expressed as
+of observing a authentic change, often expressed as
 \textit{p}-value. These DEGs are
@@
 		\begin{minipage}{17 cm}
-	\includegraphics[width=1.4\linewidth]{Git/MSF_Paper/newestMergedNetwork}
+	\includegraphics[width=1.4\linewidth]{newestMergedNetwork.png}
 	\caption{Shows connected modulated sub-graphs from 6hpi, where nodes are genes and edges are the interactions.
@@
 		\begin{minipage}{17 cm}
-			\includegraphics[width=16cm]{Git/MSF_Paper/DEGvsMSF2}
+			\includegraphics[width=16cm]{DEGvsMSF2.png}
 			\caption{Graph shows the percentage of DEG analysis genes and MSF identified sub-graph genes recall rates 100 times for the three different time points after the addition of Poisson distributed noise.}

### 2018-03-19 MSF_Paper.tex
create make file; implement F1000 style; fix reference; write abstract; externalize reference

--- 2018-03-17 MSF_Paper_FA.tex
@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/Tgfb1} 
+		\includegraphics[width=10.0cm]{Tgfb1.png}
 		\caption{Input}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep1}
+		\includegraphics[width=10.0cm]{TGFBstep1.png}
 		\caption{Step 1}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep2} 
+		\includegraphics[width=10.0cm]{TGFBstep2.png}
 		\caption{Step 2}
@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep3}
+		\includegraphics[width=10.0cm]{TGFBstep3.png}
 		\caption{Step 3}
@@
 		\begin{minipage}{17 cm}
-			\includegraphics[width=16cm]{Documents/PhDThesis/MSFPaper/DEGvsMSF2}
+			\includegraphics[width=16cm]{DEGvsMSF2.png}
 			\caption{Graph shows the percentage of DEG analysis genes and MSF identified sub-graph genes recall rates 100 times for the three different time points after the addition of Poisson distributed noise.}
@@
 		\begin{minipage}{17 cm}
-	\includegraphics[width=1.05\linewidth]{ModulatesSubPaths/MSFData/Ebola6H/SeperatedMergedNetwork}
+	\includegraphics[width=1.05\linewidth]{SeperatedMergedNetwork.png}
 	\caption{Shows the three modulated sub-graphs identified by MSF at 6hpi, where nodes are genes and edges are the interactions.

### 2018-03-20 MSF_Paper_FA.tex
firest few paragraphs

--- 2018-03-17 MSF_Paper_FA.tex
@@
-\documentclass[twocolumn]{article}
-\usepackage{authblk}
-\usepackage{blindtext}
-\usepackage{multicol}% http://ctan.org/pkg/multicols
-\usepackage{graphicx}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Welcome to Overleaf --- just edit your article on the left,
+% and we'll compile it for you on the right. If you give
+% someone the link to this page, they can edit at the same
+% time. See the help menu above for more info. Enjoy!
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% For more detailed article preparation guidelines, please see:
+% http://f1000research.com/author-guidelines
+
+\documentclass[10pt,a4paper,twocolumn]{article}
+\usepackage{f1000_styles}
+
+%% Default: numerical citations
+\usepackage[numbers]{natbib}
+
+%% Uncomment this lines for superscript citations instead
+% \usepackage[super]{natbib}
+
+%% Uncomment these lines for author-year citations instead
+% \usepackage[round]{natbib}
+% \let\cite\citep
+
 \usepackage{subcaption}
-\usepackage{cite}
-\usepackage[table,xcdraw]{xcolor}
+\usepackage{hyperref}
 \newcommand{\FA}[1]{\begingroup\color{magenta}#1\endgroup}
@@
 
-\author[1]{Mariam R Farman}
+\author[1]{Mariam R. Farman}
 \author[1,2]{Fabian Amman}
-\author[1]{Ivo L Hofacker}
-\affil[1]{Institute for Theoretical Chemistry,Theoretical Biochemistry Group,University of Vienna,Austria}
-\affil[2]{Department of Chromosome Biology, Max F. Perutz Laboratories,University of Vienna,Austria}
+\author[1]{Ivo L. Hofacker}
+\affil[1]{Institute for Theoretical Chemistry,Theoretical Biochemistry Group,University of Vienna, Austria}
+\affil[2]{Department of Chromosome Biology, Max F. Perutz Laboratories,University of Vienna, Austria}
 
 \maketitle
+\thispagestyle{fancy}
 
 \begin{abstract}
-Differentially expressed genes play an important role in giving
-insights to a phenotype. These DEGs are then used to identify the
-pathways altered during changing biological conditions. We developed a
-novel approach to find the maximal significantly dis-regulated
-sub-graphs or cluster of genes from the cell signaling network, giving
-these sub-graphs an overall significance of modulation by combining the
-individual \textit{p}-values of the genes.
+
+High throughput techniques such as RNA-seq or microarray analysis have
+proven to be invaluable for the characterization of global
+transcriptional gene activity changes due to external stimuli or
+diseases. Differential gene expression analysis (DGEA) is the first
+step in the course of data interpretation, typically producing lists
+of of dozens to thousands of differentially expressed genes. To
+further guide the interpretation of these lists, different pathway
+analysis approaches have been developed. Most (\TODO{all?}) of which
+classify genes into set of genes, i.e. a pathway, which do interact
+with each other function in a common biological process. Regardless of
+technical differences, all these methods do not properly account for
+cross talk between different pathways and rely on binary
+classification into differentially expressed gene and unaffected genes
+based on an arbitrarily set p-value cut-off.
+
+To overcome this limitation, we developed a novel approach to identify
+concertedly modulated sub-graphs in the global cell signaling network,
+based on the DGEA results of all genes tested. Thereby, expression
+patterns of genes are integrated according to the topology of their
+interactions and allow potentially to read the flow of information
+from the perturbation source to the effectors. The described software,
+names \texttt{Modulated Sub-graph Finder} is freely available at
+\url{http//:path/to/git}.
+
 \end{abstract}
@@
 
-\section*{Background}
+\clearpage
+
+\section*{Introduction}
 
 High throughput sequencing techniques have been widely used to yield
-the differentially expressed genes (DEG)~\cite{DEG}. To this end,
-changes in transcript abundance are measured, e.g.~by next generation
+differentially expressed genes (DEG)~\cite{DEG}. To this end, changes
+in transcript abundance are measured, e.g.~by next generation
 sequencing techniques, and interpreted as an indicator of differential
-expression of genes. DEGs can be used to get
-insights into the mechanism underlying differences between conditions of samples, such as
-healthy versus diseased. Differential gene
-expression analysis (DGEA) informs about the magnitude of expression changes
-between the conditions which are often expressed as fold change, sign of fold change and the confidence level
-of observing a authentic change, often expressed as
-\textit{p}-value. These DEGs are
-then used as information converted from data, this information is further interpreted to knowledge to extract meaningful insights, for example the genes that
-could be important for a particular condition or maybe the target gene
-of an infection or disease. To this end, pathway-based analysis has become an
-important tool to further interpret the results of a DGEA and to
-acquire understandings of the perturbations in a biological
-system. Biological pathways are sets of genes and their interactions forming a functional
-unit. DEGs help to identify pathways or networks that may be altered
-during a change of condition providing important information about
-diseases and its treatment process~\cite{Khatri2012}. Pathway-based
-methods use predefined pathways or networks such as
-KEGG~\cite{Kegg} and Reactome~\cite{Reactome}, the expression
-measurements of the genes obtained from DGEA in combination with statistical methods and algorithms to
-identify specifically modulated pathways and processes~\cite{Campos}.
+expression of genes. DEGs can be used to get insights into the
+mechanism underlying differences between conditions of samples, such
+as healthy versus diseased. Differential gene expression analysis
+(DGEA) informs about the magnitude of expression changes between the
+conditions which are often expressed as fold change, sign of fold
+change and the confidence level of observing an authentic change,
+often expressed as \textit{p}-value. These DEGs information is further
+interpreted to extract meaningful biological insights. For example,
+genes that could be involved in the response to a particular stimuli
+or maybe the cause of a disease. To this end, pathway-based analysis
+has become an important tool to further interpret the results of a
+DGEA and to acquire understandings of the perturbations in a
+biological system. Biological pathways are sets of genes and their
+interactions forming a functional unit. DEGs help to identify pathways
+or networks that may be altered during a change of condition providing
+important information about diseases and its treatment
+process~\cite{Khatri2012}. Pathway-based methods use predefined
+pathways or networks such as KEGG~\cite{Kegg} and
+Reactome~\cite{Reactome}, the expression measurements of the genes
+obtained from DGEA in combination with statistical methods and
+algorithms to identify specifically modulated pathways and
+processes~\cite{Campos}.
 
-Predefined pathways are sets of genes and their interactions categorized in a functional unit differently by different databases known as pathways.
-For example, KEGG (Kyoto Encyclopedia of Genes and Genomes) is a database resource for understanding the functionality and utilities of the biological system.
-KEGG pathways is a branch of KEGG database that has a collection of manually drawn pathway maps representing the molecular interaction, reaction and relation networks of cellular functions~\cite{Kegg}.
-Similarly, Reactome is an open-source, manually curated, peer-reviewed and well known database for signaling and metabolic molecules with their interactions formed into different  biological pathways for nineteen species~\cite{Reactome}.
+Well established resources for pathway annotation are KEGG (Kyoto
+Encyclopedia of Genes and Genomes)\cite{Kegg} and
+Reactome~\cite{Reactome}. KEGG pathways is a branch of KEGG database
+that has a collection of manually drawn pathway maps representing the
+molecular interaction, reaction and relation networks of cellular
+functions. Similarly, Reactome is an open-source, manually curated,
+peer-reviewed database for signaling and metabolic molecules with
+their interactions formed into different biological pathways for
+nineteen species~\cite{Reactome}. Both provide predefined pathways
+which are sets of genes and their interactions categorized into
+functional units. Starting from a gene interaction network, genes are
+labeled according their role in a specific biological process. In this
+sense a particular gene can be assigned to different pathways. E.g.,
+the human gene STAT1 is associated with 24 different pathways in the
+pathway annotation curated by
+KEGG~\footnote{\url{http://www.genome.jp/dbget-bin/www_bget?hsa:6772}}
+and in 12 different pathways in the Reactome data
+base~\footnote{\url{http://www.reactome.org/content/detail/R-HSA-629622}}.
+Although carefully produced, the assignment of genes to those
+predefined pathway unites can be considered to be subjective to some
+degree and suffers from observatinal bias (\TODO{citation needed}).
+
+\TODO{FA until here}.
 
@@
 analysis), FCS (Functional class scoring) and pathway topology based
-methods. ORA is the first and the most basic method of pathway
-analysis. It uses a DEG list with user defined cut-off for the log-fold change and
-\textit{p}-value (most commonly using absolute log-fold change
-$\geq$ 2 and \textit{p}-value $\leq$ 0.05). Subsequently, sets of genes associated
-with annotated pathways are tested for being over-represented in the
-set of DEGs. To this end, hyper-geometric
-distribution, chi-square tests, binomial probability or the Fisher’s
-exact test are used. Thereby the information of the topology of genes in the
-pathways are ignored~\cite{Bayer}. Furthermore, ORA assumes that the biological
-pathways are independent of each other and ignores the fact that they cross-talk and overlap~\cite{Khatri2012,Campos}.
+methods. All of which aim in finding in a subset of genes, e.g.,
+significantly differentially expressed genes, genes associated with a
+certain pathway more often than expected given the total set of
+examined genes, e.g. the genome background.  ORA is the first and the
+most basic method of pathway analysis. It uses a DEG list with user
+defined cut-off for the log-fold change and \textit{p}-value (most
+commonly using absolute log-fold change $\geq$ 2 and \textit{p}-value
+$\leq$ 0.05). Subsequently, sets of genes associated with annotated
+pathways are tested for being over-represented in the set of DEGs. To
+this end, hyper-geometric distribution, chi-square tests, binomial
+probability or the Fisher’s exact test are used. Thereby the
+information of the topology of genes in the pathways are
+ignored~\cite{Bayer}. Furthermore, ORA assumes that the biological
+pathways are independent of each other and ignores the fact that they
+cross-talk and overlap~\cite{Khatri2012,Campos}.
 
-Unlike ORA, FCS has no artificial cut-off to define DEG list. FCS works in three
-step, first it calculates the gene-level statistics including correlation of molecular measurements  using differential expression of individual genes, ANOVA, t-test and  Z-score.
-In the second step the statistics of individual genes in a pathway are transformed to an individual pathway-level statistic commonly using Kolmogorov-Smirnov statistic, mean or median.
-Finally the statistical significance of the pathway-level statistics is assessed.
-Although FCS covers some of the
-limitations from ORA, it still lacks the topology of genes in a pathway, cross-talk and
-overlap of the pathways~\cite{Khatri2012,Campos}. Pathway topology
-based methods are similar to FCS except that they consider the
-topology of each gene during the gene-level
-statistics but still are unable to link different pathways~\cite{Khatri2012}.
+Unlike ORA, FCS has no artificial cut-off to define DEG list. FCS
+works in three step, first it calculates the gene-level statistics
+including correlation of molecular measurements using differential
+expression of individual genes, ANOVA, t-test and Z-score. In the
+second step the statistics of individual genes in a pathway are
+transformed to an individual pathway-level statistic commonly using
+Kolmogorov-Smirnov statistic, mean or median. Finally the statistical
+significance of the pathway-level statistics is assessed. Although FCS
+covers some of the limitations from ORA, it still lacks the topology
+of genes in a pathway, cross-talk and overlap of the
+pathways~\cite{Khatri2012,Campos}. Pathway topology based methods are
+similar to FCS except that they consider the topology of each gene
+during the gene-level statistics but still are unable to link
+different pathways~\cite{Khatri2012}.
 
@@
 pathway annotation resources available to gain additional functional
-insights from basic DGEA. To this, we
-start with the presupposition that expression of neighboring genes
-within a functional pathway are not independent from each
-other. Rather, they are often regulating each others expression or are
-part of the same regulon~\cite{Michalak}. We
-understand that the categorization of links between genes into labeled
-pathways is often an arbitrary one, given the extensive cross talk
-between different pathways. Although these categories have proven to be
-useful in many situations, they force a certain perspective onto the
-interpretation of novel data. Based on these two principles, we aim to
-find sub-graphs of connected genes within cell signaling network which exhibit as a whole
+insights from basic DGEA. To this, we start with the presupposition
+that expression of neighboring genes within a functional pathway are
+not independent from each other. Rather, they are often regulating
+each others expression or are part of the same
+regulon~\cite{Michalak}. We understand that the categorization of
+links between genes into labeled pathways is often an arbitrary one,
+given the extensive cross talk between different pathways. Although
+these categories have proven to be useful in many situations, they
+force a certain perspective onto the interpretation of novel
+data. Based on these two principles, we aim to find sub-graphs of
+connected genes within cell signaling network which exhibit as a whole
 significant differential expression changes. As input, information on
@@
 implemented as a java program, which was named Modulated Sub-graph
-Finder (\texttt{MSF}). \texttt{MSF} would help transform the information obtained from DGEA into understandable knowledge of signal transduction of genes. \texttt{MSF} is freely accessible on github.
+Finder (\texttt{MSF}). \texttt{MSF} would help transform the
+information obtained from DGEA into understandable knowledge of signal
+transduction of genes. \texttt{MSF} is freely accessible on github.
 
@@
 find the modulated sub-graphs of gene interactions from the cell
-signaling network.
-\texttt{MSF} does not use predefined gene sets in functional units, it builds up the sub-graphs from scratch using the whole cell signaling network from Reactome.
-The network has nodes corresponding to genes and edges representing their directed regulating interactions.
+signaling network. \texttt{MSF} does not use predefined gene sets in
+functional units, it builds up the sub-graphs from scratch using the
+whole cell signaling network from Reactome. The network has nodes
+corresponding to genes and edges representing their directed
+regulating interactions.
 
 \texttt{MSF} uses the individual gene's \textit{p}-values generated
-from the DGEA. The \textit{p}-value
-expresses the probability that the null hypothesis of unmodified gene
-expression can be rejected for a given statistical model. To find
-significantly modulated sub-graphs individual \textit{p}-values of the
-vicinal genes in the global network are combined into a single
-combined \textit{p}-value, using a statistical method for combining
-dependent \textit{p}-values described by Hartung~\cite{Hartung}. Hartung uses the inverse of standard normal distribution function, individual gene \textit{p}-values are
-first transformed to their corresponding normal score. Then using these normal scores, the correlation between genes is calculated, the normal scores and correlation are applied to
-the inverse normal function to calculate the individual
-\textit{p}-value for a sub-graph. An
-overall \textit{p}-value of a sub-graph will express the significance
-of all genes in the sub-graph being modulated together.
+from the DGEA. The \textit{p}-value expresses the probability that the
+null hypothesis of unmodified gene expression can be rejected for a
+given statistical model. To find significantly modulated sub-graphs
+individual \textit{p}-values of the vicinal genes in the global
+network are combined into a single combined \textit{p}-value, using a
+statistical method for combining dependent \textit{p}-values described
+by Hartung~\cite{Hartung}. Hartung uses the inverse of standard normal
+distribution function, individual gene \textit{p}-values are first
+transformed to their corresponding normal score. Then using these
+normal scores, the correlation between genes is calculated, the normal
+scores and correlation are applied to the inverse normal function to
+calculate the individual \textit{p}-value for a sub-graph. An overall
+\textit{p}-value of a sub-graph will express the significance of all
+genes in the sub-graph being modulated together.
 
@@
 \texttt{MSF} applies a three steps heuristic as described in the
-following.The pipeline of identification of modulated sub-graphs from a network by \texttt{MSF} is depicted in
-Fig (1).
+following.The pipeline of identification of modulated sub-graphs from
+a network by \texttt{MSF} is depicted in Fig (1).
 \newline
@@
 associated with the lowest (most significant) \textit{p}-value from
-the network. It tries to extend the
-sub-graph to the most significant neighboring gene. A single combined
-\textit{p}-value is calculated for the two genes. If the combined
-\textit{p}-value is smaller than the minimal individual gene's
-\textit{p}-value, the extended sub-graph is accepted.
-Then the next most significant neighboring gene is added to the sub-graph and the combined \textit{p}-value is calculated, if the new combined \textit{p}-value of three genes is smaller than the combined \textit{p}-value of the first two genes, the extended sub-graph is accepted.
-This step is iteratively repeated until no further extension is accepted.
-In this case the process starts over with all remaining genes not yet in a
-significantly modulated sub-graph. This step identifies all the trivial
-sub-graphs that are modulated in the whole network~(Fig.~1b).\newline
+the network. It tries to extend the sub-graph to the most significant
+neighboring gene. A single combined \textit{p}-value is calculated for
+the two genes. If the combined \textit{p}-value is smaller than the
+minimal individual gene's \textit{p}-value, the extended sub-graph is
+accepted. Then the next most significant neighboring gene is added to
+the sub-graph and the combined \textit{p}-value is calculated, if the
+new combined \textit{p}-value of three genes is smaller than the
+combined \textit{p}-value of the first two genes, the extended
+sub-graph is accepted. This step is iteratively repeated until no
+further extension is accepted. In this case the process starts over
+with all remaining genes not yet in a significantly modulated
+sub-graph. This step identifies all the trivial sub-graphs that are
+modulated in the whole network~(Fig.~1b).\newline
 
@@
 they could further be extended beyond the immediate neighborhood.
-This is done by testing all possible extension by N genes (default two genes) for
-all genes in the sub-graph. Again, this step is iteratively repeated until no
-further genes are added to the significant differentially expressed
-sub-graphs. This steps bridges small gaps of genes without a clear
-differential signal in the DGEA~(Fig.~1c).\newline
+This is done by testing all possible extension by N genes (default two
+genes) for all genes in the sub-graph. Again, this step is iteratively
+repeated until no further genes are added to the significant
+differentially expressed sub-graphs. This steps bridges small gaps of
+genes without a clear differential signal in the
+DGEA~(Fig.~1c).\newline
 
@@
 After detection and extension of the modulated sub-graphs, they are
-tested if combined sub-graph score is better than on their own.
-The merging of the two sub-graphs is done by depth first search.
-If the two sub-graphs merge with the connector of at most N genes (default two genes) and the
-combined \textit{p}-value of the merged sub-graph including the
-bridging genes in between is less than the individual
+tested if combined sub-graph score is better than on their own. The
+merging of the two sub-graphs is done by depth first search. If the
+two sub-graphs merge with the connector of at most N genes (default
+two genes) and the combined \textit{p}-value of the merged sub-graph
+including the bridging genes in between is less than the individual
 \textit{p}-values of the two sub-graphs, the two sub-graphs are merged
-together to one big modulated sub-graph~(Fig.~1d). This step is repeated iteratively until no sub-graphs could be merged.\newline
+together to one big modulated sub-graph~(Fig.~1d). This step is
+repeated iteratively until no sub-graphs could be merged.\newline
 
@@
 In a last post processing step \texttt{MSF} identifies the trigger
-points of the modulated sub-graphs. These trigger genes are the sources
-of the sub-graphs with only outgoing edges. These genes can be
-interpreted as the possible entry points of perturbation from where the
-stimulus causes downstream effects. In the same spirit the most
+points of the modulated sub-graphs. These trigger genes are the
+sources of the sub-graphs with only outgoing edges. These genes can be
+interpreted as the possible entry points of perturbation from where
+the stimulus causes downstream effects. In the same spirit the most
 downstream genes of the modulated sub-graph are identified and defined
@@
 integrated information within the signal transduction network is set
-to action. Genes in circular loops in the sub-graphs have no source or sink.
+to action. Genes in circular loops in the sub-graphs have no source or
+sink.
 
@@
 
-To assess the robustness and stability of our method, Poisson distributed noise was added to the sample read counts.
-Then DGEA was carried out on the disturbed data with the same parameters as for the native data, followed by analysis with \texttt{MSF}.
-Using this criteria, the goal was to spot almost
-identical modulated sub-graphs in the real and noisy data.
-This procedure was carried out 100 times and every time the genes from the sub-graphs identified from noisy data were compared to the genes of sub-graphs identified from the native data.
-This was also done for the DEG obtained for each run, a cutoff was used to compare the retrieved DEG from noisy data to the native data.
-The robustness of \texttt{MSF} and the DEG analysis for the three time-points are shown in~(Fig.~2).
-We were very stringent in adding noise to the counts, which could be seen by the DEG analysis robustness.
-For the last time-point (48hpi) 84\% genes of the sub-graphs from native data were retrieved, for 24hpi robustness was about 79\%.
-Robustness was 71\% at 6hpi, this could be due to that the infection merely started and the signal transdcution was weak as compared to the later time-points.
+To assess the robustness and stability of our method, Poisson
+distributed noise was added to the sample read counts. Then DGEA was
+carried out on the disturbed data with the same parameters as for the
+native data, followed by analysis with \texttt{MSF}. Using this
+criteria, the goal was to spot almost identical modulated sub-graphs
+in the real and noisy data. This procedure was carried out 100 times
+and every time the genes from the sub-graphs identified from noisy
+data were compared to the genes of sub-graphs identified from the
+native data. This was also done for the DEG obtained for each run, a
+cutoff was used to compare the retrieved DEG from noisy data to the
+native data.  The robustness of \texttt{MSF} and the DEG analysis for
+the three time-points are shown in~(Fig.~2). We were very stringent in
+adding noise to the counts, which could be seen by the DEG analysis
+robustness. For the last time-point (48hpi) 84\% genes of the
+sub-graphs from native data were retrieved, for 24hpi robustness was
+about 79\%. Robustness was 71\% at 6hpi, this could be due to that the
+infection merely started and the signal transdcution was weak as
+compared to the later time-points.
 
-Simulation testing was also conducted on \texttt{MSF}, the objective was to find less sources than observed in the sub-graphs identified by \texttt{MSF} from native data.
-Simulation results showed a chance of less than 10 percent to get more sources than observed.
+Simulation testing was also conducted on \texttt{MSF}, the objective
+was to find less sources than observed in the sub-graphs identified by
+\texttt{MSF} from native data. Simulation results showed a chance of
+less than 10 percent to get more sources than observed.
 
@@
 set of primary human monocyte-derived macrophages (MDMs) infected with
-Ebola virus (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV)
-belongs to the Filoviridea family; filamentous, enveloped and single
-stranded RNA viruses. EBOV causes hemorrhagic fever in humans,
-inducing the host innate and adaptive immune response to be unable to
-control virus infection~\cite{Prins}. Until now there are no approved
-antiviral drugs for the treatment of Ebola virus infection~\cite{Konde,Rhein}. 
-The initial targets of EBOV are the macrophages and
-dendritic immune cells~\cite{Falasca,Rhein}. EBOV inhibits the critical
-innate immune response of the host, which includes the activation of
-alpha/beta interferon (IFN-$\alpha /
-\beta$)~\cite{Prins,Konde,Cardenas}. The Ebola viral protein VP35
-targets the host type~I interferon IFN-$\alpha / \beta$ to block the
-early antiviral immune response
+Ebola virus (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV) belongs to
+the Filoviridea family; filamentous, enveloped and single stranded RNA
+viruses. EBOV causes hemorrhagic fever in humans, inducing the host
+innate and adaptive immune response to be unable to control virus
+infection~\cite{Prins}. Until now there are no approved antiviral
+drugs for the treatment of Ebola virus infection~\cite{Konde,Rhein}.
+The initial targets of EBOV are the macrophages and dendritic immune
+cells~\cite{Falasca,Rhein}. EBOV inhibits the critical innate immune
+response of the host, which includes the activation of alpha/beta
+interferon (IFN-$\alpha / \beta$)~\cite{Prins,Konde,Cardenas}. The
+Ebola viral protein VP35 targets the host type~I interferon
+IFN-$\alpha / \beta$ to block the early antiviral immune response
 ~\cite{Prins,Konde,Falasca,Cardenas,Olejnik}. It has been proposed
-that IFN-$\alpha / \beta$  should be
-tested against Ebola for its antiviral activity through clinical trials~\cite{Konde}.
-The aim of testing the EBOV infection data with \texttt{MSF} was to identify the modulated sub-graphs and check if \texttt{MSF} is able to identify the IFN-$\alpha / \beta$ gene as one of the sources.
+that IFN-$\alpha / \beta$ should be tested against Ebola for its
+antiviral activity through clinical trials~\cite{Konde}. The aim of
+testing the EBOV infection data with \texttt{MSF} was to identify the
+modulated sub-graphs and check if \texttt{MSF} is able to identify the
+IFN-$\alpha / \beta$ gene as one of the sources.
 
-EBOV infection count data was downloaded from GEO (GSE84188).
-Differential gene expression analysis was performed on the count data with edgeR package (version 3.4.2)~\cite{edgeR}.
-The DEG list generated by edgeR was used as an input for \texttt{MSF}.
-The EBOV infection data has three time-points 6, 24  and 48 hour post infection (hpi).
-At 6hpi, five modulated sub-graphs were identified.
-Most of the genes part of the sub-graphs were chemokines (CXCL10, CCL8) and Interleukin genes (IL6, IL27, IL23) which serve as a subset of CD4+ T helper
-cells.
-IFNB1 and IFNA1 were both identified as two of the possible sources.
-Most of the sources identified by \texttt{MSF} were type~I interferon induced genes.
-At 24hpi nine modulated sub-graphs were identified, IFNA1 was identified as one of the sources.
-For the last time-point again IFNB1 and IFNA1 were identified as the two sources out of the possible sources from seven modulated sub-graphs identified.
+EBOV infection count data was downloaded from GEO
+(GSE84188). Differential gene expression analysis was performed on the
+count data with edgeR package (version 3.4.2)~\cite{edgeR}. The DEG
+list generated by edgeR was used as an input for \texttt{MSF}. The
+EBOV infection data has three time-points 6, 24 and 48 hour post
+infection (hpi). At 6hpi, five modulated sub-graphs were
+identified. Most of the genes part of the sub-graphs were chemokines
+(CXCL10, CCL8) and Interleukin genes (IL6, IL27, IL23) which serve as
+a subset of CD4+ T helper cells. IFNB1 and IFNA1 were both identified
+as two of the possible sources. Most of the sources identified by
+\texttt{MSF} were type~I interferon induced genes. At 24hpi nine
+modulated sub-graphs were identified, IFNA1 was identified as one of
+the sources. For the last time-point again IFNB1 and IFNA1 were
+identified as the two sources out of the possible sources from seven
+modulated sub-graphs identified.
 
@@
 genes of Ebola infection. We were able to successfully identify IFNA1
-as a source in all three time-points and INFB1 in two of the time-points. Although IFNA1 and IFNB1 were two of the most
-significant gene in the later time points, \texttt{MSF} was able to detect them as
+as a source in all three time-points and INFB1 in two of the
+time-points. Although IFNA1 and IFNB1 were two of the most significant
+gene in the later time points, \texttt{MSF} was able to detect them as
 a source in the early time-point when the genes were not significant
-based on the individual DGEA alone.
-Identifying these sources will reduce the search space for potential target genes and can help the biologist as the starting point of clinical testing for drugs and vaccines against an infection.
+based on the individual DGEA alone. Identifying these sources will
+reduce the search space for potential target genes and can help the
+biologist as the starting point of clinical testing for drugs and
+vaccines against an infection.
 
@@
 
- Three main modulated sub-graphs identified by \texttt{MSF} at 6hpi are shown in~(Fig.~3).
-The graph shows different colored nodes as genes associated to KEGG pathways and how they cross talk.
-The KEGG association and visualization on the modulated sub-graphs was done using StringApp~\cite{StringApp} in Cytoscape~\cite{Cyto}.
-The graph unveils how more than one stimulus coincides to prevent the progression of infection.
-Sub-graph 1 shows toll-like receptor, cytokine, chemokine and jak-stat genes leading to TNF and eventually going into apoptosis.
-Sub-graph 2 shows Extra-cellular matrix (ECM) receptor interaction genes which are suggested to be in interaction with Ebola glycoprotein (GP)~\cite{Veljkovic}.
-ECM receptor, Focal adhesion and P13k-Akt signaling genes are the sources from one end of the sub-graph.
-P13k-Akt signaling goes to FoxO signaling which eventually leads to cell cycle, which at later time-points is disrupted.
-From the other end of sub-graph 2 toll-like recpetor combined with cytosolic DNA sensing and cytokine-chemokine signaling activates Jak-stat which leads to the start of apoptosis with a activation of a few apoptosis genes.
-Cytokine is also seen to sink into natural killer mediated cytotoxicity.
-Sub-graph 3 shows how some of the cytokine and chemokine genes activate the natural killer mediated cytotoxicity and apoptosis.
+ Three main modulated sub-graphs identified by \texttt{MSF} at 6hpi
+ are shown in~(Fig.~3). The graph shows different colored nodes as
+ genes associated to KEGG pathways and how they cross talk. The KEGG
+ association and visualization on the modulated sub-graphs was done
+ using StringApp~\cite{StringApp} in Cytoscape~\cite{Cyto}. The graph
+ unveils how more than one stimulus coincides to prevent the
+ progression of infection. Sub-graph 1 shows toll-like receptor,
+ cytokine, chemokine and jak-stat genes leading to TNF and eventually
+ going into apoptosis. Sub-graph 2 shows Extra-cellular matrix (ECM)
+ receptor interaction genes which are suggested to be in interaction
+ with Ebola glycoprotein (GP)~\cite{Veljkovic}. ECM receptor, Focal
+ adhesion and P13k-Akt signaling genes are the sources from one end of
+ the sub-graph. P13k-Akt signaling goes to FoxO signaling which
+ eventually leads to cell cycle, which at later time-points is
+ disrupted. From the other end of sub-graph 2 toll-like recpetor
+ combined with cytosolic DNA sensing and cytokine-chemokine signaling
+ activates Jak-stat which leads to the start of apoptosis with a
+ activation of a few apoptosis genes. Cytokine is also seen to sink
+ into natural killer mediated cytotoxicity. Sub-graph 3 shows how some
+ of the cytokine and chemokine genes activate the natural killer
+ mediated cytotoxicity and apoptosis.
 
@@
 
-To further understand the functional role of the sub-graphs identified by \texttt{MSF}, we linked the  sub-graphs to the predefined KEGG pathways and Reactome
-pathways through gene enrichment analysis.
-We used Cytoscape~\cite{Cyto} plugins to identify the enriched patwhays, CytoKegg~\cite{Cytokegg} for KEGG and Reactome FI~\cite{Reactome} for Reactome.
-We compared the pathways enriched in \texttt{MSF} identified sub-graphs to pathway analysis by SPIA~\cite{Tarca} for KEGG and Reactome pathway enrichment analysis for Reactome.
+To further understand the functional role of the sub-graphs identified
+by \texttt{MSF}, we linked the sub-graphs to the predefined KEGG
+pathways and Reactome pathways through gene enrichment analysis. We
+used Cytoscape~\cite{Cyto} plugins to identify the enriched patwhays,
+CytoKegg~\cite{Cytokegg} for KEGG and Reactome FI~\cite{Reactome} for
+Reactome. We compared the pathways enriched in \texttt{MSF} identified
+sub-graphs to pathway analysis by SPIA~\cite{Tarca} for KEGG and
+Reactome pathway enrichment analysis for Reactome.
 
-At 6hpi Cytokegg identified the sub-graph genes to be enriched in Toll-like receptor signaling, TNF signaling, IL-17 signaling and NF-kappa signaling pathways, as also stated by~\cite{Olejnik}.
-The earliest response EBOV induces on cytokine is via TLR4-mediated signaling~\cite{Olejnik}, Gene enrichment analysis of sub-graph genes at 6hpi showed toll-like receptor signaling being most significantly dis-regulated pathway, this important signaling pathway was not identified by SPIA along with RIG-I-like receptor signaling pathway, TNF signaling pathway and IL-17 signaling (Table 1 in the supplement material).
-On the later time-points \texttt{MSF} and SPIA showed consensus on the important pathways e.g Chemokine signaling, Cytokine-Cytokine receptor interaction, NF-kappa B signaling, Cytosolic DNA-sensing and RIG-I-like receptor signaling as being dis-regulated (Table 2-3 in the supplement).
+At 6hpi Cytokegg identified the sub-graph genes to be enriched in
+Toll-like receptor signaling, TNF signaling, IL-17 signaling and
+NF-kappa signaling pathways, as also stated by~\cite{Olejnik}. The
+earliest response EBOV induces on cytokine is via TLR4-mediated
+signaling~\cite{Olejnik}, Gene enrichment analysis of sub-graph genes
+at 6hpi showed toll-like receptor signaling being most significantly
+dis-regulated pathway, this important signaling pathway was not
+identified by SPIA along with RIG-I-like receptor signaling pathway,
+TNF signaling pathway and IL-17 signaling (Table 1 in the supplement
+material). On the later time-points \texttt{MSF} and SPIA showed
+consensus on the important pathways e.g Chemokine signaling,
+Cytokine-Cytokine receptor interaction, NF-kappa B signaling,
+Cytosolic DNA-sensing and RIG-I-like receptor signaling as being
+dis-regulated (Table 2-3 in the supplement).
 
-Likewise we compared gene enrichment results of sub-graph genes identified by \texttt{MSF} using Reactome FI to the Reactome predefined pathway enrichment analysis.
-Although \texttt{MSF} and Reactome showed good agreement on the dis-regulated pathways e.g Signaling by Interleukins, Interleukin-10 signaling, Cytokine Signaling in Immune system, Chemokine receptors bind chemokines, RIG-I/MDA5 mediated induction of IFN-alpha/beta pathways, NF-kB activation, Reactome as well could not identify Toll-like receptor signaling (Table 1-3 in the supplement material) .
-This shows that \texttt{MSF} is able to identify the important modulated genes in the sub-graphs at the earliest time-point of infection.
+Likewise we compared gene enrichment results of sub-graph genes
+identified by \texttt{MSF} using Reactome FI to the Reactome
+predefined pathway enrichment analysis. Although \texttt{MSF} and
+Reactome showed good agreement on the dis-regulated pathways e.g
+Signaling by Interleukins, Interleukin-10 signaling, Cytokine
+Signaling in Immune system, Chemokine receptors bind chemokines,
+RIG-I/MDA5 mediated induction of IFN-alpha/beta pathways, NF-kB
+activation, Reactome as well could not identify Toll-like receptor
+signaling (Table 1-3 in the supplement material) . This shows that
+\texttt{MSF} is able to identify the important modulated genes in the
+sub-graphs at the earliest time-point of infection.
 
@@
 
-\textbf{MSF} is a fast and robust tool to find modulated sub-graphs from a given network.
-So far, it is the only tool from existing approaches that builds the modulated sub-graph from scratch using the network and DEG, where the predefined pathways could be seen connected to each other.
-The Robustness analysis shows that \textbf{MSF} is fairly robust against noise and thus finds core modulated sub-graphs even if the DGEA is not carried out vigilantly.
-Alongside with \textbf{MSF}, we provide a good directed cell signaling network file, which is a subset of interactions from Reactome Functional interactions (FIs)
-Version 2016 .
+\textbf{MSF} is a fast and robust tool to find modulated sub-graphs
+from a given network. So far, it is the only tool from existing
+approaches that builds the modulated sub-graph from scratch using the
+network and DEG, where the predefined pathways could be seen connected
+to each other. The Robustness analysis shows that \textbf{MSF} is
+fairly robust against noise and thus finds core modulated sub-graphs
+even if the DGEA is not carried out vigilantly. Alongside with
+\textbf{MSF}, we provide a good directed cell signaling network file,
+which is a subset of interactions from Reactome Functional
+interactions (FIs) Version 2016 .
 
@@
 
-The Ebola infection RNAseq dataset analyzed during the current study are
-available in the GEO repository (GSE84188). The cell signaling
+The Ebola infection RNAseq dataset analyzed during the current study
+are available in the GEO repository (GSE84188). The cell signaling
 network file used is from Reactome Functional interactions (FIs)
@@
 \subsection*{Author contributions}
-FA proposed the idea. MF implemented the method. MF designed the benchmark
-data set. MF analyzed the benchmark results. MF, FA, and IH contributed to
-the manuscript writing. MF produced all figures. All authors read and
-approved the final manuscript.
+FA proposed the idea. MF implemented the method. MF designed the
+benchmark data set. MF analyzed the benchmark results. MF, FA, and IH
+contributed to the manuscript writing. MF produced all figures. All
+authors read and approved the final manuscript.
 
@@
 
-This work was funded by the FWF (“Fonds zur Forderung der wissenschaftlichen Forschung”) within the project Internationalen Kooperationsprojektes - Intl cooperation Project (Joint Project - Lead Agency Verfahren) with the project number (I 1988-B22).
-The grant was assigned to Ivo Hofacker
+This work was funded by the FWF (“Fonds zur Forderung der
+wissenschaftlichen Forschung”) within the project Internationalen
+Kooperationsprojektes - Intl cooperation Project (Joint Project - Lead
+Agency Verfahren) with the project number (I 1988-B22). The grant was
+assigned to Ivo Hofacker
 
@@
 
-We thank all our colleagues who provided insight and expertise that greatly assisted the research and all reviewers for their constructive criticism.
+We thank all our colleagues who provided insight and expertise that
+greatly assisted the research and all reviewers for their constructive
+criticism.
 
@@
 
-The supplementary file contains tables of comparison between \textbf{MSF} identified sub-graph gene enrichment with SPIA and Reactome pathway analysis. 
-\clearpage
+The supplementary file contains tables of comparison between
+\textbf{MSF} identified sub-graph gene enrichment with SPIA and
+Reactome pathway analysis.
 
-\begin{thebibliography}{9}
+{\small\bibliographystyle{unsrtnat}
+\bibliography{MSF_Paper_FA}}
 
-	\bibitem{DEG}
-	Malone, J. H., \& Oliver, B. (2011). Microarrays, deep sequencing and the true measure of the transcriptome. BMC Biology, 9(1), 34. doi:10.1186/1741-7007-9-34
-	\bibitem{Khatri2012}
-	Khatri, P., Sirota, M., \& Butte, A. J. (2012). Ten Years of Pathway Analysis: Current Approaches and Outstanding Challenges. PLoS Computational Biology, 8(2). doi:10.1371/journal.pcbi.1002375
-	\bibitem{Kegg}
-	Kanehisa, M. (2000). KEGG: Kyoto Encyclopedia of Genes and Genomes. Nucleic Acids Research, 28(1), 27-30. doi:10.1093/nar/28.1.27
-	\bibitem{Reactome}
-	Croft, D., \& Croft, D. (2010). Reactome: A database of biological pathways. Nature Precedings. doi:10.1038/npre.2010.5025
-	\bibitem{Campos}
-	García-Campos, M. A., Espinal-Enríquez, J., \& Hernández-Lemus, E. (2015). Pathway Analysis: State of the Art. Frontiers in Physiology, 6. doi:10.3389/fphys.2015.00383 
-	\bibitem{Bayer}
-	Bayerlová, M., Jung, K., Kramer, F., Klemm, F., Bleckmann, A., \& Beißbarth, T.
-(2015).
-Comparative study on gene set and pathway topology-based enrichment methods.
-BMC Bioinformatics, 16(1).
-doi:10.1186/s12859-015-0751-5
-	\bibitem{Hartung}
-	Hartung, J. (1999). A Note on Combining Dependent Tests of Significance. Biometrical Journal, 41(7), 849-855. doi:10.1002/(sici)1521-4036(199911)41:7<849::aid-bimj849>3.0.co;2-t 5
-
-	\bibitem {Prins}
-	Prins, K., Cardenas, W.
-and Basler, C.
-(2009).
-Ebola Virus Protein VP35 Impairs the Function of Interferon Regulatory Factor-Activating Kinases IKK and TBK-1.
-Journal of Virology, 83(7), pp.3069-3077.
-
-	\bibitem {Konde}
-	Konde, M., Baker, D., Traore, F., Sow, M., Camara, A., Barry, A., Mara, D., Barry, A., Cone, M., Kaba, I., Richard, A., Beavogui, A., Günther, S., Pintilie, M.
-and Fish, E.
-(2017).
-Interferon β-1a for the treatment of Ebola virus disease: A historically controlled, single-arm proof-of-concept trial.
-PLOS ONE, 12(2).
-
-	\bibitem {Rhein}
-	Rhein, B., Powers, L., Rogers, K., Anantpadma, M., Singh, B., Sakurai, Y., Bair, T., Miller-Hunt, C., Sinn, P., Davey, R., Monick, M.
-and Maury, W.
-(2015).
-Interferon-γ Inhibits Ebola Virus Infection.
-PLOS Pathogens, 11(11), p.e1005263.
-
-	\bibitem {Falasca}
-	Falasca, L., Agrati, C., Petrosillo, N., Di Caro, A., Capobianchi, M., Ippolito, G. and Piacentini, M. (2015).
-	 Molecular mechanisms of Ebola virus pathogenesis: focus on cell death. Cell Death and Differentiation, 22(8), pp.1250-1259.
-
-	\bibitem {Cardenas}
-	Cardenas, W., Loo, Y., Gale, M., Hartman, A., Kimberlin, C., Martinez-Sobrido, L., Saphire, E.
-and Basler, C.
-(2006).
-Ebola Virus VP35 Protein Binds Double-Stranded RNA and Inhibits Alpha/Beta Interferon Production Induced by RIG-I Signaling.
-Journal of Virology, 80(11), pp.5168-5178.
-
-	\bibitem {Olejnik}
-	Olejnik, J., Forero, A., Deflubé, L., Hume, A., Manhart, W., Nishida, A., Marzi, A., Katze, M., Ebihara, H., Rasmussen, A.
-and Mühlberger, E.
-(2017).
-Ebolaviruses Associated with Differential Pathogenicity Induce Distinct Host Responses in Human Macrophages.
-Journal of Virology, 91(11), pp.e00179-17.
-
-	\bibitem {Dilley}
-	Dilley, K., Voorhies, A., Luthra, P., Puri, V., Stockwell, T., Lorenzi, H., Basler, C.
-and Shabman, R.
-(2017).
-The Ebola virus VP35 protein binds viral immunostimulatory and host RNAs identified through deep sequencing.
-PLOS ONE, 12(6), p.e0178717.
-
-	\bibitem{Michalak}
-	Michalak, P. (2008). Coexpression, coregulation, and cofunctionality of neighboring genes in eukaryotic genomes. Genomics, 91(3), 243-248. doi:10.1016/j.ygeno.2007.11.002
-
-	\bibitem{Cyto}
-	http://www.cytoscape.org/
-
-	\bibitem{Cytokegg}
-	http://apps.cytoscape.org/apps/cytokegg
-
-	\bibitem{Reactome}
-	https://wiki.reactome.org/index.php?title=ReactomeFIViz
-
-	\bibitem{Tarca}
-	Tarca, A.
-L., Draghici, S., Khatri, P., Hassan, S.
-S., Mittal, P., Kim, J., … Romero, R.
-(2009).
-A novel signaling pathway impact analysis.
-Bioinformatics, 25(1), 75–82.
-http://doi.org/10.1093/bioinformatics/btn577
-
-	\bibitem{StringApp}
-	http://apps.cytoscape.org/apps/stringapp
-
-	\bibitem{Veljkovic}
-	Veljkovic, V., Glisic, S., Muller, C.
-P., Scotch, M., Branch, D.
-R., Perovic, V.
-R., … Colombatti, A.
-(2015).
-In silico analysis suggests interaction between Ebola virus and the extracellular matrix.
-Frontiers in Microbiology, 6, 135.
-http://doi.org/10.3389/fmicb.2015.00135
-
-	\bibitem{edgeR}
-	Robinson, M.
-D., Mccarthy, D.
-J., \& Smyth, G.
-K.
-(2009).
-EdgeR: a Bioconductor package for differential expression analysis of digital gene expression data.
-Bioinformatics, 26(1), 139-140.
-doi:10.1093/bioinformatics/btp616
-
-\end{thebibliography}
+\bigskip
 
@@
 		\centering
-		\includegraphics[width=10.0cm]{Documents/PhDThesis/MSFPaper/TGFBstep3}
+		\includegraphics[width=10.0cm]{TGFBstep3.png}
 		\caption{Step 3}
 	\end{subfigure}
-	\caption{Overview of \texttt{MSF} process showing 3 step toy example to identify the modulated sub-graphs.
-(a) Input: showing a network of genes with green nodes as genes with \textit{p}-values $< $ 0.05, and blue nodes are genes with \textit{p}-values $>$ 0.05, (b) Step 1: genes circled red show the two initial modulated sub-graphs identified, sub-graph 1 with genes M,K,C and sub-graph 2 with genes S,V,T, (c) Step 2: shows the modulated sub-graph 1 being extended to M,K,C,A,D while no extension in sub-graph 2 (d) Step 3: shows both the modulated sub-graphs merging together with the addition of 3 genes Z,W,P.}
+	\caption{Overview of \texttt{MSF} process showing 3 step toy
+          example to identify the modulated sub-graphs. (a) Input:
+          showing a network of genes with green nodes as genes with
+          \textit{p}-values $< $ 0.05, and blue nodes are genes with
+          \textit{p}-values $>$ 0.05, (b) Step 1: genes circled red
+          show the two initial modulated sub-graphs identified,
+          sub-graph 1 with genes M,K,C and sub-graph 2 with genes
+          S,V,T, (c) Step 2: shows the modulated sub-graph 1 being
+          extended to M,K,C,A,D while no extension in sub-graph 2 (d)
+          Step 3: shows both the modulated sub-graphs merging together
+          with the addition of 3 genes Z,W,P.}
 			\end{minipage}
@@
 		\begin{minipage}{17 cm}
-			\includegraphics[width=16cm]{Documents/PhDThesis/MSFPaper/DEGvsMSF2}
-			\caption{Graph shows the percentage of DEG analysis genes and MSF identified sub-graph genes recall rates 100 times for the three different time points after the addition of Poisson distributed noise.}
+			\includegraphics[width=16cm]{DEGvsMSF2.png}
+			\caption{Graph shows the percentage of DEG
+                          analysis genes and MSF identified sub-graph
+                          genes recall rates 100 times for the three
+                          different time points after the addition of
+                          Poisson distributed noise.}
 			\label{label}
@@
 		\begin{minipage}{17 cm}
-	\includegraphics[width=1.05\linewidth]{ModulatesSubPaths/MSFData/Ebola6H/SeperatedMergedNetwork}
-	\caption{Shows the three modulated sub-graphs identified by MSF at 6hpi, where nodes are genes and edges are the interactions.
-The node coloring is associated to KEGG pathways referring to the colors in the legend.
-The graph edges are from Reactome.
-Important genes to EBOV infection as from literature are enlarged in the graph.
-The signal flow corresponding to each sub-graph is depicted as simple flow chart against each sub-graph.}
+	\includegraphics[width=1.05\linewidth]{SeperatedMergedNetwork.png}
+	\caption{Shows the three modulated sub-graphs identified by
+          MSF at 6hpi, where nodes are genes and edges are the
+          interactions. The node coloring is associated to KEGG
+          pathways referring to the colors in the legend. The graph
+          edges are from Reactome. Important genes to EBOV infection
+          as from literature are enlarged in the graph. The signal
+          flow corresponding to each sub-graph is depicted as simple
+          flow chart against each sub-graph.}
 		\end{minipage}

### 2018-03-20 MSF_Paper_FA.tex
a few paragraphs more

@@
 
-\TODO{FA until here}.
-
 Existing pathway-based analysis approaches use different research
@@
 pathway categories are returned. All of this can be helpful to
-understand the cause end effect of a stimulus and might inform about
+understand the cause and effect of a stimulus and might inform about
 potential points of intervention. The proposed algorithm was
-implemented as a java program, which was named Modulated Sub-graph
-Finder (\texttt{MSF}). \texttt{MSF} would help transform the
+implemented as a java program, which was named \texttt{Modulated Sub-graph
+Finder} (abbreviated \texttt{MSF}). \texttt{MSF} can help transform the
 information obtained from DGEA into understandable knowledge of signal
-transduction of genes. \texttt{MSF} is freely accessible on github.
+transduction of genes and thereby being a valuable complement to existing
+pathway based methods. \texttt{MSF} is freely accessible on github.
+
+\TODO{FA until here}.
 

### 2018-03-20 MSF_Paper_FA.tex
MSF_Paper_FA.tex

@@
 
-\TODO{FA until here}.
-
 \section*{Methods}
-\texttt{MSF} is developed as a novel heuristic approach in Java to
-find the modulated sub-graphs of gene interactions from the cell
-signaling network. \texttt{MSF} does not use predefined gene sets in
-functional units, it builds up the sub-graphs from scratch using the
-whole cell signaling network from Reactome. The network has nodes
-corresponding to genes and edges representing their directed
-regulating interactions.
+\texttt{MSF} is developed as a novel heuristic approach to
+find concertedly modulated sub-graphs in networks of biological interactions.
+\texttt{MSF} does not use predefined gene sets grouped into
+functional units, but rather relies purely on the network of interacting genes.
+The inputed network consists of nodes corresponding to genes and edges
+representing interactions. Furthermore it utilizes comprehensive results from a
+differential gene expression analysis to discover the sub-graphs, or modules, which
+are as a whole modulated.
 
@@
 scores and correlation are applied to the inverse normal function to
-calculate the individual \textit{p}-value for a sub-graph. An overall
-\textit{p}-value of a sub-graph will express the significance of all
-genes in the sub-graph being modulated together.
+calculate the combined \textit{p}-value for all genes examined, namely the
+examined sub-graph. The combined \textit{p}-value of a sub-graph will
+express the significance of all genes in the sub-graph being modulated
+together. Thereby, the information from the different genes are used as,
+although not independent, replicated measurement of the behavior of the whole
+sub-graph. This potentially increases the power to detect also significant
+sub-graphs consisting of genes which are not significant on there own.
+\newline
 
@@
 \texttt{MSF} applies a three steps heuristic as described in the
-following.The pipeline of identification of modulated sub-graphs from
-a network by \texttt{MSF} is depicted in Fig (1).
-\newline
+following. The proceeding identification of modulated sub-graphs from
+a network by \texttt{MSF} is presented graphical and as pseudocode in
+Fig.~\ref{fig:algorithm}. \TODO{add pseudocode or better replace figure with pseudocode}
+
+\begin{figure}[ht!]
+		\centering
+	\fbox{
+		\begin{minipage}{8 cm}
+
+	\begin{subfigure}{.6\textwidth}
+		\centering
+		\includegraphics[width=6.5cm]{Tgfb1.png}
+		\caption{Input}
+	\end{subfigure}
+	\begin{subfigure}{.5\textwidth}
+		\centering
+		\includegraphics[width=6.5cm]{TGFBstep1.png}
+		\caption{Step 1}
+	\end{subfigure}
+	\begin{subfigure}{.5\textwidth}
+		\centering
+		\includegraphics[width=6.5cm]{TGFBstep2.png}
+		\caption{Step 2}
+	\end{subfigure}
+	\begin{subfigure}{.5\textwidth}
+		\centering
+		\includegraphics[width=6.5cm]{TGFBstep3.png}
+		\caption{Step 3}
+	\end{subfigure}
+	\caption{Overview of \texttt{MSF} process steps exemplified on a simplified toy
+          example. (a) Input:
+          showing a network of genes with green nodes as genes with
+          \textit{p}-values $< $ 0.05, and blue nodes are genes with
+          \textit{p}-values $>$ 0.05, (b) Step 1: genes circled red
+          show the two initial modulated sub-graphs identified,
+          sub-graph 1 with genes M,K,C and sub-graph 2 with genes
+          S,V,T, (c) Step 2: shows the modulated sub-graph 1 being
+          extended to M,K,C,A,D while no extension in sub-graph 2 (d)
+          Step 3: shows both the modulated sub-graphs merging together
+          with the addition of 3 genes Z,W,P.\label{fig:algorithm}}
+			\end{minipage}
+}
+
+\end{figure}
 
@@
 \texttt{MSF} constructs the first sub-graph starting with the genes
-associated with the lowest (most significant) \textit{p}-value from
-the network. It tries to extend the sub-graph to the most significant
-neighboring gene. A single combined \textit{p}-value is calculated for
+associated with the lowest (most significant) \textit{p}-value deduced from
+the DGEA. From this seed the sub-graph is attempted to be extended with its
+neighboring genes, starting with the most significant one. A single combined
+\textit{p}-value is calculated for
 the two genes. If the combined \textit{p}-value is smaller than the
 minimal individual gene's \textit{p}-value, the extended sub-graph is
-accepted. Then the next most significant neighboring gene is added to
-the sub-graph and the combined \textit{p}-value is calculated, if the
-new combined \textit{p}-value of three genes is smaller than the
-combined \textit{p}-value of the first two genes, the extended
-sub-graph is accepted. This step is iteratively repeated until no
+accepted. This step is iteratively repeated until no
 further extension is accepted. In this case the process starts over
@@
 sub-graph. This step identifies all the trivial sub-graphs that are
-modulated in the whole network~(Fig.~1b).\newline
+modulated in the whole network (Fig.~\ref{fig:algorithm}b).\newline
 
@@
 they could further be extended beyond the immediate neighborhood.
-This is done by testing all possible extension by N genes (default two
-genes) for all genes in the sub-graph. Again, this step is iteratively
+This is done by testing all possible extension paths up to N genes
+for all genes in the sub-graph. Again, this step is iteratively
 repeated until no further genes are added to the significant
@@
 genes without a clear differential signal in the
-DGEA~(Fig.~1c).\newline
+DGEA~(Fig.~\ref{fig:algorithm}c).\newline
 
@@
 \textit{p}-values of the two sub-graphs, the two sub-graphs are merged
-together to one big modulated sub-graph~(Fig.~1d). This step is
+together to one big modulated sub-graph~(Fig.~\ref{fig:algorithm}d). This step is
 repeated iteratively until no sub-graphs could be merged.\newline
@@
 integrated information within the signal transduction network is set
-to action. Genes in circular loops in the sub-graphs have no source or
-sink.
+to action. Due to circular loops not all sub-graphs must have sources or
+sinks.
 
@@
 To assess the robustness and stability of our method, Poisson
-distributed noise was added to the sample read counts. Then DGEA was
+distributed noise was added to the sample read counts \TODO{more details;what samples;what DGEA}. Then DGEA was
 carried out on the disturbed data with the same parameters as for the
-native data, followed by analysis with \texttt{MSF}. Using this
-criteria, the goal was to spot almost identical modulated sub-graphs
-in the real and noisy data. This procedure was carried out 100 times
+native data, followed by analysis with \texttt{MSF}. This procedure was carried out 100 times
 and every time the genes from the sub-graphs identified from noisy
@@
 native data.  The robustness of \texttt{MSF} and the DEG analysis for
-the three time-points are shown in~(Fig.~2). We were very stringent in
+the three time-points are shown in~(Fig.~2) \TODO{which time points}. We were very stringent in
 adding noise to the counts, which could be seen by the DEG analysis
-robustness. For the last time-point (48hpi) 84\% genes of the
-sub-graphs from native data were retrieved, for 24hpi robustness was
-about 79\%. Robustness was 71\% at 6hpi, this could be due to that the
+robustness. For the last time-point (48h~p.i.) 84~\% genes of the
+sub-graphs from native data were retrieved, for 24~h~p.i. robustness was
+about 79~\%. Robustness was 71~\% at 6~h~p.i., this could be due to that the
 infection merely started and the signal transdcution was weak as
@@
 \texttt{MSF} from native data. Simulation results showed a chance of
-less than 10 percent to get more sources than observed.
+less than 10 percent to get more sources than observed \TODO{we agreed on massively expanding this section or skip it alltogether}.
 
@@
 modulated sub-graphs and check if \texttt{MSF} is able to identify the
-IFN-$\alpha / \beta$ gene as one of the sources.
+IFN-$\alpha / \beta$ gene as one of the sources. \TODO{why this data set?}
 
@@
 count data with edgeR package (version 3.4.2)~\cite{edgeR}. The DEG
-list generated by edgeR was used as an input for \texttt{MSF}. The
-EBOV infection data has three time-points 6, 24 and 48 hour post
-infection (hpi). At 6hpi, five modulated sub-graphs were
-identified. Most of the genes part of the sub-graphs were chemokines
+list generated by edgeR was used as an input for \texttt{MSF} \TODO{above you said MSF does not use a DEG gene list}. The
+EBOV infection experiments describe the course of infection at three time-points 6, 24 and 48 hour post
+infection (h~p.i.). For the earliest time point at 6 h~p.i., five modulated sub-graphs were
+identified \TODO{with how many genes each?}. Most of the genes part of the sub-graphs were chemokines
 (CXCL10, CCL8) and Interleukin genes (IL6, IL27, IL23) which serve as
@@
 as two of the possible sources. Most of the sources identified by
-\texttt{MSF} were type~I interferon induced genes. At 24hpi nine
+\texttt{MSF} were type~I interferon induced genes. At 24h~p.i. nine
 modulated sub-graphs were identified, IFNA1 was identified as one of
@@
 modulated sub-graphs identified.
+
+\TODO{bis da!!}
 
@@
 
-\subsection*{Modulated sub-graphs at 6hpi}
+\subsection*{Modulated sub-graphs at 6h~p.i.}
 
- Three main modulated sub-graphs identified by \texttt{MSF} at 6hpi
+ Three main modulated sub-graphs identified by \texttt{MSF} at 6h~p.i.
  are shown in~(Fig.~3). The graph shows different colored nodes as
@@
 
-At 6hpi Cytokegg identified the sub-graph genes to be enriched in
+At 6h~p.i. Cytokegg identified the sub-graph genes to be enriched in
 Toll-like receptor signaling, TNF signaling, IL-17 signaling and
@@
 signaling~\cite{Olejnik}, Gene enrichment analysis of sub-graph genes
-at 6hpi showed toll-like receptor signaling being most significantly
+at 6h~p.i. showed toll-like receptor signaling being most significantly
 dis-regulated pathway, this important signaling pathway was not
@@
 
-\begin{figure*}[ht!]
-		\centering
-	\fbox{
-		\begin{minipage}{17 cm}
-
-	\begin{subfigure}{.5\textwidth}
-		\centering
-		\includegraphics[width=10.0cm]{Tgfb1.png}
-		\caption{Input}
-	\end{subfigure}
-	\begin{subfigure}{.5\textwidth}
-		\centering
-		\includegraphics[width=10.0cm]{TGFBstep1.png}
-		\caption{Step 1}
-	\end{subfigure}
-	\begin{subfigure}{.5\textwidth}
-		\centering
-		\includegraphics[width=10.0cm]{TGFBstep2.png}
-		\caption{Step 2}
-	\end{subfigure}
-	\begin{subfigure}{.5\textwidth}
-		\centering
-		\includegraphics[width=10.0cm]{TGFBstep3.png}
-		\caption{Step 3}
-	\end{subfigure}
-	\caption{Overview of \texttt{MSF} process showing 3 step toy
-          example to identify the modulated sub-graphs. (a) Input:
-          showing a network of genes with green nodes as genes with
-          \textit{p}-values $< $ 0.05, and blue nodes are genes with
-          \textit{p}-values $>$ 0.05, (b) Step 1: genes circled red
-          show the two initial modulated sub-graphs identified,
-          sub-graph 1 with genes M,K,C and sub-graph 2 with genes
-          S,V,T, (c) Step 2: shows the modulated sub-graph 1 being
-          extended to M,K,C,A,D while no extension in sub-graph 2 (d)
-          Step 3: shows both the modulated sub-graphs merging together
-          with the addition of 3 genes Z,W,P.}
-			\end{minipage}
-}
-\end{figure*}
-
 \begin{figure*}[p]
@@
 	\caption{Shows the three modulated sub-graphs identified by
-          MSF at 6hpi, where nodes are genes and edges are the
+          MSF at 6h~p.i., where nodes are genes and edges are the
           interactions. The node coloring is associated to KEGG

### 2018-03-22 MSF_Paper_FA.tex
Corrections

@@
 names \texttt{Modulated Sub-graph Finder} is freely available at
-\url{http//:path/to/git}.
+\url{https://github.com/MariamFarman/Modulated-SubGraph-Finder}.
 
@@
 
-\subsection*{Robustness}
-
-To assess the robustness and stability of our method, Poisson
-distributed noise was added to the sample read counts \TODO{more details;what samples;what DGEA}. Then DGEA was
-carried out on the disturbed data with the same parameters as for the
-native data, followed by analysis with \texttt{MSF}. This procedure was carried out 100 times
-and every time the genes from the sub-graphs identified from noisy
-data were compared to the genes of sub-graphs identified from the
-native data. This was also done for the DEG obtained for each run, a
-cutoff was used to compare the retrieved DEG from noisy data to the
-native data.  The robustness of \texttt{MSF} and the DEG analysis for
-the three time-points are shown in~(Fig.~2) \TODO{which time points}. We were very stringent in
-adding noise to the counts, which could be seen by the DEG analysis
-robustness. For the last time-point (48h~p.i.) 84~\% genes of the
-sub-graphs from native data were retrieved, for 24~h~p.i. robustness was
-about 79~\%. Robustness was 71~\% at 6~h~p.i., this could be due to that the
-infection merely started and the signal transdcution was weak as
-compared to the later time-points.
-
-Simulation testing was also conducted on \texttt{MSF}, the objective
-was to find less sources than observed in the sub-graphs identified by
-\texttt{MSF} from native data. Simulation results showed a chance of
-less than 10 percent to get more sources than observed \TODO{we agreed on massively expanding this section or skip it alltogether}.
-
 \subsection*{Case Study}
@@
 that IFN-$\alpha / \beta$ should be tested against Ebola for its
-antiviral activity through clinical trials~\cite{Konde}. The aim of
-testing the EBOV infection data with \texttt{MSF} was to identify the
+antiviral activity through clinical trials~\cite{Konde}.
+This data was chosen for analysis because it was the most recent EBOV infection data and the infection was carried out in human macrophages, which would help see the innate immune response of host.
+The aim of
+testing this data with \texttt{MSF} was to identify the
 modulated sub-graphs and check if \texttt{MSF} is able to identify the
-IFN-$\alpha / \beta$ gene as one of the sources. \TODO{why this data set?}
+IFN-$\alpha / \beta$ gene as one of the sources.
 
@@
 count data with edgeR package (version 3.4.2)~\cite{edgeR}. The DEG
-list generated by edgeR was used as an input for \texttt{MSF} \TODO{above you said MSF does not use a DEG gene list}. The
+analysis results generated by edgeR were used as input for \texttt{MSF}. The
 EBOV infection experiments describe the course of infection at three time-points 6, 24 and 48 hour post
-infection (h~p.i.). For the earliest time point at 6 h~p.i., five modulated sub-graphs were
-identified \TODO{with how many genes each?}. Most of the genes part of the sub-graphs were chemokines
+infection (hpi). For the earliest time point at 6 hpi, five modulated sub-graphs were
+identified with 41, 107, and 18 genes. Two of the five sub-graphs were less than 4 genes long. Most of the genes part of the sub-graphs were cytokines, chemokines
 (CXCL10, CCL8) and Interleukin genes (IL6, IL27, IL23) which serve as
@@
 as two of the possible sources. Most of the sources identified by
-\texttt{MSF} were type~I interferon induced genes. At 24h~p.i. nine
+\texttt{MSF} were type~I interferon induced genes. At 24 hpi nine
 modulated sub-graphs were identified, IFNA1 was identified as one of
@@
 
-\subsection*{Modulated sub-graphs at 6h~p.i.}
+\subsection*{Robustness}
 
- Three main modulated sub-graphs identified by \texttt{MSF} at 6h~p.i.
- are shown in~(Fig.~3). The graph shows different colored nodes as
- genes associated to KEGG pathways and how they cross talk. The KEGG
+To assess the robustness and stability of our method, Poisson
+distributed noise was added to the read counts at the 3 time-points. Then DGEA was
+carried out on the disturbed data with the same parameters as for the
+native data using edgeR, followed by analysis with \texttt{MSF}. This procedure was carried out 100 times
+and every time the genes from the sub-graphs identified from noisy
+data were compared to the genes of sub-graphs identified from the
+native data. This was also done for the DEG obtained for each run, a
+cutoff was used to compare the retrieved DEG from noisy data to the
+native data.  The robustness of \texttt{MSF} and the DEG analysis for
+the time-point 6, 24, and 48 hpi are shown in Fig.~\ref{fig:DEGvsMSF}. We were very stringent in
+adding noise to the counts, which could be seen by the DEG analysis
+robustness. For the last time-point (48h~p.i.) 84~\% genes of the
+sub-graphs from native data were retrieved, for 24~h~p.i. robustness was
+about 79~\%. Robustness was 71~\% at 6~h~p.i., this could be due to that the
+infection merely started and the signal transdcution was weak as
+compared to the later time-points.
+
+Simulation testing was also conducted on \texttt{MSF}.
+For each time-point, one \texttt{MSF} identified sub-graph was taken.
+A random graph was generated with the same number of genes as the sub-graphs and the predicted sources were counted.
+The objective
+was to find less sources in the simulated sub-graphs than observed in the real sub-graphs identified by
+\texttt{MSF}. Simulation results showed a chance of $<$10$\%$
+of the 1000 random sub-graphs generated to get more sources than observed.
+
+\subsection*{Modulated sub-graphs at 6 hpi}
+
+ Three main modulated sub-graphs identified by \texttt{MSF} at 6 hpi
+ are shown in Fig.~\ref{fig:Sub-graph6hpi}. The graph shows different colored nodes as
+ genes associated to different KEGG pathways and how they cross talk. The KEGG
  association and visualization on the modulated sub-graphs was done
@@
 CytoKegg~\cite{Cytokegg} for KEGG and Reactome FI~\cite{Reactome} for
-Reactome. We compared the pathways enriched in \texttt{MSF} identified
+Reactome. A comparison to other tools was performed on the same dataset using each software's default parameters. We compared the pathways enriched in \texttt{MSF} identified
 sub-graphs to pathway analysis by SPIA~\cite{Tarca} for KEGG and
@@
 
-At 6h~p.i. Cytokegg identified the sub-graph genes to be enriched in
+At 6 hpi Cytokegg identified the sub-graph genes to be enriched in
 Toll-like receptor signaling, TNF signaling, IL-17 signaling and
@@
 signaling~\cite{Olejnik}, Gene enrichment analysis of sub-graph genes
-at 6h~p.i. showed toll-like receptor signaling being most significantly
+at 6 hpi showed toll-like receptor signaling being most significantly
 dis-regulated pathway, this important signaling pathway was not
 identified by SPIA along with RIG-I-like receptor signaling pathway,
-TNF signaling pathway and IL-17 signaling (Table 1 in the supplement
-material). On the later time-points \texttt{MSF} and SPIA showed
+TNF signaling pathway and IL-17 signaling. On the later time-points \texttt{MSF} and SPIA showed
 consensus on the important pathways e.g Chemokine signaling,
@@
 Cytosolic DNA-sensing and RIG-I-like receptor signaling as being
-dis-regulated (Table 2-3 in the supplement).
+dis-regulated.
 
@@
 activation, Reactome as well could not identify Toll-like receptor
-signaling (Table 1-3 in the supplement material) . This shows that
+signaling. This shows that
 \texttt{MSF} is able to identify the important modulated genes in the
-sub-graphs at the earliest time-point of infection.
+sub-graphs at the earliest time-point of infection. A summary of the comparison between the different tools is given in the supplement Table 1.
 
@@
                           different time points after the addition of
-                          Poisson distributed noise.}
+                          Poisson distributed noise.\label{fig:DEGvsMSF}}
 			\label{label}
@@
 	\caption{Shows the three modulated sub-graphs identified by
-          MSF at 6h~p.i., where nodes are genes and edges are the
+          MSF at 6 hpi, where nodes are genes and edges are the
           interactions. The node coloring is associated to KEGG
@@
           flow corresponding to each sub-graph is depicted as simple
-          flow chart against each sub-graph.}
+          flow chart against each sub-graph.\label{fig:Sub-graph6hpi}}
 		\end{minipage}

### 2018-03-27 MSF_Paper_FA.tex
Corrections

@@
 
-\subsection*{Robustness}
-
-To assess the robustness and stability of our method, Poisson
-distributed noise was added to the read counts at the 3 time-points. Then DGEA was
-carried out on the disturbed data with the same parameters as for the
-native data using edgeR, followed by analysis with \texttt{MSF}. This procedure was carried out 100 times
-and every time the genes from the sub-graphs identified from noisy
-data were compared to the genes of sub-graphs identified from the
-native data. This was also done for the DEG obtained for each run, a
-cutoff was used to compare the retrieved DEG from noisy data to the
-native data.  The robustness of \texttt{MSF} and the DEG analysis for
-the time-point 6, 24, and 48 hpi are shown in Fig.~\ref{fig:DEGvsMSF}. We were very stringent in
-adding noise to the counts, which could be seen by the DEG analysis
-robustness. For the last time-point (48h~p.i.) 84~\% genes of the
-sub-graphs from native data were retrieved, for 24~h~p.i. robustness was
-about 79~\%. Robustness was 71~\% at 6~h~p.i., this could be due to that the
-infection merely started and the signal transdcution was weak as
-compared to the later time-points.
-
-Simulation testing was also conducted on \texttt{MSF}.
-For each time-point, one \texttt{MSF} identified sub-graph was taken.
-A random graph was generated with the same number of genes as the sub-graphs and the predicted sources were counted.
-The objective
-was to find less sources in the simulated sub-graphs than observed in the real sub-graphs identified by
-\texttt{MSF}. Simulation results showed a chance of $<$10$\%$
-of the 1000 random sub-graphs generated to get more sources than observed.
-
 \subsection*{Modulated sub-graphs at 6 hpi}
@@
  mediated cytotoxicity and apoptosis.
+
+ \subsection*{Robustness}
+
+ To assess the robustness and stability of our method, Poisson
+ distributed noise was added to the read counts at the 3 time-points. Then DGEA was
+ carried out on the disturbed data with the same parameters as for the
+ native data using edgeR, followed by analysis with \texttt{MSF}. This procedure was carried out 100 times
+ and every time the genes from the sub-graphs identified from noisy
+ data were compared to the genes of sub-graphs identified from the
+ native data. This was also done for the DEG obtained for each run, a
+ cutoff was used to compare the retrieved DEG from noisy data to the
+ native data.  The robustness of \texttt{MSF} and the DEG analysis for
+ the time-point 6, 24, and 48 hpi are shown in Fig.~\ref{fig:DEGvsMSF}. We were very stringent in
+ adding noise to the counts, which could be seen by the DEG analysis
+ robustness. For the last time-point (48h~p.i.) 84~\% genes of the
+ sub-graphs from native data were retrieved, for 24~h~p.i. robustness was
+ about 79~\%. Robustness was 71~\% at 6~h~p.i., this could be due to that the
+ infection merely started and the signal transdcution was weak as
+ compared to the later time-points.
+
+ To test \texttt{MSF's} correctness simulation was also conducted.
+For each time-point, one \texttt{MSF} identified sub-graph was taken.
+A random graph was generated with the same number of genes as the sub-graphs using the same interaction file and the identified sources were counted.
+The objective
+ was to find less sources in the simulated sub-graphs than observed in the real sub-graphs identified by
+ \texttt{MSF}. Simulation results showed a chance of $<$10$\%$
+ of the 1000 random sub-graphs generated to get more sources than observed.
 
@@
 
-\begin{figure*}[p]
-	\centering
-	\fbox{
-		\begin{minipage}{17 cm}
-			\includegraphics[width=16cm]{DEGvsMSF2.png}
-			\caption{Graph shows the percentage of DEG
-                          analysis genes and MSF identified sub-graph
-                          genes recall rates 100 times for the three
-                          different time points after the addition of
-                          Poisson distributed noise.\label{fig:DEGvsMSF}}
-			\label{label}
-		\end{minipage}
-	}
-\end{figure*}
-
 \begin{figure*}
@@
 
+\begin{figure*}[p]
+	\centering
+	\fbox{
+		\begin{minipage}{17 cm}
+			\includegraphics[width=16cm]{DEGvsMSF2.png}
+			\caption{Graph shows the percentage of DEG
+                          analysis genes and MSF identified sub-graph
+                          genes recall rates 100 times for the three
+                          different time points after the addition of
+                          Poisson distributed noise.\label{fig:DEGvsMSF}}
+			\label{label}
+		\end{minipage}
+	}
+\end{figure*}
+
 \end{document}

### 2018-03-28 MSF_Paper_FA.tex
Update and Pseudocode

@@
 from the perturbation source to the effectors. The described software,
-names \texttt{Modulated Sub-graph Finder} is freely available at
+named \texttt{Modulated Sub-graph Finder} is freely available at
 \url{https://github.com/MariamFarman/Modulated-SubGraph-Finder}.
@@
 To reduce the complexity to score all possible connected sub-graphs
-\texttt{MSF} applies a three steps heuristic as described in the
+\texttt{MSF} applies a four steps heuristic as described in the
 following. The proceeding identification of modulated sub-graphs from
-a network by \texttt{MSF} is presented graphical and as pseudocode in
-Fig.~\ref{fig:algorithm}. \TODO{add pseudocode or better replace figure with pseudocode}
-
-\begin{figure}[ht!]
-		\centering
-	\fbox{
-		\begin{minipage}{8 cm}
-
-	\begin{subfigure}{.6\textwidth}
-		\centering
-		\includegraphics[width=6.5cm]{Tgfb1.png}
-		\caption{Input}
-	\end{subfigure}
-	\begin{subfigure}{.5\textwidth}
-		\centering
-		\includegraphics[width=6.5cm]{TGFBstep1.png}
-		\caption{Step 1}
-	\end{subfigure}
-	\begin{subfigure}{.5\textwidth}
-		\centering
-		\includegraphics[width=6.5cm]{TGFBstep2.png}
-		\caption{Step 2}
-	\end{subfigure}
-	\begin{subfigure}{.5\textwidth}
-		\centering
-		\includegraphics[width=6.5cm]{TGFBstep3.png}
-		\caption{Step 3}
-	\end{subfigure}
-	\caption{Overview of \texttt{MSF} process steps exemplified on a simplified toy
-          example. (a) Input:
-          showing a network of genes with green nodes as genes with
-          \textit{p}-values $< $ 0.05, and blue nodes are genes with
-          \textit{p}-values $>$ 0.05, (b) Step 1: genes circled red
-          show the two initial modulated sub-graphs identified,
-          sub-graph 1 with genes M,K,C and sub-graph 2 with genes
-          S,V,T, (c) Step 2: shows the modulated sub-graph 1 being
-          extended to M,K,C,A,D while no extension in sub-graph 2 (d)
-          Step 3: shows both the modulated sub-graphs merging together
-          with the addition of 3 genes Z,W,P.\label{fig:algorithm}}
-			\end{minipage}
-}
-
-\end{figure}
+a network by \texttt{MSF} is presented as pseudocode in flowchart in
+Fig.~\ref{fig:pseudocodemsf}.
 
@@
 antiviral activity through clinical trials~\cite{Konde}.
-This data was chosen for analysis because it was the most recent EBOV infection data and the infection was carried out in human macrophages, which would help see the innate immune response of host.
+This data was chosen for analysis because it was the most recent EBOV infection data and the infection was carried out in human macrophages, which would help us comprehend the innate immune response of host.
 The aim of
@@
 \texttt{MSF} were type~I interferon induced genes. At 24 hpi nine
-modulated sub-graphs were identified, IFNA1 was identified as one of
-the sources. For the last time-point again IFNB1 and IFNA1 were
-identified as the two sources out of the possible sources from seven
-modulated sub-graphs identified.
+modulated sub-graphs were identified with four main sub-graphs consisting of 38, 85, 148 and 167 genes. IFNA1 was identified as one of
+the sources in the most significantly modulated sub-graph.
+For the last time-point 48 hpi, eleven modulated sub-graphs were identified.
+Eight of the sub-graphs were less than seven genes and main sub-graphs had 39, 71 and 210 genes.
+IFNB1 and IFNA1 were
+identified as the two sources out of the possible sources.
 
@@
 a source in the early time-point when the genes were not significant
-based on the individual DGEA alone. Identifying these sources will
+based on the individual DGEA alone. Identifying the possible sources will
 reduce the search space for potential target genes and can help the
@@
  Three main modulated sub-graphs identified by \texttt{MSF} at 6 hpi
- are shown in Fig.~\ref{fig:Sub-graph6hpi}. The graph shows different colored nodes as
+ are shown in Fig.~\ref{fig:Sub-graph6hpi}. the figure shows two forms of graphs, one graph has different colored nodes as
  genes associated to different KEGG pathways and how they cross talk. The KEGG
  association and visualization on the modulated sub-graphs was done
- using StringApp~\cite{StringApp} in Cytoscape~\cite{Cyto}. The graph
+ using StringApp~\cite{StringApp} in Cytoscape~\cite{Cyto}.
+StringApp can retrieve functional enrichment for Gene Ontology terms, KEGG Pathways, and protein domains for a user-specified network.
+The other graph like a flow chart,
  unveils how more than one stimulus coincides to prevent the
@@
  To test \texttt{MSF's} correctness simulation was also conducted.
-For each time-point, one \texttt{MSF} identified sub-graph was taken.
-A random graph was generated with the same number of genes as the sub-graphs using the same interaction file and the identified sources were counted.
+For each time-point, one \texttt{MSF} identified sub-graph was selected and  using simulation random graph was generated with the same number of genes as the sub-graphs utilizing the same interaction file.
+The simulation identified sources were counted.
 The objective
@@
 \texttt{MSF} is able to identify the important modulated genes in the
-sub-graphs at the earliest time-point of infection. A summary of the comparison between the different tools is given in the supplement Table 1.
+sub-graphs at the earliest time-point of infection.
 
@@
 		\begin{minipage}{17 cm}
+		\includegraphics[width=1.0\linewidth]{PseudoCodeMSF}
+		\caption{Shows the algorithm of MSF with four steps to find the modulated sub-graphs and identify sources and sinks.}
+		\label{fig:pseudocodemsf}
+	\end{minipage}
+}
+
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\fbox{
+		\begin{minipage}{17 cm}
 			\includegraphics[width=1.05\linewidth]{SeperatedMergedNetwork.png}

### 2018-04-03 MSF_Paper_FA.tex
Algo and updated tex

@@
 following. The proceeding identification of modulated sub-graphs from
-a network by \texttt{MSF} is presented as pseudocode in flowchart in
+a network by \texttt{MSF} is presented as a graphical pseudo code  in
 Fig.~\ref{fig:pseudocodemsf}.
@@
 sub-graph. This step identifies all the trivial sub-graphs that are
-modulated in the whole network (Fig.~\ref{fig:algorithm}b).\newline
+modulated in the whole network.\newline
 
@@
 In the next step initial modulated sub-graphs are used to check if
-they could further be extended beyond the immediate neighborhood.
-This is done by testing all possible extension paths up to N genes
+they could further be extended beyond the immediate neighborhood. This is done by testing all possible extension paths up to N genes
 for all genes in the sub-graph. Again, this step is iteratively
@@
 genes without a clear differential signal in the
-DGEA~(Fig.~\ref{fig:algorithm}c).\newline
+DGEA.\newline
 
@@
 \textit{p}-values of the two sub-graphs, the two sub-graphs are merged
-together to one big modulated sub-graph~(Fig.~\ref{fig:algorithm}d). This step is
+together to one big modulated sub-graph. This step is
 repeated iteratively until no sub-graphs could be merged.\newline
@@
 response of the host, which includes the activation of alpha/beta
-interferon (IFN-$\alpha / \beta$)~\cite{Prins,Konde,Cardenas}. The
-Ebola viral protein VP35 targets the host type~I interferon
-IFN-$\alpha / \beta$ to block the early antiviral immune response
-~\cite{Prins,Konde,Falasca,Cardenas,Olejnik}. It has been proposed
+interferon (IFN-$\alpha / \beta$)~\cite{Prins,Konde,Cardenas}. It has been proposed
 that IFN-$\alpha / \beta$ should be tested against Ebola for its
 antiviral activity through clinical trials~\cite{Konde}.
-This data was chosen for analysis because it was the most recent EBOV infection data and the infection was carried out in human macrophages, which would help us comprehend the innate immune response of host.
+Since the data was most recent and the  infection was carried out in human macrophages, it helped us comprehend the innate immune response of host.
 The aim of
@@
  Three main modulated sub-graphs identified by \texttt{MSF} at 6 hpi
- are shown in Fig.~\ref{fig:Sub-graph6hpi}. the figure shows two forms of graphs, one graph has different colored nodes as
- genes associated to different KEGG pathways and how they cross talk. The KEGG
+ are shown in Fig.~\ref{fig:Sub-graph6hpi}. It shows two forms of graphs, one graph has different colored nodes as
+ genes associated to different KEGG pathways. The KEGG
  association and visualization on the modulated sub-graphs was done
@@
 StringApp can retrieve functional enrichment for Gene Ontology terms, KEGG Pathways, and protein domains for a user-specified network.
-The other graph like a flow chart,
+The other graph is a flow chart, that
  unveils how more than one stimulus coincides to prevent the
@@
  cytokine, chemokine and jak-stat genes leading to TNF and eventually
- going into apoptosis. Sub-graph 2 shows Extra-cellular matrix (ECM)
- receptor interaction genes which are suggested to be in interaction
+ going into apoptosis. From one end  of sub-graph 2, Extra-cellular matrix (ECM)
+ receptor interaction genes are modulated, which are suggested to be in interaction
  with Ebola glycoprotein (GP)~\cite{Veljkovic}. ECM receptor, Focal
- adhesion and P13k-Akt signaling genes are the sources from one end of
- the sub-graph. P13k-Akt signaling goes to FoxO signaling which
- eventually leads to cell cycle, which at later time-points is
- disrupted. From the other end of sub-graph 2 toll-like recpetor
+ adhesion and P13k-Akt signaling genes are the sources from this end of
+ the sub-graph, with P13k-Akt signaling going to FoxO signaling which
+ eventually leads to cell cycle, that at later time-points is
+ seen to be disrupted. From the other end of sub-graph 2 toll-like recpetor
  combined with cytosolic DNA sensing and cytokine-chemokine signaling
  activates Jak-stat which leads to the start of apoptosis with a
- activation of a few apoptosis genes. Cytokine is also seen to sink
+ few apoptosis genes being modulated. Cytokine is also seen to sink
  into natural killer mediated cytotoxicity. Sub-graph 3 shows how some
@@
  To assess the robustness and stability of our method, Poisson
- distributed noise was added to the read counts at the 3 time-points. Then DGEA was
+ distributed noise was added to the read counts of the three time-points. Then DGEA was
  carried out on the disturbed data with the same parameters as for the
@@
  robustness. For the last time-point (48h~p.i.) 84~\% genes of the
- sub-graphs from native data were retrieved, for 24~h~p.i. robustness was
- about 79~\%. Robustness was 71~\% at 6~h~p.i., this could be due to that the
+ sub-graphs from native data were retrieved, for 24~h~p.i. recall rate was
+ 79~\%. Robustness was 71~\% at 6~h~p.i., this could be due to that the
  infection merely started and the signal transdcution was weak as
@@
 For each time-point, one \texttt{MSF} identified sub-graph was selected and using simulation random graph was generated with the same number of genes as the sub-graphs utilizing the same interaction file.
-The simulation identified sources were counted.
+The simulation identified sources were counted from the simulated sub-graph.
 The objective
@@
 
-To further understand the functional role of the sub-graphs identified
-by \texttt{MSF}, we linked the sub-graphs to the predefined KEGG
-pathways and Reactome pathways through gene enrichment analysis. We
+ A comparison to other tools was performed on the same dataset using each software's default parameters. We compared the gene enrichment analysis in \texttt{MSF} identified
+sub-graphs to pathway analysis by SPIA~\cite{Tarca} for KEGG and
+Reactome pathway enrichment analysis for Reactome.  We
 used Cytoscape~\cite{Cyto} plugins to identify the enriched patwhays,
 CytoKegg~\cite{Cytokegg} for KEGG and Reactome FI~\cite{Reactome} for
-Reactome. A comparison to other tools was performed on the same dataset using each software's default parameters. We compared the pathways enriched in \texttt{MSF} identified
-sub-graphs to pathway analysis by SPIA~\cite{Tarca} for KEGG and
-Reactome pathway enrichment analysis for Reactome.
+Reactome for \texttt{MSF} identified
+sub-graphs.
 
@@
 signaling. This shows that
-\texttt{MSF} is able to identify the important modulated genes in the
-sub-graphs at the earliest time-point of infection.
+\texttt{MSF} is able to identify the important modulated genes of the early infection stage.
 
@@
 
-The discussion should include the implications of the article results
-in view of prior work in this field.
+MSF is able to identify the connected modulated genes, showing how the signal transmission flows between the different predefined pathways.
+MSF is able to identify the important genes at the early stage of infection which at the later stage of infection became the most modulated genes.
+Even with the addition of noise MSF identifies the important modulated core set of genes.
+The sources identified by MSF can assist in the discovery of new drug targets.
+One limitation for MSF would be an incomplete interaction network.
+For the unknown or less known genes, the interactions are still not well known and so are usually missing in the biological network in different databases.
 
@@
 		\begin{minipage}{17 cm}
-		\includegraphics[width=1.0\linewidth]{PseudoCodeMSF}
+		\includegraphics[width=1.0\linewidth]{AlgorithmMSF.png}
 		\caption{Shows the algorithm of MSF with four steps to find the modulated sub-graphs and identify sources and sinks.}

### 2018-04-10 MSF_Paper_FA.tex
Supplement removed
revised the text untill Comparison

@@
 High throughput techniques such as RNA-seq or microarray analysis have
-proven to be invaluable for the characterization of global
-transcriptional gene activity changes due to external stimuli or
-diseases. Differential gene expression analysis (DGEA) is the first
-step in the course of data interpretation, typically producing lists
-of of dozens to thousands of differentially expressed genes. To
-further guide the interpretation of these lists, different pathway
-analysis approaches have been developed. Most (\TODO{all?}) of which
-classify genes into set of genes, i.e. a pathway, which do interact
-with each other function in a common biological process. Regardless of
-technical differences, all these methods do not properly account for
-cross talk between different pathways and rely on binary
-classification into differentially expressed gene and unaffected genes
-based on an arbitrarily set p-value cut-off.
+proven to be invaluable for the characterization of global transcriptional
+gene activity changes due to external stimuli or diseases. Differential
+gene expression analysis (DGEA) is the first step in the course of data
+interpretation, typically producing lists of of dozens to thousands of
+differentially expressed genes. To further guide the interpretation of
+these lists, different pathway analysis approaches have been
+developed. These tools typically classify genes into set of genes, i.e. a
+pathway, which interact with each other and function in a common biological
+process. Regardless of technical differences, all these methods do not
+properly account for cross talk between different pathways and rely on
+binary classification into differentially expressed gene and unaffected
+genes based on an arbitrarily set p-value cut-off.
 
@@
 from the perturbation source to the effectors. The described software,
-named \texttt{Modulated Sub-graph Finder} is freely available at
+named \texttt{Modulated Sub-graph Finder} (\texttt{MSF}) is freely available at
 \url{https://github.com/MariamFarman/Modulated-SubGraph-Finder}.
@@
 
-Differentially expressed genes, Pathway analysis, Combining \textit{p}-value, Cell signaling network, Modulated Sub-graphs
+Differential gene expression analysis; pathway analysis; combining \textit{p}-value; cell signaling network;
 
@@
 Encyclopedia of Genes and Genomes)\cite{Kegg} and
-Reactome~\cite{Reactome}. KEGG pathways is a branch of KEGG database
-that has a collection of manually drawn pathway maps representing the
+Reactome~\cite{Reactome}. KEGG pathways is a branch of KEGG database that
+hosts a collection of manually drawn pathway maps representing the
 molecular interaction, reaction and relation networks of cellular
 functions. Similarly, Reactome is an open-source, manually curated,
-peer-reviewed database for signaling and metabolic molecules with
-their interactions formed into different biological pathways for
-nineteen species~\cite{Reactome}. Both provide predefined pathways
-which are sets of genes and their interactions categorized into
-functional units. Starting from a gene interaction network, genes are
-labeled according their role in a specific biological process. In this
-sense a particular gene can be assigned to different pathways. E.g.,
-the human gene STAT1 is associated with 24 different pathways in the
-pathway annotation curated by
-KEGG~\footnote{\url{http://www.genome.jp/dbget-bin/www_bget?hsa:6772}}
-and in 12 different pathways in the Reactome data
+peer-reviewed database for signaling and metabolic molecules with their
+interactions formed into different biological pathways for nineteen
+species~\cite{Reactome}. Both provide predefined pathways which are sets of
+genes and their interactions categorized into functional units. Starting
+from a gene interaction network, genes are labeled according their role in
+a specific biological process. In this sense a particular gene can be
+assigned to different pathways. E.g., the human gene STAT1 is associated
+with 24 different pathways in the pathway annotation curated by
+KEGG~\footnote{\url{http://www.genome.jp/dbget-bin/www_bget?hsa:6772}} and
+in 12 different pathways in the Reactome data
 base~\footnote{\url{http://www.reactome.org/content/detail/R-HSA-629622}}.
-Although carefully produced, the assignment of genes to those
-predefined pathway unites can be considered to be subjective to some
-degree and suffers from observatinal bias (\TODO{citation needed}).
+Although carefully produced, the assignment of genes to those predefined
+pathway unites can be considered to be subjective to some degree and
+suffers from observational bias~\cite{schnoes2013biases}.
 
-Existing pathway-based analysis approaches use different research
-designs, which can be categorized into ORA (Over-representation
-analysis), FCS (Functional class scoring) and pathway topology based
-methods. All of which aim in finding in a subset of genes, e.g.,
-significantly differentially expressed genes, genes associated with a
-certain pathway more often than expected given the total set of
-examined genes, e.g. the genome background.  ORA is the first and the
-most basic method of pathway analysis. It uses a DEG list with user
-defined cut-off for the log-fold change and \textit{p}-value (most
-commonly using absolute log-fold change $\geq$ 2 and \textit{p}-value
-$\leq$ 0.05). Subsequently, sets of genes associated with annotated
-pathways are tested for being over-represented in the set of DEGs. To
-this end, hyper-geometric distribution, chi-square tests, binomial
-probability or the Fisher’s exact test are used. Thereby the
-information of the topology of genes in the pathways are
-ignored~\cite{Bayer}. Furthermore, ORA assumes that the biological
-pathways are independent of each other and ignores the fact that they
-cross-talk and overlap~\cite{Khatri2012,Campos}.
+Existing pathway-based analysis approaches use different research designs,
+which can be categorized into ORA (Over-representation analysis), FCS
+(Functional class scoring) and pathway topology based methods. All of which
+aim in finding in a subset of genes, e.g., significantly differentially
+expressed genes, genes associated with a certain pathway more often than
+expected given the total set of examined genes, e.g.~the whole genome
+background.  ORA is the first and the most basic method of pathway
+analysis. It uses a DEG list with user defined cut-off for the log-fold
+change and \textit{p}-value (most commonly using absolute log-fold change
+$\geq$ 2 and \textit{p}-value $\leq$ 0.05). Subsequently, sets of genes
+associated with annotated pathways are tested for being over-represented in
+the set of DEGs. To this end, hyper-geometric distribution, chi-square
+tests, binomial probability or the Fisher’s exact test are used. Thereby
+the information of the topology of genes in the pathways are
+ignored~\cite{Bayer}. Furthermore, ORA assumes that the biological pathways
+are independent of each other and ignores the fact that they cross-talk and
+overlap~\cite{Khatri2012,Campos}.
 
-Unlike ORA, FCS has no artificial cut-off to define DEG list. FCS
-works in three step, first it calculates the gene-level statistics
-including correlation of molecular measurements using differential
-expression of individual genes, ANOVA, t-test and Z-score. In the
-second step the statistics of individual genes in a pathway are
-transformed to an individual pathway-level statistic commonly using
-Kolmogorov-Smirnov statistic, mean or median. Finally the statistical
-significance of the pathway-level statistics is assessed. Although FCS
-covers some of the limitations from ORA, it still lacks the topology
-of genes in a pathway, cross-talk and overlap of the
-pathways~\cite{Khatri2012,Campos}. Pathway topology based methods are
-similar to FCS except that they consider the topology of each gene
-during the gene-level statistics but still are unable to link
-different pathways~\cite{Khatri2012}.
+Unlike ORA, FCS has no artificial cut-off to define DEG list. FCS works in
+three step, first it calculates the gene-level statistics including
+correlation of molecular measurements using differential expression of
+individual genes, ANOVA, t-test and Z-score. In the second step the
+statistics of individual genes in a pathway are transformed to an
+individual pathway-level statistic commonly using Kolmogorov-Smirnov
+statistic, mean or median. Finally the statistical significance of the
+pathway-level statistics is assessed. Although FCS covers some of the
+limitations from ORA, it still lacks the topology of genes in a pathway,
+cross-talk and overlap of the pathways~\cite{Khatri2012,Campos}. Pathway
+topology based methods are similar to FCS except that they consider the
+topology of each gene during the gene-level statistics but still don't aim
+to link different functional pathways~\cite{Khatri2012}.
 
-On these grounds we propose a novel approach to make use of the rich
-pathway annotation resources available to gain additional functional
-insights from basic DGEA. To this, we start with the presupposition
-that expression of neighboring genes within a functional pathway are
-not independent from each other. Rather, they are often regulating
-each others expression or are part of the same
-regulon~\cite{Michalak}. We understand that the categorization of
-links between genes into labeled pathways is often an arbitrary one,
-given the extensive cross talk between different pathways. Although
-these categories have proven to be useful in many situations, they
-force a certain perspective onto the interpretation of novel
-data. Based on these two principles, we aim to find sub-graphs of
-connected genes within cell signaling network which exhibit as a whole
-significant differential expression changes. As input, information on
-functional links between genes provided by e.g. KEGG or Reactome and
-information on the differential expression status of single genes
-resulting from a DGEA, are required. As a result the analysis returns
-sub-graphs and their joint confidence scores, reflecting how the
-perturbation is migrated through the network. Furthermore, the entry
-points of perturbation in the networks and overlap with conventional
-pathway categories are returned. All of this can be helpful to
-understand the cause and effect of a stimulus and might inform about
-potential points of intervention. The proposed algorithm was
-implemented as a java program, which was named \texttt{Modulated Sub-graph
-Finder} (abbreviated \texttt{MSF}). \texttt{MSF} can help transform the
-information obtained from DGEA into understandable knowledge of signal
-transduction of genes and thereby being a valuable complement to existing
-pathway based methods. \texttt{MSF} is freely accessible on github.
+On these grounds we propose a novel approach to make use of the rich gene
+and protein interaction annotation resources available to gain additional
+functional insights from basic DGEA. To this, we start with the
+presupposition that expression of neighboring genes within a functional
+pathway are not independent from each other. Rather, they are often
+regulating each others expression or are part of the same
+regulon~\cite{Michalak}. We understand that the categorization of links
+between genes into labeled pathways is often an arbitrary one, given the
+extensive cross talk between different pathways. Although these categories
+have proven to be useful in many situations, they force a certain
+perspective onto the interpretation of novel data. Based on these two
+principles, we aim to find sub-graphs of connected genes within cell
+signaling network which exhibit as a whole significant differential
+expression changes. This approach differes in two main aspects from common
+pathway analysis tools. First, it does not aim to identify functional
+pathways enriched in differentially expressed genes, but detects sub-graphs
+or branches in a network graph (potentially spanning more than one
+functionally grouped pathway) which is coherently modulated. Second, it
+aims to improve the DGEA on the gene level, by collecting the information
+of neighboring genes, which as a whole might exhibit prominent enough
+signal to be called, again as a whole, significantly modulated.
+
+As input, information on functional links between genes provided by
+e.g.~KEGG or Reactome and information on the differential expression status
+of single genes resulting from a DGEA, are required. As a result the
+analysis returns sub-graphs and their joint confidence scores, reflecting
+how the perturbation is migrated through the network. Furthermore, the
+entry points of perturbation in the networks and overlap with conventional
+pathway categories are returned. The output is prepared for convenient
+visualisation with the \TODO{StringApp, available as a Cytoscape plug-in.}
+
+All of this can be helpful to understand the cause and effect of a stimulus
+and might inform about potential points of intervention. The proposed
+algorithm was implemented as a java program, which was named
+\texttt{Modulated Sub-graph Finder} (abbreviated
+\texttt{MSF}). \texttt{MSF} can help transform the information obtained
+from DGEA into comprehensible knowledge of signal transduction of genes and
+thereby being a valuable complement to existing pathway based
+methods. \texttt{MSF} is freely accessible on github \TODO{which licence?}.
 
@@
 \texttt{MSF} applies a four steps heuristic as described in the
-following. The proceeding identification of modulated sub-graphs from
-a network by \texttt{MSF} is presented as a graphical pseudo code  in
-Fig.~\ref{fig:pseudocodemsf}.
+following. The proceeding identification of modulated sub-graphs from a
+network by \texttt{MSF} is presented as a flowchart diagram
+(Fig.~\ref{fig:pseudocodemsf}).
 
@@
 the DGEA. From this seed the sub-graph is attempted to be extended with its
-neighboring genes, starting with the most significant one. A single combined
-\textit{p}-value is calculated for
-the two genes. If the combined \textit{p}-value is smaller than the
-minimal individual gene's \textit{p}-value, the extended sub-graph is
-accepted. This step is iteratively repeated until no
-further extension is accepted. In this case the process starts over
-with all remaining genes not yet in a significantly modulated
-sub-graph. This step identifies all the trivial sub-graphs that are
-modulated in the whole network.\newline
+neighboring genes, starting with the most significant one. A single
+combined \textit{p}-value is calculated for the two genes. If the combined
+\textit{p}-value is smaller than the minimal individual gene's
+\textit{p}-value, the extended sub-graph is accepted. This step is
+iteratively repeated until no further extension is accepted. In this case
+the process starts over with all remaining genes not yet in a significantly
+modulated sub-graph. This step identifies all the trivial sub-graphs that
+are modulated in the whole network.\newline
 
@@
 
-In the next step initial modulated sub-graphs are used to check if
-they could further be extended beyond the immediate neighborhood. This is done by testing all possible extension paths up to N genes
-for all genes in the sub-graph. Again, this step is iteratively
-repeated until no further genes are added to the significant
-differentially expressed sub-graphs. This steps bridges small gaps of
-genes without a clear differential signal in the
-DGEA.\newline
+In the next step initial modulated sub-graphs are used to check if they
+could further be extended beyond the immediate neighborhood. This is done
+by testing all possible extension paths up to N genes for all genes in the
+sub-graph. Again, this step is iteratively repeated until no further genes
+are added to the significant differentially expressed sub-graphs. This
+steps bridges small gaps of genes without a clear differential signal in
+the DGEA.\newline
 
@@
 
-After detection and extension of the modulated sub-graphs, they are
-tested if combined sub-graph score is better than on their own. The
-merging of the two sub-graphs is done by depth first search. If the
-two sub-graphs merge with the connector of at most N genes (default
-two genes) and the combined \textit{p}-value of the merged sub-graph
-including the bridging genes in between is less than the individual
-\textit{p}-values of the two sub-graphs, the two sub-graphs are merged
-together to one big modulated sub-graph. This step is
-repeated iteratively until no sub-graphs could be merged.\newline
+After detection and extension of the modulated sub-graphs, they are tested
+if combined sub-graphs score better than on their own. The merging of the
+two sub-graphs is done by depth first search. If the two sub-graphs merge
+with the connector of at most N genes (default two genes) and the combined
+\textit{p}-value of the merged sub-graph including the bridging genes in
+between is less than the individual \textit{p}-values of the two
+sub-graphs, the two sub-graphs are merged together to one big modulated
+sub-graph. This step is repeated iteratively until no sub-graphs could be
+merged.\newline
 
@@
 
-In a last post processing step \texttt{MSF} identifies the trigger
-points of the modulated sub-graphs. These trigger genes are the
-sources of the sub-graphs with only outgoing edges. These genes can be
-interpreted as the possible entry points of perturbation from where
-the stimulus causes downstream effects. In the same spirit the most
-downstream genes of the modulated sub-graph are identified and defined
-as sinks. Sinks can be interpreted as the effectors where the
-integrated information within the signal transduction network is set
-to action. Due to circular loops not all sub-graphs must have sources or
-sinks.
+In a last post processing step \texttt{MSF} identifies the trigger points
+of the modulated sub-graphs. These trigger genes are the sources of the
+sub-graphs with only outgoing edges. These genes can be interpreted as the
+possible entry points of perturbation from where the stimulus causes
+downstream effects. In the same spirit the most downstream genes of the
+modulated sub-graph are identified and defined as sinks. Sinks can be
+interpreted as the effectors where the integrated information within the
+signal transduction network is set to action. Due to circular loops not all
+sub-graphs are guaranteed to have sources or sinks.
 
@@
 
-To demonstrate its usefulness \texttt{MSF} is applied to RNAseq data
+To demonstrate its usefulness, \texttt{MSF} is applied to an RNAseq data
 set of primary human monocyte-derived macrophages (MDMs) infected with
-Ebola virus (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV) belongs to
-the Filoviridea family; filamentous, enveloped and single stranded RNA
-viruses. EBOV causes hemorrhagic fever in humans, inducing the host
-innate and adaptive immune response to be unable to control virus
-infection~\cite{Prins}. Until now there are no approved antiviral
-drugs for the treatment of Ebola virus infection~\cite{Konde,Rhein}.
-The initial targets of EBOV are the macrophages and dendritic immune
+Ebola virus (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV) belongs to the
+Filoviridea family; filamentous, enveloped and single stranded RNA
+viruses. EBOV causes hemorrhagic fever in humans, inducing the host innate
+and adaptive immune response to be unable to control virus
+infection~\cite{Prins}. Until now, there are no approved antiviral drugs
+for the treatment of Ebola virus infection~\cite{Konde,Rhein}.  The initial
+targets of EBOV are the macrophages and dendritic immune
 cells~\cite{Falasca,Rhein}. EBOV inhibits the critical innate immune
 response of the host, which includes the activation of alpha/beta
-interferon (IFN-$\alpha / \beta$)~\cite{Prins,Konde,Cardenas}. It has been proposed
-that IFN-$\alpha / \beta$ should be tested against Ebola for its
-antiviral activity through clinical trials~\cite{Konde}.
-Since the data was most recent and the  infection was carried out in human macrophages, it helped us comprehend the innate immune response of host.
-The aim of
-testing this data with \texttt{MSF} was to identify the
-modulated sub-graphs and check if \texttt{MSF} is able to identify the
-IFN-$\alpha / \beta$ gene as one of the sources.
+interferon (IFN-$\alpha / \beta$)~\cite{Prins,Konde,Cardenas}. It has been
+proposed that IFN-$\alpha / \beta$ should be tested against Ebola for its
+antiviral activity through clinical trials~\cite{Konde}. \TODO{I do not get
+this sentence: Since the data was most recent and the infection was
+carried out in human macrophages, it helped us comprehend the innate
+immune response of host}. The aim of testing this data with \texttt{MSF} was
+to identify the modulated sub-graphs and check if \texttt{MSF} is able to
+identify the IFN-$\alpha / \beta$ gene as one of the sources, and show that
+the capability to close ``gaps'' by using information of neighboring genes
+is effective.
 
-EBOV infection count data was downloaded from GEO
-(GSE84188). Differential gene expression analysis was performed on the
-count data with edgeR package (version 3.4.2)~\cite{edgeR}. The DEG
-analysis results generated by edgeR were used as input for \texttt{MSF}. The
-EBOV infection experiments describe the course of infection at three time-points 6, 24 and 48 hour post
-infection (hpi). For the earliest time point at 6 hpi, five modulated sub-graphs were
-identified with 41, 107, and 18 genes. Two of the five sub-graphs were less than 4 genes long. Most of the genes part of the sub-graphs were cytokines, chemokines
-(CXCL10, CCL8) and Interleukin genes (IL6, IL27, IL23) which serve as
-a subset of CD4+ T helper cells. IFNB1 and IFNA1 were both identified
-as two of the possible sources. Most of the sources identified by
-\texttt{MSF} were type~I interferon induced genes. At 24 hpi nine
-modulated sub-graphs were identified with four main sub-graphs consisting of 38, 85, 148 and 167 genes. IFNA1 was identified as one of
-the sources in the most significantly modulated sub-graph.
-For the last time-point 48 hpi, eleven modulated sub-graphs were identified.
-Eight of the sub-graphs were less than seven genes and main sub-graphs had 39, 71 and 210 genes.
-IFNB1 and IFNA1 were
-identified as the two sources out of the possible sources.
+\TODO{TODO: If MSF is not applied but only significant genes are mapped on
+  the reactom data set, into how many sub-graphs would the data fall apart?
+  in other words: how many genes are only significant because they are part
+  of a significant sub-graph but are not on there own?}
 
-\TODO{bis da!!}
+EBOV infection count data was downloaded from GEO (GSE84188). Differential
+gene expression analysis was performed on the count data with edgeR package
+(version 3.4.2)~\cite{edgeR}. The DEG analysis results generated by edgeR
+were used as input for \texttt{MSF}. The EBOV infection experiments
+describe the course of infection at three time-points 6, 24 and 48 hour
+post infection (hpi). For the earliest time point at 6~hpi, five modulated
+sub-graphs were identified with 41, 107, and 18 genes. Two of the five
+sub-graphs were less than 4 genes long. Most of the genes part of the
+sub-graphs were cytokines, chemokines (CXCL10, CCL8) and Interleukin genes
+(IL6, IL27, IL23) which serve as a subset of CD4+ T helper cells. IFNB1 and
+IFNA1 were both identified as two of the possible sources. Most of the
+sources identified by \texttt{MSF} were type~I interferon induced genes. At
+24~hpi nine modulated sub-graphs were identified with four main sub-graphs
+consisting of 38, 85, 148 and 167 genes. IFNA1 was identified as one of the
+sources in the most significantly modulated sub-graph. For the last
+time-point 48~hpi, eleven modulated sub-graphs were identified. Eight of
+the sub-graphs were less than seven genes and main sub-graphs had 39, 71
+and 210 genes. IFNB1 and IFNA1 were identified as the two sources out of
+the possible sources.
 
@@
 
-\subsection*{Modulated sub-graphs at 6 hpi}
+\subsection*{Modulated sub-graphs at 6~hpi}
 
- Three main modulated sub-graphs identified by \texttt{MSF} at 6 hpi
- are shown in Fig.~\ref{fig:Sub-graph6hpi}. It shows two forms of graphs, one graph has different colored nodes as
- genes associated to different KEGG pathways. The KEGG
- association and visualization on the modulated sub-graphs was done
- using StringApp~\cite{StringApp} in Cytoscape~\cite{Cyto}.
-StringApp can retrieve functional enrichment for Gene Ontology terms, KEGG Pathways, and protein domains for a user-specified network.
-The other graph is a flow chart, that
- unveils how more than one stimulus coincides to prevent the
- progression of infection. Sub-graph 1 shows toll-like receptor,
- cytokine, chemokine and jak-stat genes leading to TNF and eventually
- going into apoptosis. From one end  of sub-graph 2, Extra-cellular matrix (ECM)
- receptor interaction genes are modulated, which are suggested to be in interaction
- with Ebola glycoprotein (GP)~\cite{Veljkovic}. ECM receptor, Focal
- adhesion and P13k-Akt signaling genes are the sources from this end of
- the sub-graph, with P13k-Akt signaling going to FoxO signaling which
- eventually leads to cell cycle, that at later time-points is
- seen to be disrupted. From the other end of sub-graph 2 toll-like recpetor
- combined with cytosolic DNA sensing and cytokine-chemokine signaling
- activates Jak-stat which leads to the start of apoptosis with a
- few apoptosis genes being modulated. Cytokine is also seen to sink
- into natural killer mediated cytotoxicity. Sub-graph 3 shows how some
- of the cytokine and chemokine genes activate the natural killer
- mediated cytotoxicity and apoptosis.
+ Three main modulated sub-graphs identified by \texttt{MSF} at 6~hpi are
+ shown in Fig.~\ref{fig:Sub-graph6hpi}. The gene based graphs on the right
+ hand side, represent the immediate output of the \texttt{MSF}-analysis,
+ visualised by StringApp~\cite{StringApp} in Cytoscape~\cite{Cyto}. Each
+ node represents a gene part of a modualted sub-graph, whereby the
+ associated colors code the functional annotation deduced from KEGG
+ Pathways. The crosstalk between the pathways and also the mutliple
+ employment of many genes is evident. The more schematic drawing on the
+ right side represents the effortlessly deduced flow of information between
+ the sensors and effectors in this particular example.
+
+ In more detail, sub-graph~1 (top) shows how the activation of toll-like
+ receptor, cytokine, chemokine and jak-stat genes lead via TNF into
+ apoptosis. The next significant sub-grap (sub-graph~2: middle) reveals how
+ information from the Extra-cellular matrix (ECM) receptor, which are
+ reported to interact with Ebola glycoprotein (GP)~\cite{Veljkovic},
+ chemokines and cytokines, and cytosolic DNA sensing, is integrated into
+ again modulation of apoptosis pathway. Eventually, sub-graph~3 (bottom)
+ demonstrate how INFA1 and INFB1 modulate once more, via only a few
+ intermediat steps, the aptoptotic response of the cell.
+
+ This show cast example might advertise with how little effort complex data
+ cat be interpreted, help to apprehend the dynamics of the underlying
+ processes and suggest testable hypothesis and potential points of
+ intervention.
 
@@
 
- To assess the robustness and stability of our method, Poisson
- distributed noise was added to the read counts of the three time-points. Then DGEA was
- carried out on the disturbed data with the same parameters as for the
- native data using edgeR, followed by analysis with \texttt{MSF}. This procedure was carried out 100 times
- and every time the genes from the sub-graphs identified from noisy
- data were compared to the genes of sub-graphs identified from the
- native data. This was also done for the DEG obtained for each run, a
- cutoff was used to compare the retrieved DEG from noisy data to the
- native data.  The robustness of \texttt{MSF} and the DEG analysis for
- the time-point 6, 24, and 48 hpi are shown in Fig.~\ref{fig:DEGvsMSF}. We were very stringent in
- adding noise to the counts, which could be seen by the DEG analysis
- robustness. For the last time-point (48h~p.i.) 84~\% genes of the
- sub-graphs from native data were retrieved, for 24~h~p.i. recall rate was
- 79~\%. Robustness was 71~\% at 6~h~p.i., this could be due to that the
- infection merely started and the signal transdcution was weak as
- compared to the later time-points.
+ To assess the robustness and stability of our method, Poisson distributed
+ noise was added to the read counts of the three time-points data set, used
+ in the previous analysis. Then DGEA was carried out on the disturbed data
+ with the same parameters as for the native data using edgeR, followed by
+ analysis with \texttt{MSF}. This procedure was carried out 100 times.
+ Every time the genes from the modulated sub-graphs identified from noisy
+ data were compared to the genes of sub-graphs identified from the native
+ data. This was also done for the DEG obtained for each run, a cutoff
+ \TODO{what cutoff?} was used to compare the retrieved DEG from noisy data
+ to the native data.  The robustness of \texttt{MSF} and the DEG analysis
+ for the time-point 6, 24, and 48~hpi are shown in
+ Fig.~\ref{fig:DEGvsMSF}. The procedure how data noise was modeled can be
+ considered as rather stringent, which is already reflected limited recall
+ rate in the edgeR based DGEA, between \TODO{X and Y~\%}. For
+ \texttt{MSF}-analysis the observed median recall rates lay between 71~\%
+ (6~hpi) and 84~\% (48~hpi).
 
- To test \texttt{MSF's} correctness simulation was also conducted.
-For each time-point, one \texttt{MSF} identified sub-graph was selected and using simulation random graph was generated with the same number of genes as the sub-graphs utilizing the same interaction file.
-The simulation identified sources were counted from the simulated sub-graph.
-The objective
- was to find less sources in the simulated sub-graphs than observed in the real sub-graphs identified by
- \texttt{MSF}. Simulation results showed a chance of $<$10$\%$
- of the 1000 random sub-graphs generated to get more sources than observed.
+ \TODO{This paragraph makes no sense to me:
+ To test \texttt{MSF's} correctness simulation was also conducted. For each
+ time-point, one \texttt{MSF} identified sub-graph was selected and using
+ simulation random graph was generated with the same number of genes as the
+ sub-graphs utilizing the same interaction file. The simulation identified
+ sources were counted from the simulated sub-graph. The objective was to
+ find less sources in the simulated sub-graphs than observed in the real
+ sub-graphs identified by \texttt{MSF}. Simulation results showed a chance
+ of $<$10$\%$ of the 1000 random sub-graphs generated to get more sources
+ than observed.}
+
+\TODO{bis da!!}
 
@@
 
- A comparison to other tools was performed on the same dataset using each software's default parameters. We compared the gene enrichment analysis in \texttt{MSF} identified
-sub-graphs to pathway analysis by SPIA~\cite{Tarca} for KEGG and
-Reactome pathway enrichment analysis for Reactome.  We
-used Cytoscape~\cite{Cyto} plugins to identify the enriched patwhays,
-CytoKegg~\cite{Cytokegg} for KEGG and Reactome FI~\cite{Reactome} for
-Reactome for \texttt{MSF} identified
-sub-graphs.
+ A comparison to other tools was performed on the same dataset using each
+ software's default parameters. We compared the gene enrichment analysis in
+ \texttt{MSF} identified sub-graphs to pathway analysis by
+ SPIA~\cite{Tarca} for KEGG and Reactome pathway enrichment analysis for
+ Reactome.  We used Cytoscape~\cite{Cyto} plugins to identify the enriched
+ patwhays, CytoKegg~\cite{Cytokegg} for KEGG and Reactome
+ FI~\cite{Reactome} for Reactome for \texttt{MSF} identified sub-graphs.
 
 At 6 hpi Cytokegg identified the sub-graph genes to be enriched in
-Toll-like receptor signaling, TNF signaling, IL-17 signaling and
-NF-kappa signaling pathways, as also stated by~\cite{Olejnik}. The
-earliest response EBOV induces on cytokine is via TLR4-mediated
-signaling~\cite{Olejnik}, Gene enrichment analysis of sub-graph genes
-at 6 hpi showed toll-like receptor signaling being most significantly
-dis-regulated pathway, this important signaling pathway was not
-identified by SPIA along with RIG-I-like receptor signaling pathway,
-TNF signaling pathway and IL-17 signaling. On the later time-points \texttt{MSF} and SPIA showed
-consensus on the important pathways e.g Chemokine signaling,
-Cytokine-Cytokine receptor interaction, NF-kappa B signaling,
-Cytosolic DNA-sensing and RIG-I-like receptor signaling as being
-dis-regulated.
+Toll-like receptor signaling, TNF signaling, IL-17 signaling and NF-kappa
+signaling pathways, as also stated by~\cite{Olejnik}. The earliest response
+EBOV induces on cytokine is via TLR4-mediated signaling~\cite{Olejnik},
+Gene enrichment analysis of sub-graph genes at 6 hpi showed toll-like
+receptor signaling being most significantly dis-regulated pathway, this
+important signaling pathway was not identified by SPIA along with
+RIG-I-like receptor signaling pathway, TNF signaling pathway and IL-17
+signaling. On the later time-points \texttt{MSF} and SPIA showed consensus
+on the important pathways e.g Chemokine signaling, Cytokine-Cytokine
+receptor interaction, NF-kappa B signaling, Cytosolic DNA-sensing and
+RIG-I-like receptor signaling as being dis-regulated.
 
-Likewise we compared gene enrichment results of sub-graph genes
-identified by \texttt{MSF} using Reactome FI to the Reactome
-predefined pathway enrichment analysis. Although \texttt{MSF} and
-Reactome showed good agreement on the dis-regulated pathways e.g
-Signaling by Interleukins, Interleukin-10 signaling, Cytokine
-Signaling in Immune system, Chemokine receptors bind chemokines,
-RIG-I/MDA5 mediated induction of IFN-alpha/beta pathways, NF-kB
-activation, Reactome as well could not identify Toll-like receptor
-signaling. This shows that
-\texttt{MSF} is able to identify the important modulated genes of the early infection stage.
+Likewise we compared gene enrichment results of sub-graph genes identified
+by \texttt{MSF} using Reactome FI to the Reactome predefined pathway
+enrichment analysis. Although \texttt{MSF} and Reactome showed good
+agreement on the dis-regulated pathways e.g Signaling by Interleukins,
+Interleukin-10 signaling, Cytokine Signaling in Immune system, Chemokine
+receptors bind chemokines, RIG-I/MDA5 mediated induction of IFN-alpha/beta
+pathways, NF-kB activation, Reactome as well could not identify Toll-like
+receptor signaling. This shows that \texttt{MSF} is able to identify the
+important modulated genes of the early infection stage.
 
@@
 
-MSF is able to identify the connected modulated genes, showing how the signal transmission flows between the different predefined pathways.
-MSF is able to identify the important genes at the early stage of infection which at the later stage of infection became the most modulated genes.
-Even with the addition of noise MSF identifies the important modulated core set of genes.
-The sources identified by MSF can assist in the discovery of new drug targets.
-One limitation for MSF would be an incomplete interaction network.
-For the unknown or less known genes, the interactions are still not well known and so are usually missing in the biological network in different databases.
+MSF is able to identify the connected modulated genes, showing how the
+signal transmission flows between the different predefined pathways. MSF is
+able to identify the important genes at the early stage of infection which
+at the later stage of infection became the most modulated genes. Even with
+the addition of noise MSF identifies the important modulated core set of
+genes. The sources identified by MSF can assist in the discovery of new
+drug targets. One limitation for MSF would be an incomplete interaction
+network. For the unknown or less known genes, the interactions are still
+not well known and so are usually missing in the biological network in
+different databases.
 
@@
 
-The Ebola infection RNAseq dataset analyzed during the current study
-are available in the GEO repository (GSE84188). The cell signaling
-network file used is from Reactome Functional interactions (FIs)
-Version 2016
+The Ebola infection RNAseq dataset analyzed during the current study are
+available in the GEO repository (GSE84188). The cell signaling network file
+used is from Reactome Functional interactions (FIs) Version 2016
 
 \subsection*{Author contributions}
-FA proposed the idea. MF implemented the method. MF designed the
-benchmark data set. MF analyzed the benchmark results. MF, FA, and IH
-contributed to the manuscript writing. MF produced all figures. All
-authors read and approved the final manuscript.
+FA proposed the idea. MF implemented the method. MF designed the benchmark
+data set. MF analyzed the benchmark results. MF, FA, and IH contributed to
+the manuscript writing. MF produced all figures. All authors read and
+approved the final manuscript.
 
@@
 
-We thank all our colleagues who provided insight and expertise that
-greatly assisted the research and all reviewers for their constructive
-criticism.
-
-\subsection*{Supplementary Material}
-
-The supplementary file contains tables of comparison between
-\textbf{MSF} identified sub-graph gene enrichment with SPIA and
-Reactome pathway analysis.
+We thank all our colleagues who provided insight and expertise that greatly
+assisted the research and all reviewers for their constructive criticism.
 

### 2018-04-13 MSF_Paper_FA.tex
review to the end; but not done yet

@@
 \author[1]{Mariam R. Farman}
+\author[1]{Ivo L. Hofacker}
 \author[1,2]{Fabian Amman}
-\author[1]{Ivo L. Hofacker}
 \affil[1]{Institute for Theoretical Chemistry,Theoretical Biochemistry Group,University of Vienna, Austria}
@@
 Encyclopedia of Genes and Genomes)\cite{Kegg} and
-Reactome~\cite{Reactome}. KEGG pathways is a branch of KEGG database that
-hosts a collection of manually drawn pathway maps representing the
-molecular interaction, reaction and relation networks of cellular
+Reactome~\cite{Reactome}. KEGG pathways is a branch of KEGG database
+that hosts a collection of manually drawn pathway maps representing
+the molecular interaction, reaction and relation networks of cellular
 functions. Similarly, Reactome is an open-source, manually curated,
-peer-reviewed database for signaling and metabolic molecules with their
-interactions formed into different biological pathways for nineteen
-species~\cite{Reactome}. Both provide predefined pathways which are sets of
-genes and their interactions categorized into functional units. Starting
-from a gene interaction network, genes are labeled according their role in
-a specific biological process. In this sense a particular gene can be
-assigned to different pathways. E.g., the human gene STAT1 is associated
-with 24 different pathways in the pathway annotation curated by
-KEGG~\footnote{\url{http://www.genome.jp/dbget-bin/www_bget?hsa:6772}} and
-in 12 different pathways in the Reactome data
+peer-reviewed database for signaling and metabolic molecules with
+their interactions formed into different biological pathways for
+nineteen species~\cite{Reactome}. Both provide predefined pathways
+which are sets of genes and their interactions categorized into
+functional units. Starting from a gene interaction network, genes are
+labeled according their role in a specific biological process. In this
+sense a particular gene can be assigned to different pathways. E.g.,
+the human gene STAT1 is associated with 24 different pathways in the
+pathway annotation curated by
+KEGG~\footnote{\url{http://www.genome.jp/dbget-bin/www_bget?hsa:6772}}
+and in 12 different pathways in the Reactome data
 base~\footnote{\url{http://www.reactome.org/content/detail/R-HSA-629622}}.
-Although carefully produced, the assignment of genes to those predefined
-pathway unites can be considered to be subjective to some degree and
-suffers from observational bias~\cite{schnoes2013biases}.
+Although carefully produced, the assignment of genes to those
+predefined pathway unites can be considered to be subjective to some
+degree and suffers from observational bias~\cite{schnoes2013biases}.
 
-Existing pathway-based analysis approaches use different research designs,
-which can be categorized into ORA (Over-representation analysis), FCS
-(Functional class scoring) and pathway topology based methods. All of which
-aim in finding in a subset of genes, e.g., significantly differentially
-expressed genes, genes associated with a certain pathway more often than
-expected given the total set of examined genes, e.g.~the whole genome
-background.  ORA is the first and the most basic method of pathway
-analysis. It uses a DEG list with user defined cut-off for the log-fold
-change and \textit{p}-value (most commonly using absolute log-fold change
-$\geq$ 2 and \textit{p}-value $\leq$ 0.05). Subsequently, sets of genes
-associated with annotated pathways are tested for being over-represented in
-the set of DEGs. To this end, hyper-geometric distribution, chi-square
-tests, binomial probability or the Fisher’s exact test are used. Thereby
-the information of the topology of genes in the pathways are
-ignored~\cite{Bayer}. Furthermore, ORA assumes that the biological pathways
-are independent of each other and ignores the fact that they cross-talk and
-overlap~\cite{Khatri2012,Campos}.
+Existing pathway-based analysis approaches use different research
+designs, which can be categorized into ORA (Over-representation
+analysis), FCS (Functional class scoring) and pathway topology based
+methods. All of which aim in finding in a subset of genes, e.g.,
+significantly differentially expressed genes, genes associated with a
+certain pathway more often than expected given the total set of
+examined genes, e.g.~the whole genome background.  ORA is the first
+and the most basic method of pathway analysis. It uses a DEG list with
+user defined cut-off for the log-fold change and \textit{p}-value
+(most commonly using absolute log-fold change $\geq$ 2 and
+\textit{p}-value $\leq$ 0.05). Subsequently, sets of genes associated
+with annotated pathways are tested for being over-represented in the
+set of DEGs. To this end, hyper-geometric distribution, chi-square
+tests, binomial probability or the Fisher’s exact test are
+used. Thereby the information of the topology of genes in the pathways
+are ignored~\cite{Bayer}. Furthermore, ORA assumes that the biological
+pathways are independent of each other and ignores the fact that they
+cross-talk and overlap~\cite{Khatri2012,Campos}.
 
-Unlike ORA, FCS has no artificial cut-off to define DEG list. FCS works in
-three step, first it calculates the gene-level statistics including
-correlation of molecular measurements using differential expression of
-individual genes, ANOVA, t-test and Z-score. In the second step the
-statistics of individual genes in a pathway are transformed to an
-individual pathway-level statistic commonly using Kolmogorov-Smirnov
-statistic, mean or median. Finally the statistical significance of the
-pathway-level statistics is assessed. Although FCS covers some of the
-limitations from ORA, it still lacks the topology of genes in a pathway,
-cross-talk and overlap of the pathways~\cite{Khatri2012,Campos}. Pathway
-topology based methods are similar to FCS except that they consider the
-topology of each gene during the gene-level statistics but still don't aim
-to link different functional pathways~\cite{Khatri2012}.
+Unlike ORA, FCS has no artificial cut-off to define DEG list. FCS
+works in three step, first it calculates the gene-level statistics
+including correlation of molecular measurements using differential
+expression of individual genes, ANOVA, t-test and Z-score. In the
+second step the statistics of individual genes in a pathway are
+transformed to an individual pathway-level statistic commonly using
+Kolmogorov-Smirnov statistic, mean or median. Finally the statistical
+significance of the pathway-level statistics is assessed. Although FCS
+covers some of the limitations from ORA, it still lacks the topology
+of genes in a pathway, cross-talk and overlap of the
+pathways~\cite{Khatri2012,Campos}. Pathway topology based methods are
+similar to FCS except that they consider the topology of each gene
+during the gene-level statistics but still don't aim to link different
+functional pathways~\cite{Khatri2012}.
 
-On these grounds we propose a novel approach to make use of the rich gene
-and protein interaction annotation resources available to gain additional
-functional insights from basic DGEA. To this, we start with the
-presupposition that expression of neighboring genes within a functional
-pathway are not independent from each other. Rather, they are often
-regulating each others expression or are part of the same
-regulon~\cite{Michalak}. We understand that the categorization of links
-between genes into labeled pathways is often an arbitrary one, given the
-extensive cross talk between different pathways. Although these categories
-have proven to be useful in many situations, they force a certain
-perspective onto the interpretation of novel data. Based on these two
-principles, we aim to find sub-graphs of connected genes within cell
-signaling network which exhibit as a whole significant differential
-expression changes. This approach differes in two main aspects from common
-pathway analysis tools. First, it does not aim to identify functional
-pathways enriched in differentially expressed genes, but detects sub-graphs
-or branches in a network graph (potentially spanning more than one
-functionally grouped pathway) which is coherently modulated. Second, it
-aims to improve the DGEA on the gene level, by collecting the information
-of neighboring genes, which as a whole might exhibit prominent enough
-signal to be called, again as a whole, significantly modulated.
+On these grounds we propose a novel approach to make use of the rich
+gene and protein interaction annotation resources available to gain
+additional functional insights from basic DGEA. To this, we start with
+the presupposition that expression of neighboring genes within a
+functional pathway are not independent from each other. Rather, they
+are often regulating each others expression or are part of the same
+regulon~\cite{Michalak}. We understand that the categorization of
+links between genes into labeled pathways is often an arbitrary one,
+given the extensive cross talk between different pathways. Although
+these categories have proven to be useful in many situations, they
+force a certain perspective onto the interpretation of novel
+data. Based on these two principles, we aim to find sub-graphs of
+connected genes within cell signaling network which exhibit as a whole
+significant differential expression changes. This approach differes in
+two main aspects from common pathway analysis tools. First, it does
+not aim to identify functional pathways enriched in differentially
+expressed genes, but detects sub-graphs or branches in a network graph
+(potentially spanning more than one functionally grouped pathway)
+which is coherently modulated. Second, it aims to improve the DGEA on
+the gene level, by collecting the information of neighboring genes,
+which as a whole might exhibit prominent enough signal to be called,
+again as a whole, significantly modulated.
 
 As input, information on functional links between genes provided by
-e.g.~KEGG or Reactome and information on the differential expression status
-of single genes resulting from a DGEA, are required. As a result the
-analysis returns sub-graphs and their joint confidence scores, reflecting
-how the perturbation is migrated through the network. Furthermore, the
-entry points of perturbation in the networks and overlap with conventional
-pathway categories are returned. The output is prepared for convenient
-visualisation with the \TODO{StringApp, available as a Cytoscape plug-in.}
+e.g.~KEGG or Reactome and information on the differential expression
+status of single genes resulting from a DGEA, are required. As a
+result the analysis returns sub-graphs and their joint confidence
+scores, reflecting how the perturbation is migrated through the
+network. Furthermore, the entry points of perturbation in the networks
+and overlap with conventional pathway categories are returned. The
+output is prepared in a directed adjacency file, convenient for
+visualization with the StringApp~\cite{StringApp}, available as a
+Cytoscape plug-in~\cite{Cyto}.
 
-All of this can be helpful to understand the cause and effect of a stimulus
-and might inform about potential points of intervention. The proposed
-algorithm was implemented as a java program, which was named
+All of this can be helpful to understand the cause and effect of a
+stimulus and might inform about potential points of intervention. The
+proposed algorithm was implemented as a java program, which was named
 \texttt{Modulated Sub-graph Finder} (abbreviated
-\texttt{MSF}). \texttt{MSF} can help transform the information obtained
-from DGEA into comprehensible knowledge of signal transduction of genes and
-thereby being a valuable complement to existing pathway based
-methods. \texttt{MSF} is freely accessible on github \TODO{which licence?}.
+\texttt{MSF}). \texttt{MSF} can help transform the information
+obtained from DGEA into comprehensible knowledge of signal
+transduction of genes andl thereby being a valuable complement to
+existing pathway based methods. \texttt{MSF} is freely accessible on
+github under the terms of the Creative Commons Attribution 4.0
+International License.
 
 \section*{Methods}
-\texttt{MSF} is developed as a novel heuristic approach to
-find concertedly modulated sub-graphs in networks of biological interactions.
-\texttt{MSF} does not use predefined gene sets grouped into
-functional units, but rather relies purely on the network of interacting genes.
-The inputed network consists of nodes corresponding to genes and edges
-representing interactions. Furthermore it utilizes comprehensive results from a
-differential gene expression analysis to discover the sub-graphs, or modules, which
-are as a whole modulated.
+\texttt{MSF} is developed as a novel heuristic approach to find
+concertedly modulated sub-graphs in networks of biological
+interactions.  \texttt{MSF} does not use predefined gene sets grouped
+into functional units, but rather relies purely on the network of
+interacting genes.  The inputed network consists of nodes
+corresponding to genes and edges representing
+interactions. Furthermore it utilizes comprehensive results from a
+differential gene expression analysis to discover the sub-graphs, or
+modules, which are as a whole modulated.
 
@@
 scores and correlation are applied to the inverse normal function to
-calculate the combined \textit{p}-value for all genes examined, namely the
-examined sub-graph. The combined \textit{p}-value of a sub-graph will
-express the significance of all genes in the sub-graph being modulated
-together. Thereby, the information from the different genes are used as,
-although not independent, replicated measurement of the behavior of the whole
-sub-graph. This potentially increases the power to detect also significant
-sub-graphs consisting of genes which are not significant on there own.
+calculate the combined \textit{p}-value for all genes examined, namely
+the examined sub-graph. The combined \textit{p}-value of a sub-graph
+will express the significance of all genes in the sub-graph being
+modulated together. Thereby, the information from the different genes
+are used as, although not independent, replicated measurement of the
+behavior of the whole sub-graph. This potentially increases the power
+to detect also significant sub-graphs consisting of genes which are
+not significant on there own.
 \newline
@@
 \texttt{MSF} applies a four steps heuristic as described in the
-following. The proceeding identification of modulated sub-graphs from a
-network by \texttt{MSF} is presented as a flowchart diagram
+following. The proceeding identification of modulated sub-graphs from
+a network by \texttt{MSF} is presented as a flowchart diagram
 (Fig.~\ref{fig:pseudocodemsf}).
@@
 \texttt{MSF} constructs the first sub-graph starting with the genes
-associated with the lowest (most significant) \textit{p}-value deduced from
-the DGEA. From this seed the sub-graph is attempted to be extended with its
-neighboring genes, starting with the most significant one. A single
-combined \textit{p}-value is calculated for the two genes. If the combined
-\textit{p}-value is smaller than the minimal individual gene's
-\textit{p}-value, the extended sub-graph is accepted. This step is
-iteratively repeated until no further extension is accepted. In this case
-the process starts over with all remaining genes not yet in a significantly
-modulated sub-graph. This step identifies all the trivial sub-graphs that
-are modulated in the whole network.\newline
+associated with the lowest (most significant) \textit{p}-value deduced
+from the DGEA. From this seed the sub-graph is attempted to be
+extended with its neighboring genes, starting with the most
+significant one. A single combined \textit{p}-value is calculated for
+the two genes. If the combined \textit{p}-value is smaller than the
+minimal individual gene's \textit{p}-value, the extended sub-graph is
+accepted. This step is iteratively repeated until no further extension
+is accepted. In this case the process starts over with all remaining
+genes not yet in a significantly modulated sub-graph. This step
+identifies all the trivial sub-graphs that are modulated in the whole
+network.\newline
 
@@
 
-In the next step initial modulated sub-graphs are used to check if they
-could further be extended beyond the immediate neighborhood. This is done
-by testing all possible extension paths up to N genes for all genes in the
-sub-graph. Again, this step is iteratively repeated until no further genes
-are added to the significant differentially expressed sub-graphs. This
-steps bridges small gaps of genes without a clear differential signal in
-the DGEA.\newline
+In the next step initial modulated sub-graphs are used to check if
+they could further be extended beyond the immediate neighborhood. This
+is done by testing all possible extension paths up to N genes for all
+genes in the sub-graph. Again, this step is iteratively repeated until
+no further genes are added to the significant differentially expressed
+sub-graphs. This steps bridges small gaps of genes without a clear
+differential signal in the DGEA.\newline
 
@@
 
-In a last post processing step \texttt{MSF} identifies the trigger points
-of the modulated sub-graphs. These trigger genes are the sources of the
-sub-graphs with only outgoing edges. These genes can be interpreted as the
-possible entry points of perturbation from where the stimulus causes
-downstream effects. In the same spirit the most downstream genes of the
-modulated sub-graph are identified and defined as sinks. Sinks can be
-interpreted as the effectors where the integrated information within the
-signal transduction network is set to action. Due to circular loops not all
-sub-graphs are guaranteed to have sources or sinks.
+In a last post processing step \texttt{MSF} identifies the trigger
+points of the modulated sub-graphs. These trigger genes are the
+sources of the sub-graphs with only outgoing edges. These genes can be
+interpreted as the possible entry points of perturbation from where
+the stimulus causes downstream effects. In the same spirit the most
+downstream genes of the modulated sub-graph are identified and defined
+as sinks. Sinks can be interpreted as the effectors where the
+integrated information within the signal transduction network is set
+to action. Due to circular loops not all sub-graphs are guaranteed to
+have sources or sinks.
 
@@
 
-To demonstrate its usefulness, \texttt{MSF} is applied to an RNAseq data
-set of primary human monocyte-derived macrophages (MDMs) infected with
-Ebola virus (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV) belongs to the
-Filoviridea family; filamentous, enveloped and single stranded RNA
-viruses. EBOV causes hemorrhagic fever in humans, inducing the host innate
-and adaptive immune response to be unable to control virus
-infection~\cite{Prins}. Until now, there are no approved antiviral drugs
-for the treatment of Ebola virus infection~\cite{Konde,Rhein}.  The initial
-targets of EBOV are the macrophages and dendritic immune
+To demonstrate its usefulness, \texttt{MSF} is applied to an RNAseq
+data set of primary human monocyte-derived macrophages (MDMs) infected
+with Ebola virus (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV) belongs
+to the Filoviridea family; filamentous, enveloped and single stranded
+RNA viruses. EBOV causes hemorrhagic fever in humans, inducing the
+host innate and adaptive immune response to be unable to control virus
+infection~\cite{Prins}. Until now, there are no approved antiviral
+drugs for the treatment of Ebola virus infection~\cite{Konde,Rhein}.
+The initial targets of EBOV are the macrophages and dendritic immune
 cells~\cite{Falasca,Rhein}. EBOV inhibits the critical innate immune
 response of the host, which includes the activation of alpha/beta
-interferon (IFN-$\alpha / \beta$)~\cite{Prins,Konde,Cardenas}. It has been
-proposed that IFN-$\alpha / \beta$ should be tested against Ebola for its
-antiviral activity through clinical trials~\cite{Konde}. \TODO{I do not get
-this sentence: Since the data was most recent and the infection was
-carried out in human macrophages, it helped us comprehend the innate
-immune response of host}. The aim of testing this data with \texttt{MSF} was
-to identify the modulated sub-graphs and check if \texttt{MSF} is able to
-identify the IFN-$\alpha / \beta$ gene as one of the sources, and show that
-the capability to close ``gaps'' by using information of neighboring genes
-is effective.
+interferon (IFN-$\alpha / \beta$)~\cite{Prins,Konde,Cardenas}. It has
+been proposed that IFN-$\alpha / \beta$ should be tested against Ebola
+for its antiviral activity through clinical trials~\cite{Konde}. Ebola
+infection data was selected to test the approach because it has been
+well recognized in the literature for the last several decades. It
+would be easy to cross verify the results of \texttt{MSF} with the
+vast literature present on Ebola infection.  The aim of testing this
+data with \texttt{MSF} was to identify the modulated sub-graphs and
+check if \texttt{MSF} is able to identify the IFN-$\alpha / \beta$
+gene as one of the sources, and show that the capability to close
+``gaps'' by using information of neighboring genes is effective.
 
-\TODO{TODO: If MSF is not applied but only significant genes are mapped on
-  the reactom data set, into how many sub-graphs would the data fall apart?
-  in other words: how many genes are only significant because they are part
-  of a significant sub-graph but are not on there own?}
-
-EBOV infection count data was downloaded from GEO (GSE84188). Differential
-gene expression analysis was performed on the count data with edgeR package
-(version 3.4.2)~\cite{edgeR}. The DEG analysis results generated by edgeR
-were used as input for \texttt{MSF}. The EBOV infection experiments
-describe the course of infection at three time-points 6, 24 and 48 hour
-post infection (hpi). For the earliest time point at 6~hpi, five modulated
-sub-graphs were identified with 41, 107, and 18 genes. Two of the five
-sub-graphs were less than 4 genes long. Most of the genes part of the
-sub-graphs were cytokines, chemokines (CXCL10, CCL8) and Interleukin genes
-(IL6, IL27, IL23) which serve as a subset of CD4+ T helper cells. IFNB1 and
-IFNA1 were both identified as two of the possible sources. Most of the
-sources identified by \texttt{MSF} were type~I interferon induced genes. At
-24~hpi nine modulated sub-graphs were identified with four main sub-graphs
-consisting of 38, 85, 148 and 167 genes. IFNA1 was identified as one of the
-sources in the most significantly modulated sub-graph. For the last
-time-point 48~hpi, eleven modulated sub-graphs were identified. Eight of
-the sub-graphs were less than seven genes and main sub-graphs had 39, 71
-and 210 genes. IFNB1 and IFNA1 were identified as the two sources out of
-the possible sources.
+EBOV infection count data was downloaded from GEO (GSE84188)
+\TODO{citation needed};\TODO{What was infected?}. Differential gene
+expression analysis was performed on the count data with edgeR package
+(version 3.4.2)~\cite{edgeR}. The DEG analysis results generated by
+edgeR were used as input for \texttt{MSF}. The EBOV infection
+experiments describe the course of infection at three time-points 6,
+24 and 48 hour post infection (hpi). For the earliest time point at
+6~hpi, five modulated sub-graphs were identified with 41, 107, and 18
+genes. Two of the five sub-graphs were less than 4 genes long. Most of
+the genes part of the sub-graphs were cytokines, chemokines (CXCL10,
+CCL8) and Interleukin genes (IL6, IL27, IL23) which serve as a subset
+of CD4+ T helper cells. IFNB1 and IFNA1 were both identified as two of
+the possible sources. Most of the sources identified by \texttt{MSF}
+were type~I interferon induced genes. At 24~hpi nine modulated
+sub-graphs were identified with four main sub-graphs consisting of 38,
+85, 148 and 167 genes. IFNA1 was identified as one of the sources in
+the most significantly modulated sub-graph. For the last time-point
+48~hpi, eleven modulated sub-graphs were identified. Eight of the
+sub-graphs were less than seven genes and main sub-graphs had 39, 71
+and 210 genes. IFNB1 and IFNA1 were identified as the two sources out
+of the possible sources. Table 1 shows if \texttt{MSF} does not
+consider the complete DEG results and uses different cut offs for
+significant genes from DEG results, how the signal transduction would
+fall apart into meaningless small sub-graphs and single unconnected
+genes.
 
@@
 a source in the early time-point when the genes were not significant
-based on the individual DGEA alone. Identifying the possible sources will
-reduce the search space for potential target genes and can help the
-biologist as the starting point of clinical testing for drugs and
+based on the individual DGEA alone. Identifying the possible sources
+will reduce the search space for potential target genes and can help
+the biologist as the starting point of clinical testing for drugs and
 vaccines against an infection.
+
+\TODO{I hoped that you would not just add the table as discussed but
+  would also add some discussion in the main text explaining what the
+  table~\ref{tab:rawVsHartung} illustrates!}
 
@@
 
- Three main modulated sub-graphs identified by \texttt{MSF} at 6~hpi are
- shown in Fig.~\ref{fig:Sub-graph6hpi}. The gene based graphs on the right
- hand side, represent the immediate output of the \texttt{MSF}-analysis,
- visualised by StringApp~\cite{StringApp} in Cytoscape~\cite{Cyto}. Each
- node represents a gene part of a modualted sub-graph, whereby the
- associated colors code the functional annotation deduced from KEGG
- Pathways. The crosstalk between the pathways and also the mutliple
- employment of many genes is evident. The more schematic drawing on the
- right side represents the effortlessly deduced flow of information between
- the sensors and effectors in this particular example.
+ Three main modulated sub-graphs identified by \texttt{MSF} at 6~hpi
+ are shown in Fig.~\ref{fig:Sub-graph6hpi}. The gene based graphs on
+ the right hand side, represent the immediate output of the
+ \texttt{MSF}-analysis, visualised by StringApp~\cite{StringApp} in
+ Cytoscape~\cite{Cyto}. Each node represents a gene part of a
+ modualted sub-graph, whereby the associated colors code the
+ functional annotation deduced from KEGG Pathways. The crosstalk
+ between the pathways and also the mutliple employment of many genes
+ is evident. The more schematic drawing on the right side represents
+ the effortlessly deduced flow of information between the sensors and
+ effectors in this particular example.
 
- In more detail, sub-graph~1 (top) shows how the activation of toll-like
- receptor, cytokine, chemokine and jak-stat genes lead via TNF into
- apoptosis. The next significant sub-grap (sub-graph~2: middle) reveals how
- information from the Extra-cellular matrix (ECM) receptor, which are
- reported to interact with Ebola glycoprotein (GP)~\cite{Veljkovic},
- chemokines and cytokines, and cytosolic DNA sensing, is integrated into
- again modulation of apoptosis pathway. Eventually, sub-graph~3 (bottom)
- demonstrate how INFA1 and INFB1 modulate once more, via only a few
- intermediat steps, the aptoptotic response of the cell.
+ In more detail, sub-graph~1 (top) shows how the activation of
+ toll-like receptor, cytokine, chemokine and jak-stat genes lead via
+ TNF into apoptosis. The next significant sub-grap (sub-graph~2:
+ middle) reveals how information from the Extra-cellular matrix (ECM)
+ receptor, which are reported to interact with Ebola glycoprotein
+ (GP)~\cite{Veljkovic}, chemokines and cytokines, and cytosolic DNA
+ sensing, is integrated into again modulation of apoptosis
+ pathway. Eventually, sub-graph~3 (bottom) demonstrate how INFA1 and
+ INFB1 modulate once more, via only a few intermediat steps, the
+ aptoptotic response of the cell.
 
- This show cast example might advertise with how little effort complex data
- cat be interpreted, help to apprehend the dynamics of the underlying
- processes and suggest testable hypothesis and potential points of
- intervention.
+ This show cast example might advertise with how little effort complex
+ data cat be interpreted, help to apprehend the dynamics of the
+ underlying processes and suggest testable hypothesis and potential
+ points of intervention.
 
@@
 
- To assess the robustness and stability of our method, Poisson distributed
- noise was added to the read counts of the three time-points data set, used
- in the previous analysis. Then DGEA was carried out on the disturbed data
- with the same parameters as for the native data using edgeR, followed by
- analysis with \texttt{MSF}. This procedure was carried out 100 times.
- Every time the genes from the modulated sub-graphs identified from noisy
- data were compared to the genes of sub-graphs identified from the native
- data. This was also done for the DEG obtained for each run, a cutoff
- \TODO{what cutoff?} was used to compare the retrieved DEG from noisy data
- to the native data.  The robustness of \texttt{MSF} and the DEG analysis
- for the time-point 6, 24, and 48~hpi are shown in
- Fig.~\ref{fig:DEGvsMSF}. The procedure how data noise was modeled can be
- considered as rather stringent, which is already reflected limited recall
- rate in the edgeR based DGEA, between \TODO{X and Y~\%}. For
- \texttt{MSF}-analysis the observed median recall rates lay between 71~\%
- (6~hpi) and 84~\% (48~hpi).
+ To assess the robustness and stability of our method, Poisson
+ distributed noise was added to the read counts of the three
+ time-points data set, used in the previous analysis. Then DGEA was
+ carried out on the disturbed data with the same parameters as for the
+ native data using edgeR, followed by analysis with \texttt{MSF}. This
+ procedure was carried out 100 times.  Every time the genes from the
+ modulated sub-graphs identified from noisy data were compared to the
+ genes of sub-graphs identified from the native data. This was also
+ done for the DEG obtained for each run, a cutoff of FDR less than
+ 0.05 was used to compare the retrieved DEG from noisy data to the
+ native data.  The robustness of \texttt{MSF} and the DEG analysis for
+ the time-point 6, 24, and 48~hpi are shown in
+ Fig.~\ref{fig:DEGvsMSF}. The procedure how data noise was modeled can
+ be considered as rather stringent, which is already reflected limited
+ recall rate in the edgeR based DGEA, between \TODO{X and Y~\%}. For
+ \texttt{MSF}-analysis the observed median recall rates lay between
+ 71~\% (6~hpi) and 84~\% (48~hpi).
 
- \TODO{This paragraph makes no sense to me:
- To test \texttt{MSF's} correctness simulation was also conducted. For each
- time-point, one \texttt{MSF} identified sub-graph was selected and using
- simulation random graph was generated with the same number of genes as the
- sub-graphs utilizing the same interaction file. The simulation identified
- sources were counted from the simulated sub-graph. The objective was to
- find less sources in the simulated sub-graphs than observed in the real
- sub-graphs identified by \texttt{MSF}. Simulation results showed a chance
- of $<$10$\%$ of the 1000 random sub-graphs generated to get more sources
- than observed.}
+\subsection*{Comparison to existing pathway analysis tools}
 
-\TODO{bis da!!}
-
-\subsection*{Comparison}
-
- A comparison to other tools was performed on the same dataset using each
- software's default parameters. We compared the gene enrichment analysis in
+ A comparison to other tools was performed on the same dataset using
+ each software's default parameters. We compared the gene enrichment
+ analysis \TODO{does MSF realy do a ``gene enrichment analysis''?} in
  \texttt{MSF} identified sub-graphs to pathway analysis by
- SPIA~\cite{Tarca} for KEGG and Reactome pathway enrichment analysis for
- Reactome.  We used Cytoscape~\cite{Cyto} plugins to identify the enriched
- patwhays, CytoKegg~\cite{Cytokegg} for KEGG and Reactome
- FI~\cite{Reactome} for Reactome for \texttt{MSF} identified sub-graphs.
+ SPIA~\cite{Tarca} for KEGG and Reactome pathway enrichment analysis
+ for Reactome.  We used Cytoscape~\cite{Cyto} plugins to identify the
+ enriched patwhays, CytoKegg~\cite{Cytokegg} for KEGG and Reactome
+ FI~\cite{Reactome} for Reactome for \texttt{MSF} identified
+ sub-graphs. \TODO{I do not get those two sentences. What do those
+   tools do? How do you compare the results of those tools?}
 
 At 6 hpi Cytokegg identified the sub-graph genes to be enriched in
-Toll-like receptor signaling, TNF signaling, IL-17 signaling and NF-kappa
-signaling pathways, as also stated by~\cite{Olejnik}. The earliest response
-EBOV induces on cytokine is via TLR4-mediated signaling~\cite{Olejnik},
-Gene enrichment analysis of sub-graph genes at 6 hpi showed toll-like
-receptor signaling being most significantly dis-regulated pathway, this
-important signaling pathway was not identified by SPIA along with
-RIG-I-like receptor signaling pathway, TNF signaling pathway and IL-17
-signaling. On the later time-points \texttt{MSF} and SPIA showed consensus
-on the important pathways e.g Chemokine signaling, Cytokine-Cytokine
-receptor interaction, NF-kappa B signaling, Cytosolic DNA-sensing and
-RIG-I-like receptor signaling as being dis-regulated.
+Toll-like receptor signaling, TNF signaling, IL-17 signaling and
+NF-kappa signaling pathways, as also stated by~\cite{Olejnik}. The
+earliest response EBOV induces on cytokine is via TLR4-mediated
+signaling~\cite{Olejnik}, Gene enrichment analysis of sub-graph genes
+at 6 hpi showed toll-like receptor signaling being most significantly
+dis-regulated pathway, this important signaling pathway was not
+identified by SPIA along with RIG-I-like receptor signaling pathway,
+TNF signaling pathway and IL-17 signaling. On the later time-points
+\texttt{MSF} and SPIA showed consensus on the important pathways e.g
+Chemokine signaling, Cytokine-Cytokine receptor interaction, NF-kappa
+B signaling, Cytosolic DNA-sensing and RIG-I-like receptor signaling
+as being dis-regulated.
 
-Likewise we compared gene enrichment results of sub-graph genes identified
-by \texttt{MSF} using Reactome FI to the Reactome predefined pathway
-enrichment analysis. Although \texttt{MSF} and Reactome showed good
-agreement on the dis-regulated pathways e.g Signaling by Interleukins,
-Interleukin-10 signaling, Cytokine Signaling in Immune system, Chemokine
-receptors bind chemokines, RIG-I/MDA5 mediated induction of IFN-alpha/beta
-pathways, NF-kB activation, Reactome as well could not identify Toll-like
-receptor signaling. This shows that \texttt{MSF} is able to identify the
-important modulated genes of the early infection stage.
+Likewise we compared gene enrichment results of sub-graph genes
+identified by \texttt{MSF} using Reactome FI to the Reactome
+predefined pathway enrichment analysis. Although \texttt{MSF} and
+Reactome showed good agreement on the dis-regulated pathways e.g
+Signaling by Interleukins, Interleukin-10 signaling, Cytokine
+Signaling in Immune system, Chemokine receptors bind chemokines,
+RIG-I/MDA5 mediated induction of IFN-alpha/beta pathways, NF-kB
+activation, Reactome as well could not identify Toll-like receptor
+signaling. This shows that \texttt{MSF} is able to identify the
+important modulated genes of the early infection stage. \TODO{The
+  whole paragraph is hard to read and get the essentials. I suggest
+  you describe in this section what the different tools work,
+  especially what are the differences between them. For the results I
+  suggest a plot; maybe something in the line of upset or a venn
+  diagram, if appropriate.}
 
@@
 
-MSF is able to identify the connected modulated genes, showing how the
-signal transmission flows between the different predefined pathways. MSF is
-able to identify the important genes at the early stage of infection which
-at the later stage of infection became the most modulated genes. Even with
-the addition of noise MSF identifies the important modulated core set of
-genes. The sources identified by MSF can assist in the discovery of new
-drug targets. One limitation for MSF would be an incomplete interaction
-network. For the unknown or less known genes, the interactions are still
-not well known and so are usually missing in the biological network in
-different databases.
+Classic pathway analysis tools aim to detect in lists of significantly
+deregulated genes enriched associations with pathway genes categorized
+by their biological function and their interactions. Thereby,
+depending on the tool, the internal pathway topology is considered or
+neglected alltogether. The here presented tool, \texttt{MSF}, employs
+a different approach, by aiming to detect sub-graphs in whole gene
+regulatory networks which are significantly deregulated in a concerted
+manner. To this end, neighboring genes in the user provided network
+are tested for jointly common regulation. Exploiting that each gene's
+abundance, although not independent from its neighbors, is measured
+repeatedly on its own, sensitivity can be increased by our applied
+p-value meta-analyse, namely Hartung's method. This potentially
+enables to call just not significant modulated genes based on the DGEA
+to be convincingly called to be part of a deregulated gene group.
+Furthermore, it allows to identify connected sub-graphs, representing
+the propagation of gene regulation perturbation in the input
+network. A better understanding of this propagation, especially the
+critical spots such as sensors, effectors, or intermediate bottlenecks
+and hubs, facilitates the projection of potential intervention points,
+e.g. for drug development. Since \texttt{MSF} only uses interaction
+information in gene regulation network, but not the functional
+grouping of the genes into functional pathways, it is especially
+adapted to discover so called cross-talk between such pathways.
 
@@
 
-\textbf{MSF} is a fast and robust tool to find modulated sub-graphs
-from a given network. So far, it is the only tool from existing
-approaches that builds the modulated sub-graph from scratch using the
-network and DEG, where the predefined pathways could be seen connected
-to each other. The Robustness analysis shows that \textbf{MSF} is
-fairly robust against noise and thus finds core modulated sub-graphs
-even if the DGEA is not carried out vigilantly. Alongside with
-\textbf{MSF}, we provide a good directed cell signaling network file,
-which is a subset of interactions from Reactome Functional
-interactions (FIs) Version 2016 .
+\texttt{MSF} is a fast, robust and easy to use tool to find
+concertedly modulated sub-graphs in a given network. Its
+implementation in java enables it use across many operating
+systems. It requires as input the results from a differential gene
+expression analysis in the appropriate file format. So far the raw
+output from edgeR and DESeq2~\cite{love2014moderated} are supported
+\TODO{did you already implement a input option for
+  DESeq2?}. Furthermore, a gene network has to be provided \TODO{do
+  you check at some point if the IDs between the DGEA input and the
+  gene network fit?} in the format of a simple adjacency list \TODO{is
+  this true?}. For convenience \texttt{MSF} is distributed with a
+prefiltered version of all directed cell signalling interactions from
+Reactome Functional interactions (FIs) Version 2016. \TODO{citation
+  needed} ;\TODO{did you check if reactome comes with the appropriate
+  licence that we allowed to redistribute this file? this is very
+  important otherwise we need good laywers} ;\TODO{In the case study
+  section you never mention this file and especially the prefiltering
+  of it? wouldn t you consider this crucial?}
 
@@
 
-The Ebola infection RNAseq dataset analyzed during the current study are
-available in the GEO repository (GSE84188). The cell signaling network file
-used is from Reactome Functional interactions (FIs) Version 2016
+The Ebola infection RNAseq dataset analyzed during the current study
+are available in the GEO repository (GSE84188) \TODO{citation
+  needed}. The cell signaling network file used is from Reactome
+Functional interactions (FIs) Version 2016 \TODO{citation needed}.
 
-\subsection*{Author contributions}
-FA proposed the idea. MF implemented the method. MF designed the benchmark
-data set. MF analyzed the benchmark results. MF, FA, and IH contributed to
-the manuscript writing. MF produced all figures. All authors read and
-approved the final manuscript.
+\subsection*{Data and Software Availability}
+
+\TODO{see
+  \url{https://f1000research.com/for-authors/article-guidelines/software-tool-articles}
+  what has to be stated there}
+
+\section*{Author contributions}
+Conceptualization: FA, ILH; Funding Acquisition: ILH; Resources: ILH;
+Software: MRF; Writing – Original Draft Preparation: MRF; Writing –
+Review \& Editing: FA, ILH. All authors read and approved the final
+manuscript.
 
@@
 Agency Verfahren) with the project number (I 1988-B22). The grant was
-assigned to Ivo Hofacker
+assigned to ILH. FA was funded by the Austrian Science Fund (FWF)
+project SFB F43.
 
@@
 
-We thank all our colleagues who provided insight and expertise that greatly
-assisted the research and all reviewers for their constructive criticism.
+We thank all our colleagues who provided insight and expertise that
+greatly assisted the research and all reviewers for their constructive
+criticism.\TODO{I'd suggest name those colleagues by name or remove
+  the section.}
 
@@
 		\includegraphics[width=1.0\linewidth]{AlgorithmMSF.png}
-		\caption{Shows the algorithm of MSF with four steps to find the modulated sub-graphs and identify sources and sinks.}
+		\caption{Graphical representation of the \texttt{MSF}
+                  heuristical approach to detect modulated sub-graphs
+                  in a global gene regulatory network without
+                  exhaustively testing all connected sub-graphs.}
 		\label{fig:pseudocodemsf}
@@
 			\includegraphics[width=1.05\linewidth]{SeperatedMergedNetwork.png}
-			\caption{Shows the three modulated sub-graphs identified by
-				MSF at 6 hpi, where nodes are genes and edges are the
-				interactions. The node coloring is associated to KEGG
-				pathways referring to the colors in the legend. The graph
-				edges are from Reactome. Important genes to EBOV infection
-				as from literature are enlarged in the graph. The signal
-				flow corresponding to each sub-graph is depicted as simple
-				flow chart against each sub-graph.\label{fig:Sub-graph6hpi}}
+			\caption{Visualisation of the three modulated
+                          sub-graphs identified by MSF at 6~h after
+                          EBOV infection in gene detail and their
+                          epitomized representation to depict the flow
+                          of perturbation in the directed network. The
+                          node coloring is associated to KEGG pathways
+                          referring to the colors in the legend. The
+                          graph edges are from Reactome. Important
+                          genes to EBOV infection as from literature
+                          are enlarged in the
+                          graph.\label{fig:Sub-graph6hpi}}
 		\end{minipage}
@@
 			\includegraphics[width=16cm]{DEGvsMSF2.png}
-			\caption{Graph shows the percentage of DEG
-                          analysis genes and MSF identified sub-graph
-                          genes recall rates 100 times for the three
-                          different time points after the addition of
-                          Poisson distributed noise.\label{fig:DEGvsMSF}}
+			\caption{The percentage of DEG analysis genes
+                          and MSF identified sub-graph genes recall
+                          rates for the three different time points of
+                          EBOV infection data for 100 simulation where
+                          Poisson distributed noise was added to the
+                          experimental deduced reads per gene
+                          counts.\label{fig:DEGvsMSF}}
 			\label{label}
@@
 
+\begin{table*}[]
+	\centering
+	\caption{Comparison of connected sub-graphs of modulated genes
+          in the global network between the analysis results of
+          \texttt{MSF} and mapping the raw list of differentially
+          expressed genes from the standard edgeR analysis, applying
+          different p-value cut-offs, onto the network. }
+	\label{tab:rawVsHartung}
+	\begin{tabular}{lll}
+		\hline
+		\multicolumn{1}{|l|}{}                            & \multicolumn{1}{c|}{\textbf{Total number of}}       & \multicolumn{1}{c|}{\textbf{Number of connected}}      \\ 
+                \multicolumn{1}{|l|}{}                            & \multicolumn{1}{c|}{\textbf{genes in network}}      & \multicolumn{1}{c|}{\textbf{sub-graph in network}}     \\ \hline
+		\multicolumn{1}{|l|}{\textbf{6hpi}}               & \multicolumn{1}{l|}{}                                                & \multicolumn{1}{c|}{}                 \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{166}                                             & \multicolumn{1}{c|}{3}                \\ \hline
+		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{123}                                             & \multicolumn{1}{c|}{39}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{123}                                             & \multicolumn{1}{c|}{47}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{112}                                             & \multicolumn{1}{c|}{53}               \\ \hline
+		\multicolumn{1}{|l|}{\textbf{24hpi}}              & \multicolumn{1}{c|}{}                                                & \multicolumn{1}{c|}{}                 \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{438}                                             & \multicolumn{1}{c|}{4}                \\ \hline
+		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{325}                                             & \multicolumn{1}{c|}{76}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{305}                                             & \multicolumn{1}{c|}{102}              \\ \hline
+		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{262}                                             & \multicolumn{1}{c|}{110}              \\ \hline
+		\multicolumn{1}{|l|}{\textbf{48hpi}}              & \multicolumn{1}{c|}{}                                                & \multicolumn{1}{c|}{}                 \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{320}                                             & \multicolumn{1}{c|}{3}                \\ \hline
+		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{226}                                             & \multicolumn{1}{c|}{39}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{207}                                             & \multicolumn{1}{c|}{62}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{173}                                             & \multicolumn{1}{c|}{67}               \\ \hline
+	\end{tabular}
+\end{table*}
+
 \end{document}

### 2018-04-29 MSF_Paper_FA.tex
Comparison correction with figure

@@
 connected genes within cell signaling network which exhibit as a whole
-significant differential expression changes. This approach differes in
+significant differential expression changes. This approach differs in
 two main aspects from common pathway analysis tools. First, it does
@@
 obtained from DGEA into comprehensible knowledge of signal
-transduction of genes andl thereby being a valuable complement to
+transduction of genes and thereby being a valuable complement to
 existing pathway based methods. \texttt{MSF} is freely accessible on
@@
 into functional units, but rather relies purely on the network of
-interacting genes.  The inputed network consists of nodes
+interacting genes.  The inputted network consists of nodes
 corresponding to genes and edges representing
@@
 
-EBOV infection count data was downloaded from GEO (GSE84188)
-\TODO{citation needed};\TODO{What was infected?}. Differential gene
+EBOV infection count data was downloaded from GEO (GSE84188). Differential gene
 expression analysis was performed on the count data with edgeR package
 (version 3.4.2)~\cite{edgeR}. The DEG analysis results generated by
-edgeR were used as input for \texttt{MSF}. The EBOV infection
+edgeR were used as input for \texttt{MSF}. Directed cell signaling interactions were filtered from
+Reactome Functional interactions (FIs) Version 2016~\cite{Cytokegg}, which was used as a second input for \texttt{MSF}.
+
+The EBOV infection
 experiments describe the course of infection at three time-points 6,
@@
 and 210 genes. IFNB1 and IFNA1 were identified as the two sources out
-of the possible sources. Table 1 shows if \texttt{MSF} does not
-consider the complete DEG results and uses different cut offs for
-significant genes from DEG results, how the signal transduction would
-fall apart into meaningless small sub-graphs and single unconnected
-genes.
+of the possible sources.
 
@@
 
-\TODO{I hoped that you would not just add the table as discussed but
-  would also add some discussion in the main text explaining what the
-  table~\ref{tab:rawVsHartung} illustrates!}
+Table~\ref{tab:rawVsHartung} shows if \texttt{MSF} does not
+consider the complete DEG results and uses different cut offs for
+significant genes from DEG results, how the signal transduction would
+fall apart into incomprehensible small sub-graphs and single unconnected
+genes.
 
@@
  the right hand side, represent the immediate output of the
- \texttt{MSF}-analysis, visualised by StringApp~\cite{StringApp} in
+ \texttt{MSF}-analysis, visualized by StringApp~\cite{StringApp} in
  Cytoscape~\cite{Cyto}. Each node represents a gene part of a
- modualted sub-graph, whereby the associated colors code the
- functional annotation deduced from KEGG Pathways. The crosstalk
- between the pathways and also the mutliple employment of many genes
+ modulated sub-graph, whereby the associated colors code the
+ functional annotation deduced from KEGG Pathways. The cross-talk
+ between the pathways and also the multiple employment of many genes
  is evident. The more schematic drawing on the right side represents
@@
  toll-like receptor, cytokine, chemokine and jak-stat genes lead via
- TNF into apoptosis. The next significant sub-grap (sub-graph~2:
+ TNF into apoptosis. The next significant sub-graph (sub-graph~2:
  middle) reveals how information from the Extra-cellular matrix (ECM)
@@
  pathway. Eventually, sub-graph~3 (bottom) demonstrate how INFA1 and
- INFB1 modulate once more, via only a few intermediat steps, the
- aptoptotic response of the cell.
+ INFB1 modulate once more, via only a few intermediate steps, the
+ apoptotic response of the cell.
 
@@
  be considered as rather stringent, which is already reflected limited
- recall rate in the edgeR based DGEA, between \TODO{X and Y~\%}. For
+ recall rate in the edgeR based DGEA, between 84~\% (6~hpi) and 93~\% (24~hpi). For
  \texttt{MSF}-analysis the observed median recall rates lay between
@@
 
-\subsection*{Comparison to existing pathway analysis tools}
+\subsection*{Comparison to Reactome pathway analysis tool}
 
- A comparison to other tools was performed on the same dataset using
- each software's default parameters. We compared the gene enrichment
- analysis \TODO{does MSF realy do a ``gene enrichment analysis''?} in
- \texttt{MSF} identified sub-graphs to pathway analysis by
- SPIA~\cite{Tarca} for KEGG and Reactome pathway enrichment analysis
- for Reactome.  We used Cytoscape~\cite{Cyto} plugins to identify the
- enriched patwhays, CytoKegg~\cite{Cytokegg} for KEGG and Reactome
- FI~\cite{Reactome} for Reactome for \texttt{MSF} identified
- sub-graphs. \TODO{I do not get those two sentences. What do those
-   tools do? How do you compare the results of those tools?}
+  Gene enrichment analysis was performed on \texttt{MSF} identified sub-graphs for each time-point using Reactome Analyze data tool~\cite{Reactome}.
+Reactome's over-representation analysis tests whether certain Reactome pathways are enriched by the list of genes submitted.
+For comparison the DEG results from edgeR were filtered using a cut off of adjusted p-value < 0.1.
+This subset of DEG list was used for gene enrichment analysis.
+Enriched pathways were further filtered by confidence score p-values, pathways having p-value less $<$ 0.05 for cut off DEG list and $>$ 1$^{-10}$ for  pathway enrichment in \texttt{MSF} identified sub-graph genes.
+Out of 282 pathways dis-regulated in the two comparison analysis with three time-points each, pathways that were present in at least three time-points are shown in Fig~\ref{fig:msfvsreactome}.
 
-At 6 hpi Cytokegg identified the sub-graph genes to be enriched in
-Toll-like receptor signaling, TNF signaling, IL-17 signaling and
-NF-kappa signaling pathways, as also stated by~\cite{Olejnik}. The
-earliest response EBOV induces on cytokine is via TLR4-mediated
-signaling~\cite{Olejnik}, Gene enrichment analysis of sub-graph genes
-at 6 hpi showed toll-like receptor signaling being most significantly
-dis-regulated pathway, this important signaling pathway was not
-identified by SPIA along with RIG-I-like receptor signaling pathway,
-TNF signaling pathway and IL-17 signaling. On the later time-points
-\texttt{MSF} and SPIA showed consensus on the important pathways e.g
-Chemokine signaling, Cytokine-Cytokine receptor interaction, NF-kappa
-B signaling, Cytosolic DNA-sensing and RIG-I-like receptor signaling
-as being dis-regulated.
-
-Likewise we compared gene enrichment results of sub-graph genes
-identified by \texttt{MSF} using Reactome FI to the Reactome
-predefined pathway enrichment analysis. Although \texttt{MSF} and
-Reactome showed good agreement on the dis-regulated pathways e.g
-Signaling by Interleukins, Interleukin-10 signaling, Cytokine
-Signaling in Immune system, Chemokine receptors bind chemokines,
-RIG-I/MDA5 mediated induction of IFN-alpha/beta pathways, NF-kB
-activation, Reactome as well could not identify Toll-like receptor
-signaling. This shows that \texttt{MSF} is able to identify the
-important modulated genes of the early infection stage. \TODO{The
-  whole paragraph is hard to read and get the essentials. I suggest
-  you describe in this section what the different tools work,
-  especially what are the differences between them. For the results I
-  suggest a plot; maybe something in the line of upset or a venn
-  diagram, if appropriate.}
+  The comparison shows most of the important pathways known from literature to be dis-regulated by Ebola infection are enriched in both the enrichment analysis.
+Toll-Like receptor signaling pathway when interacts with EBOV glycoprotein (GP), triggers the activation of cytokines~\cite{Olejnik}.
+This pathway was not identified when a p-value cut off DEG list was analyzed for enrichment.
+Due to \texttt{MSF} considering the complete DEG results, even the weak signal at the earliest time-point was detected; for example Toll-like receptor signaling.
 
@@
 depending on the tool, the internal pathway topology is considered or
-neglected alltogether. The here presented tool, \texttt{MSF}, employs
+neglected all together. The here presented tool, \texttt{MSF}, employs
 a different approach, by aiming to detect sub-graphs in whole gene
@@
 repeatedly on its own, sensitivity can be increased by our applied
-p-value meta-analyse, namely Hartung's method. This potentially
+p-value meta-analysis, namely Hartung's method. This potentially
 enables to call just not significant modulated genes based on the DGEA
@@
 expression analysis in the appropriate file format. So far the raw
-output from edgeR and DESeq2~\cite{love2014moderated} are supported
-\TODO{did you already implement a input option for
-  DESeq2?}. Furthermore, a gene network has to be provided \TODO{do
-  you check at some point if the IDs between the DGEA input and the
-  gene network fit?} in the format of a simple adjacency list \TODO{is
-  this true?}. For convenience \texttt{MSF} is distributed with a
-prefiltered version of all directed cell signalling interactions from
-Reactome Functional interactions (FIs) Version 2016. \TODO{citation
-  needed} ;\TODO{did you check if reactome comes with the appropriate
-  licence that we allowed to redistribute this file? this is very
-  important otherwise we need good laywers} ;\TODO{In the case study
-  section you never mention this file and especially the prefiltering
-  of it? wouldn t you consider this crucial?}
+output from edgeR~\cite{edgeR} and DESeq2~\cite{love2014moderated} are supported
+. Furthermore, a gene network has to be provided in the format of a directed adjacency list.
 
@@
 The Ebola infection RNAseq dataset analyzed during the current study
-are available in the GEO repository (GSE84188) \TODO{citation
-  needed}. The cell signaling network file used is from Reactome
-Functional interactions (FIs) Version 2016 \TODO{citation needed}.
+are available in the GEO repository (GSE84188)~\cite{Olejnik}. The cell signaling network file used is from Reactome
+Functional interactions (FIs) Version 2016~\cite{Cytokegg}
 
-\subsection*{Data and Software Availability}
+\subsection*{Software Availability}
 
-\TODO{see
-  \url{https://f1000research.com/for-authors/article-guidelines/software-tool-articles}
-  what has to be stated there}
-
+Source code: https://github.com/MariamFarman/Modulated-SubGraph-Finder \newline
+Software license: Creative Commons Attribution 4.0 International License
 \section*{Author contributions}
@@
 project SFB F43.
-
-\subsection*{Acknowledgements}
-
-We thank all our colleagues who provided insight and expertise that
-greatly assisted the research and all reviewers for their constructive
-criticism.\TODO{I'd suggest name those colleagues by name or remove
-  the section.}
 
@@
 
+\begin{figure*}[p]
+	\centering
+	\fbox{
+		\begin{minipage}{10 cm}
+	\includegraphics[width=1.0\linewidth]{MSFvsReactome}
+	\caption{The heatmap shows Reactome pathway enrichment for MSF identifed sub-graphs for the three time-points and the Reactome pathway enrichment for subset of DEG with adjusted p-value cut off of < 0.1.
+Green color shows the pathway being enriched and red color shows pathway not found as enriched.}
+	\label{fig:msfvsreactome}
+\end{minipage}
+}
+\end{figure*}
+
 \begin{table*}[]

### 2018-05-07 MSF_Paper_FA.tex
editing of several parts

@@
 gene expression analysis (DGEA) is the first step in the course of data
-interpretation, typically producing lists of of dozens to thousands of
+interpretation, typically producing lists of dozens to thousands of
 differentially expressed genes. To further guide the interpretation of
 these lists, different pathway analysis approaches have been
-developed. These tools typically classify genes into set of genes, i.e. a
-pathway, which interact with each other and function in a common biological
-process. Regardless of technical differences, all these methods do not
-properly account for cross talk between different pathways and rely on
-binary classification into differentially expressed gene and unaffected
-genes based on an arbitrarily set p-value cut-off.
+developed. These tools typically rely on the classification of genes into
+set of genes, i.e. a pathway, based on the interactions between the genes
+and their function in a common biological process. Regardless of technical
+differences, these methods do not properly account for cross talk between
+different pathways and rely on binary separation into differentially
+expressed gene and unaffected genes based on an arbitrarily set p-value
+cut-off.
 
@@
 concertedly modulated sub-graphs in the global cell signaling network,
-based on the DGEA results of all genes tested. Thereby, expression
-patterns of genes are integrated according to the topology of their
-interactions and allow potentially to read the flow of information
-from the perturbation source to the effectors. The described software,
-named \texttt{Modulated Sub-graph Finder} (\texttt{MSF}) is freely available at
+based on the DGEA results of all genes tested. Thereby, expression patterns
+of genes are integrated according to the topology of their interactions and
+allow potentially to read the flow of information from the perturbation
+source to the effectors. The described software, named \texttt{Modulated
+  Sub-graph Finder} (\texttt{MSF}) is freely available at
 \url{https://github.com/MariamFarman/Modulated-SubGraph-Finder}.
@@
 High throughput sequencing techniques have been widely used to yield
-differentially expressed genes (DEG)~\cite{DEG}. To this end, changes
-in transcript abundance are measured, e.g.~by next generation
-sequencing techniques, and interpreted as an indicator of differential
-expression of genes. DEGs can be used to get insights into the
-mechanism underlying differences between conditions of samples, such
-as healthy versus diseased. Differential gene expression analysis
-(DGEA) informs about the magnitude of expression changes between the
-conditions which are often expressed as fold change, sign of fold
-change and the confidence level of observing an authentic change,
-often expressed as \textit{p}-value. These DEGs information is further
-interpreted to extract meaningful biological insights. For example,
-genes that could be involved in the response to a particular stimuli
-or maybe the cause of a disease. To this end, pathway-based analysis
-has become an important tool to further interpret the results of a
-DGEA and to acquire understandings of the perturbations in a
-biological system. Biological pathways are sets of genes and their
-interactions forming a functional unit. DEGs help to identify pathways
-or networks that may be altered during a change of condition providing
+differentially expressed genes (DEG)~\cite{DEG}. To this end, changes in
+transcript abundance are measured, e.g.~by next generation sequencing
+techniques, and interpreted as an indicator of differential expression of
+genes. DEGs can be used to get insights into the mechanism underlying
+differences between conditions of samples, such as healthy versus
+diseased. Differential gene expression analysis (DGEA) informs about the
+magnitude of expression changes between the conditions which are often
+expressed as fold change, sign of fold change and the confidence level of
+observing an authentic change, often expressed as \textit{p}-value. These
+DEGs information is further interpreted to extract meaningful biological
+insights. For example, genes that could be involved in the response to a
+particular stimuli or maybe the cause of a disease. To this end,
+pathway-based analysis has become an important tool to further interpret
+the results of a DGEA and to acquire understandings of the perturbations in
+a biological system. Biological pathways are sets of genes and their
+interactions forming a functional unit. DEGs help to identify pathways or
+networks that may be altered during a change of condition providing
 important information about diseases and its treatment
-process~\cite{Khatri2012}. Pathway-based methods use predefined
-pathways or networks such as KEGG~\cite{Kegg} and
-Reactome~\cite{Reactome}, the expression measurements of the genes
-obtained from DGEA in combination with statistical methods and
-algorithms to identify specifically modulated pathways and
-processes~\cite{Campos}.
+process~\cite{Khatri2012}. Pathway-based methods use predefined pathways or
+networks such as KEGG~\cite{Kegg} and Reactome~\cite{Reactome}, the
+expression measurements of the genes obtained from DGEA in combination with
+statistical methods and algorithms to identify specifically modulated
+pathways and processes~\cite{Campos}.
 
@@
 Encyclopedia of Genes and Genomes)\cite{Kegg} and
-Reactome~\cite{Reactome}. KEGG pathways is a branch of KEGG database
-that hosts a collection of manually drawn pathway maps representing
-the molecular interaction, reaction and relation networks of cellular
+Reactome~\cite{Reactome}. KEGG pathways is a branch of KEGG database that
+hosts a collection of manually drawn pathway maps representing the
+molecular interaction, reaction and relation networks of cellular
 functions. Similarly, Reactome is an open-source, manually curated,
-peer-reviewed database for signaling and metabolic molecules with
-their interactions formed into different biological pathways for
-nineteen species~\cite{Reactome}. Both provide predefined pathways
-which are sets of genes and their interactions categorized into
-functional units. Starting from a gene interaction network, genes are
-labeled according their role in a specific biological process. In this
-sense a particular gene can be assigned to different pathways. E.g.,
-the human gene STAT1 is associated with 24 different pathways in the
-pathway annotation curated by
-KEGG~\footnote{\url{http://www.genome.jp/dbget-bin/www_bget?hsa:6772}}
-and in 12 different pathways in the Reactome data
+peer-reviewed database for signaling and metabolic molecules with their
+interactions formed into different biological pathways for nineteen
+species~\cite{Reactome}. Both provide predefined pathways which are sets of
+genes and their interactions categorized into functional units. Starting
+from a gene interaction network, genes are labeled according their role in
+a specific biological process. In this sense a particular gene can be
+assigned to different pathways. E.g., the human gene STAT1 is associated
+with 24 different pathways in the pathway annotation curated by
+KEGG~\footnote{\url{http://www.genome.jp/dbget-bin/www_bget?hsa:6772}} and
+in 12 different pathways in the Reactome data
 base~\footnote{\url{http://www.reactome.org/content/detail/R-HSA-629622}}.
-Although carefully produced, the assignment of genes to those
-predefined pathway unites can be considered to be subjective to some
-degree and suffers from observational bias~\cite{schnoes2013biases}.
+Although carefully produced, the assignment of genes to those predefined
+pathway unites can be considered to be subjective to some degree and
+suffers from observational bias~\cite{schnoes2013biases}.
 
-Existing pathway-based analysis approaches use different research
-designs, which can be categorized into ORA (Over-representation
-analysis), FCS (Functional class scoring) and pathway topology based
-methods. All of which aim in finding in a subset of genes, e.g.,
-significantly differentially expressed genes, genes associated with a
-certain pathway more often than expected given the total set of
-examined genes, e.g.~the whole genome background.  ORA is the first
-and the most basic method of pathway analysis. It uses a DEG list with
-user defined cut-off for the log-fold change and \textit{p}-value
-(most commonly using absolute log-fold change $\geq$ 2 and
-\textit{p}-value $\leq$ 0.05). Subsequently, sets of genes associated
-with annotated pathways are tested for being over-represented in the
-set of DEGs. To this end, hyper-geometric distribution, chi-square
-tests, binomial probability or the Fisher’s exact test are
-used. Thereby the information of the topology of genes in the pathways
-are ignored~\cite{Bayer}. Furthermore, ORA assumes that the biological
+Existing pathway-based analysis approaches use different research designs,
+which can be categorized into ORA (Over-representation analysis), FCS
+(Functional class scoring) and pathway topology based methods. All of which
+aim in finding in a subset of genes, e.g., significantly differentially
+expressed genes, genes associated with a certain pathway more often than
+expected given the total set of examined genes, e.g.~the whole genome
+background.  ORA is the first and the most basic method of pathway
+analysis. It uses a DEG list with user defined cut-off for the log-fold
+change and \textit{p}-value (most commonly using absolute log-fold change
+$\geq$ 2 and \textit{p}-value $\leq$ 0.05). Subsequently, sets of genes
+associated with annotated pathways are tested for being over-represented in
+the set of DEGs. To this end, hyper-geometric distribution, chi-square
+tests, binomial probability or the Fisher’s exact test are used. Thereby
+the information of the topology of genes in the pathways is
+neglected~\cite{Bayer}. Furthermore, ORA assumes that the biological
 pathways are independent of each other and ignores the fact that they
@@
 
-Unlike ORA, FCS has no artificial cut-off to define DEG list. FCS
-works in three step, first it calculates the gene-level statistics
-including correlation of molecular measurements using differential
-expression of individual genes, ANOVA, t-test and Z-score. In the
-second step the statistics of individual genes in a pathway are
-transformed to an individual pathway-level statistic commonly using
-Kolmogorov-Smirnov statistic, mean or median. Finally the statistical
-significance of the pathway-level statistics is assessed. Although FCS
-covers some of the limitations from ORA, it still lacks the topology
-of genes in a pathway, cross-talk and overlap of the
-pathways~\cite{Khatri2012,Campos}. Pathway topology based methods are
-similar to FCS except that they consider the topology of each gene
-during the gene-level statistics but still don't aim to link different
-functional pathways~\cite{Khatri2012}.
+Unlike ORA, FCS has no artificial cut-off to define DEG list. FCS works in
+three step, first it calculates the gene-level statistics including
+correlation of molecular measurements using differential expression of
+individual genes, ANOVA, t-test and Z-score. In the second step the
+statistics of individual genes in a pathway are transformed to an
+individual pathway-level statistic commonly using Kolmogorov-Smirnov
+statistic, mean or median. Finally the statistical significance of the
+pathway-level statistics is assessed. Although FCS covers some of the
+limitations from ORA, it still lacks the topology of genes in a pathway,
+cross-talk and overlap of the pathways~\cite{Khatri2012,Campos}. Pathway
+topology based methods are similar to FCS except that they consider the
+topology of each gene during the gene-level statistics but still don't aim
+to link different functional pathways~\cite{Khatri2012}.
 
-On these grounds we propose a novel approach to make use of the rich
-gene and protein interaction annotation resources available to gain
-additional functional insights from basic DGEA. To this, we start with
-the presupposition that expression of neighboring genes within a
-functional pathway are not independent from each other. Rather, they
-are often regulating each others expression or are part of the same
-regulon~\cite{Michalak}. We understand that the categorization of
-links between genes into labeled pathways is often an arbitrary one,
-given the extensive cross talk between different pathways. Although
-these categories have proven to be useful in many situations, they
-force a certain perspective onto the interpretation of novel
-data. Based on these two principles, we aim to find sub-graphs of
-connected genes within cell signaling network which exhibit as a whole
-significant differential expression changes. This approach differs in
-two main aspects from common pathway analysis tools. First, it does
-not aim to identify functional pathways enriched in differentially
-expressed genes, but detects sub-graphs or branches in a network graph
-(potentially spanning more than one functionally grouped pathway)
-which is coherently modulated. Second, it aims to improve the DGEA on
-the gene level, by collecting the information of neighboring genes,
-which as a whole might exhibit prominent enough signal to be called,
-again as a whole, significantly modulated.
+On these grounds we propose a novel approach to make use of the rich gene
+and protein interaction annotation resources available to gain additional
+functional insights from basic DGEA. To this, we start with the
+presupposition that expression of neighboring genes within a functional
+pathway are not independent from each other. Rather, they are often
+regulating each others expression or are part of the same
+regulon~\cite{Michalak}. We understand that the categorization of links
+between genes into labeled pathways is often an arbitrary one, given the
+extensive cross talk between different pathways. Although these categories
+have proven to be useful in many situations, they force a certain
+perspective onto the interpretation of novel data. Based on these two
+principles, we aim to find sub-graphs of connected genes within cell
+signaling network which exhibit as a whole significant differential
+expression changes. This approach differs in two main aspects from common
+pathway analysis tools. First, it does not aim to identify functional
+pathways enriched in differentially expressed genes, but detects sub-graphs
+or branches in a network graph (potentially spanning more than one
+functionally grouped pathway) which is coherently modulated. Second, it
+aims to improve the DGEA on the gene level, by collecting the information
+of neighboring genes, which as a whole might exhibit prominent enough
+signal to be called, again as a whole, significantly modulated.
 
 As input, information on functional links between genes provided by
-e.g.~KEGG or Reactome and information on the differential expression
-status of single genes resulting from a DGEA, are required. As a
-result the analysis returns sub-graphs and their joint confidence
-scores, reflecting how the perturbation is migrated through the
-network. Furthermore, the entry points of perturbation in the networks
-and overlap with conventional pathway categories are returned. The
-output is prepared in a directed adjacency file, convenient for
-visualization with the StringApp~\cite{StringApp}, available as a
-Cytoscape plug-in~\cite{Cyto}.
+e.g.~KEGG or Reactome and information on the differential expression status
+of single genes resulting from a DGEA, are required. As a result the
+analysis returns sub-graphs and their joint confidence scores, reflecting
+how the perturbation is migrated through the network. Furthermore, the
+entry points of perturbation in the networks and overlap with conventional
+pathway categories are returned. The output is prepared in a directed
+adjacency file, convenient for visualization, e.g., with
+StringApp~\cite{StringApp}, available as a Cytoscape plug-in~\cite{Cyto}.
 
-All of this can be helpful to understand the cause and effect of a
-stimulus and might inform about potential points of intervention. The
-proposed algorithm was implemented as a java program, which was named
+All of this can be helpful to understand the cause and effect of a stimulus
+and might inform about potential points of intervention. The proposed
+algorithm was implemented as a java program, which was named
 \texttt{Modulated Sub-graph Finder} (abbreviated
-\texttt{MSF}). \texttt{MSF} can help transform the information
-obtained from DGEA into comprehensible knowledge of signal
-transduction of genes and thereby being a valuable complement to
-existing pathway based methods. \texttt{MSF} is freely accessible on
-github under the terms of the Creative Commons Attribution 4.0
-International License.
+\texttt{MSF}). \texttt{MSF} can help transform the information obtained
+from DGEA into comprehensible knowledge of signal transduction of genes and
+thereby being a valuable complement to existing pathway based
+methods. \texttt{MSF} is freely accessible on github under the terms of the
+Creative Commons Attribution 4.0 International License.
 
 \section*{Methods}
-\texttt{MSF} is developed as a novel heuristic approach to find
-concertedly modulated sub-graphs in networks of biological
-interactions.  \texttt{MSF} does not use predefined gene sets grouped
-into functional units, but rather relies purely on the network of
-interacting genes.  The inputted network consists of nodes
-corresponding to genes and edges representing
+\texttt{MSF} is developed as a novel heuristic approach to find concertedly
+modulated sub-graphs in networks of biological interactions.  \texttt{MSF}
+does not use predefined gene sets grouped into functional units, but rather
+relies purely on the network of interacting genes.  The inputted network
+consists of nodes corresponding to genes and edges representing
 interactions. Furthermore it utilizes comprehensive results from a
@@
 
-\texttt{MSF} uses the individual gene's \textit{p}-values generated
-from the DGEA. The \textit{p}-value expresses the probability that the
-null hypothesis of unmodified gene expression can be rejected for a
-given statistical model. To find significantly modulated sub-graphs
-individual \textit{p}-values of the vicinal genes in the global
-network are combined into a single combined \textit{p}-value, using a
-statistical method for combining dependent \textit{p}-values described
-by Hartung~\cite{Hartung}. Hartung uses the inverse of standard normal
-distribution function, individual gene \textit{p}-values are first
-transformed to their corresponding normal score. Then using these
-normal scores, the correlation between genes is calculated, the normal
-scores and correlation are applied to the inverse normal function to
-calculate the combined \textit{p}-value for all genes examined, namely
-the examined sub-graph. The combined \textit{p}-value of a sub-graph
-will express the significance of all genes in the sub-graph being
-modulated together. Thereby, the information from the different genes
-are used as, although not independent, replicated measurement of the
-behavior of the whole sub-graph. This potentially increases the power
-to detect also significant sub-graphs consisting of genes which are
-not significant on there own.
+\texttt{MSF} uses the individual gene's \textit{p}-values generated from
+the DGEA. The \textit{p}-value expresses the probability that the null
+hypothesis of unmodified gene expression can't be rejected for a given
+statistical model. To find significantly modulated sub-graphs individual
+\textit{p}-values of the vicinal genes in the global network are combined
+into a single combined \textit{p}-value, using a statistical method for
+combining dependent \textit{p}-values described by
+Hartung~\cite{Hartung}. Hartung's method uses the inverse of standard
+normal distribution function, individual gene \textit{p}-values are first
+transformed to their corresponding normal score. Then using these normal
+scores, the correlation between genes is calculated, the normal scores and
+correlation are applied to the inverse normal function to calculate the
+combined \textit{p}-value for all genes examined, namely the examined
+sub-graph. The combined \textit{p}-value of a sub-graph will express the
+significance of all genes in the sub-graph being modulated
+together. Thereby, the information from the different genes are used as,
+although not independent, replicated measurement of the behavior of the
+whole sub-graph. This potentially increases the power to detect also
+significant sub-graphs consisting of genes which are not significant on
+there own.
 \newline
@@
 \texttt{MSF} applies a four steps heuristic as described in the
-following. The proceeding identification of modulated sub-graphs from
-a network by \texttt{MSF} is presented as a flowchart diagram
+following. The proceeding identification of modulated sub-graphs from a
+network by \texttt{MSF} is presented as a flowchart diagram
 (Fig.~\ref{fig:pseudocodemsf}).
@@
 \texttt{MSF} constructs the first sub-graph starting with the genes
-associated with the lowest (most significant) \textit{p}-value deduced
-from the DGEA. From this seed the sub-graph is attempted to be
-extended with its neighboring genes, starting with the most
-significant one. A single combined \textit{p}-value is calculated for
-the two genes. If the combined \textit{p}-value is smaller than the
-minimal individual gene's \textit{p}-value, the extended sub-graph is
-accepted. This step is iteratively repeated until no further extension
-is accepted. In this case the process starts over with all remaining
-genes not yet in a significantly modulated sub-graph. This step
-identifies all the trivial sub-graphs that are modulated in the whole
-network.\newline
+associated with the lowest (most significant) \textit{p}-value deduced from
+the DGEA. From this seed the sub-graph is attempted to be extended with its
+neighboring genes, starting with the most significant one. A single
+combined \textit{p}-value is calculated for the two genes. If the combined
+\textit{p}-value is smaller than the minimal individual gene's
+\textit{p}-value, the extended sub-graph is accepted. This step is
+iteratively repeated until no further extension is accepted. In this case
+the process starts over with all remaining genes not yet in a significantly
+modulated sub-graph. This step identifies all the trivial sub-graphs that
+are modulated in the whole network.\newline
 
@@
 
-In the next step initial modulated sub-graphs are used to check if
-they could further be extended beyond the immediate neighborhood. This
-is done by testing all possible extension paths up to N genes for all
-genes in the sub-graph. Again, this step is iteratively repeated until
-no further genes are added to the significant differentially expressed
-sub-graphs. This steps bridges small gaps of genes without a clear
+In the next step initial modulated sub-graphs are used to check if they
+could further be extended beyond the immediate neighborhood. This is done
+by testing all possible extension paths up to \emph{N} genes for all genes
+in the sub-graph. Again, this step is iteratively repeated until no further
+genes are added to the significant differentially expressed
+sub-graphs. This step bridges small gaps of genes without a clear
 differential signal in the DGEA.\newline
@@
 two sub-graphs is done by depth first search. If the two sub-graphs merge
-with the connector of at most N genes (default two genes) and the combined
-\textit{p}-value of the merged sub-graph including the bridging genes in
-between is less than the individual \textit{p}-values of the two
+with the connector of at most \emph{N} genes (default two genes) and the
+combined \textit{p}-value of the merged sub-graph including the bridging
+genes in between is less than the individual \textit{p}-values of the two
 sub-graphs, the two sub-graphs are merged together to one big modulated
@@
 
-In a last post processing step \texttt{MSF} identifies the trigger
-points of the modulated sub-graphs. These trigger genes are the
-sources of the sub-graphs with only outgoing edges. These genes can be
-interpreted as the possible entry points of perturbation from where
-the stimulus causes downstream effects. In the same spirit the most
-downstream genes of the modulated sub-graph are identified and defined
-as sinks. Sinks can be interpreted as the effectors where the
-integrated information within the signal transduction network is set
-to action. Due to circular loops not all sub-graphs are guaranteed to
-have sources or sinks.
+In a last post processing step \texttt{MSF} identifies the trigger points
+of the modulated sub-graphs. These trigger genes are the sources of the
+sub-graphs with only outgoing edges. These genes can be interpreted as the
+possible entry points of perturbation from where the stimulus causes
+downstream effects. In the same spirit the most downstream genes of the
+modulated sub-graph are identified and defined as sinks. Sinks can be
+interpreted as the effectors where the integrated information within the
+signal transduction network is set to action. Due to circular loops not all
+sub-graphs are guaranteed to have sources or sinks.
 
@@
 
-To demonstrate its usefulness, \texttt{MSF} is applied to an RNAseq
-data set of primary human monocyte-derived macrophages (MDMs) infected
-with Ebola virus (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV) belongs
-to the Filoviridea family; filamentous, enveloped and single stranded
-RNA viruses. EBOV causes hemorrhagic fever in humans, inducing the
-host innate and adaptive immune response to be unable to control virus
-infection~\cite{Prins}. Until now, there are no approved antiviral
-drugs for the treatment of Ebola virus infection~\cite{Konde,Rhein}.
-The initial targets of EBOV are the macrophages and dendritic immune
+To demonstrate its usefulness, \texttt{MSF} is applied to an RNA-seq data
+set of primary human monocyte-derived macrophages (MDMs) infected with
+Ebola virus (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV) belongs to the
+Filoviridea family; filamentous, enveloped and single stranded RNA
+viruses. EBOV causes hemorrhagic fever in humans, inducing the host innate
+and adaptive immune response to be unable to control virus
+infection~\cite{Prins}. Until now, there are no approved antiviral drugs
+for the treatment of Ebola virus infection~\cite{Konde,Rhein}.  The initial
+targets of EBOV are the macrophages and dendritic immune
 cells~\cite{Falasca,Rhein}. EBOV inhibits the critical innate immune
 response of the host, which includes the activation of alpha/beta
-interferon (IFN-$\alpha / \beta$)~\cite{Prins,Konde,Cardenas}. It has
-been proposed that IFN-$\alpha / \beta$ should be tested against Ebola
-for its antiviral activity through clinical trials~\cite{Konde}. Ebola
-infection data was selected to test the approach because it has been
-well recognized in the literature for the last several decades. It
-would be easy to cross verify the results of \texttt{MSF} with the
-vast literature present on Ebola infection.  The aim of testing this
-data with \texttt{MSF} was to identify the modulated sub-graphs and
-check if \texttt{MSF} is able to identify the IFN-$\alpha / \beta$
-gene as one of the sources, and show that the capability to close
-``gaps'' by using information of neighboring genes is effective.
+interferon (IFN-$\alpha / \beta$)~\cite{Prins,Konde,Cardenas}. It has been
+proposed that IFN-$\alpha / \beta$ should be tested against Ebola for its
+antiviral activity through clinical trials~\cite{Konde}. Ebola infection
+data was selected to test the approach because it has been well recognized
+in the literature for the last several decades \TODO{what has been
+  recognized??}.  Thereby facilitating the verification of the results of
+\texttt{MSF} with the vast literature present on Ebola
+infection. Especially, the detection of IFN-$\alpha / \beta$ as point of
+action for the virus, could be considered as an basic indicator of the
+rightness and usefulness of the approach.
 
-EBOV infection count data was downloaded from GEO (GSE84188). Differential gene
-expression analysis was performed on the count data with edgeR package
-(version 3.4.2)~\cite{edgeR}. The DEG analysis results generated by
-edgeR were used as input for \texttt{MSF}. Directed cell signaling interactions were filtered from
-Reactome Functional interactions (FIs) Version 2016~\cite{Cytokegg}, which was used as a second input for \texttt{MSF}.
+EBOV infection count data was downloaded from GEO (GSE84188). Differential
+gene expression analysis was performed on the count data with edgeR package
+(version 3.4.2)~\cite{edgeR}. The DEG analysis results generated by edgeR
+were used as input for \texttt{MSF}. Directed cell signaling interactions
+were filtered from Reactome Functional interactions (FIs) Version
+2016~\cite{Cytokegg}, which was used as a second input for \texttt{MSF}.
 
-The EBOV infection
-experiments describe the course of infection at three time-points 6,
-24 and 48 hour post infection (hpi). For the earliest time point at
-6~hpi, five modulated sub-graphs were identified with 41, 107, and 18
-genes. Two of the five sub-graphs were less than 4 genes long. Most of
-the genes part of the sub-graphs were cytokines, chemokines (CXCL10,
-CCL8) and Interleukin genes (IL6, IL27, IL23) which serve as a subset
-of CD4+ T helper cells. IFNB1 and IFNA1 were both identified as two of
-the possible sources. Most of the sources identified by \texttt{MSF}
-were type~I interferon induced genes. At 24~hpi nine modulated
-sub-graphs were identified with four main sub-graphs consisting of 38,
-85, 148 and 167 genes. IFNA1 was identified as one of the sources in
-the most significantly modulated sub-graph. For the last time-point
-48~hpi, eleven modulated sub-graphs were identified. Eight of the
-sub-graphs were less than seven genes and main sub-graphs had 39, 71
-and 210 genes. IFNB1 and IFNA1 were identified as the two sources out
-of the possible sources.
+The EBOV infection experiments describe the course of infection at three
+time-points 6, 24 and 48 hour post infection (hpi). For the earliest time
+point at 6~hpi, five modulated sub-graphs were identified with 41, 107, and
+18 genes. Two of the five sub-graphs were less than 4 genes long. Most of
+the genes part of the sub-graphs were cytokines, chemokines (CXCL10, CCL8)
+and Interleukin genes (IL6, IL27, IL23) which serve as a subset \TODO{??}
+of CD4+ T helper cells. IFNB1 and IFNA1 were both identified as two of the
+possible sources. Most of the sources identified by \texttt{MSF} were
+type~I interferon induced genes. At 24~hpi nine modulated sub-graphs were
+identified with four main \TODO{(more than X genes)} sub-graphs consisting
+of 38, 85, 148 and 167 genes. IFNA1 was identified as one of the sources in
+the most significantly modulated sub-graph. For the last time-point 48~hpi,
+eleven modulated sub-graphs were identified. Eight of the sub-graphs were
+less than seven genes and main sub-graphs had 39, 71 and 210 genes. IFNB1
+and IFNA1 were identified as the two sources out of the possible sources
+\TODO{??}.
 
-As stated earlier IFN-$\alpha / \beta$ could be one of the target
-genes of Ebola infection. We were able to successfully identify IFNA1
-as a source in all three time-points and INFB1 in two of the
-time-points. Although IFNA1 and IFNB1 were two of the most significant
-gene in the later time points, \texttt{MSF} was able to detect them as
-a source in the early time-point when the genes were not significant
-based on the individual DGEA alone. Identifying the possible sources
-will reduce the search space for potential target genes and can help
-the biologist as the starting point of clinical testing for drugs and
-vaccines against an infection.
+As stated earlier IFN-$\alpha / \beta$ was reported to be one of the target
+genes of Ebola infection. We were able to successfully identify IFNA1 as a
+source in all three time-points and INFB1 in two of the
+time-points. Although IFNA1 and IFNB1 were already among the most
+significant gene in the DGEA during the later time points, \texttt{MSF} was
+also able to detect them as a source in the very early time-point when the
+genes were not significant based on the individual DGEA alone. Identifying
+the possible sources will reduce the search space for potential target
+genes and can help the biologist as the starting point of clinical testing
+for drugs and vaccines against an infection.
 
-Table~\ref{tab:rawVsHartung} shows if \texttt{MSF} does not
-consider the complete DEG results and uses different cut offs for
-significant genes from DEG results, how the signal transduction would
-fall apart into incomprehensible small sub-graphs and single unconnected
-genes.
+Table~\ref{tab:rawVsHartung} compares the results of \texttt{MSF}, namely
+the number of detected sub-modules and their total genes numbers, to a
+simple analysis of mapping significantly modulated genes from the DGEA to
+the network and joining neighbors to modules. The numbers indicate that
+\texttt{MSF} detects less but larger sub-modules, applying its statistical
+test.  Furthermore, the dependency of the results from the p-value cutoff
+choice is demonstrated for the DGEA, which is avoided for \texttt{MSF}
+altogether.
 
@@
 
- Three main modulated sub-graphs identified by \texttt{MSF} at 6~hpi
- are shown in Fig.~\ref{fig:Sub-graph6hpi}. The gene based graphs on
- the right hand side, represent the immediate output of the
- \texttt{MSF}-analysis, visualized by StringApp~\cite{StringApp} in
- Cytoscape~\cite{Cyto}. Each node represents a gene part of a
- modulated sub-graph, whereby the associated colors code the
- functional annotation deduced from KEGG Pathways. The cross-talk
- between the pathways and also the multiple employment of many genes
- is evident. The more schematic drawing on the right side represents
- the effortlessly deduced flow of information between the sensors and
- effectors in this particular example.
+ Three main modulated sub-graphs identified by \texttt{MSF} at 6~hpi are
+ shown in Fig.~\ref{fig:Sub-graph6hpi}. The gene based graphs on the right
+ hand side, represent the immediate output of the \texttt{MSF}-analysis,
+ visualized by StringApp~\cite{StringApp} in Cytoscape~\cite{Cyto}. Each
+ node represents a gene part of a modulated sub-graph, whereby the
+ associated colors code the functional annotation deduced from KEGG
+ Pathways. The cross-talk between the pathways and also the multiple
+ employment of many genes is evident. The more schematic drawing on the
+ right side represents the effortlessly deduced flow of information between
+ the sensors and effectors in this particular example.
 
- In more detail, sub-graph~1 (top) shows how the activation of
- toll-like receptor, cytokine, chemokine and jak-stat genes lead via
- TNF into apoptosis. The next significant sub-graph (sub-graph~2:
- middle) reveals how information from the Extra-cellular matrix (ECM)
- receptor, which are reported to interact with Ebola glycoprotein
- (GP)~\cite{Veljkovic}, chemokines and cytokines, and cytosolic DNA
- sensing, is integrated into again modulation of apoptosis
- pathway. Eventually, sub-graph~3 (bottom) demonstrate how INFA1 and
- INFB1 modulate once more, via only a few intermediate steps, the
- apoptotic response of the cell.
+ In more detail, sub-graph~1 (top) shows how the activation of toll-like
+ receptor, cytokine, chemokine and jak-stat genes lead via TNF into
+ apoptosis. The next significant sub-graph (sub-graph~2: middle) reveals
+ how information from the Extra-cellular matrix (ECM) receptor, which are
+ reported to interact with Ebola glycoprotein (GP)~\cite{Veljkovic},
+ chemokines and cytokines, and cytosolic DNA sensing, is integrated into
+ again modulation of apoptosis pathway. Eventually, sub-graph~3 (bottom)
+ demonstrate how INFA1 and INFB1 modulate once more, via only a few
+ intermediate steps, the apoptotic response of the cell.
 
- This show cast example might advertise with how little effort complex
- data cat be interpreted, help to apprehend the dynamics of the
- underlying processes and suggest testable hypothesis and potential
- points of intervention.
+ This show cast example might advertise with how little effort complex data
+ can be interpreted, help to apprehend the dynamics of the underlying
+ processes and suggest testable hypothesis and potential points of
+ intervention.
 
@@
 
- To assess the robustness and stability of our method, Poisson
- distributed noise was added to the read counts of the three
- time-points data set, used in the previous analysis. Then DGEA was
- carried out on the disturbed data with the same parameters as for the
- native data using edgeR, followed by analysis with \texttt{MSF}. This
- procedure was carried out 100 times.  Every time the genes from the
- modulated sub-graphs identified from noisy data were compared to the
- genes of sub-graphs identified from the native data. This was also
- done for the DEG obtained for each run, a cutoff of FDR less than
- 0.05 was used to compare the retrieved DEG from noisy data to the
- native data.  The robustness of \texttt{MSF} and the DEG analysis for
- the time-point 6, 24, and 48~hpi are shown in
- Fig.~\ref{fig:DEGvsMSF}. The procedure how data noise was modeled can
- be considered as rather stringent, which is already reflected limited
- recall rate in the edgeR based DGEA, between 84~\% (6~hpi) and 93~\% (24~hpi). For
- \texttt{MSF}-analysis the observed median recall rates lay between
- 71~\% (6~hpi) and 84~\% (48~hpi).
+ To assess the robustness and stability of our method, Poisson distributed
+ noise was added to the read counts of the three time-points data set, used
+ in the previous analysis. Then DGEA was carried out on the disturbed data
+ with the same parameters as for the native data using edgeR, followed by
+ analysis with \texttt{MSF}. This procedure was carried out 100 times.
+ Every time the genes from the modulated sub-graphs identified from noisy
+ data were compared to the genes of sub-graphs identified from the native
+ data. This was also done for the DEG obtained for each run, a cutoff of
+ FDR less than 0.05 was used to compare the retrieved DEG from noisy data
+ to the native data.  The robustness of \texttt{MSF} and the DEG analysis
+ for the time-point 6, 24, and 48~hpi are shown in
+ Fig.~\ref{fig:DEGvsMSF}. The procedure how data noise was modeled can be
+ considered as rather stringent, which is already reflected by the limited
+ recall rate in the edgeR based DGEA, between 84~\% (6~hpi) and 93~\%
+ (24~hpi). For \texttt{MSF}-analysis the observed median recall rates lay
+ between 71~\% (6~hpi) and 84~\% (48~hpi). The better performance of the
+ pure DGEA can be explained by the fact that disturbed p-values do not
+ change the results for DGEA as long as the p-value does not wise above the
+ chosen cutoff value. In contrast, \texttt{MSF} is sensitive to p-value
+ changes across the whole range of possible values.
 
@@
 
-  Gene enrichment analysis was performed on \texttt{MSF} identified sub-graphs for each time-point using Reactome Analyze data tool~\cite{Reactome}.
-Reactome's over-representation analysis tests whether certain Reactome pathways are enriched by the list of genes submitted.
-For comparison the DEG results from edgeR were filtered using a cut off of adjusted p-value < 0.1.
-This subset of DEG list was used for gene enrichment analysis.
-Enriched pathways were further filtered by confidence score p-values, pathways having p-value less $<$ 0.05 for cut off DEG list and $>$ 1$^{-10}$ for  pathway enrichment in \texttt{MSF} identified sub-graph genes.
-Out of 282 pathways dis-regulated in the two comparison analysis with three time-points each, pathways that were present in at least three time-points are shown in Fig~\ref{fig:msfvsreactome}.
+Gene enrichment analysis was performed on genes in \texttt{MSF} identified
+sub-graphs for each time-point using Reactome Analyze data
+tool~\cite{Reactome}. Reactome's over-representation analysis tests whether
+certain Reactome pathways are enriched by the list of genes submitted. For
+comparison the DEG results from edgeR were filtered using a cut off of
+adjusted p-value < 0.1. This subset of DEG list was used for gene
+enrichment analysis. Enriched pathways were further filtered by confidence
+score p-values, pathways having p-value less $<$ 0.05 for cut off DEG list
+and $>$ 1$^{-10}$ for pathway enrichment in \texttt{MSF} identified
+sub-graph genes. Out of 282 pathways dis-regulated in the two comparison
+analysis with three time-points each, pathways that were present in at
+least three time-points are shown in Fig~\ref{fig:msfvsreactome} \TODO{rephrase; not really clear how this filtering of pathways shown was done}.
 
-  The comparison shows most of the important pathways known from literature to be dis-regulated by Ebola infection are enriched in both the enrichment analysis.
-Toll-Like receptor signaling pathway when interacts with EBOV glycoprotein (GP), triggers the activation of cytokines~\cite{Olejnik}.
-This pathway was not identified when a p-value cut off DEG list was analyzed for enrichment.
-Due to \texttt{MSF} considering the complete DEG results, even the weak signal at the earliest time-point was detected; for example Toll-like receptor signaling.
+The comparison shows most of the pathways known from literature
+to be dis-regulated by Ebola infection are enriched in both the enrichment
+analysis. Toll-Like receptor signaling pathway when interacts with EBOV
+glycoprotein (GP), triggers the activation of
+cytokines~\cite{Olejnik}. This pathway was not identified when a p-value
+cut off DEG list was analyzed for enrichment. Due to \texttt{MSF}
+considering the complete DEG results, even the weak signal at the earliest
+time-point was detected; for example Toll-like receptor signaling. \TODO{what is the point you want to say and state it clearly. this is to vague}
 
@@
 Classic pathway analysis tools aim to detect in lists of significantly
-deregulated genes enriched associations with pathway genes categorized
-by their biological function and their interactions. Thereby,
-depending on the tool, the internal pathway topology is considered or
-neglected all together. The here presented tool, \texttt{MSF}, employs
-a different approach, by aiming to detect sub-graphs in whole gene
-regulatory networks which are significantly deregulated in a concerted
-manner. To this end, neighboring genes in the user provided network
-are tested for jointly common regulation. Exploiting that each gene's
-abundance, although not independent from its neighbors, is measured
-repeatedly on its own, sensitivity can be increased by our applied
-p-value meta-analysis, namely Hartung's method. This potentially
-enables to call just not significant modulated genes based on the DGEA
-to be convincingly called to be part of a deregulated gene group.
-Furthermore, it allows to identify connected sub-graphs, representing
-the propagation of gene regulation perturbation in the input
-network. A better understanding of this propagation, especially the
-critical spots such as sensors, effectors, or intermediate bottlenecks
+deregulated genes enriched associations with pathway genes categorized by
+their biological function and their interactions. Thereby, depending on the
+tool, the internal pathway topology is considered or neglected all
+together. The here presented tool, \texttt{MSF}, employs a different
+approach, by aiming to detect sub-graphs in whole gene regulatory networks
+which are significantly deregulated in a concerted manner. To this end,
+neighboring genes in the user provided network are tested for jointly
+common regulation. Exploiting that each gene's abundance, although not
+independent from its neighbors, is measured repeatedly on its own,
+sensitivity can be increased by our applied p-value meta-analysis, namely
+Hartung's method. This potentially enables to call just not significant
+modulated genes based on the DGEA to be convincingly called to be part of a
+deregulated gene group.  Furthermore, it allows to identify connected
+sub-graphs, representing the propagation of gene regulation perturbation in
+the input network. A better understanding of this propagation, especially
+the critical spots such as sensors, effectors, or intermediate bottlenecks
 and hubs, facilitates the projection of potential intervention points,
-e.g. for drug development. Since \texttt{MSF} only uses interaction
-information in gene regulation network, but not the functional
-grouping of the genes into functional pathways, it is especially
-adapted to discover so called cross-talk between such pathways.
+e.g., for drug development. Since \texttt{MSF} only uses interaction
+information in gene regulation network, but not the functional grouping of
+the genes into functional pathways, it is especially adapted to discover so
+called cross-talk between such pathways.
 
@@
 
-\texttt{MSF} is a fast, robust and easy to use tool to find
-concertedly modulated sub-graphs in a given network. Its
-implementation in java enables it use across many operating
-systems. It requires as input the results from a differential gene
-expression analysis in the appropriate file format. So far the raw
-output from edgeR~\cite{edgeR} and DESeq2~\cite{love2014moderated} are supported
-. Furthermore, a gene network has to be provided in the format of a directed adjacency list.
+\texttt{MSF} is a fast, robust and easy to use tool to find concertedly
+modulated sub-graphs in a given network. Its implementation in java enables
+its use across many operating systems. It requires as input the results from
+a differential gene expression analysis in the appropriate file format. So
+far the raw output from edgeR~\cite{edgeR} and
+DESeq2~\cite{love2014moderated} are supported. Furthermore, a gene network
+has to be provided in the format of a directed adjacency list.
 
@@
 
-The Ebola infection RNAseq dataset analyzed during the current study
-are available in the GEO repository (GSE84188)~\cite{Olejnik}. The cell signaling network file used is from Reactome
-Functional interactions (FIs) Version 2016~\cite{Cytokegg}
+The Ebola infection RNA-seq dataset analyzed during the current study are
+available in the GEO repository (GSE84188)~\cite{Olejnik}. The cell
+signaling network file used is from Reactome Functional interactions (FIs)
+Version 2016~\cite{Cytokegg}
 
@@
 
-Source code: https://github.com/MariamFarman/Modulated-SubGraph-Finder \newline
-Software license: Creative Commons Attribution 4.0 International License
+Source code:
+https://github.com/MariamFarman/Modulated-SubGraph-Finder \newline Software
+license: Creative Commons Attribution 4.0 International License
+
 \section*{Author contributions}
 Conceptualization: FA, ILH; Funding Acquisition: ILH; Resources: ILH;
-Software: MRF; Writing – Original Draft Preparation: MRF; Writing –
-Review \& Editing: FA, ILH. All authors read and approved the final
-manuscript.
+Software: MRF; Writing – Original Draft Preparation: MRF; Writing – Review
+\& Editing: FA, ILH. All authors read and approved the final manuscript.
 
@@
 Agency Verfahren) with the project number (I 1988-B22). The grant was
-assigned to ILH. FA was funded by the Austrian Science Fund (FWF)
-project SFB F43.
+assigned to ILH. FA was funded by the Austrian Science Fund (FWF) project
+SFB F43.
 
@@
 			\includegraphics[width=16cm]{DEGvsMSF2.png}
-			\caption{The percentage of DEG analysis genes
+			\caption{\TODO{make plot smaller and consider to add DEG vor 0.1, 0.05 and 0.01 as you did before} The percentage of DEG analysis genes
                           and MSF identified sub-graph genes recall
@@
 	\includegraphics[width=1.0\linewidth]{MSFvsReactome}
-	\caption{The heatmap shows Reactome pathway enrichment for MSF identifed sub-graphs for the three time-points and the Reactome pathway enrichment for subset of DEG with adjusted p-value cut off of < 0.1.
-Green color shows the pathway being enriched and red color shows pathway not found as enriched.}
+	\caption{The heatmap shows Reactome pathway enrichment for genes in \texttt{MSF} identifed sub-graphs for the three time-points and the Reactome pathway enrichment for subset of DEG with adjusted p-value cut off of < 0.1.
+Green color shows the pathway being enriched and red color shows pathway not found as enriched.
+\TODO{this is not really a heatmap (no quantitative value encoded in the color but only qualitative value); you might want to include different p-value cutof for the DGE; in your set up you have to comapre column 1 with column 4; column 2 with column 5, and so on; that is suboptimal; the time points are not chronological ordered.}}
 	\label{fig:msfvsreactome}

### 2018-05-14 MSF_Paper_FA.tex
Figures with different Cut offs

@@
 antiviral activity through clinical trials~\cite{Konde}. Ebola infection
-data was selected to test the approach because it has been well recognized
-in the literature for the last several decades \TODO{what has been
-  recognized??}.  Thereby facilitating the verification of the results of
+data was selected to test the approach because it has been well recognized for the last several decades, and vast literature is available for the pathogenesis of Ebola.
+Thereby facilitating the verification of the results of
 \texttt{MSF} with the vast literature present on Ebola
@@
 the genes part of the sub-graphs were cytokines, chemokines (CXCL10, CCL8)
-and Interleukin genes (IL6, IL27, IL23) which serve as a subset \TODO{??}
-of CD4+ T helper cells. IFNB1 and IFNA1 were both identified as two of the
+and Interleukin genes (IL6, IL27, IL23). IFNB1 and IFNA1 were both identified as two of the
 possible sources. Most of the sources identified by \texttt{MSF} were
 type~I interferon induced genes. At 24~hpi nine modulated sub-graphs were
-identified with four main \TODO{(more than X genes)} sub-graphs consisting
-of 38, 85, 148 and 167 genes. IFNA1 was identified as one of the sources in
+identified with four main sub-graphs consisting
+of 38, 85, 148 and 167 genes, others being smaller than 10 genes. IFNA1 was identified as one of the sources in
 the most significantly modulated sub-graph. For the last time-point 48~hpi,
@@
 less than seven genes and main sub-graphs had 39, 71 and 210 genes. IFNB1
-and IFNA1 were identified as the two sources out of the possible sources
-\TODO{??}.
+and IFNA1 were identified as the two sources out of the total sources.
 
@@
  data were compared to the genes of sub-graphs identified from the native
- data. This was also done for the DEG obtained for each run, a cutoff of
- FDR less than 0.05 was used to compare the retrieved DEG from noisy data
- to the native data.  The robustness of \texttt{MSF} and the DEG analysis
+ data. This was also done for the DEG obtained for each run, using three different cutoffs of
+ FDR 0.01, 0.05 and 0.1. The robustness of \texttt{MSF} and the DEG analysis
  for the time-point 6, 24, and 48~hpi are shown in
@@
  considered as rather stringent, which is already reflected by the limited
- recall rate in the edgeR based DGEA, between 84~\% (6~hpi) and 93~\%
+ recall rate in the edgeR based DGEA, between 68~\% (6~hpi) and 93~\%
  (24~hpi). For \texttt{MSF}-analysis the observed median recall rates lay
@@
 
-Gene enrichment analysis was performed on genes in \texttt{MSF} identified
-sub-graphs for each time-point using Reactome Analyze data
+Gene enrichment analysis was performed using Reactome Analyze data
 tool~\cite{Reactome}. Reactome's over-representation analysis tests whether
-certain Reactome pathways are enriched by the list of genes submitted. For
-comparison the DEG results from edgeR were filtered using a cut off of
-adjusted p-value < 0.1. This subset of DEG list was used for gene
-enrichment analysis. Enriched pathways were further filtered by confidence
-score p-values, pathways having p-value less $<$ 0.05 for cut off DEG list
-and $>$ 1$^{-10}$ for pathway enrichment in \texttt{MSF} identified
-sub-graph genes. Out of 282 pathways dis-regulated in the two comparison
-analysis with three time-points each, pathways that were present in at
-least three time-points are shown in Fig~\ref{fig:msfvsreactome} \TODO{rephrase; not really clear how this filtering of pathways shown was done}.
+certain Reactome pathways are enriched by the list of genes submitted.Genes from \texttt{MSF} identified
+sub-graphs for each time-point were analyzed for gene enrichment analysis.  For
+comparison the DEG results from edgeR were filtered using three different cut off of
+adjusted p-value 0.01, 0.05 and 0.1. This three subsets of DEG list were used for gene
+enrichment analysis. The compression of \texttt{MSF}  and the three subset DEG list is shown in Fig~\ref{fig:msfvsreactome}
 
@@
 analysis. Toll-Like receptor signaling pathway when interacts with EBOV
-glycoprotein (GP), triggers the activation of
-cytokines~\cite{Olejnik}. This pathway was not identified when a p-value
-cut off DEG list was analyzed for enrichment. Due to \texttt{MSF}
-considering the complete DEG results, even the weak signal at the earliest
-time-point was detected; for example Toll-like receptor signaling. \TODO{what is the point you want to say and state it clearly. this is to vague}
+glycoprotein (GP), it triggers the activation of
+cytokines~\cite{Olejnik}. Toll-like receptor pathway is expected to be dis-regulated in the early stage of infection,  this pathway was not identified as significantly dis-regulated when p-value
+cut off DEG lists were analyzed for enrichment.
+Ten toll-like receptor cascades were identified as dis-regulated from gene enrichmant analysis of \texttt{MSF} identified genes, not a single one of these pathway was shown to be dis-regulated in DEG cut off lists.
+Since \texttt{MSF}
+considers the complete DEG results, even the weak signal at the earliest
+time-point was detected; for example Toll-like receptor signaling.
 
@@
 
-The Ebola infection RNA-seq dataset analyzed during the current study are
+Tlhe Ebola infection RNA-seq dataset analyzed during the current study are
 available in the GEO repository (GSE84188)~\cite{Olejnik}. The cell
@@
 	\fbox{
-		\begin{minipage}{17 cm}
-			\includegraphics[width=16cm]{DEGvsMSF2.png}
-			\caption{\TODO{make plot smaller and consider to add DEG vor 0.1, 0.05 and 0.01 as you did before} The percentage of DEG analysis genes
+		\begin{minipage}{12 cm}
+			\includegraphics[width=12cm]{DEGvsMSF.png}
+			\caption{Shows the percentage of DEG analysis (3 different cut-offs) genes
                           and MSF identified sub-graph genes recall
@@
 		\begin{minipage}{10 cm}
-	\includegraphics[width=1.0\linewidth]{MSFvsReactome}
-	\caption{The heatmap shows Reactome pathway enrichment for genes in \texttt{MSF} identifed sub-graphs for the three time-points and the Reactome pathway enrichment for subset of DEG with adjusted p-value cut off of < 0.1.
-Green color shows the pathway being enriched and red color shows pathway not found as enriched.
-\TODO{this is not really a heatmap (no quantitative value encoded in the color but only qualitative value); you might want to include different p-value cutof for the DGE; in your set up you have to comapre column 1 with column 4; column 2 with column 5, and so on; that is suboptimal; the time points are not chronological ordered.}}
+	\includegraphics[width=1.0\linewidth]{MSFvsReactomeUpsetPlot}
+	\caption{The Upset plot shows the number of common pathways between different time-point and cut offs. }
 	\label{fig:msfvsreactome}

### 2018-05-15 MSF_Paper_FA.tex
last editing

@@
 process~\cite{Khatri2012}. Pathway-based methods use predefined pathways or
-networks such as KEGG~\cite{Kegg} and Reactome~\cite{Reactome}, the
+networks such as KEGG~\cite{Kegg} and Reactome~\cite{Reactome2}, the
 expression measurements of the genes obtained from DGEA in combination with
@@
 Encyclopedia of Genes and Genomes)\cite{Kegg} and
-Reactome~\cite{Reactome}. KEGG pathways is a branch of KEGG database that
+Reactome~\cite{Reactome2}. KEGG pathways is a branch of KEGG database that
 hosts a collection of manually drawn pathway maps representing the
@@
 interactions formed into different biological pathways for nineteen
-species~\cite{Reactome}. Both provide predefined pathways which are sets of
+species~\cite{Reactome2}. Both provide predefined pathways which are sets of
 genes and their interactions categorized into functional units. Starting
@@
 (Functional class scoring) and pathway topology based methods. All of which
-aim in finding in a subset of genes, e.g., significantly differentially
+aim in finding in a subset of genes, i.e., significantly differentially
 expressed genes, genes associated with a certain pathway more often than
@@
 the set of DEGs. To this end, hyper-geometric distribution, chi-square
-tests, binomial probability or the Fisher’s exact test are used. Thereby
+tests, binomial probability or the Fisher’s exact test are used~\cite{Bayer}. Thereby
 the information of the topology of genes in the pathways is
-neglected~\cite{Bayer}. Furthermore, ORA assumes that the biological
+neglected. Furthermore, ORA assumes that the biological
 pathways are independent of each other and ignores the fact that they
@@
 statistic, mean or median. Finally the statistical significance of the
-pathway-level statistics is assessed. Although FCS covers some of the
+pathway-level statistics is assessed. A popular implementation of FCS is \texttt{GSEA}~\cite{subramanian2005gene}.
+Although FCS covers some of the
 limitations from ORA, it still lacks the topology of genes in a pathway,
@@
 analysis returns sub-graphs and their joint confidence scores, reflecting
-how the perturbation is migrated through the network. Furthermore, the
+how the perturbation is migrating through the network. Furthermore, the
 entry points of perturbation in the networks and overlap with conventional
@@
 adjacency file, convenient for visualization, e.g., with
-StringApp~\cite{StringApp}, available as a Cytoscape plug-in~\cite{Cyto}.
+StringApp~\cite{StringApp}, which is available as a Cytoscape plug-in~\cite{Cyto}.
 
@@
 downstream effects. In the same spirit the most downstream genes of the
-modulated sub-graph are identified and defined as sinks. Sinks can be
+modulated sub-graph are identified and referred to as sinks. Sinks can be
 interpreted as the effectors where the integrated information within the
@@
 antiviral activity through clinical trials~\cite{Konde}. Ebola infection
-data was selected to test the approach because it has been well recognized for the last several decades, and vast literature is available for the pathogenesis of Ebola.
+data was selected to test our novel method because it has been well recognized for the last several decades, and vast literature is available for the pathogenesis of Ebola.
 Thereby facilitating the verification of the results of
@@
 Table~\ref{tab:rawVsHartung} compares the results of \texttt{MSF}, namely
-the number of detected sub-modules and their total genes numbers, to a
+the number of detected sub-modules and their total gene numbers, to a
 simple analysis of mapping significantly modulated genes from the DGEA to
@@
  pure DGEA can be explained by the fact that disturbed p-values do not
- change the results for DGEA as long as the p-value does not wise above the
+ change the results for DGEA as long as the p-value does not rise above the
  chosen cutoff value. In contrast, \texttt{MSF} is sensitive to p-value
@@
 Gene enrichment analysis was performed using Reactome Analyze data
-tool~\cite{Reactome}. Reactome's over-representation analysis tests whether
-certain Reactome pathways are enriched by the list of genes submitted.Genes from \texttt{MSF} identified
+tool~\cite{Reactome2}. Reactome's over-representation analysis tests whether
+certain Reactome pathways are enriched by the list of genes submitted. Genes from \texttt{MSF} identified
 sub-graphs for each time-point were analyzed for gene enrichment analysis.  For
@@
 	\fbox{
-		\begin{minipage}{10 cm}
+		\begin{minipage}{15 cm}
 	\includegraphics[width=1.0\linewidth]{MSFvsReactomeUpsetPlot}

### 2018-05-22 MSF_Paper_FA.tex
Add files via upload

--- 2018-05-14 MSF_Paper_FA.tex
@@
 \usepackage{hyperref}
+\usepackage{pdfpages}
 \newcommand{\FA}[1]{\begingroup\color{magenta}#1\endgroup}
@@
 
+\noindent
 Source code:
@@
 	\includegraphics[width=1.0\linewidth]{MSFvsReactomeUpsetPlot}
-	\caption{The Upset plot shows the number of common pathways between different time-point and cut offs. }
+	\caption{The Upset plot shows the number of shared pathways between different time-point and cut offs.
+All the different toll-like receptor signaling pathways identified from \texttt{MSF} identified genes are marked.
+}
 	\label{fig:msfvsreactome}

### 2018-07-31 MSF_Paper_FA.tex
changes and comments

@@
 \newcommand{\FA}[1]{\begingroup\color{magenta}#1\endgroup}
+\newcommand{\ILH}[1]{\begingroup\color{blue}#1\endgroup}
 \newcommand{\TODO}[1]{\begingroup\color{red}#1\endgroup}
@@
 developed. These tools typically rely on the classification of genes into
-set of genes, i.e. a pathway, based on the interactions between the genes
+sets of genes, such as pathways, based on the interactions between the genes
 and their function in a common biological process. Regardless of technical
@@
 (Functional class scoring) and pathway topology based methods. All of which
-aim in finding in a subset of genes, e.g., significantly differentially
+aim to find a subset of genes, e.g., significantly differentially
 expressed genes, genes associated with a certain pathway more often than
 expected given the total set of examined genes, e.g.~the whole genome
-background.  ORA is the first and the most basic method of pathway
+background.  ORA \cite{} is the first and the most basic method of pathway
 analysis. It uses a DEG list with user defined cut-off for the log-fold
@@
 
-Unlike ORA, FCS has no artificial cut-off to define DEG list. FCS works in
-three step, first it calculates the gene-level statistics including
+Unlike ORA, FCS has no artificial cut-off to define a DEG list. FCS works in
+three steps. First it calculates the gene-level statistics including
 correlation of molecular measurements using differential expression of
@@
 pathway-level statistics is assessed. Although FCS covers some of the
-limitations from ORA, it still lacks the topology of genes in a pathway,
+limitations from ORA, it still ignores the topology of genes in a pathway,
 cross-talk and overlap of the pathways~\cite{Khatri2012,Campos}. Pathway
@@
 does not use predefined gene sets grouped into functional units, but rather
-relies purely on the network of interacting genes.  The inputted network
+relies purely on the network of interacting genes.  The input network
 consists of nodes corresponding to genes and edges representing
@@
 To reduce the complexity to score all possible connected sub-graphs
-\texttt{MSF} applies a four steps heuristic as described in the
+\texttt{MSF} applies a four step heuristic as described in the
 following. The proceeding identification of modulated sub-graphs from a
@@
 associated with the lowest (most significant) \textit{p}-value deduced from
-the DGEA. From this seed the sub-graph is attempted to be extended with its
+the DGEA. From this seed it tries to extend the sub-graph by adding directly
 neighboring genes, starting with the most significant one. A single
-combined \textit{p}-value is calculated for the two genes. If the combined
-\textit{p}-value is smaller than the minimal individual gene's
-\textit{p}-value, the extended sub-graph is accepted. This step is
+combined \textit{p}-value is calculated for the extended sub-graph.
+If the combined \textit{p}-value is smaller than the \textit{p}-value of
+the original sub-graph,
+the extended sub-graph is accepted. This step is
 iteratively repeated until no further extension is accepted. In this case
 the process starts over with all remaining genes not yet in a significantly
-modulated sub-graph. This step identifies all the trivial sub-graphs that
+modulated sub-graph. This step identifies all simple sub-graphs that
 are modulated in the whole network.\newline
@@
 
-In the next step initial modulated sub-graphs are used to check if they
-could further be extended beyond the immediate neighborhood. This is done
-by testing all possible extension paths up to \emph{N} genes for all genes
-in the sub-graph. Again, this step is iteratively repeated until no further
-genes are added to the significant differentially expressed
+In the next step, we check if any of the initial modulated sub-graphs 
+could further be extended by adding more than one gene at a time. This is done
+by testing all possible extension paths up to \emph{N} (usually 2) genes at
+all nodes in the sub-graph. Again, this step is iteratively repeated until
+no further genes are added to the significant differentially expressed
 sub-graphs. This step bridges small gaps of genes without a clear
@@
 if combined sub-graphs score better than on their own. The merging of the
-two sub-graphs is done by depth first search. If the two sub-graphs merge
+two sub-graphs is done by depth first search \ILH{unclear}. If the two sub-graphs merge
 with the connector of at most \emph{N} genes (default two genes) and the
@@
 action for the virus, could be considered as an basic indicator of the
-rightness and usefulness of the approach.
+correctness and usefulness of the approach.
 
@@
 time-points 6, 24 and 48 hour post infection (hpi). For the earliest time
-point at 6~hpi, five modulated sub-graphs were identified with 41, 107, and
-18 genes. Two of the five sub-graphs were less than 4 genes long. Most of
-the genes part of the sub-graphs were cytokines, chemokines (CXCL10, CCL8)
-and Interleukin genes (IL6, IL27, IL23). IFNB1 and IFNA1 were both identified as two of the
-possible sources. Most of the sources identified by \texttt{MSF} were
-type~I interferon induced genes. At 24~hpi nine modulated sub-graphs were
+point at 6~hpi, three large modulated sub-graphs were identified with 41,
+107, and 18 genes were identified, as well as two with less than 4 genes.
+The modulated sub-graphs consist predominantly of cytokines, chemokines
+(CXCL10, CCL8) and Interleukin genes (IL6, IL27, IL23).
+IFNB1 and IFNA1 were both identified as two of the possible sources
+\ILH{(in all subgraphs?)}.
+Most of the sources identified by \texttt{MSF} were
+type~I interferon induced genes \ILH{(link to some supplementary
+  material with identified sub-graphs)}.
+At 24~hpi nine modulated sub-graphs were
 identified with four main sub-graphs consisting
@@
 less than seven genes and main sub-graphs had 39, 71 and 210 genes. IFNB1
-and IFNA1 were identified as the two sources out of the total sources.
+and IFNA1 were identified as \TODO{the two sources out of the total sources}.
 
@@
 
- To assess the robustness and stability of our method, Poisson distributed
- noise was added to the read counts of the three time-points data set, used
- in the previous analysis. Then DGEA was carried out on the disturbed data
+A potential concern is how noise in the gene expression measurements
+affects our analysis. To assess the robustness and stability of our method,
+we therefore added Poisson distributed noise
+to the read counts of the three time-points data set, used above.
+Then DGEA was carried out on the disturbed data
 with the same parameters as for the native data using edgeR, followed by
@@
 pure DGEA can be explained by the fact that disturbed p-values do not
- change the results for DGEA as long as the p-value does not wise above the
+change the results for DGEA as long as the p-value does not rise above the
 chosen cutoff value. In contrast, \texttt{MSF} is sensitive to p-value
@@
 cut off DEG lists were analyzed for enrichment.
-Ten toll-like receptor cascades were identified as dis-regulated from gene enrichmant analysis of \texttt{MSF} identified genes, not a single one of these pathway was shown to be dis-regulated in DEG cut off lists.
+Ten toll-like receptor cascades were identified as dis-regulated from gene enrichment analysis of \texttt{MSF} identified genes, not a single one of these pathway was shown to be dis-regulated in DEG cut off lists.
 Since \texttt{MSF}
@@
 
-Tlhe Ebola infection RNA-seq dataset analyzed during the current study are
+The Ebola infection RNA-seq data set analyzed during the current study are
 available in the GEO repository (GSE84188)~\cite{Olejnik}. The cell
@@
 
-This work was funded by the FWF (“Fonds zur Forderung der
+This work was funded by the FWF (“Fonds zur F{\"o}rderung der
 wissenschaftlichen Forschung”) within the project Internationalen
@@
 \end{document}
+
+% LocalWords:  MSF Mariam Farman Ivo Perutz microarray DGEA differentially
+% LocalWords:  effectors DEGs KEGG Reactome STAT1 FCS ANOVA Kolmogorov IFN
+% LocalWords:  Smirnov regulon StringApp Cytoscape transduction github FIs
+% LocalWords:  vicinal Hartung Hartung's iteratively monocyte macrophages
+% LocalWords:  MDMs GSE84188 EBOV Filoviridea dendritic pathogenesis edgeR
+% LocalWords:  hpi cytokines chemokines CXCL10 CCL8 IL6 IL27 IL23 IFNB1
+% LocalWords:  IFNA1 subgraphs INFB1 cytokine chemokine jak TNF apoptosis
+% LocalWords:  ECM glycoprotein cytosolic INFA1 apoptotic Reactome's ILH
+% LocalWords:  DESeq2 MRF FWF Fonds zur Forderung der wissenschaftlichen

### 2018-07-31 MSF_Paper_FA.tex
add tag to Ivos comment

@@
 
-Existing pathway-based analysis approaches use different research designs,
-which can be categorized into ORA (Over-representation analysis), FCS
-(Functional class scoring) and pathway topology based methods. All of which
-aim to find a subset of genes, e.g., significantly differentially
-expressed genes, genes associated with a certain pathway more often than
-expected given the total set of examined genes, e.g.~the whole genome
-background.  ORA \cite{} is the first and the most basic method of pathway
-analysis. It uses a DEG list with user defined cut-off for the log-fold
-change and \textit{p}-value (most commonly using absolute log-fold change
-$\geq$ 2 and \textit{p}-value $\leq$ 0.05). Subsequently, sets of genes
-associated with annotated pathways are tested for being over-represented in
-the set of DEGs. To this end, hyper-geometric distribution, chi-square
-tests, binomial probability or the Fisher’s exact test are used. Thereby
-the information of the topology of genes in the pathways is
-neglected~\cite{Bayer}. Furthermore, ORA assumes that the biological
-pathways are independent of each other and ignores the fact that they
-cross-talk and overlap~\cite{Khatri2012,Campos}.
+Existing pathway-based analysis approaches use different research
+designs, which can be categorized into ORA (Over-representation
+analysis), FCS (Functional class scoring) and pathway topology based
+methods. All of which aim to find a subset of genes, e.g.,
+significantly differentially expressed genes, genes associated with a
+certain pathway more often than expected given the total set of
+examined genes, e.g.~the whole genome background.  \ILH{ORA \cite{} is
+  the first and the most basic method of pathway analysis.} It uses a
+DEG list with user defined cut-off for the log-fold change and
+\textit{p}-value (most commonly using absolute log-fold change $\geq$
+2 and \textit{p}-value $\leq$ 0.05). Subsequently, sets of genes
+associated with annotated pathways are tested for being
+over-represented in the set of DEGs. To this end, hyper-geometric
+distribution, chi-square tests, binomial probability or the Fisher’s
+exact test are used. Thereby the information of the topology of genes
+in the pathways is neglected~\cite{Bayer}. Furthermore, ORA assumes
+that the biological pathways are independent of each other and ignores
+the fact that they cross-talk and overlap~\cite{Khatri2012,Campos}.
 

### 2018-08-09 MSF_Paper_FA.tex
Supplement ref

@@
 certain pathway more often than expected given the total set of
-examined genes, e.g.~the whole genome background.  \ILH{ORA \cite{} is
-  the first and the most basic method of pathway analysis.} It uses a
+examined genes, e.g.~the whole genome background. ORA is
+  the first and the most basic method of pathway analysis~\cite{Campos}. It uses a
 DEG list with user defined cut-off for the log-fold change and
@@
 could further be extended by adding more than one gene at a time. This is done
-by testing all possible extension paths up to \emph{N} (usually 2) genes at
+by testing all possible extension paths up to \emph{N} (default 2) genes at
 all nodes in the sub-graph. Again, this step is iteratively repeated until
@@
 After detection and extension of the modulated sub-graphs, they are tested
-if combined sub-graphs score better than on their own. The merging of the
-two sub-graphs is done by depth first search \ILH{unclear}. If the two sub-graphs merge
-with the connector of at most \emph{N} genes (default two genes) and the
+if combined sub-graphs score better than on their own.
+The merging of the sub-graphs is done by depth first search traversal from the first sub-graph to the second sub-graph.
+If the two sub-graphs merge
+with the connector of at most \emph{N} genes (default 1 gene) and the
 combined \textit{p}-value of the merged sub-graph including the bridging
@@
 IFNB1 and IFNA1 were both identified as two of the possible sources
-\ILH{(in all subgraphs?)}.
+in one of the sub-graph identified with 18 genes.
 Most of the sources identified by \texttt{MSF} were
-type~I interferon induced genes \ILH{(link to some supplementary
-  material with identified sub-graphs)}.
-At 24~hpi nine modulated sub-graphs were
+type~I interferon induced genes (Supplement material 6H).
+At 24~hpi eight modulated sub-graphs were
 identified with four main sub-graphs consisting
-of 38, 85, 148 and 167 genes, others being smaller than 10 genes. IFNA1 was identified as one of the sources in
-the most significantly modulated sub-graph. For the last time-point 48~hpi,
-eleven modulated sub-graphs were identified. Eight of the sub-graphs were
-less than seven genes and main sub-graphs had 39, 71 and 210 genes. IFNB1
-and IFNA1 were identified as \TODO{the two sources out of the total sources}.
+of 27, 101, 134 and 194 genes, others being smaller than 10 genes. IFNB1
+and IFNA1 were identified as the two sources out of the total sources. For the last time-point 48~hpi,
+six modulated sub-graphs were identified. Three of the sub-graphs were
+less than nine genes and main sub-graphs had 122, 191 and 202 genes. IFNB1 was identified as one of the sources in
+the most significantly modulated sub-graph (Supplment).
 
@@
 \noindent
-Source code:
-https://github.com/MariamFarman/Modulated-SubGraph-Finder \newline Software
+Source code: \newline
+https://github.com/Modulated-Subgraph-Finder/MSF \newline Software
 license: Creative Commons Attribution 4.0 International License
+
+\subsection*{Supplement Material}
+
+\noindent
+https://github.com/Modulated-Subgraph-Finder/MSF
 
@@
 		\multicolumn{1}{|l|}{\textbf{24hpi}}              & \multicolumn{1}{c|}{}                                                & \multicolumn{1}{c|}{}                 \\ \hline
-		\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{438}                                             & \multicolumn{1}{c|}{4}                \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{456}                                             & \multicolumn{1}{c|}{4}                \\ \hline
 		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{325}                                             & \multicolumn{1}{c|}{76}               \\ \hline
@@
 		\multicolumn{1}{|l|}{\textbf{48hpi}}              & \multicolumn{1}{c|}{}                                                & \multicolumn{1}{c|}{}                 \\ \hline
-		\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{320}                                             & \multicolumn{1}{c|}{3}                \\ \hline
-		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{226}                                             & \multicolumn{1}{c|}{39}               \\ \hline
-		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{207}                                             & \multicolumn{1}{c|}{62}               \\ \hline
-		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{173}                                             & \multicolumn{1}{c|}{67}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{515}                                             & \multicolumn{1}{c|}{3}                \\ \hline
+		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{392}                                             & \multicolumn{1}{c|}{43}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{366}                                             & \multicolumn{1}{c|}{74}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{306}                                             & \multicolumn{1}{c|}{106}               \\ \hline
 	\end{tabular}

### 2018-08-09 MSF_Paper_FA.tex
Update MSF_Paper_FA.tex

@@
   Sub-graph Finder} (\texttt{MSF}) is freely available at
-\url{https://github.com/MariamFarman/Modulated-SubGraph-Finder}.
+\url{https://github.com/Modulated-Subgraph-Finder/MSF}.
 
@@
 less than nine genes and main sub-graphs had 122, 191 and 202 genes. IFNB1 was identified as one of the sources in
-the most significantly modulated sub-graph (Supplment).
+the most significantly modulated sub-graph (Supplement).
 

### 2019-02-19 MSF_Paper_FA.tex
Updates

@@
 different pathways and rely on binary separation into differentially
-expressed gene and unaffected genes based on an arbitrarily set p-value
+expressed gene and unaffected genes based on an arbitrarily set \textit{p}-value
 cut-off.
@@
 of genes are integrated according to the topology of their interactions and
-allow potentially to read the flow of information from the perturbation
-source to the effectors. The described software, named \texttt{Modulated
+allow potentially to read the flow of information and identify the effectors. The described software, named \texttt{Modulated
   Sub-graph Finder} (\texttt{MSF}) is freely available at
@@
 High throughput sequencing techniques have been widely used to yield
-differentially expressed genes (DEG)~\cite{DEG}. To this end, changes in
+differentially expressed genes (DEG)~\cite{DEG}. The changes in
 transcript abundance are measured, e.g.~by next generation sequencing
-techniques, and interpreted as an indicator of differential expression of
-genes. DEGs can be used to get insights into the mechanism underlying
-differences between conditions of samples, such as healthy versus
-diseased. Differential gene expression analysis (DGEA) informs about the
-magnitude of expression changes between the conditions which are often
-expressed as fold change, sign of fold change and the confidence level of
-observing an authentic change, often expressed as \textit{p}-value. These
-DEGs information is further interpreted to extract meaningful biological
+techniques and interpreted as an indicator of differential expression of
+genes.
+DEGs can be used to gain insights into the mechanisms underlying differences between conditions of samples, such as healthy versus infection.
+Differential gene expression analysis (DGEA) informs about the
+magnitude of expression changes, which are often
+expressed as log-fold change. The sign of log-fold change and the confidence level of
+observing an authentic change, often expressed as \textit{p}-value. The information from DEGs is further interpreted to extract meaningful biological
 insights. For example, genes that could be involved in the response to a
-particular stimuli or maybe the cause of a disease. To this end,
-pathway-based analysis has become an important tool to further interpret
+particular stimulus or may be the cause of an infection. To this end, pathway-based analysis has become an important tool to further interpret
 the results of a DGEA and to acquire understandings of the perturbations in
-a biological system. Biological pathways are sets of genes and their
+a biological system. These pathway-based methods use predefined pathways or
+networks which are sets of genes with their
 interactions forming a functional unit. DEGs help to identify pathways or
-networks that may be altered during a change of condition providing
+networks that may be altered during an infection providing
 important information about diseases and its treatment
-process~\cite{Khatri2012}. Pathway-based methods use predefined pathways or
-networks such as KEGG~\cite{Kegg} and Reactome~\cite{Reactome}, the
+process~\cite{Khatri2012}. The
 expression measurements of the genes obtained from DGEA in combination with
-statistical methods and algorithms to identify specifically modulated
+statistical methods and the predefined pathways are used to identify specifically modulated
 pathways and processes~\cite{Campos}.
@@
 Well established resources for pathway annotation are KEGG (Kyoto
-Encyclopedia of Genes and Genomes)\cite{Kegg} and
-Reactome~\cite{Reactome}. KEGG pathways is a branch of KEGG database that
+Encyclopedia of Genes and Genomes)~\cite{Kegg} and Reactome~\cite{Reactome}. KEGG pathways is a branch of KEGG database that
 hosts a collection of manually drawn pathway maps representing the
@@
 peer-reviewed database for signaling and metabolic molecules with their
-interactions formed into different biological pathways for nineteen
-species~\cite{Reactome}. Both provide predefined pathways which are sets of
-genes and their interactions categorized into functional units. Starting
-from a gene interaction network, genes are labeled according their role in
-a specific biological process. In this sense a particular gene can be
-assigned to different pathways. E.g., the human gene STAT1 is associated
-with 24 different pathways in the pathway annotation curated by
-KEGG~\footnote{\url{http://www.genome.jp/dbget-bin/www_bget?hsa:6772}} and
-in 12 different pathways in the Reactome data
-base~\footnote{\url{http://www.reactome.org/content/detail/R-HSA-629622}}.
-Although carefully produced, the assignment of genes to those predefined
-pathway unites can be considered to be subjective to some degree and
-suffers from observational bias~\cite{schnoes2013biases}.
+interactions formed into different biological pathways~\cite{Reactome}. Both provide predefined pathways which are sets of
+genes and their interactions categorized into functional units.
 
@@
 of neighboring genes, which as a whole might exhibit prominent enough
-signal to be called, again as a whole, significantly modulated.
+signal to be called; again as a whole, significantly modulated. All of this can be helpful to understand the cause and effect of a stimulus and might inform about potential points of intervention.
 
@@
 
-All of this can be helpful to understand the cause and effect of a stimulus
-and might inform about potential points of intervention. The proposed
-algorithm was implemented as a java program, which was named
+ The proposed algorithm is named
 \texttt{Modulated Sub-graph Finder} (abbreviated
@@
 \section*{Methods}
-\texttt{MSF} is developed as a novel heuristic approach to find concertedly
+
+\texttt{MSF} was implemented as a java program. It is developed as a novel heuristic approach to find concertedly
 modulated sub-graphs in networks of biological interactions.  \texttt{MSF}
@@
 Hartung~\cite{Hartung}. Hartung's method uses the inverse of standard
-normal distribution function, individual gene \textit{p}-values are first
-transformed to their corresponding normal score. Then using these normal
-scores, the correlation between genes is calculated, the normal scores and
-correlation are applied to the inverse normal function to calculate the
-combined \textit{p}-value for all genes examined, namely the examined
-sub-graph. The combined \textit{p}-value of a sub-graph will express the
+normal distribution function. Using the inverse normal cumulative distribution function $\Phi^{-1}$, individual gene \textit{p}-values $T_{i}$ are 
+transformed to their corresponding normal score $t_{i}=\Phi^{-1}(1-T_{i})$ that is uniformly distributed on (0,1). Then using these normal
+scores, the correlation between genes is calculated $Cov(t_{i},t_{j})=\rho$. The normal scores and
+correlation are applied to the weighted inverse normal function to calculate the
+combined \textit{p}-value $t(\rho)$ for all genes examined, namely the examined
+sub-graph
+\newline
+ \begin{center}
+ 	$t(\rho)=\frac{\sum_{i=1}^{n}\lambda_i t_{i} }{\sqrt{(1-\rho) \sum_{i=1}^{n} \lambda^{2}+\rho(\sum_{i=1}^{n} \lambda_i)^{2}}}$
+ \end{center}
+
+  Lambda $\lambda$ be the weights for each gene. The combined \textit{p}-value  $t(\rho)$ of a sub-graph will express the
 significance of all genes in the sub-graph being modulated
@@
 network by \texttt{MSF} is presented as a flowchart diagram
-(Fig.~\ref{fig:pseudocodemsf}).
+(Fig.~\ref{fig:pseudocodemsf}). \newline
 
-\textbf{Initial Modulated sub-graphs}
+\textbf{Initializing modulated sub-graphs}
 
@@
 the DGEA. From this seed it tries to extend the sub-graph by adding directly
-neighboring genes, starting with the most significant one. A single
+neighboring genes, starting with the next most significant one. A single
 combined \textit{p}-value is calculated for the extended sub-graph.
@@
 iteratively repeated until no further extension is accepted. In this case
-the process starts over with all remaining genes not yet in a significantly
+the process starts over with all remaining genes not yet in the significantly
 modulated sub-graph. This step identifies all simple sub-graphs that
@@
 
-\textbf{Extending Modulated sub-graphs}
+\textbf{Extending modulated sub-graphs}
 
 In the next step, we check if any of the initial modulated sub-graphs 
-could further be extended by adding more than one gene at a time. This is done
-by testing all possible extension paths up to \emph{N} (default 2) genes at
-all nodes in the sub-graph. Again, this step is iteratively repeated until
+could further be extended.
+This is done when the sub-graph could not be extended any more to the direct neighboring gene due to a bad \textit{p}-value, but the neighboring gene to the direct neighboring gene has a good \textit{p}-value.
+This was required to connect together good \textit{p}-value gene but incorporating a single bad \textit{p}-value gene.
+If the combined \textit{p}-value of the initial modulated sub-graph and the extension genes is smaller than the \textit{p}-value of the initial sub-graph the extension is accepted.
+All possible extension paths up to \emph{3} (default 2) genes at
+all nodes in the sub-graph are tested. Again, this step is iteratively repeated until
 no further genes are added to the significant differentially expressed
@@
 
-\textbf{Merging Modulated sub-graphs}
+\textbf{Merging modulated sub-graphs}
 
 After detection and extension of the modulated sub-graphs, they are tested
-if combined sub-graphs score better than on their own.
-The merging of the sub-graphs is done by depth first search traversal from the first sub-graph to the second sub-graph.
+if combining the different sub-graphs score better than on their own.
+The merging of the sub-graphs is done by depth first search traversal from one sub-graph to the other sub-graph.
 If the two sub-graphs merge
-with the connector of at most \emph{N} genes (default 1 gene) and the
+with the connector of at most \emph{3} genes (default 2 gene) and the
 combined \textit{p}-value of the merged sub-graph including the bridging
@@
 
-\textbf{Finding Sources \& Sinks}
+\textbf{Finding sources \& sinks}
 
@@
 possible entry points of perturbation from where the stimulus causes
-downstream effects. In the same spirit the most downstream genes of the
-modulated sub-graph are identified and defined as sinks. Sinks can be
-interpreted as the effectors where the integrated information within the
-signal transduction network is set to action. Due to circular loops not all
+downstream effects.
+Each individual source is given an impact score, depending on the number of down stream genes effected in the corresponding sub-graph, equipping with a way to choose the most significant genes for unknown phenotypes or biological processes.
+In the same spirit the most downstream genes of the
+modulated sub-graph are identified and defined as sinks. Due to loops not all
 sub-graphs are guaranteed to have sources or sinks.
+
+\textbf{MSF input}
+
+As input \texttt{MSF} requires the results from
+a DGEA in the appropriate file format. So
+far the raw output from edgeR~\cite{edgeR} and
+DESeq2~\cite{love2014moderated} are supported. Furthermore, a gene network
+has to be provided in the format of a directed adjacency list.
+
+\textbf{MSF output}
+
+\texttt{MSF} generates a directed network file as an output, with complete directed interactions of all modulated sub-graphs identified.
+This file could be imported into Cytoscape~\cite{Cyto} for visualization.
+Second output file is the source \& sink file.
+It contains for all the modulated sub-graphs, sources and sinks identified with the source impact score.
+The third output file, source weightage is provided as a node attribute file.
+It could be imported into Cytoscape to see the impact score of sources as the size of the node and log-fold change could be used to see how the regulation of different gene hubs in the modulated sub-graphs look.
 
@@
 
-\subsection*{Case Study}
+\subsection*{Case study}
 
-To demonstrate its usefulness, \texttt{MSF} is applied to an RNA-seq data
+To demonstrate its usefulness, \texttt{MSF} is applied to a RNA-seq data
 set of primary human monocyte-derived macrophages (MDMs) infected with
-Ebola virus (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV) belongs to the
-Filoviridea family; filamentous, enveloped and single stranded RNA
+Ebola virus (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV) belongs to the Filoviridea family: filamentous, enveloped and single stranded RNA
 viruses. EBOV causes hemorrhagic fever in humans, inducing the host innate
 and adaptive immune response to be unable to control virus
-infection~\cite{Prins}. Until now, there are no approved antiviral drugs
+infection~\cite{Prins}. Currently, there are no approved antiviral drugs
 for the treatment of Ebola virus infection~\cite{Konde,Rhein}.  The initial
@@
 antiviral activity through clinical trials~\cite{Konde}. Ebola infection
-data was selected to test the approach because it has been well recognized for the last several decades, and vast literature is available for the pathogenesis of Ebola.
-Thereby facilitating the verification of the results of
+data was selected to test the approach because it has been well recognized for the last several decades, and vast literature is available for the pathogenesis of Ebola, thereby, facilitating the verification of the results of
 \texttt{MSF} with the vast literature present on Ebola
@@
 
-EBOV infection count data was downloaded from GEO (GSE84188). Differential
+EBOV infection count data was downloaded from GEO (GSE84188), it describes the course of infection at three
+time-points 6, 24 and 48 hour post infection (hpi). Differential
 gene expression analysis was performed on the count data with edgeR package
-(version 3.4.2)~\cite{edgeR}. The DEG analysis results generated by edgeR
-were used as input for \texttt{MSF}. Directed cell signaling interactions
+(version 3.4.2)~\cite{edgeR} using an upper-quartile normalization. The DEG analysis results generated by edgeR
+were used as input for \texttt{MSF}. Cell signaling interactions
 were filtered from Reactome Functional interactions (FIs) Version
-2016~\cite{Cytokegg}, which was used as a second input for \texttt{MSF}.
+2016~\cite{Cytokegg} for only direct interactions, which was used as a second input for \texttt{MSF}.
 
-The EBOV infection experiments describe the course of infection at three
-time-points 6, 24 and 48 hour post infection (hpi). For the earliest time
-point at 6~hpi, three large modulated sub-graphs were identified with 41,
-107, and 18 genes were identified, as well as two with less than 4 genes.
+For the earliest time point at 6~hpi, three large modulated sub-graphs were identified with 42,
+139, and 69 genes.
 The modulated sub-graphs consist predominantly of cytokines, chemokines
-(CXCL10, CCL8) and Interleukin genes (IL6, IL27, IL23).
+(CXCL10, CCL8, CXCL9, CXCL11, CXCR4, CCR7, CCL4L1, CCL3L1, CCL4, CCL8, CCL20, CCL3, CCL19) and Interleukin genes (IL6, IL27, IL23).
 IFNB1 and IFNA1 were both identified as two of the possible sources
-in one of the sub-graph identified with 18 genes.
+in the most significantly modulated sub-graph identified with 42 genes. The impact score of IFNB1 is 14.5\% and IFNA1 is 8.7\% for the sub-graph they belong to.
 Most of the sources identified by \texttt{MSF} were
 type~I interferon induced genes (Supplement material 6H).
-At 24~hpi eight modulated sub-graphs were
+At 24~hpi seven modulated sub-graphs were
 identified with four main sub-graphs consisting
-of 27, 101, 134 and 194 genes, others being smaller than 10 genes. IFNB1
-and IFNA1 were identified as the two sources out of the total sources. For the last time-point 48~hpi,
+of 61, 222, 130 and 242 genes, others being smaller than 6 genes. Again, IFNB1
+and IFNA1 were identified as two sources out of the total sources with 3.9\% and 1.6\% impact score. For the last time-point 48~hpi,
 six modulated sub-graphs were identified. Three of the sub-graphs were
-less than nine genes and main sub-graphs had 122, 191 and 202 genes. IFNB1 was identified as one of the sources in
-the most significantly modulated sub-graph (Supplement).
+less than ten genes and main sub-graphs had 217, 224 and 276 genes. IFNB1 and IFNA1 were identified as sources in
+the most significantly modulated sub-graph with an impact score of 2.8\% and 3.7\% (Supplement).
 
 As stated earlier IFN-$\alpha / \beta$ was reported to be one of the target
-genes of Ebola infection. We were able to successfully identify IFNA1 as a
-source in all three time-points and INFB1 in two of the
-time-points. Although IFNA1 and IFNB1 were already among the most
-significant gene in the DGEA during the later time points, \texttt{MSF} was
+genes of Ebola infection. We were able to successfully identify IFNA1 and IFNB1 as
+sources in all three Ebola infection time-points. Although IFNA1 and IFNB1 were already among the most
+significant genes in the DGEA during the later time points, \texttt{MSF} was
 also able to detect them as a source in the very early time-point when the
@@
 the network and joining neighbors to modules. The numbers indicate that
-\texttt{MSF} detects less but larger sub-modules, applying its statistical
-test.  Furthermore, the dependency of the results from the p-value cutoff
+\texttt{MSF} detects less but larger and meaningful sub-modules, applying its statistical
+test. Furthermore, the dependency of the results from the \textit{p}-value cutoff
 choice is demonstrated for the DGEA, which is avoided for \texttt{MSF}
 altogether.
+It showcases how applying different cut-offs to the \textit{p}-value of genes from edgeR to the sub-graphs identified by \texttt{MSF} breaks the larger sub-graphs to many smaller unconnected sub-graphs, many of which are single genes.
 
@@
 
-In more detail, sub-graph~1 (top) shows how the activation of toll-like
-receptor, cytokine, chemokine and jak-stat genes lead via TNF into
-apoptosis. The next significant sub-graph (sub-graph~2: middle) reveals
+In more detail, sub-graph~1 (bottom) shows how the activation of toll-like
+receptor, cytokine, chemokine activating jak-stat and mapk gene, together with TNF they lead into apoptosis. The next significant sub-graph (sub-graph~2: top right) reveals
 how information from the Extra-cellular matrix (ECM) receptor, which are
@@
 chemokines and cytokines, and cytosolic DNA sensing, is integrated into
-again modulation of apoptosis pathway. Eventually, sub-graph~3 (bottom)
+again modulation of apoptosis pathway. Eventually, sub-graph~3 (top left)
 demonstrate how INFA1 and INFB1 modulate once more, via only a few
-intermediate steps, the apoptotic response of the cell.
+intermediate steps, the apoptotic response of the cell. cAMP signaling genes activates platelet genes.
 
-This show cast example might advertise with how little effort complex data
+This display case might advertise with how little effort complex data
 can be interpreted, help to apprehend the dynamics of the underlying
@@
 between 71~\% (6~hpi) and 84~\% (48~hpi). The better performance of the
-pure DGEA can be explained by the fact that disturbed p-values do not
-change the results for DGEA as long as the p-value does not rise above the
-chosen cutoff value. In contrast, \texttt{MSF} is sensitive to p-value
+pure DGEA can be explained by the fact that disturbed \textit{p}-values do not
+change the results for DGEA as long as the \textit{p}-value does not rise above the
+chosen cutoff value. In contrast, \texttt{MSF} is sensitive to \textit{p}-value
 changes across the whole range of possible values.
 
-\subsection*{Comparison to Reactome pathway analysis tool}
+\subsection*{Benchmark}
 
-Gene enrichment analysis was performed using Reactome Analyze data
-tool~\cite{Reactome}. Reactome's over-representation analysis tests whether
-certain Reactome pathways are enriched by the list of genes submitted.Genes from \texttt{MSF} identified
-sub-graphs for each time-point were analyzed for gene enrichment analysis.  For
-comparison the DEG results from edgeR were filtered using three different cut off of
-adjusted p-value 0.01, 0.05 and 0.1. This three subsets of DEG list were used for gene
-enrichment analysis. The compression of \texttt{MSF}  and the three subset DEG list is shown in Fig~\ref{fig:msfvsreactome}
+\textbf{jActiveModules}
 
-The comparison shows most of the pathways known from literature
+jActiveModules~\cite{jActiveModules} is a plugin in Cytoscape that searches for molecular interaction network to find expression activated sub-networks.
+The method used to find the expression activated sub-networks is close to the method used in \texttt{MSF}.
+
+The first time-point of Ebola infection data was analyzed using jActiveModule (Version 3.2.1) to compare the modulated sub-graphs identified by \texttt{MSF} and jActiveModule.
+The input files were same for both tools, number of modules parameter was set to 5.
+jActiveModule identified 5 modules, the module with the highest pathScore was selected for comparison with \texttt{MSF} identied modulated sub-graphs.
+The module consisting of a single graph with 314 genes.
+While \texttt{MSF} identified three directed modulated sub-graphs with 42, 69 and 139 genes.
+The overlap of the common genes identified between \texttt{MSF} and jActiveModule is shown in Fig~\ref{fig:MSFvsJActiveModule}.
+It is difficult to state which tool identifies more accurate set of genes as modulated, since there is no golden standard data example to refer.
+Although jActiveModule identifies single connected module, \texttt{MSF} provides directionality, with the identification of possible perturbation sources of the sub-graphs.
+
+\textbf{Reactome pathway analysis}
+
+Gene enrichment analysis was performed using Reactome analyze data
+tool~\cite{Reactome} (version 67). Reactome's over-representation analysis tool tests whether
+certain Reactome pathways are enriched for the list of genes submitted to it. Genes from \texttt{MSF} identified
+sub-graphs for each time-point were analyzed for gene enrichment using this tool.  For
+comparison the DEG results from edgeR for the three time-points were filtered using the cut-off of
+adjusted \textit{p}-value > 0.05. This DEG list was used for gene
+enrichment analysis. The compression of \texttt{MSF} identified sub-graphs gene list and the DEG list analysis is shown in Fig~\ref{fig:msfvsreactome}
+
+All enriched pathways with a cut-off of \textit{p}-value > 0.05 for \texttt{MSF} and DEG list for the three time-points were selected. The comparison shows most of the pathways known from literature
 to be dis-regulated by Ebola infection are enriched in both the enrichment
@@
 glycoprotein (GP), it triggers the activation of
-cytokines~\cite{Olejnik}. Toll-like receptor pathway is expected to be dis-regulated in the early stage of infection,  this pathway was not identified as significantly dis-regulated when p-value
-cut off DEG lists were analyzed for enrichment.
-Ten toll-like receptor cascades were identified as dis-regulated from gene enrichment analysis of \texttt{MSF} identified genes, not a single one of these pathway was shown to be dis-regulated in DEG cut off lists.
+cytokines~\cite{Olejnik}.
+Toll-like receptor pathway is expected to be dis-regulated in the early stage of infection, this pathway was not identified as significantly dis-regulated when \textit{p}-value
+cut off DEG list was analyzed for enrichment.
+Nine toll-like receptor cascades TLR10, TLR2, TLR3, TLR4, TLR5, TLR7/8, TLR9, TLR1:TLR2 and TLR6:TLR2 were identified as dis-regulated from gene enrichment analysis of \texttt{MSF} identified sub-graph genes, not a single one of these cascade was shown to be dis-regulated in pathway enrichment analysis from DEG cut-off list.
 Since \texttt{MSF}
@@
 time-point was detected; for example Toll-like receptor signaling.
+While \texttt{MSF} is able to catch weak signals, it does not provide information about the functional relationships among genes like Reactome tool.
+
+\textbf{Gene set enrichment analysis (GSEA)}
+
+Gene set enrichment analysis (GSEA)~\cite{Subramanian15545} is a method to identify classes of genes or proteins that are over-represented in a large set of genes or proteins.
+GSEA uses statistical approaches to identify significantly enriched or depleted groups of genes.
+The complete DEG list from DGEA of the first time-point was analyzed using bioconductor package GSEABase (version 1.44.0).
+GSEA was able to identify Toll like recpetor, Chemokine signaling pathway, Cytosolic DNA-sensing pathway, Jak-STAT signaling pathway, RIG-I-like receptor signaling pathway and apoptosis as the highest ranked pathways.
+Although GSEA identified the important pathways for Ebola infection, it did not show how the cross-talk and topology of the genes between the pathways identifed.
+Like \texttt{MSF}, GSEA uses complete DEG list without using any cut-off, reasoning why Ebola infection pathways showed up even with weak signal genes in it.
 
@@
 independent from its neighbors, is measured repeatedly on its own,
-sensitivity can be increased by our applied p-value meta-analysis, namely
+sensitivity can be increased by our applied \textit{p}-value meta-analysis, namely
 Hartung's method. This potentially enables to call just not significant
@@
 the input network. A better understanding of this propagation, especially
-the critical spots such as sensors, effectors, or intermediate bottlenecks
-and hubs, facilitates the projection of potential intervention points,
+the critical spots such as sensors, effectors, and hubs, facilitates the projection of potential intervention points,
 e.g., for drug development. Since \texttt{MSF} only uses interaction
@@
 
-\texttt{MSF} is a fast, robust and easy to use tool to find concertedly
+\texttt{MSF} is a fast and easy to use tool to find concertedly
 modulated sub-graphs in a given network. Its implementation in java enables
-its use across many operating systems. It requires as input the results from
-a differential gene expression analysis in the appropriate file format. So
-far the raw output from edgeR~\cite{edgeR} and
-DESeq2~\cite{love2014moderated} are supported. Furthermore, a gene network
-has to be provided in the format of a directed adjacency list.
+its use across many operating systems e.g. linux and windows .
 
@@
 
-\subsection*{Software Availability}
+\subsection*{Software availability}
 
@@
 
-\subsection*{Supplement Material}
+\subsection*{Supplement material}
 
@@
                   heuristical approach to detect modulated sub-graphs
-                  in a global gene regulatory network without
-                  exhaustively testing all connected sub-graphs.}
+                  in a global gene regulatory network.}
 		\label{fig:pseudocodemsf}
@@
 	\fbox{
-		\begin{minipage}{17 cm}
-			\includegraphics[width=1.05\linewidth]{SeperatedMergedNetwork.png}
+		\begin{minipage}{18 cm}
+			\includegraphics[width=1.0\linewidth]{DirectedMerged_Network.png}
 			\caption{Visualisation of the three modulated
-                          sub-graphs identified by MSF at 6~h after
-                          EBOV infection in gene detail and their
-                          epitomized representation to depict the flow
-                          of perturbation in the directed network. The
+                           directed sub-graphs identified by \texttt{MSF} at 6~h after
+                          EBOV infection in gene detail. The
                           node coloring is associated to KEGG pathways
                           referring to the colors in the legend. The
-                          graph edges are from Reactome. Important
-                          genes to EBOV infection as from literature
-                          are enlarged in the
-                          graph.\label{fig:Sub-graph6hpi}}
+                          graph edges are from Reactome. \label{fig:Sub-graph6hpi}}
 		\end{minipage}
@@
 			\caption{Shows the percentage of DEG analysis (3 different cut-offs) genes
-                          and MSF identified sub-graph genes recall
+                          and \texttt{MSF} identified sub-graph genes recall
                           rates for the three different time points of
@@
 
+\begin{figure*}
+		\centering
+	\fbox{
+		\begin{minipage}{12 cm}
+			\includegraphics[width=0.9\linewidth]{MSFvsJActiveModule.png}
+			\caption{The Venn diagram shows the common genes identified as modulated from \texttt{MSF} identified sub-graphs and jActiveModule identified module.}
+			\label{fig:MSFvsJActiveModule}
+		\end{minipage}
+	}
+\end{figure*}
+
 \begin{figure*}[p]
@@
 		\begin{minipage}{10 cm}
-	\includegraphics[width=1.0\linewidth]{MSFvsReactomeUpsetPlot}
-	\caption{The Upset plot shows the number of shared pathways between different time-point and cut offs.
-All the different toll-like receptor signaling pathways identified from \texttt{MSF} identified genes are marked.
-}
+	\includegraphics[width=1.0\linewidth]{ReactomeVsMSF.png}
+	\caption{The Upset plot shows the number of shared pathways between \texttt{MSF} identified sub-graph gene list and DEG cut-off list for the 3 time-points.
+All the different toll-like receptor cascades are shown to be in the bar with 164 shared pathways only between \texttt{MSF} at different time-points.}
 	\label{fig:msfvsreactome}
@@
 	\caption{Comparison of connected sub-graphs of modulated genes
-          in the global network between the analysis results of
-          \texttt{MSF} and mapping the raw list of differentially
-          expressed genes from the standard edgeR analysis, applying
-          different p-value cut-offs, onto the network. }
+          in the global network identified after the analysis with
+          \texttt{MSF} and applying different \textit{p}-value cut-offs to gene from edgeR onto the network. }
 	\label{tab:rawVsHartung}
@@
 		\multicolumn{1}{|l|}{\textbf{6hpi}}               & \multicolumn{1}{l|}{}                                                & \multicolumn{1}{c|}{}                 \\ \hline
-		\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{166}                                             & \multicolumn{1}{c|}{3}                \\ \hline
-		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{123}                                             & \multicolumn{1}{c|}{39}               \\ \hline
-		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{123}                                             & \multicolumn{1}{c|}{47}               \\ \hline
-		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{112}                                             & \multicolumn{1}{c|}{53}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{250}                                             & \multicolumn{1}{c|}{3}                \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{166}                                             & \multicolumn{1}{c|}{87}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{152}                                             & \multicolumn{1}{c|}{89}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{125}                                             & \multicolumn{1}{c|}{76}               \\ \hline
 		\multicolumn{1}{|l|}{\textbf{24hpi}}              & \multicolumn{1}{c|}{}                                                & \multicolumn{1}{c|}{}                 \\ \hline
-		\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{456}                                             & \multicolumn{1}{c|}{4}                \\ \hline
-		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{325}                                             & \multicolumn{1}{c|}{76}               \\ \hline
-		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{305}                                             & \multicolumn{1}{c|}{102}              \\ \hline
-		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{262}                                             & \multicolumn{1}{c|}{110}              \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{656}                                             & \multicolumn{1}{c|}{7}                \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{457}                                             & \multicolumn{1}{c|}{183}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{418}                                             & \multicolumn{1}{c|}{198}              \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{332}                                             & \multicolumn{1}{c|}{216}              \\ \hline
 		\multicolumn{1}{|l|}{\textbf{48hpi}}              & \multicolumn{1}{c|}{}                                                & \multicolumn{1}{c|}{}                 \\ \hline
-		\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{515}                                             & \multicolumn{1}{c|}{3}                \\ \hline
-		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{392}                                             & \multicolumn{1}{c|}{43}               \\ \hline
-		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{366}                                             & \multicolumn{1}{c|}{74}               \\ \hline
-		\multicolumn{1}{|l|}{edgeR (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{306}                                             & \multicolumn{1}{c|}{106}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{744}                                             & \multicolumn{1}{c|}{6}                \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{514}                                             & \multicolumn{1}{c|}{189}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{468}                                             & \multicolumn{1}{c|}{206}               \\ \hline
+		\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{363}                                             & \multicolumn{1}{c|}{241}               \\ \hline
 	\end{tabular}

### 2019-02-20 MSF_Paper_FA.tex
Operation added

@@
 
-\textbf{MSF input}
-
-As input \texttt{MSF} requires the results from
-a DGEA in the appropriate file format. So
-far the raw output from edgeR~\cite{edgeR} and
-DESeq2~\cite{love2014moderated} are supported. Furthermore, a gene network
-has to be provided in the format of a directed adjacency list.
-
 	\textbf{MSF output}
@@
 It could be imported into Cytoscape to see the impact score of sources as the size of the node and log-fold change could be used to see how the regulation of different gene hubs in the modulated sub-graphs look.
+
+	\textbf{Operation}
+	MSF requires Java version 8 and JDK 1.8.
+The few package dependencies are already been added to the release.
+MSF runs fast on a standard laptop computer and so it has normal system requirements.
+To run MSF, the user must provide two text files, one containing the DGEA and the second one containing the interactions in an adjacency format file.
+Example files and a detailed tutorial to use MSF has been provided on github https://github.
+com/Modulated-Subgraph-Finder/MSF.
 
@@
 	modulated sub-graphs in a given network. Its implementation in java enables
-its use across many operating systems e.g. linux and windows .
+	its use across many operating systems e.g. linux and windows. So far the raw output from edgeR~\cite{edgeR} and DESeq2~\cite{love2014moderated} are supported.
 
@@
 	https://github.com/Modulated-Subgraph-Finder/MSF \newline Software
-license: Creative Commons Attribution 4.0 International License
+	Software license: MIT license.
 
 	\subsection*{Supplement material}
-
-\noindent
-https://github.com/Modulated-Subgraph-Finder/MSF
+	Supplementary material is available form GitHub: https://github.com/Modulated-Subgraph-Finder/MSF
 

### 2019-02-26 MSF_Paper_FA.tex
Latest Version of Manuscript

@@
 
-	Well established resources for pathway annotation are KEGG (Kyoto
+	Well established resources for pathway annotations are KEGG (Kyoto
 	Encyclopedia of Genes and Genomes)~\cite{Kegg} and Reactome~\cite{Reactome}. KEGG pathways is a branch of KEGG database that
@@
 	topology based methods are similar to FCS except that they consider the
-	topology of each gene during the gene-level statistics but still don't aim
-	to link different functional pathways~\cite{Khatri2012}.
+	topology of each gene during the gene-level statistics but still lacks to aim
+	to link the different functional pathways~\cite{Khatri2012}.
 
 	On these grounds we propose a novel approach to make use of the rich gene
-	and protein interaction annotation resources available to gain additional
+	and protein interaction annotation resources available, to gain additional
 	functional insights from basic DGEA. To this, we start with the
@@
 	principles, we aim to find sub-graphs of connected genes within cell
-	signaling network which exhibit as a whole significant differential
+	signaling network, which exhibit as a whole significant differential
 	expression changes. This approach differs in two main aspects from common
@@
 
-	Lambda $\lambda$ be the weights for each gene. The combined \textit{p}-value  $t(\rho)$ of a sub-graph will express the
+	Lambda $\lambda$ be the weights for each gene, currently all genes have equal weight i.e. 1. The combined \textit{p}-value  $t(\rho)$ of a sub-graph will express the
 	significance of all genes in the sub-graph being modulated
@@
 	could further be extended.
-This is done when the sub-graph could not be extended any more to the direct neighboring gene due to a bad \textit{p}-value, but the neighboring gene to the direct neighboring gene has a good \textit{p}-value.
-This was required to connect together good \textit{p}-value gene but incorporating a single bad \textit{p}-value gene.
+This is done when the sub-graph could not be extended any more to the direct neighboring gene due to a high \textit{p}-value, but the neighboring gene to the direct neighboring gene has a low \textit{p}-value.
+This was required to connect together low \textit{p}-value gene but incorporating a single high \textit{p}-value gene.
 If the combined \textit{p}-value of the initial modulated sub-graph and the extension genes is smaller than the \textit{p}-value of the initial sub-graph the extension is accepted.
@@
 In the same spirit the most downstream genes of the
-	modulated sub-graph are identified and defined as sinks. Due to loops not all
+	modulated sub-graphs are identified and defined as sinks. Due to loops not all
 	sub-graphs are guaranteed to have sources or sinks.
@@
 
-	\texttt{MSF} generates a directed network file as an output, with complete directed interactions of all modulated sub-graphs identified.
+	\texttt{MSF} generates a directed network file as an output, containing complete directed interactions of all modulated sub-graphs identified.
 This file could be imported into Cytoscape~\cite{Cyto} for visualization.
-Second output file is the source \& sink file.
-It contains for all the modulated sub-graphs, sources and sinks identified with the source impact score.
-The third output file, source weightage is provided as a node attribute file.
+Second output is the source \& sink file.
+It contains for all the modulated sub-graphs, sources and sinks identified with the source impact scores.
+The third output, source weightage is provided as a node attribute file.
 It could be imported into Cytoscape to see the impact score of sources as the size of the node and log-fold change could be used to see how the regulation of different gene hubs in the modulated sub-graphs look.
@@
 	\textbf{Operation}
+
 	MSF requires Java version 8 and JDK 1.8.
@@
 MSF runs fast on a standard laptop computer and so it has normal system requirements.
-To run MSF, the user must provide two text files, one containing the DGEA and the second one containing the interactions in an adjacency format file.
+To run MSF, the user must provide two text files, one containing the DGEA and the second one containing directed interactions in an adjacency format file.
 Example files and a detailed tutorial to use MSF has been provided on github https://github.
@@
 
-	To demonstrate its usefulness, \texttt{MSF} is applied to a RNA-seq data
+	To demonstrate its usefulness, \texttt{MSF} is applied to RNA-seq data
 	set of primary human monocyte-derived macrophages (MDMs) infected with
@@
 	infection. Especially, the detection of IFN-$\alpha / \beta$ as point of
-	action for the virus, could be considered as an basic indicator of the
+	action for the virus, could be considered as a basic indicator of the
 	correctness and usefulness of the approach.
@@
 	Most of the sources identified by \texttt{MSF} were
-	type~I interferon induced genes (Supplement material 6H).
+	type~I interferon induced genes (Supplement material).
 	At 24~hpi seven modulated sub-graphs were
@@
 	less than ten genes and main sub-graphs had 217, 224 and 276 genes. IFNB1 and IFNA1 were identified as sources in
-	the most significantly modulated sub-graph with an impact score of 2.8\% and 3.7\% (Supplement).
+	the most significantly modulated sub-graph with an impact score of 2.8\% and 3.7\% (Supplement material).
 
@@
 	Three main modulated sub-graphs identified by \texttt{MSF} at 6~hpi are
-	shown in Fig.~\ref{fig:Sub-graph6hpi}. The gene based graphs on the right
-	hand side, represent the immediate output of the \texttt{MSF}-analysis,
+	shown in Fig.~\ref{fig:Sub-graph6hpi}. The graphs represent the immediate output of the \texttt{MSF}-analysis,
 	visualized by StringApp~\cite{StringApp} in Cytoscape~\cite{Cyto}. Each
@@
 	Pathways. The cross-talk between the pathways and also the multiple
-	employment of many genes is evident. The more schematic drawing on the
-	right side represents the effortlessly deduced flow of information between
-	the sensors and effectors in this particular example.
-
+	employment of many genes is evident. The flow of information between
+	the sensors and effectors in this particular example could easily be deduced.
 	In more detail, sub-graph~1 (bottom) shows how the activation of toll-like
-	receptor, cytokine, chemokine activating jak-stat and mapk gene, together with TNF they lead into apoptosis. The next significant sub-graph (sub-graph~2: top right) reveals
+	receptor, cytokine, chemokine activating jak-stat and mapk genes, together with TNF leads into apoptosis. The next significant sub-graph (sub-graph~2: top right) reveals
 	how information from the Extra-cellular matrix (ECM) receptor, which are
-	reported to interact with Ebola glycoprotein (GP)~\cite{Veljkovic},
-	chemokines and cytokines, and cytosolic DNA sensing, is integrated into
-	again modulation of apoptosis pathway. Eventually, sub-graph~3 (top left)
-	demonstrate how INFA1 and INFB1 modulate once more, via only a few
-	intermediate steps, the apoptotic response of the cell. cAMP signaling genes activates platelet genes.
+	reported to interact with Ebola glycoprotein (GP)~\cite{Veljkovic}, chemokines, cytokines, and cytosolic DNA sensing are directly or indirectly controlling cell growth, differentiation, proliferation and apoptosis.
+It suggests that dysregulation of these pathways is responsible for modulation of apoptosis.
+Eventually, sub-graph~3 (top left)
+	demonstrates how INFA1 and INFB1 modulates once more, via only a few
+	intermediate steps, the apoptotic response of the cell. On the other side cAMP signaling genes activates platelet genes.
 
@@
 	data were compared to the genes of sub-graphs identified from the native
-	data. This was also done for the DEG obtained for each run, using three different cutoffs of
-	FDR 0.01, 0.05 and 0.1. The robustness of \texttt{MSF} and the DEG analysis
-	for the time-point 6, 24, and 48~hpi are shown in
+	data. The robustness of \texttt{MSF} analysis
+	for the time-point 6, 24, and 48~hpi is shown in
 	Fig.~\ref{fig:DEGvsMSF}. The procedure how data noise was modeled can be
-	considered as rather stringent, which is already reflected by the limited
-	recall rate in the edgeR based DGEA, between 68~\% (6~hpi) and 93~\%
-	(24~hpi). For \texttt{MSF}-analysis the observed median recall rates lay
-	between 71~\% (6~hpi) and 84~\% (48~hpi). The better performance of the
-	pure DGEA can be explained by the fact that disturbed \textit{p}-values do not
-	change the results for DGEA as long as the \textit{p}-value does not rise above the
-	chosen cutoff value. In contrast, \texttt{MSF} is sensitive to \textit{p}-value
-	changes across the whole range of possible values.
+	considered as rather stringent as \texttt{MSF} is sensitive to \textit{p}-value
+	changes across the whole range of possible values. The observed median recall rates lay
+	between 71~\% (6~hpi) and 84~\% (48~hpi).
 
@@
 	jActiveModules~\cite{jActiveModules} is a plugin in Cytoscape that searches for molecular interaction network to find expression activated sub-networks.
-The method used to find the expression activated sub-networks is close to the method used in \texttt{MSF}.
+The method used to score the expression activated sub-networks is close to the method used in \texttt{MSF}.
 
@@
 The input files were same for both tools, number of modules parameter was set to 5.
-jActiveModule identified 5 modules, the module with the highest pathScore was selected for comparison with \texttt{MSF} identied modulated sub-graphs.
-The module consisting of a single graph with 314 genes.
+jActiveModule identified 5 modules, the module with the highest pathScore was selected for comparison with \texttt{MSF} identified modulated sub-graphs.
+The module consists of a single graph with 314 genes.
 While \texttt{MSF} identified three directed modulated sub-graphs with 42, 69 and 139 genes.
@@
 GSEA uses statistical approaches to identify significantly enriched or depleted groups of genes.
-The complete DEG list from DGEA of the first time-point was analyzed using bioconductor package GSEABase (version 1.44.0).
+The complete DEG list from DGEA of the first time-point 6 hpi was analyzed using bioconductor package GSEABase (version 1.44.0).
 GSEA was able to identify Toll like recpetor, Chemokine signaling pathway, Cytosolic DNA-sensing pathway, Jak-STAT signaling pathway, RIG-I-like receptor signaling pathway and apoptosis as the highest ranked pathways.
-Although GSEA identified the important pathways for Ebola infection, it did not show how the cross-talk and topology of the genes between the pathways identifed.
-Like \texttt{MSF}, GSEA uses complete DEG list without using any cut-off, reasoning why Ebola infection pathways showed up even with weak signal genes in it.
+Although GSEA identified the important pathways for Ebola infection, it did not show how the cross-talk and topology of genes between the pathways identifed.
+\texttt{MSF} and GSEA uses complete DEG list without any cut-off, is why pathways important for Ebola infection showed up even with weak signal genes in it.
 
@@
 				\caption{Visualisation of the three modulated
-					directed sub-graphs identified by \texttt{MSF} at 6~h after
-					EBOV infection in gene detail. The
+					directed sub-graphs identified by \texttt{MSF} at 6~hpi. The
 					node coloring is associated to KEGG pathways
@@
 		\fbox{
-			\begin{minipage}{12 cm}
-				\includegraphics[width=12cm]{DEGvsMSF.png}
-				\caption{Shows the percentage of DEG analysis (3 different cut-offs) genes
-					and \texttt{MSF} identified sub-graph genes recall
+			\begin{minipage}{13 cm}
+				\includegraphics[width=13cm]{MSF_Robustness.png}
+				\caption{Shows the percentage of \texttt{MSF} identified sub-graph genes recall
 					rates for the three different time points of
@@
 		\fbox{
-			\begin{minipage}{10 cm}
-				\includegraphics[width=1.0\linewidth]{ReactomeVsMSF.png}
+			\begin{minipage}{15 cm}
+				\includegraphics[width=1.01\linewidth]{ReactomeVsMSF.png}
 				\caption{The Upset plot shows the number of shared pathways between \texttt{MSF} identified sub-graph gene list and DEG cut-off list for the 3 time-points.
@@
 			in the global network identified after the analysis with
-			\texttt{MSF} and applying different \textit{p}-value cut-offs to gene from edgeR onto the network. }
+			\texttt{MSF} and applying different \textit{p}-value cut-offs from edgeR to genes in MSF identified modulated sub-graphs. }
 		\label{tab:rawVsHartung}
@@
 			\multicolumn{1}{|l|}{}                            & \multicolumn{1}{c|}{\textbf{Total number of}}       & \multicolumn{1}{c|}{\textbf{Number of connected}}      \\
-			\multicolumn{1}{|l|}{}                            & \multicolumn{1}{c|}{\textbf{genes in network}}      & \multicolumn{1}{c|}{\textbf{sub-graph in network}}     \\ \hline
+			\multicolumn{1}{|l|}{}                            & \multicolumn{1}{c|}{\textbf{genes in networks}}      & \multicolumn{1}{c|}{\textbf{sub-graph in network}}     \\ \hline
 			\multicolumn{1}{|l|}{\textbf{6hpi}}               & \multicolumn{1}{l|}{}                                                & \multicolumn{1}{c|}{}                 \\ \hline
 			\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{250}                                             & \multicolumn{1}{c|}{3}                \\ \hline
-			\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{166}                                             & \multicolumn{1}{c|}{87}               \\ \hline
-			\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{152}                                             & \multicolumn{1}{c|}{89}               \\ \hline
-			\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{125}                                             & \multicolumn{1}{c|}{76}               \\ \hline
+			\multicolumn{1}{|l|}{p-value $\leq$ 0.1}  & \multicolumn{1}{c|}{166}                                             & \multicolumn{1}{c|}{87}               \\ \hline
+			\multicolumn{1}{|l|}{p-value $\leq$ 0.05} & \multicolumn{1}{c|}{152}                                             & \multicolumn{1}{c|}{89}               \\ \hline
+			\multicolumn{1}{|l|}{p-value $\leq$ 0.01} & \multicolumn{1}{c|}{125}                                             & \multicolumn{1}{c|}{76}               \\ \hline
 			\multicolumn{1}{|l|}{\textbf{24hpi}}              & \multicolumn{1}{c|}{}                                                & \multicolumn{1}{c|}{}                 \\ \hline
 			\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{656}                                             & \multicolumn{1}{c|}{7}                \\ \hline
-			\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{457}                                             & \multicolumn{1}{c|}{183}               \\ \hline
-			\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{418}                                             & \multicolumn{1}{c|}{198}              \\ \hline
-			\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{332}                                             & \multicolumn{1}{c|}{216}              \\ \hline
+			\multicolumn{1}{|l|}{p-value $\leq$ 0.1}  & \multicolumn{1}{c|}{457}                                             & \multicolumn{1}{c|}{183}               \\ \hline
+			\multicolumn{1}{|l|}{p-value $\leq$ 0.05} & \multicolumn{1}{c|}{418}                                             & \multicolumn{1}{c|}{198}              \\ \hline
+			\multicolumn{1}{|l|}{p-value $\leq$ 0.01} & \multicolumn{1}{c|}{332}                                             & \multicolumn{1}{c|}{216}              \\ \hline
 			\multicolumn{1}{|l|}{\textbf{48hpi}}              & \multicolumn{1}{c|}{}                                                & \multicolumn{1}{c|}{}                 \\ \hline
 			\multicolumn{1}{|l|}{edgeR + MSF}                 & \multicolumn{1}{c|}{744}                                             & \multicolumn{1}{c|}{6}                \\ \hline
-			\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.1)}  & \multicolumn{1}{c|}{514}                                             & \multicolumn{1}{c|}{189}               \\ \hline
-			\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.05)} & \multicolumn{1}{c|}{468}                                             & \multicolumn{1}{c|}{206}               \\ \hline
-			\multicolumn{1}{|l|}{edgeR + MSF (p-value $\leq$ 0.01)} & \multicolumn{1}{c|}{363}                                             & \multicolumn{1}{c|}{241}               \\ \hline
+			\multicolumn{1}{|l|}{p-value $\leq$ 0.1}  & \multicolumn{1}{c|}{514}                                             & \multicolumn{1}{c|}{189}               \\ \hline
+			\multicolumn{1}{|l|}{p-value $\leq$ 0.05} & \multicolumn{1}{c|}{468}                                             & \multicolumn{1}{c|}{206}               \\ \hline
+			\multicolumn{1}{|l|}{p-value $\leq$ 0.01} & \multicolumn{1}{c|}{363}                                             & \multicolumn{1}{c|}{241}               \\ \hline
 		\end{tabular}

### 2019-02-28 MSF_Paper_FA.tex
comments FA

@@
 \usepackage{pdfpages}
+\usepackage[normalem]{ulem}
 \newcommand{\FA}[1]{\begingroup\color{magenta}#1\endgroup}
@@
 	topology of each gene during the gene-level statistics but still lacks to aim
-	to link the different functional pathways~\cite{Khatri2012}.
+	to link the different functional pathways~\cite{Khatri2012}. \FA{mention jActiveModules somewhere in the intro}
 
@@
 	entry points of perturbation in the networks and overlap with conventional
-	pathway categories are returned. The output is prepared in a directed
+	pathway categories are returned \FA{mention impact score and reliablilty of sources here}. The output is prepared in a directed
 	adjacency file, convenient for visualization, e.g., with
@@
 
-	In the next step, we check if any of the initial modulated sub-graphs 
-	could further be extended.
-This is done when the sub-graph could not be extended any more to the direct neighboring gene due to a high \textit{p}-value, but the neighboring gene to the direct neighboring gene has a low \textit{p}-value.
-This was required to connect together low \textit{p}-value gene but incorporating a single high \textit{p}-value gene.
-If the combined \textit{p}-value of the initial modulated sub-graph and the extension genes is smaller than the \textit{p}-value of the initial sub-graph the extension is accepted.
-All possible extension paths up to \emph{3} (default 2) genes at
-	all nodes in the sub-graph are tested. Again, this step is iteratively repeated until
-	no further genes are added to the significant differentially expressed
-	sub-graphs. This step bridges small gaps of genes without a clear
-	differential signal in the DGEA.\newline
+	In the next step, we check if any of the initial modulated
+        sub-graphs could further be extended \FA{beyond the immediate
+          neighbouring genes. Instead of testing single genes and
+          their compatibility to be added, groups of genes are
+          considered.}
+
+        \sout{This is done when the sub-graph could not be extended any more
+        to the direct neighboring gene due to a high \textit{p}-value,
+        but the neighboring gene to the direct neighboring gene has a
+        low \textit{p}-value. This was required to connect together
+        low \textit{p}-value gene but incorporating a single high
+        \textit{p}-value gene.} If the combined \textit{p}-value of the
+        initial modulated sub-graph and the extension genes is smaller
+        than the \textit{p}-value of the initial sub-graph the
+        extension is accepted. All possible extension paths up to
+        \emph{3} (default 2) genes at all nodes in the sub-graph are
+        tested. Again, this step is iteratively repeated until no
+        further genes are added to the significant differentially
+        expressed sub-graphs. This step bridges small gaps of genes
+        without a clear differential signal in the DGEA.\newline
 
@@
 
-	After detection and extension of the modulated sub-graphs, they are tested
-	if combining the different sub-graphs score better than on their own.
-The merging of the sub-graphs is done by depth first search traversal from one sub-graph to the other sub-graph.
-If the two sub-graphs merge
-	with the connector of at most \emph{3} genes (default 2 gene) and the
-	combined \textit{p}-value of the merged sub-graph including the bridging
-	genes in between is less than the individual \textit{p}-values of the two
-	sub-graphs, the two sub-graphs are merged together to one big modulated
-	sub-graph. This step is repeated iteratively until no sub-graphs could be
-	merged.\newline
+        \FA{After detection and extension of the modulated sub-graphs,
+          each pair of so far identified sub-graphs is tested if its
+          combination scores better than each on its own.}
+        \sout{After detection and extension of the modulated
+          sub-graphs, they are tested if combining the different
+          sub-graphs score better than on their own.} The merging of
+        the sub-graphs is done by depth first search traversal from
+        one sub-graph to the other sub-graph. If the two sub-graphs
+        merge with the connector of at most \emph{3} genes (default 2
+        gene) and the combined \textit{p}-value of the merged
+        sub-graph including the bridging genes in between is less than
+        the individual \textit{p}-values of the two sub-graphs, the
+        two sub-graphs are merged together to one big\FA{ger}
+        modulated sub-graph. This step is repeated iteratively until
+        no sub-graphs \sout{could be merged}\FA{can be merged
+          anymore}.\newline
 
@@
 
-	In a last post processing step \texttt{MSF} identifies the trigger points
-	of the modulated sub-graphs. These trigger genes are the sources of the
-	sub-graphs with only outgoing edges. These genes can be interpreted as the
-	possible entry points of perturbation from where the stimulus causes
-	downstream effects.
-Each individual source is given an impact score, depending on the number of down stream genes effected in the corresponding sub-graph, equipping with a way to choose the most significant genes for unknown phenotypes or biological processes.
-In the same spirit the most downstream genes of the
-	modulated sub-graphs are identified and defined as sinks. Due to loops not all
-	sub-graphs are guaranteed to have sources or sinks.
+	In a last post processing step \texttt{MSF} identifies the
+        trigger points of the modulated sub-graphs. These trigger
+        genes are the sources of the sub-graphs with only outgoing
+        edges. These genes can be interpreted as the possible entry
+        points of perturbation from where the stimulus causes
+        downstream effects. Each individual source is given an impact
+        score, \FA{expressing the relative number of downstream genes
+          within the corresponding sub-graph directly connected by
+          directed links. This score can be interpreted as an upper
+          limit of how much of the sub-graph's perturbation could have
+          been introduced by the respective source, and thereby could
+          be helpful to prioritize different identified sources for
+          larger sub-graphs.}\sout{depending on the number of down
+          stream genes effected in the corresponding sub-graph,
+          equipping with a way to choose the most significant genes
+          for unknown phenotypes or biological processes.} In the same
+        spirit \FA{as sources}, the most downstream genes of the
+        modulated sub-graphs are identified and defined as sinks. Due
+        to loops not all sub-graphs are guaranteed to have sources or
+        sinks.
+
+        \TODO{?what about source reliability?}
 
@@
 
-	\texttt{MSF} generates a directed network file as an output, containing complete directed interactions of all modulated sub-graphs identified.
-This file could be imported into Cytoscape~\cite{Cyto} for visualization.
-Second output is the source \& sink file.
-It contains for all the modulated sub-graphs, sources and sinks identified with the source impact scores.
-The third output, source weightage is provided as a node attribute file.
-It could be imported into Cytoscape to see the impact score of sources as the size of the node and log-fold change could be used to see how the regulation of different gene hubs in the modulated sub-graphs look.
+	\texttt{MSF} generates a directed network file as an output,
+        containing complete directed interactions of all modulated
+        sub-graphs identified. This file could be imported into
+        Cytoscape~\cite{Cyto} for visualization. \sout{Second output
+          is the source \& sink file. It contains} \FA{Additionally, a
+          file containing details on all sources and sinks for all
+          modulated sub-graphs is reported. Furthermore, for
+          facilitated visualisation in Cytoscape, a node attribute
+          file is provided, containing the source weightage and the
+          log-fold chances of all considered genes}. \sout{The third output,
+        source weightage is provided as a node attribute file. It
+        could be imported into Cytoscape to see the impact score of
+        sources as the size of the node and log-fold change could be
+        used to see how the regulation of different gene hubs in the
+        modulated sub-graphs look.}
 
@@
 
-	MSF requires Java version 8 and JDK 1.8.
-The few package dependencies are already been added to the release.
-MSF runs fast on a standard laptop computer and so it has normal system requirements.
-To run MSF, the user must provide two text files, one containing the DGEA and the second one containing directed interactions in an adjacency format file.
-Example files and a detailed tutorial to use MSF has been provided on github https://github.
-com/Modulated-Subgraph-Finder/MSF.
+	MSF requires Java version 8 and JDK 1.8. The few package
+        dependencies are already been added to the release. MSF runs
+        fast on a standard laptop computer and so it has normal system
+        requirements \TODO{What is fast and what is normal?}. To run
+        MSF, the user must provide two text files, one containing the
+        DGEA \TODO{DESeq2 edgR?} and the second one containing
+        directed interactions in an adjacency format file. Example
+        files and a detailed tutorial to use MSF has been provided on
+        github https://github. com/Modulated-Subgraph-Finder/MSF.
 
@@
 
-	To demonstrate its usefulness, \texttt{MSF} is applied to RNA-seq data
-	set of primary human monocyte-derived macrophages (MDMs) infected with
-	Ebola virus (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV) belongs to the Filoviridea family: filamentous, enveloped and single stranded RNA
-	viruses. EBOV causes hemorrhagic fever in humans, inducing the host innate
-	and adaptive immune response to be unable to control virus
-	infection~\cite{Prins}. Currently, there are no approved antiviral drugs
-	for the treatment of Ebola virus infection~\cite{Konde,Rhein}.  The initial
-	targets of EBOV are the macrophages and dendritic immune
-	cells~\cite{Falasca,Rhein}. EBOV inhibits the critical innate immune
-	response of the host, which includes the activation of alpha/beta
-	interferon (IFN-$\alpha / \beta$)~\cite{Prins,Konde,Cardenas}. It has been
-	proposed that IFN-$\alpha / \beta$ should be tested against Ebola for its
-	antiviral activity through clinical trials~\cite{Konde}. Ebola infection
-	data was selected to test the approach because it has been well recognized for the last several decades, and vast literature is available for the pathogenesis of Ebola, thereby, facilitating the verification of the results of
+	To demonstrate its usefulness, \texttt{MSF} is applied to
+        RNA-seq data set of primary human monocyte-derived macrophages
+        (MDMs) infected with Ebola virus
+        (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV) belongs to the
+        Filoviridea family: filamentous, enveloped and single stranded
+        RNA viruses. EBOV causes hemorrhagic fever in humans, inducing
+        the host innate and adaptive immune response to be unable to
+        control virus infection~\cite{Prins}. Currently, there are no
+        approved antiviral drugs for the treatment of Ebola virus
+        infection~\cite{Konde,Rhein}.  The initial targets of EBOV are
+        the macrophages and dendritic immune
+        cells~\cite{Falasca,Rhein}. EBOV inhibits the critical innate
+        immune response of the host, which includes the activation of
+        alpha/beta interferon (IFN-$\alpha /
+        \beta$)~\cite{Prins,Konde,Cardenas}. It has been proposed that
+        IFN-$\alpha / \beta$ should be tested against Ebola for its
+        antiviral activity through clinical trials~\cite{Konde}. Ebola
+        infection data was selected to test the approach because it
+        has been well recognized for the last several decades, and
+        vast literature is available for the pathogenesis of Ebola,
+        thereby, facilitating the verification of the results of
         \texttt{MSF} with the vast literature present on Ebola
-	infection. Especially, the detection of IFN-$\alpha / \beta$ as point of
-	action for the virus, could be considered as a basic indicator of the
-	correctness and usefulness of the approach.
+        infection. Especially, the detection of IFN-$\alpha / \beta$
+        as point of action for the virus, could be considered as a
+        basic indicator of the correctness and usefulness of the
+        approach.
 
-	EBOV infection count data was downloaded from GEO (GSE84188), it describes the course of infection at three
-	time-points 6, 24 and 48 hour post infection (hpi). Differential
-	gene expression analysis was performed on the count data with edgeR package
-	(version 3.4.2)~\cite{edgeR} using an upper-quartile normalization. The DEG analysis results generated by edgeR
-	were used as input for \texttt{MSF}. Cell signaling interactions
-	were filtered from Reactome Functional interactions (FIs) Version
-	2016~\cite{Cytokegg} for only direct interactions, which was used as a second input for \texttt{MSF}.
+	EBOV infection count data was downloaded from GEO (GSE84188),
+        it describes the course of infection at three time-points 6,
+        24 and 48 hour post infection (hpi). Differential gene
+        expression analysis was performed on the count data with edgeR
+        package (version 3.4.2)~\cite{edgeR} using an upper-quartile
+        normalization. The DEG analysis results generated by edgeR
+        were used as input for \texttt{MSF}. Cell signaling
+        interactions were filtered from Reactome Functional
+        interactions (FIs) Version 2016~\cite{Cytokegg} for only
+        direct interactions, which was used as a second input for
+        \texttt{MSF}.
 
-	For the earliest time point at 6~hpi, three large modulated sub-graphs were identified with 42,
-	139, and 69 genes.
-	The modulated sub-graphs consist predominantly of cytokines, chemokines
-	(CXCL10, CCL8, CXCL9, CXCL11, CXCR4, CCR7, CCL4L1, CCL3L1, CCL4, CCL8, CCL20, CCL3, CCL19) and Interleukin genes (IL6, IL27, IL23).
-	IFNB1 and IFNA1 were both identified as two of the possible sources
-	in the most significantly modulated sub-graph identified with 42 genes. The impact score of IFNB1 is 14.5\% and IFNA1 is 8.7\% for the sub-graph they belong to.
-	Most of the sources identified by \texttt{MSF} were
-	type~I interferon induced genes (Supplement material).
-	At 24~hpi seven modulated sub-graphs were
-	identified with four main sub-graphs consisting
-	of 61, 222, 130 and 242 genes, others being smaller than 6 genes. Again, IFNB1
-	and IFNA1 were identified as two sources out of the total sources with 3.9\% and 1.6\% impact score. For the last time-point 48~hpi,
-	six modulated sub-graphs were identified. Three of the sub-graphs were
-	less than ten genes and main sub-graphs had 217, 224 and 276 genes. IFNB1 and IFNA1 were identified as sources in
-	the most significantly modulated sub-graph with an impact score of 2.8\% and 3.7\% (Supplement material).
+	For the earliest time point at 6~hpi, three large modulated
+        sub-graphs were identified with 42, 139, and 69 genes.  The
+        modulated sub-graphs consist predominantly of cytokines,
+        chemokines (CXCL10, CCL8, CXCL9, CXCL11, CXCR4, CCR7, CCL4L1,
+        CCL3L1, CCL4, CCL8, CCL20, CCL3, CCL19) and Interleukin genes
+        (IL6, IL27, IL23).  IFNB1 and IFNA1 were both identified as
+        two of the possible sources in the most significantly
+        modulated sub-graph identified with 42 genes. The impact score
+        of IFNB1 is 14.5\% and IFNA1 is 8.7\% for the sub-graph they
+        belong to.  Most of the sources identified by \texttt{MSF}
+        were type~I interferon induced genes (Supplement material).
+        At 24~hpi seven modulated sub-graphs were identified with four
+        main sub-graphs consisting of 61, 222, 130 and 242 genes,
+        others being smaller than 6 genes. Again, IFNB1 and IFNA1 were
+        identified as two sources out of the total sources with 3.9\%
+        and 1.6\% impact score. For the last time-point 48~hpi, six
+        modulated sub-graphs were identified. Three of the sub-graphs
+        were less than ten genes and main sub-graphs had 217, 224 and
+        276 genes. IFNB1 and IFNA1 were identified as sources in the
+        most significantly modulated sub-graph with an impact score of
+        2.8\% and 3.7\% (Supplement material). \TODO{was the mentioned
+          impact score the highest of all sources? How many sources
+          are there?}
 
-	As stated earlier IFN-$\alpha / \beta$ was reported to be one of the target
-	genes of Ebola infection. We were able to successfully identify IFNA1 and IFNB1 as
-	sources in all three Ebola infection time-points. Although IFNA1 and IFNB1 were already among the most
-	significant genes in the DGEA during the later time points, \texttt{MSF} was
-	also able to detect them as a source in the very early time-point when the
-	genes were not significant based on the individual DGEA alone. Identifying
-	the possible sources will reduce the search space for potential target
-	genes and can help the biologist as the starting point of clinical testing
-	for drugs and vaccines against an infection.
+	As stated earlier IFN-$\alpha / \beta$ was reported to be one
+        of the target genes of Ebola infection. We were able to
+        successfully identify IFNA1 and IFNB1 as sources in all three
+        Ebola infection time-points. Although IFNA1 and IFNB1 were
+        already among the most significant genes in the DGEA during
+        the later time points, \texttt{MSF} was also able to detect
+        them as a source in the very early time-point when the genes
+        were not significant based on the individual DGEA
+        alone. Identifying the possible sources will reduce the search
+        space for potential target genes and can help the biologist as
+        the starting point of clinical testing for drugs and vaccines
+        against an infection.
 
-	Table~\ref{tab:rawVsHartung} compares the results of \texttt{MSF}, namely
-	the number of detected sub-modules and their total genes numbers, to a
-	simple analysis of mapping significantly modulated genes from the DGEA to
-	the network and joining neighbors to modules. The numbers indicate that
-	\texttt{MSF} detects less but larger and meaningful sub-modules, applying its statistical
-	test. Furthermore, the dependency of the results from the \textit{p}-value cutoff
-	choice is demonstrated for the DGEA, which is avoided for \texttt{MSF}
-	altogether.
-It showcases how applying different cut-offs to the \textit{p}-value of genes from edgeR to the sub-graphs identified by \texttt{MSF} breaks the larger sub-graphs to many smaller unconnected sub-graphs, many of which are single genes.
+	Table~\ref{tab:rawVsHartung} compares the results of
+        \texttt{MSF}, namely the number of detected sub-modules and
+        their total genes numbers, to a simple analysis of mapping
+        significantly modulated genes from the DGEA to the network and
+        joining neighbors to modules. The numbers indicate that
+        \texttt{MSF} detects less but larger and meaningful \TODO{?}
+        sub-modules, applying its statistical test. Furthermore, the
+        dependency of the results from the \textit{p}-value cutoff
+        choice is demonstrated for the DGEA, which is avoided for
+        \texttt{MSF} altogether. It showcases how applying different
+        cut-offs to the \textit{p}-value of genes from edgeR to the
+        sub-graphs identified by \texttt{MSF} breaks the larger
+        sub-graphs to many smaller unconnected sub-graphs, many of
+        which are single genes.
 
@@
 
-	Three main modulated sub-graphs identified by \texttt{MSF} at 6~hpi are
-	shown in Fig.~\ref{fig:Sub-graph6hpi}. The graphs represent the immediate output of the \texttt{MSF}-analysis,
-	visualized by StringApp~\cite{StringApp} in Cytoscape~\cite{Cyto}. Each
-	node represents a gene part of a modulated sub-graph, whereby the
-	associated colors code the functional annotation deduced from KEGG
-	Pathways. The cross-talk between the pathways and also the multiple
-	employment of many genes is evident. The flow of information between
-	the sensors and effectors in this particular example could easily be deduced.
-	In more detail, sub-graph~1 (bottom) shows how the activation of toll-like
-	receptor, cytokine, chemokine activating jak-stat and mapk genes, together with TNF leads into apoptosis. The next significant sub-graph (sub-graph~2: top right) reveals
-	how information from the Extra-cellular matrix (ECM) receptor, which are
-	reported to interact with Ebola glycoprotein (GP)~\cite{Veljkovic}, chemokines, cytokines, and cytosolic DNA sensing are directly or indirectly controlling cell growth, differentiation, proliferation and apoptosis.
-It suggests that dysregulation of these pathways is responsible for modulation of apoptosis.
-Eventually, sub-graph~3 (top left)
-	demonstrates how INFA1 and INFB1 modulates once more, via only a few
-	intermediate steps, the apoptotic response of the cell. On the other side cAMP signaling genes activates platelet genes.
+	Three main modulated sub-graphs identified by \texttt{MSF} at
+        6~hpi are shown in Fig.~\ref{fig:Sub-graph6hpi}. The graphs
+        represent the immediate output of the \texttt{MSF}-analysis,
+        visualized by StringApp~\cite{StringApp} in
+        Cytoscape~\cite{Cyto}. Each node represents a gene part of a
+        modulated sub-graph, whereby the associated colors code the
+        functional annotation deduced from KEGG Pathways. The
+        cross-talk between the pathways and also the multiple
+        employment of many genes is evident. The flow of information
+        between the sensors and effectors can be perceived given the
+        directionality of each interaction, indicated by arrows
+        \sout{in this particular example could easily be deduced}.  In
+        more detail, sub-graph~1 (bottom) shows how the activation of
+        toll-like receptor, cytokine, chemokine activating jak-stat
+        and mapk genes, together with TNF leads into apoptosis. The
+        next significant sub-graph (sub-graph~2: top right) reveals
+        how information from the Extra-cellular matrix (ECM) receptor,
+        which are reported to interact with Ebola glycoprotein
+        (GP)~\cite{Veljkovic}, chemokines, cytokines, and cytosolic
+        DNA sensing are directly or indirectly controlling cell
+        growth, differentiation, proliferation and apoptosis. It
+        suggests that dysregulation of these pathways is responsible
+        for modulation of apoptosis. Eventually, sub-graph~3 (top
+        left) demonstrates how INFA1 and INFB1 modulates once more,
+        via only a few intermediate steps, the apoptotic response of
+        the cell. On the other side cAMP signaling genes activates
+        platelet genes.
 
-	This display case might advertise with how little effort complex data
-	can be interpreted, help to apprehend the dynamics of the underlying
-	processes and suggest testable hypothesis and potential points of
+	This display case might advertise with how little effort
+        complex data can be \FA{processed and prepared for
+          interpretation by the domain expert, to } \sout{interpreted,
+          help to} apprehend the dynamics of the underlying processes
+        and suggest testable hypothesis and potential points of
         intervention.
@@
 
-	A potential concern is how noise in the gene expression measurements
-	affects our analysis. To assess the robustness and stability of our method,
-	we therefore added Poisson distributed noise
-	to the read counts of the three time-points data set, used above.
-	Then DGEA was carried out on the disturbed data
-	with the same parameters as for the native data using edgeR, followed by
-	analysis with \texttt{MSF}. This procedure was carried out 100 times.
-	Every time the genes from the modulated sub-graphs identified from noisy
-	data were compared to the genes of sub-graphs identified from the native
-	data. The robustness of \texttt{MSF} analysis
-	for the time-point 6, 24, and 48~hpi is shown in
-	Fig.~\ref{fig:DEGvsMSF}. The procedure how data noise was modeled can be
-	considered as rather stringent as \texttt{MSF} is sensitive to \textit{p}-value
-	changes across the whole range of possible values. The observed median recall rates lay
+	A potential concern is how noise in the gene expression
+        measurements affects our analysis. To assess the robustness
+        and stability of our method, we therefore added Poisson
+        distributed noise to the read counts of the three time-points
+        data set, used above.  Then DGEA was carried out on the
+        disturbed data with the same parameters as for the native data
+        using edgeR, followed by analysis with \texttt{MSF}. This
+        procedure was carried out 100 times.  Every time the genes
+        from the modulated sub-graphs identified from noisy data were
+        compared to the genes of sub-graphs identified from the native
+        data. The robustness of \texttt{MSF} analysis for the
+        time-point 6, 24, and 48~hpi is shown in
+        Fig.~\ref{fig:DEGvsMSF}. The procedure how data noise was
+        modeled can be considered as rather stringent as \texttt{MSF}
+        is sensitive to \textit{p}-value changes across the whole
+        range of possible values. The observed median recall rates lay
         between 71~\% (6~hpi) and 84~\% (48~hpi).
@@
 
-	jActiveModules~\cite{jActiveModules} is a plugin in Cytoscape that searches for molecular interaction network to find expression activated sub-networks.
-The method used to score the expression activated sub-networks is close to the method used in \texttt{MSF}.
+	jActiveModules~\cite{jActiveModules} \TODO{citation missing in
+          bib} is a plugin in Cytoscape that searches for molecular
+        interaction network to find expression activated
+        sub-networks. The method used to score the expression
+        activated sub-networks is close to the method used in
+        \texttt{MSF}. \TODO{what are the differences?}
 
-	The first time-point of Ebola infection data was analyzed using jActiveModule (Version 3.2.1) to compare the modulated sub-graphs identified by \texttt{MSF} and jActiveModule.
-The input files were same for both tools, number of modules parameter was set to 5.
-jActiveModule identified 5 modules, the module with the highest pathScore was selected for comparison with \texttt{MSF} identified modulated sub-graphs.
-The module consists of a single graph with 314 genes.
-While \texttt{MSF} identified three directed modulated sub-graphs with 42, 69 and 139 genes.
-The overlap of the common genes identified between \texttt{MSF} and jActiveModule is shown in Fig~\ref{fig:MSFvsJActiveModule}.
-It is difficult to state which tool identifies more accurate set of genes as modulated, since there is no golden standard data example to refer.
-Although jActiveModule identifies single connected module, \texttt{MSF} provides directionality, with the identification of possible perturbation sources of the sub-graphs.
+	The first time-point of Ebola infection data was analyzed
+        using jActiveModule (Version 3.2.1) to compare the modulated
+        sub-graphs identified by \texttt{MSF} and jActiveModule. The
+        input files were same for both tools, number of modules
+        parameter was set to 5 \TODO{what does this parameter
+          do?}. jActiveModule identified 5 modules, the module with
+        the highest pathScore was selected for comparison with
+        \texttt{MSF} identified modulated sub-graphs \TODO{why only 1
+          out of 5?}. The module consists of a single graph with 314
+        genes. While \texttt{MSF} identified three directed modulated
+        sub-graphs with 42, 69 and 139 genes. The overlap of the
+        common genes identified between \texttt{MSF} and jActiveModule
+        is shown in Fig~\ref{fig:MSFvsJActiveModule}. \sout{It is
+          difficult to state which tool identifies more accurate set
+          of genes as modulated, since there is no golden standard
+          data example to refer.} Although jActiveModule identifies
+        single connected module, \texttt{MSF} provides directionality,
+        with the identification of possible perturbation sources of
+        the sub-graphs. \TODO{find reasons why jActiveModules gives so
+          much bigger moduls by comparing the details of the method!}
 
@@
 
-	Gene enrichment analysis was performed using Reactome analyze data
-	tool~\cite{Reactome} (version 67). Reactome's over-representation analysis tool tests whether
-	certain Reactome pathways are enriched for the list of genes submitted to it. Genes from \texttt{MSF} identified
-	sub-graphs for each time-point were analyzed for gene enrichment using this tool.  For
-	comparison the DEG results from edgeR for the three time-points were filtered using the cut-off of
-	adjusted \textit{p}-value > 0.05. This DEG list was used for gene
-	enrichment analysis. The compression of \texttt{MSF} identified sub-graphs gene list and the DEG list analysis is shown in Fig~\ref{fig:msfvsreactome}
+	Gene enrichment analysis was performed using Reactome analyze
+        data tool~\cite{Reactome} (version 67) \TODO{on which
+          data?}. Reactome's over-representation analysis tool tests
+        whether certain Reactome pathways are enriched for the list of
+        genes submitted to it. Genes from \texttt{MSF} identified
+        sub-graphs for each time-point were analyzed for gene
+        enrichment using this tool.  For comparison the DEG results
+        from edgeR for the three time-points were filtered using the
+        cut-off of adjusted \textit{p}-value \sout{>}\FA{<} 0.05. This
+        DEG list was used for gene enrichment analysis. The
+        compression of \texttt{MSF} identified sub-graphs gene list
+        and the DEG list analysis is shown in
+        Fig~\ref{fig:msfvsreactome}
 
-	All enriched pathways with a cut-off of \textit{p}-value > 0.05 for \texttt{MSF} and DEG list for the three time-points were selected. The comparison shows most of the pathways known from literature
-	to be dis-regulated by Ebola infection are enriched in both the enrichment
-	analysis. Toll-Like receptor signaling pathway when interacts with EBOV
-	glycoprotein (GP), it triggers the activation of
-	cytokines~\cite{Olejnik}.
-Toll-like receptor pathway is expected to be dis-regulated in the early stage of infection, this pathway was not identified as significantly dis-regulated when \textit{p}-value
-	cut off DEG list was analyzed for enrichment.
-Nine toll-like receptor cascades TLR10, TLR2, TLR3, TLR4, TLR5, TLR7/8, TLR9, TLR1:TLR2 and TLR6:TLR2 were identified as dis-regulated from gene enrichment analysis of \texttt{MSF} identified sub-graph genes, not a single one of these cascade was shown to be dis-regulated in pathway enrichment analysis from DEG cut-off list.
-Since \texttt{MSF}
-	considers the complete DEG results, even the weak signal at the earliest
-	time-point was detected; for example Toll-like receptor signaling.
-While \texttt{MSF} is able to catch weak signals, it does not provide information about the functional relationships among genes like Reactome tool.
+	All enriched pathways with a cut-off of \textit{p}-value
+        \sout{>}\FA{<} 0.05 0.05 for \texttt{MSF} and DEG list for the
+        three time-points were selected. The comparison shows most of
+        the pathways known from literature to be dis-regulated by
+        Ebola infection are enriched in both the enrichment
+        analysis. Toll-Like receptor signaling pathway when interacts
+        with EBOV glycoprotein (GP), it triggers the activation of
+        cytokines~\cite{Olejnik}. Toll-like receptor pathway is
+        expected to be dis-regulated in the early stage of infection,
+        this pathway was not identified as significantly dis-regulated
+        when \textit{p}-value cut off DEG list was analyzed for
+        enrichment. Nine toll-like receptor cascades TLR10, TLR2,
+        TLR3, TLR4, TLR5, TLR7/8, TLR9, TLR1:TLR2 and TLR6:TLR2 were
+        identified as dis-regulated from gene enrichment analysis of
+        \texttt{MSF} identified sub-graph genes, not a single one of
+        these cascade was shown to be dis-regulated in pathway
+        enrichment analysis from DEG cut-off list. Since \texttt{MSF}
+        considers the complete DEG results, even the weak signal at
+        the earliest time-point was detected; for example Toll-like
+        receptor signaling. While \texttt{MSF} is able to catch weak
+        signals, it does not provide information about the functional
+        relationships among genes like Reactome tool.
 
@@
 
-	Gene set enrichment analysis (GSEA)~\cite{Subramanian15545} is a method to identify classes of genes or proteins that are over-represented in a large set of genes or proteins.
-GSEA uses statistical approaches to identify significantly enriched or depleted groups of genes.
-The complete DEG list from DGEA of the first time-point 6 hpi was analyzed using bioconductor package GSEABase (version 1.44.0).
-GSEA was able to identify Toll like recpetor, Chemokine signaling pathway, Cytosolic DNA-sensing pathway, Jak-STAT signaling pathway, RIG-I-like receptor signaling pathway and apoptosis as the highest ranked pathways.
-Although GSEA identified the important pathways for Ebola infection, it did not show how the cross-talk and topology of genes between the pathways identifed.
-\texttt{MSF} and GSEA uses complete DEG list without any cut-off, is why pathways important for Ebola infection showed up even with weak signal genes in it.
+	Gene set enrichment analysis (GSEA)~\cite{Subramanian15545}
+        \TODO{reference is missing in bib} is a method to identify
+        classes of genes or proteins that are over-represented in a
+        large set of genes or proteins. GSEA uses statistical
+        approaches to identify significantly enriched or depleted
+        groups of genes. The complete DEG list from DGEA of the first
+        time-point 6 hpi was analyzed using bioconductor package
+        GSEABase (version 1.44.0). GSEA was able to identify
+        Toll\FA{-}like recpetor, chemokine signaling pathway,
+        cytosolic DNA-sensing pathway, Jak-STAT signaling pathway,
+        RIG-I-like receptor signaling pathway and apoptosis as the
+        highest ranked pathways. Although GSEA identified the
+        important pathways for Ebola infection, it did not show how
+        the cross-talk and topology of genes between the pathways
+        identifed. \texttt{MSF} and GSEA uses complete DEG list
+        without any cut-off, is why pathways important for Ebola
+        infection showed up even with weak signal genes in it.
 
@@
 
-	Classic pathway analysis tools aim to detect in lists of significantly
-	deregulated genes enriched associations with pathway genes categorized by
-	their biological function and their interactions. Thereby, depending on the
-	tool, the internal pathway topology is considered or neglected all
-	together. The here presented tool, \texttt{MSF}, employs a different
-	approach, by aiming to detect sub-graphs in whole gene regulatory networks
-	which are significantly deregulated in a concerted manner. To this end,
-	neighboring genes in the user provided network are tested for jointly
-	common regulation. Exploiting that each gene's abundance, although not
-	independent from its neighbors, is measured repeatedly on its own,
-	sensitivity can be increased by our applied \textit{p}-value meta-analysis, namely
-	Hartung's method. This potentially enables to call just not significant
-	modulated genes based on the DGEA to be convincingly called to be part of a
-	deregulated gene group.  Furthermore, it allows to identify connected
-	sub-graphs, representing the propagation of gene regulation perturbation in
-	the input network. A better understanding of this propagation, especially
-	the critical spots such as sensors, effectors, and hubs, facilitates the projection of potential intervention points,
-	e.g., for drug development. Since \texttt{MSF} only uses interaction
-	information in gene regulation network, but not the functional grouping of
-	the genes into functional pathways, it is especially adapted to discover so
-	called cross-talk between such pathways.
+	Classic pathway analysis tools aim to detect in lists of
+        significantly deregulated genes enriched associations with
+        pathway genes categorized by their biological function and
+        their interactions. Depending on the tool, the internal
+        pathway topology is considered or neglected all together. The
+        here presented tool, \texttt{MSF}, employs a different
+        approach, by aiming to detect sub-graphs in whole gene
+        regulatory networks which are significantly deregulated in a
+        concerted manner. To this end, neighboring genes in the user
+        provided network are tested for jointly common
+        regulation. Exploiting that each gene's abundance, although
+        not independent from its neighbors, is measured repeatedly on
+        its own, sensitivity can be increased by our applied
+        \textit{p}-value meta-analysis, namely Hartung's method. This
+        potentially enables to call just not significant modulated
+        genes based on the DGEA to be convincingly called to be part
+        of a deregulated gene group.  Furthermore, it allows to
+        identify connected sub-graphs, representing the propagation of
+        gene regulation perturbation in the input network. A better
+        understanding of this propagation, especially the critical
+        spots such as sensors, effectors, and hubs, facilitates the
+        projection of potential intervention points, e.g., for drug
+        development. Since \texttt{MSF} only uses interaction
+        information in gene regulation network, but not the functional
+        grouping of the genes into functional pathways, it is
+        especially adapted to discover so called cross-talk between
+        such pathways.
 
@@
 
-	\texttt{MSF} is a fast and easy to use tool to find concertedly
-	modulated sub-graphs in a given network. Its implementation in java enables
-	its use across many operating systems e.g. linux and windows. So far the raw output from edgeR~\cite{edgeR} and DESeq2~\cite{love2014moderated} are supported.
+	\texttt{MSF} is a fast and easy to use tool to find
+        concertedly modulated sub-graphs in a given network. Its
+        implementation in java enables its use across many operating
+        systems e.g. linux and windows. So far the raw output from
+        edgeR~\cite{edgeR} and DESeq2~\cite{love2014moderated} are
+        supported.
 
@@
 
-	The Ebola infection RNA-seq data set analyzed during the current study are
-	available in the GEO repository (GSE84188)~\cite{Olejnik}. The cell
-	signaling network file used is from Reactome Functional interactions (FIs)
-	Version 2016~\cite{Cytokegg}
+	The Ebola infection RNA-seq data set analyzed during the
+        current study are available in the GEO repository
+        (GSE84188)~\cite{Olejnik}. The cell signaling network file
+        used is from Reactome Functional interactions (FIs) Version
+        2016~\cite{Cytokegg}
 
@@
 	This work was funded by the FWF (“Fonds zur F{\"o}rderung der
-	wissenschaftlichen Forschung”) within the project Internationalen
-	Kooperationsprojektes - Intl cooperation Project (Joint Project - Lead
-	Agency Verfahren) with the project number (I 1988-B22). The grant was
-	assigned to ILH. FA was funded by the Austrian Science Fund (FWF) project
-	SFB F43.
+        wissenschaftlichen Forschung”) within the project
+        Internationalen Kooperationsprojektes - Intl cooperation
+        Project (Joint Project - Lead Agency Verfahren) with the
+        project number (I 1988-B22). The grant was assigned to ILH. FA
+        was funded by the Austrian Science Fund (FWF) project SFB F43.
 
@@
 				\includegraphics[width=13cm]{MSF_Robustness.png}
-				\caption{Shows the percentage of \texttt{MSF} identified sub-graph genes recall
-					rates for the three different time points of
-					EBOV infection data for 100 simulation where
-					Poisson distributed noise was added to the
+				\caption{\FA{Recall rates for genes in
+                                  }\sout{Shows the percentage of}
+                                  \texttt{MSF} identified sub-graph
+                                  \sout{genes recall rates} for the
+                                  three different time points of EBOV
+                                  infection data for 100
+                                  simulation\FA{s} where Poisson
+                                  distributed noise was added to the
                                   experimental deduced reads per gene
@@
 				\includegraphics[width=1.01\linewidth]{ReactomeVsMSF.png}
-				\caption{The Upset plot shows the number of shared pathways between \texttt{MSF} identified sub-graph gene list and DEG cut-off list for the 3 time-points.
-All the different toll-like receptor cascades are shown to be in the bar with 164 shared pathways only between \texttt{MSF} at different time-points.}
+				\caption{The Upset plot shows the
+                                  number of shared pathways between
+                                  \texttt{MSF} identified sub-graph
+                                  gene list and DEG cut-off list for
+                                  the three time-points. All
+                                  \sout{the} \FA{XXX} different
+                                  toll-like receptor cascades are
+                                  \sout{shown to be} in the \FA{set
+                                    of}\sout{bar with} 164 shared
+                                  pathways only between \texttt{MSF}
+                                  at different time-points.}
 				\label{fig:msfvsreactome}
@@
 		\centering
-		\caption{Comparison of connected sub-graphs of modulated genes
-			in the global network identified after the analysis with
-			\texttt{MSF} and applying different \textit{p}-value cut-offs from edgeR to genes in MSF identified modulated sub-graphs. }
+		\caption{Comparison of connected sub-graphs of
+                  modulated genes in the global network identified
+                  after the analysis with \texttt{MSF} and applying
+                  different \textit{p}-value cut-offs from edgeR to
+                  genes in MSF identified modulated sub-graphs. }
 		\label{tab:rawVsHartung}

### 2019-03-04 MSF_Paper_FA.tex
Fabian suggestions added

@@
 \usepackage[normalem]{ulem}
+\newcommand{\DONE}[1]{\begingroup\color{blue}#1\endgroup}
 \newcommand{\FA}[1]{\begingroup\color{magenta}#1\endgroup}
@@
 	topology of each gene during the gene-level statistics but still lacks to aim
-	to link the different functional pathways~\cite{Khatri2012}. \FA{mention jActiveModules somewhere in the intro}
+	to link the different functional pathways~\cite{Khatri2012}.
 
-	On these grounds we propose a novel approach to make use of the rich gene
+	\DONE{Recently systems biology approaches are also helping illuminate the structural and functional design of a cell~\cite{mitra_carvunis_ramesh_ideker_2013}.
+Mitra and colleagues discuss the different systems biology approaches to identify and measure the modules of inter-molecular interactions steering a particular biological process.
+These approaches are classified into four categories, identification of "active modules" with molecular profiles, identification of "conserved modules" across species , identification of "differential modules" with multiple conditions and identification of "composite modules" with interactions of different types.
+Identification of the active modules is one of the most successful approaches, which overlays the modules or networks onto the molecular profiles (transcriptomics, genomics, proteomics) to identify the "active modules".
+To find these active modules heuristics solutions like simulated annealing (SA), greedy methods, genetic algorithms (GA), network propagation and co-clustering methods are used~\cite{mitra_carvunis_ramesh_ideker_2013}.
+JActiveModules~\cite{jActiveModules} has been the first of this kind using simulated annealing to find the active modules.}
+
+	\DONE{The motivation was to find the modulated biological processes cross-talking, for which identifying the modulated networks first with the help of systems biology was necessary.} On these grounds we propose a novel approach to make use of the rich gene
 	and protein interaction annotation resources available, to gain additional
@@
 	signaling network, which exhibit as a whole significant differential
-	expression changes. This approach differs in two main aspects from common
+	expression changes. This approach differs in two main aspecarg1ts from common
 	pathway analysis tools. First, it does not aim to identify functional
@@
 	analysis returns sub-graphs and their joint confidence scores, reflecting
-	how the perturbation is migrated through the network. Furthermore, the
+	how the perturbation is migrated through the network. Furtharg1ermore, the
 	entry points of perturbation in the networks and overlap with conventional
-	pathway categories are returned \FA{mention impact score and reliablilty of sources here}. The output is prepared in a directed
+	pathway categories are returned.
+These entry points are given impact scores showing the degree of perturbations introduced by them.
+The reliability of the entry points is checked using a T-test by investigating the significant difference between the downstream and upstream genes of each entry point.
+The output is prepared in a directed
 	adjacency file, convenient for visualization, e.g., with
@@
 	In the next step, we check if any of the initial modulated
-        sub-graphs could further be extended \FA{beyond the immediate
+        sub-graphs could further be extended beyond the immediate
           neighbouring genes. Instead of testing single genes and
           their compatibility to be added, groups of genes are
-          considered.}
-
-        \sout{This is done when the sub-graph could not be extended any more
-        to the direct neighboring gene due to a high \textit{p}-value,
-        but the neighboring gene to the direct neighboring gene has a
-        low \textit{p}-value. This was required to connect together
-        low \textit{p}-value gene but incorporating a single high
-        \textit{p}-value gene.} If the combined \textit{p}-value of the
+          considered. If the combined \textit{p}-value of the
         initial modulated sub-graph and the extension genes is smaller
@@
 
-        \FA{After detection and extension of the modulated sub-graphs,
+        After detection and extension of the modulated sub-graphs,
           each pair of so far identified sub-graphs is tested if its
-          combination scores better than each on its own.}
-        \sout{After detection and extension of the modulated
-          sub-graphs, they are tested if combining the different
-          sub-graphs score better than on their own.} The merging of
+          combination scores better than each on its own. The merging of
         the sub-graphs is done by depth first search traversal from
@@
         the individual \textit{p}-values of the two sub-graphs, the
-        two sub-graphs are merged together to one big\FA{ger}
+        two sub-graphs are merged together to one bigger
         modulated sub-graph. This step is repeated iteratively until
-        no sub-graphs \sout{could be merged}\FA{can be merged
-          anymore}.\newline
+        no sub-graphs can be merged
+          anymore.\newline
 
@@
         downstream effects. Each individual source is given an impact
-        score, \FA{expressing the relative number of downstream genes
+        score, expressing the relative number of downstream genes
           within the corresponding sub-graph directly connected by
@@
           be helpful to prioritize different identified sources for
-          larger sub-graphs.}\sout{depending on the number of down
-          stream genes effected in the corresponding sub-graph,
-          equipping with a way to choose the most significant genes
-          for unknown phenotypes or biological processes.} In the same
-        spirit \FA{as sources}, the most downstream genes of the
+          larger sub-graphs. In the same
+        spirit as sources, the most downstream genes of the
         modulated sub-graphs are identified and defined as sinks. Due
@@
         sinks.
-
-        \TODO{?what about source reliability?}
+The reliability of each source is inspected using a T-test.
+The significant difference between the two groups, genes downstream the source to the genes upstream of the source is determined.
+This would help to assess if the source identified is reliable and marks the border between two different regulation regimes.
 
@@
         sub-graphs identified. This file could be imported into
-        Cytoscape~\cite{Cyto} for visualization. \sout{Second output
-          is the source \& sink file. It contains} \FA{Additionally, a
+        Cytoscape~\cite{Cyto} for visualization. Additionally, a
           file containing details on all sources and sinks for all
@@
           file is provided, containing the source weightage and the
-          log-fold chances of all considered genes}. \sout{The third output,
-        source weightage is provided as a node attribute file. It
-        could be imported into Cytoscape to see the impact score of
-        sources as the size of the node and log-fold change could be
-        used to see how the regulation of different gene hubs in the
-        modulated sub-graphs look.}
+          log-fold chances of all considered genes.
 
@@
 
-	MSF requires Java version 8 and JDK 1.8. The few package
-        dependencies are already been added to the release. MSF runs
-        fast on a standard laptop computer and so it has normal system
-        requirements \TODO{What is fast and what is normal?}. To run
+	\DONE{The only system requirements to run MSF are Java version 8 and JDK 1.8. The few package
+		dependencies are already been added to the release. To run
 		MSF, the user must provide two text files, one containing the
-        DGEA \TODO{DESeq2 edgR?} and the second one containing
+		DGEA from edgeR~\cite{edgeR} or DESeq2~\cite{love2014moderated}} and  the second one containing
         directed interactions in an adjacency format file. Example
@@
         modulated sub-graph identified with 42 genes. The impact score
-        of IFNB1 is 14.5\% and IFNA1 is 8.7\% for the sub-graph they
-        belong to.  Most of the sources identified by \texttt{MSF}
+        of IFNB1 is 14.5\%, \DONE{the highest impact score and IFNA1 is 8.7\%, in top 5 highest impact scores in the sub-graph they
+        	belong to} . Most of the sources identified by \texttt{MSF}
         were type~I interferon induced genes (Supplement material).
@@
         identified as two sources out of the total sources with 3.9\% 
-        and 1.6\% impact score. For the last time-point 48~hpi, six
+        and 1.6\% impact score. \DONE{IFNB1 was one of the top 5 impact score sources for the corresponding sub-graph}. For the last time-point 48~hpi, six
         modulated sub-graphs were identified. Three of the sub-graphs
@@
         most significantly modulated sub-graph with an impact score of
-        2.8\% and 3.7\% (Supplement material). \TODO{was the mentioned
-          impact score the highest of all sources? How many sources
-          are there?}
+        2.8\% and 3.7\%, \DONE{but not the highest ranked sources} (Supplement material).
 
@@
         joining neighbors to modules. The numbers indicate that
-        \texttt{MSF} detects less but larger and meaningful \TODO{?}
+        \texttt{MSF} detects less but larger and interpretable
         sub-modules, applying its statistical test. Furthermore, the
@@
         functional annotation deduced from KEGG Pathways. The
-        cross-talk between the pathways and also the multiple
+        cross-talk between the pathways and also the multiplhttps://github.com/Modulated-Subgraph-Finder/MSF/tree/master/bine
         employment of many genes is evident. The flow of information
         between the sensors and effectors can be perceived given the
-        directionality of each interaction, indicated by arrows
-        \sout{in this particular example could easily be deduced}.  In
+        directionality of each interaction, indicated by arrows.  In
         more detail, sub-graph~1 (bottom) shows how the activation of
@@
         for modulation of apoptosis. Eventually, sub-graph~3 (top
-        left) demonstrates how INFA1 and INFB1 modulates once more,
+        left) demonstrates how INFA1 and INFB1 modulates onchttps://github.com/Modulated-Subgraph-Finder/MSF/tree/master/bine more,
         via only a few intermediate steps, the apoptotic response of
@@
 	This display case might advertise with how little effort
-        complex data can be \FA{processed and prepared for
-          interpretation by the domain expert, to } \sout{interpreted,
-          help to} apprehend the dynamics of the underlying processes
+        complex data can be processed and prepared for
+          interpretation by the domain expert, to apprehend the dynamics of the underlying processes
         and suggest testable hypothesis and potential points of
@@
 
+	 \DONE{The purpose of benchmark is mainly to show overall capabilities of \texttt{MSF}.
+\texttt{MSF} was compared to jActivemodules~\cite{jActiveModules} since they use similar approach to find and score the modules.
+Comparison of \texttt{MSF} to Reactome~\cite{Reactome} was done because  Reactome is the standard tool used when looking for modulated pathways after DGEA.
+Since \texttt{MSF} and GSEA~\cite{Subramanian15545} use no cut-off to the DGEA, their results were also compared.}
+
 	\textbf{jActiveModules}
 
-	jActiveModules~\cite{jActiveModules} \TODO{citation missing in
-          bib} is a plugin in Cytoscape that searches for molecular
+	jActiveModules~\cite{jActiveModules} is a plugin in Cytoscape that searches for molecular
         interaction network to find expression activated
@@
         activated sub-networks is close to the method used in
-        \texttt{MSF}. \TODO{what are the differences?}
+        \texttt{MSF}.
+\DONE{The difference is in how these sub-networks are identified.
+\texttt{MSF} starts building the sub-graphs from one gene, incorporating and combining the \textit{p}-value of the next gene, with the check that the combined \textit{p}-value of new sub-graph should be better than the original.
+On the other hand jActiveModules first transforms all the gene’s \textit{p}-values to z-scores and tries to find connected sets of genes with unexpectedly high levels of differential expression, in this case high z-scores.
+The overall score of the sub-network is calculated by combining the z-scores of the genes.
+Next using their extended simulated annealing method jActiveModules finds the user defined number of highest scoring putative modules by toggling multiple nodes to merge additional components.}
 
 	The first time-point of Ebola infection data was analyzed
-        using jActiveModule (Version 3.2.1) to compare the modulated
-        sub-graphs identified by \texttt{MSF} and jActiveModule. The
+        using jActiveModules (Version 3.2.1) to compare the modulated
+        sub-graphs identified by \texttt{MSF} and jActiveModules. The
         input files were same for both tools, number of modules
-        parameter was set to 5 \TODO{what does this parameter
-          do?}. jActiveModule identified 5 modules, the module with
-        the highest pathScore was selected for comparison with
-        \texttt{MSF} identified modulated sub-graphs \TODO{why only 1
-          out of 5?}. The module consists of a single graph with 314
+        parameter was set to 5 (default).
+This would report 5 putative modules even when only one of them is good.
+Out of the 5 putative modules identified by jActiveModules, the module with the highest pathScore was selected for comparison with
+        \texttt{MSF} identified modulated sub-graphs. The module consists of a single graph with 314
         genes. While \texttt{MSF} identified three directed modulated
         sub-graphs with 42, 69 and 139 genes. The overlap of the
-        common genes identified between \texttt{MSF} and jActiveModule
-        is shown in Fig~\ref{fig:MSFvsJActiveModule}. \sout{It is
-          difficult to state which tool identifies more accurate set
-          of genes as modulated, since there is no golden standard
-          data example to refer.} Although jActiveModule identifies
+        common genes identified between \texttt{MSF} and jActiveModules
+        is shown in Fig~\ref{fig:MSFvsJActiveModule}.
+\DONE{JActiveModules identified a single module, since this module had unexpectedly high levels of differential expression.
+With the extended simulated annealing method, jActiveModules toggled multiple nodes to merge a number of components with low score into this single high-scoring component making it larger than original.} Although jActiveModules was able to identify
         single connected module, \texttt{MSF} provides directionality,
         with the identification of possible perturbation sources of
-        the sub-graphs. \TODO{find reasons why jActiveModules gives so
-          much bigger moduls by comparing the details of the method!}
+        the sub-graphs.
 
@@
 	Gene enrichment analysis was performed using Reactome analyze
-        data tool~\cite{Reactome} (version 67) \TODO{on which
-          data?}. Reactome's over-representation analysis tool tests
+        data tool~\cite{Reactome} (version 67) on the different time-points of Ebola infection data. Reactome's over-representation analysis tool tests
         whether certain Reactome pathways are enriched for the list of
@@
         from edgeR for the three time-points were filtered using the
-        cut-off of adjusted \textit{p}-value \sout{>}\FA{<} 0.05. This
+        cut-off of adjusted \textit{p}-value < 0.05. This
         DEG list was used for gene enrichment analysis. The
@@
 	All enriched pathways with a cut-off of \textit{p}-value
-        \sout{>}\FA{<} 0.05 0.05 for \texttt{MSF} and DEG list for the
+        < 0.05 for \texttt{MSF} and DEG list for the
         three time-points were selected. The comparison shows most of
@@
 	Gene set enrichment analysis (GSEA)~\cite{Subramanian15545}
-        \TODO{reference is missing in bib} is a method to identify
+        is a method to identify
         classes of genes or proteins that are over-represented in a
@@
         Toll\FA{-}like recpetor, chemokine signaling pathway,
-        cytosolic DNA-sensing pathway, Jak-STAT signaling pathway,
+        cytosolic DNA-sensing pathway, Jak-STAT signaling pathway,arg1
         RIG-I-like receptor signaling pathway and apoptosis as the
@@
 				\includegraphics[width=13cm]{MSF_Robustness.png}
-				\caption{\FA{Recall rates for genes in
-                                  }\sout{Shows the percentage of}
-                                  \texttt{MSF} identified sub-graph
-                                  \sout{genes recall rates} for the
+				\caption{Recall rates for genes in
+                                  \texttt{MSF} identified sub-graph for the
                                   three different time points of EBOV
                                   infection data for 100
-                                  simulation\FA{s} where Poisson
+                                  simulations where Poisson
                                   distributed noise was added to the
@@
                                   the three time-points. All
-                                  \sout{the} \FA{XXX} different
+                                  10 different
                                   toll-like receptor cascades are
-                                  \sout{shown to be} in the \FA{set
-                                    of}\sout{bar with} 164 shared
+                                   in the set
+                                    of 164 shared
                                   pathways only between \texttt{MSF}

### 2019-03-04 MSF_Paper_FA.tex
Fabian suggestions added

@@
 \newcommand{\FA}[1]{\begingroup\color{magenta}#1\endgroup}
-\newcommand{\ILH}[1]{\begingroup\color{blue}#1\endgroup}
+\newcommand{\ILH}[1]{\begingroup\color{green}#1\endgroup}
 \newcommand{\TODO}[1]{\begingroup\color{red}#1\endgroup}
@@
 	 \DONE{The purpose of benchmark is mainly to show overall capabilities of \texttt{MSF}.
-\texttt{MSF} was compared to jActivemodules~\cite{jActiveModules} since they use similar approach to find and score the modules.
+\texttt{MSF} was compared to jActivemodules~\cite{jActiveModules} since they use similar approach to find and score the mi am a risk takerodules.
 Comparison of \texttt{MSF} to Reactome~\cite{Reactome} was done because  Reactome is the standard tool used when looking for modulated pathways after DGEA.
@@
 On the other hand jActiveModules first transforms all the gene’s \textit{p}-values to z-scores and tries to find connected sets of genes with unexpectedly high levels of differential expression, in this case high z-scores.
-The overall score of the sub-network is calculated by combining the z-scores of the genes.
+The overall score of the sub-network is calculated by combinini am a risk takerg the z-scores of the genes.
 Next using their extended simulated annealing method jActiveModules finds the user defined number of highest scoring putative modules by toggling multiple nodes to merge additional components.}

### 2019-03-08 MSF_Paper_FA.tex
jActiveModules improved

--- 2019-03-04 MSF_Paper_FA.tex
@@
 
-	\DONE{Recently systems biology approaches are also helping illuminate the structural and functional design of a cell~\cite{mitra_carvunis_ramesh_ideker_2013}.
-Mitra and colleagues discuss the different systems biology approaches to identify and measure the modules of inter-molecular interactions steering a particular biological process.
-These approaches are classified into four categories, identification of "active modules" with molecular profiles, identification of "conserved modules" across species , identification of "differential modules" with multiple conditions and identification of "composite modules" with interactions of different types.
-Identification of the active modules is one of the most successful approaches, which overlays the modules or networks onto the molecular profiles (transcriptomics, genomics, proteomics) to identify the "active modules".
+	From another perspective, network based approaches do not categorize sets of genes into functional pathways, but they consider all interactions to be equal.
+Thereby, they avoid distinguishing arbitrarily between interactions within a pathway and interactions between pathways (i.e., cross-talk).
+With this they aim to identify sub-networks that show modulation between two conditions or upon a stimuli~\cite{mitra_carvunis_ramesh_ideker_2013}.
 To find these active modules heuristics solutions like simulated annealing (SA), greedy methods, genetic algorithms (GA), network propagation and co-clustering methods are used~\cite{mitra_carvunis_ramesh_ideker_2013}.
-JActiveModules~\cite{jActiveModules} has been the first of this kind using simulated annealing to find the active modules.}
+jActiveModules has been the first of this kind using simulated annealing to find the modulated sub-networks~\cite{jActiveModules}.
+The benefit of omitting pathways is bought by reduced interpretability of the results due to the lack of functional labels on the networks.
 
-	\DONE{The motivation was to find the modulated biological processes cross-talking, for which identifying the modulated networks first with the help of systems biology was necessary.} On these grounds we propose a novel approach to make use of the rich gene
-	and protein interaction annotation resources available, to gain additional
-	functional insights from basic DGEA. To this, we start with the
+    On these grounds we propose a novel approach to make use of the rich gene
+	and protein interaction annotation resources available and combining it to functional pathway annotations to gain additional insights from basic DGEA. To this, we start with the
 	presupposition that expression of neighboring genes within a functional
@@
 	signaling network, which exhibit as a whole significant differential
-	expression changes. This approach differs in two main aspecarg1ts from common
+	expression changes. This approach differs in two main aspects from common
 	pathway analysis tools. First, it does not aim to identify functional
@@
 	analysis returns sub-graphs and their joint confidence scores, reflecting
-	how the perturbation is migrated through the network. Furtharg1ermore, the
+	how the perturbation is migrated through the network. Furthermore, the
 	entry points of perturbation in the networks and overlap with conventional
@@
 These entry points are given impact scores showing the degree of perturbations introduced by them.
-The reliability of the entry points is checked using a T-test by investigating the significant difference between the downstream and upstream genes of each entry point.
+The reliability of the entry points is checked using a t-test on the \textit{p}-values by investigating the significant difference between the downstream and upstream genes of each entry point.
 The output is prepared in a directed
@@
 	thereby being a valuable complement to existing pathway based
-	methods.  \texttt{MSF} is freely accessible on github under the terms of the
+	methods. \texttt{MSF} is freely accessible on git hub under the terms of the
 	Creative Commons Attribution 4.0 International License.
@@
 	there own.
-	\newline
 
@@
         sub-graphs could further be extended beyond the immediate
-          neighbouring genes. Instead of testing single genes and
+          neighboring genes. Instead of testing single genes and
           their compatibility to be added, groups of genes are
@@
         sinks.
-The reliability of each source is inspected using a T-test.
+The reliability of each source is inspected using a t-test on the \textit{p}-values .
 The significant difference between the two groups, genes downstream the source to the genes upstream of the source is determined.
@@
           modulated sub-graphs is reported. Furthermore, for
-          facilitated visualisation in Cytoscape, a node attribute
-          file is provided, containing the source weightage and the
+          facilitated visualization in Cytoscape, a node attribute
+          file is provided, containing the source weight-age and the
           log-fold chances of all considered genes.
@@
 
-	\DONE{The only system requirements to run MSF are Java version 8 and JDK 1.8. The few package
+	The only system requirements to run \texttt{MSF} are Java version 8 and JDK 1.8. The few package
 		dependencies are already been added to the release. To run
-		MSF, the user must provide two text files, one containing the
-		DGEA from edgeR~\cite{edgeR} or DESeq2~\cite{love2014moderated}} and  the second one containing
+		\texttt{MSF}, the user must provide two text files, one containing the
+		DGEA and the second one containing
         directed interactions in an adjacency format file. Example
         files and a detailed tutorial to use MSF has been provided on
-        github https://github. com/Modulated-Subgraph-Finder/MSF.
+        git hub \url{https://github.com/Modulated-Subgraph-Finder/MSF}.
 
@@
 	To demonstrate its usefulness, \texttt{MSF} is applied to
-        RNA-seq data set of primary human monocyte-derived macrophages
-        (MDMs) infected with Ebola virus
-        (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV) belongs to the
+        RNA-seq data set of primary human monocyte derived macrophages
+        (MDMs) infected with Ebola virus (GSE84188)~\cite{Olejnik}. Ebola Virus (EBOV) belongs to the
         Filoviridea family: filamentous, enveloped and single stranded
@@
         sub-graphs were identified with 42, 139, and 69 genes.  The
-        modulated sub-graphs consist predominantly of cytokines,
+        modulated sub-graphs consist predominantly of Cytokines,
         chemokines (CXCL10, CCL8, CXCL9, CXCL11, CXCR4, CCR7, CCL4L1,
@@
         modulated sub-graph identified with 42 genes. The impact score
-        of IFNB1 is 14.5\%, \DONE{the highest impact score and IFNA1 is 8.7\%, in top 5 highest impact scores in the sub-graph they
-        	belong to} . Most of the sources identified by \texttt{MSF}
+        of IFNB1 is 14.5, the highest impact score and IFNA1 is 8.7, in top 5 highest impact scores in the sub-graph they belong to. Most of the sources identified by \texttt{MSF}
         were type~I interferon induced genes (Supplement material).
@@
         others being smaller than 6 genes. Again, IFNB1 and IFNA1 were
-        identified as two sources out of the total sources with 3.9\% 
-        and 1.6\% impact score. \DONE{IFNB1 was one of the top 5 impact score sources for the corresponding sub-graph}. For the last time-point 48~hpi, six
+        identified as two sources out of the total sources with 3.9 
+        and 1.6 impact score. IFNB1 was one of the top 5 impact score sources for the corresponding sub-graph. For the last time-point 48~hpi, six
         modulated sub-graphs were identified. Three of the sub-graphs
@@
         most significantly modulated sub-graph with an impact score of
-        2.8\% and 3.7\%, \DONE{but not the highest ranked sources} (Supplement material).
+        2.8 and 3.7, but not among the highest ranked sources (Supplement material).
 
@@
         functional annotation deduced from KEGG Pathways. The
-        cross-talk between the pathways and also the multiplhttps://github.com/Modulated-Subgraph-Finder/MSF/tree/master/bine
+        cross-talk between the pathways and also the multiple
         employment of many genes is evident. The flow of information
@@
         more detail, sub-graph~1 (bottom) shows how the activation of
-        toll-like receptor, cytokine, chemokine activating jak-stat
-        and mapk genes, together with TNF leads into apoptosis. The
+        Toll-like receptor, Cytokine, Chemokine activating Jak-STAT
+        and MAPK genes, together with TNF leads into apoptosis. The
         next significant sub-graph (sub-graph~2: top right) reveals
@@
         which are reported to interact with Ebola glycoprotein
-        (GP)~\cite{Veljkovic}, chemokines, cytokines, and cytosolic
+        (GP)~\cite{Veljkovic}, Chemokines, Cytokines, and Cytosolic
         DNA sensing are directly or indirectly controlling cell
@@
         for modulation of apoptosis. Eventually, sub-graph~3 (top
-        left) demonstrates how INFA1 and INFB1 modulates onchttps://github.com/Modulated-Subgraph-Finder/MSF/tree/master/bine more,
+        left) demonstrates how INFA1 and INFB1 modulates once more,
         via only a few intermediate steps, the apoptotic response of
@@
 
-	\subsection*{Robustness}
+	\subsection*{Robustness}Gene set enrichment analysis
 
@@
         range of possible values. The observed median recall rates lay
-        between 71~\% (6~hpi) and 84~\% (48~hpi).
+        between 71~\% (6~hpi) and 84~\% (48~hpi).Gene set enrichment analysis
 
@@
 
-	 \DONE{The purpose of benchmark is mainly to show overall capabilities of \texttt{MSF}.
-\texttt{MSF} was compared to jActivemodules~\cite{jActiveModules} since they use similar approach to find and score the modules.
-Comparison of \texttt{MSF} to Reactome~\cite{Reactome} was done because  Reactome is the standard tool used when looking for modulated pathways after DGEA.
-Since \texttt{MSF} and GSEA~\cite{Subramanian15545} use no cut-off to the DGEA, their results were also compared.}
+	 The purpose of the comparisons to existing tools is to show the overall capabilities of \texttt{MSF}.
+\texttt{MSF} was compared to jActiveModules~\cite{jActiveModules} since they use similar approaches to find and score modules.
+For comparison to classical pathway enrichment analysis Reactome~\cite{Reactome} and gene set enrichment analysis (GSEA)~\cite{Subramanian15545} was chosen since both are widely used and the latter does not rely on \textit{p}-value cut-off.
 
@@
         \texttt{MSF}.
-\DONE{The difference is in how these sub-networks are identified.
+The difference is in how these sub-networks are identified.
 \texttt{MSF} starts building the sub-graphs from one gene, incorporating and combining the \textit{p}-value of the next gene, with the check that the combined \textit{p}-value of new sub-graph should be better than the original.
@@
 The overall score of the sub-network is calculated by combining the z-scores of the genes.
-Next using their extended simulated annealing method jActiveModules finds the user defined number of highest scoring putative modules by toggling multiple nodes to merge additional components.}
+Next using their extended simulated annealing method jActiveModules toggles multiple nodes to merge additional components.
 
@@
         sub-graphs identified by \texttt{MSF} and jActiveModules. The
-        input files were same for both tools, number of modules
-        parameter was set to 5 (default).
-This would report 5 putative modules even when only one of them is good.
-Out of the 5 putative modules identified by jActiveModules, the module with the highest pathScore was selected for comparison with
+        input files were same for both tools. From the modules identified by jActiveModules, the module with the highest pathScore was selected for comparison with
         \texttt{MSF} identified modulated sub-graphs. The module consists of a single graph with 314
@@
         is shown in Fig~\ref{fig:MSFvsJActiveModule}.
-\DONE{JActiveModules identified a single module, since this module had unexpectedly high levels of differential expression.
-With the extended simulated annealing method, jActiveModules toggled multiple nodes to merge a number of components with low score into this single high-scoring component making it larger than original.} Although jActiveModules was able to identify
-        single connected module, \texttt{MSF} provides directionality,
+The sub-graphs identified by \texttt{MSF} are more fragmented than jActiveModules.
+Unfortunately there is no golden standard example data that could help benchmark the method.
+\texttt{MSF} provides directionality,
         with the identification of possible perturbation sources of
-        the sub-graphs.
+        the sub-graphs. The predefined pathway labels could also be seen in \texttt{MSF} identified sub-graph with little effort using StringApp~\cite{StringApp}.
 
@@
         when \textit{p}-value cut off DEG list was analyzed for
-        enrichment. Nine toll-like receptor cascades TLR10, TLR2,
+        enrichment. Nine Toll-like receptor cascades TLR10, TLR2,
         TLR3, TLR4, TLR5, TLR7/8, TLR9, TLR1:TLR2 and TLR6:TLR2 were
@@
 
-	\textbf{Gene set enrichment analysis (GSEA)}
+	\textbf{Gene set enrichment analysis }
 
@@
         GSEABase (version 1.44.0). GSEA was able to identify
-        Toll\FA{-}like recpetor, chemokine signaling pathway,
-        cytosolic DNA-sensing pathway, Jak-STAT signaling pathway,arg1
+        Toll-like receptor, chemokine signaling pathway,
+        cytosolic DNA-sensing pathway, Jak-STAT signaling pathway,
         RIG-I-like receptor signaling pathway and apoptosis as the
@@
         the cross-talk and topology of genes between the pathways
-        identifed. \texttt{MSF} and GSEA uses complete DEG list
+        identified. \texttt{MSF} and GSEA uses complete DEG list
         without any cut-off, is why pathways important for Ebola
@@
 	\subsection*{Supplement material}
-	Supplementary material is available form GitHub: https://github.com/Modulated-Subgraph-Finder/MSF
+	Supplementary material is available form Git Hub: https://github.com/Modulated-Subgraph-Finder/MSF
 
@@
 				\caption{Graphical representation of the \texttt{MSF}
-					heuristical approach to detect modulated sub-graphs
+					heuristic approach to detect modulated sub-graphs
 					in a global gene regulatory network.}
@@
 				\includegraphics[width=1.0\linewidth]{DirectedMerged_Network.png}
-				\caption{Visualisation of the three modulated
+				\caption{Visualization of the three modulated
 					directed sub-graphs identified by \texttt{MSF} at 6~hpi. The
@@
 				\includegraphics[width=0.9\linewidth]{MSFvsJActiveModule.png}
-				\caption{The Venn diagram shows the common genes identified as modulated from \texttt{MSF} identified sub-graphs and jActiveModule identified module.}
+				\caption{The Venn diagram shows the common genes identified as modulated from \texttt{MSF} identified sub-graphs and jActiveModules identified module.}
 				\label{fig:MSFvsJActiveModule}
@@
                                   10 different
-                                  toll-like receptor cascades are
+                                  Toll-like receptor cascades are
                                    in the set
@@
                   different \textit{p}-value cut-offs from edgeR to
-                  genes in MSF identified modulated sub-graphs. }
+                  genes in \texttt{MSF} identified modulated sub-graphs. }
 		\label{tab:rawVsHartung}

### 2019-03-11 MSF_Paper_FA.tex
last revision fabian

@@
 
-		To overcome this limitation, we developed a novel approach to identify
-		concertedly modulated sub-graphs in the global cell signaling network,
-		based on the DGEA results of all genes tested. Thereby, expression patterns
-		of genes are integrated according to the topology of their interactions and
-		allow potentially to read the flow of information and identify the effectors. The described software, named \texttt{Modulated
-			Sub-graph Finder} (\texttt{MSF}) is freely available at
+		To overcome this limitation, we developed a novel
+                approach to identify concertedly modulated sub-graphs
+                in the global cell signaling network, based on the
+                DGEA results of all genes tested. To this end,
+                expression patterns of genes are integrated according
+                to the topology of their interactions and allow
+                potentially to read the flow of information and
+                identify the effectors. The described software, named
+                \texttt{Modulated Sub-graph Finder} (\texttt{MSF}) is
+                freely available at
                 \url{https://github.com/Modulated-Subgraph-Finder/MSF}.
@@
 
-	Existing pathway-based analysis approaches use different research
-	designs, which can be categorized into ORA (Over-representation
-	analysis), FCS (Functional class scoring) and pathway topology based
-	methods. All of which aim to find a subset of genes, e.g.,
-	significantly differentially expressed genes, genes associated with a
-	certain pathway more often than expected given the total set of
-	examined genes, e.g.~the whole genome background. ORA is
-	the first and the most basic method of pathway analysis~\cite{Campos}. It uses a
-	DEG list with user defined cut-off for the log-fold change and
-	\textit{p}-value (most commonly using absolute log-fold change $\geq$
-	2 and \textit{p}-value $\leq$ 0.05). Subsequently, sets of genes
-	associated with annotated pathways are tested for being
-	over-represented in the set of DEGs. To this end, hyper-geometric
-	distribution, chi-square tests, binomial probability or the Fisher’s
-	exact test are used. Thereby the information of the topology of genes
-	in the pathways is neglected~\cite{Bayer}. Furthermore, ORA assumes
-	that the biological pathways are independent of each other and ignores
-	the fact that they cross-talk and overlap~\cite{Khatri2012,Campos}.
+	Existing pathway-based analysis approaches use different
+        research designs, which can be categorized into ORA
+        (Over-representation analysis), FCS (Functional class scoring)
+        and pathway topology based methods. All of which aim to find a
+        subset of genes, e.g., significantly differentially expressed
+        genes, genes associated with a certain pathway more often than
+        expected given the total set of examined genes, e.g.~the whole
+        genome background. ORA is the first and the most basic method
+        of pathway analysis~\cite{Campos}. It uses a DEG list with
+        user defined cut-off for the log-fold change and
+        \textit{p}-value (most commonly using absolute log-fold change
+        $\geq$ 2 and \textit{p}-value $\leq$ 0.05). Subsequently, sets
+        of genes associated with annotated pathways are tested for
+        being over-represented in the set of DEGs. To this end,
+        hyper-geometric distribution, chi-square tests, binomial
+        probability or the Fisher’s exact test are used, whereas, the
+        information of the topology of genes in the pathways is
+        neglected~\cite{Bayer}. Furthermore, ORA assumes that the
+        biological pathways are independent of each other and ignores
+        the fact that they cross-talk and
+        overlap~\cite{Khatri2012,Campos}.
 
@@
 
-	From another perspective, network based approaches do not categorize sets of genes into functional pathways, but they consider all interactions to be equal.
-Thereby, they avoid distinguishing arbitrarily between interactions within a pathway and interactions between pathways (i.e., cross-talk).
-With this they aim to identify sub-networks that show modulation between two conditions or upon a stimuli~\cite{mitra_carvunis_ramesh_ideker_2013}.
-To find these active modules heuristics solutions like simulated annealing (SA), greedy methods, genetic algorithms (GA), network propagation and co-clustering methods are used~\cite{mitra_carvunis_ramesh_ideker_2013}.
-jActiveModules has been the first of this kind using simulated annealing to find the modulated sub-networks~\cite{jActiveModules}.
-The benefit of omitting pathways is bought by reduced interpretability of the results due to the lack of functional labels on the networks.
+	From another perspective, network based approaches do not
+        categorize sets of genes into functional pathways, but they
+        consider all interactions to be equal. Thereby, they avoid
+        distinguishing arbitrarily between interactions within a
+        pathway and interactions between pathways (i.e.,
+        cross-talk). With this they aim to identify sub-networks that
+        show modulation between two conditions or upon a
+        stimuli~\cite{mitra_carvunis_ramesh_ideker_2013}. To find
+        these active modules heuristics solutions like simulated
+        annealing (SA), greedy methods, genetic algorithms (GA),
+        network propagation and co-clustering methods are
+        used~\cite{mitra_carvunis_ramesh_ideker_2013}. jActiveModules
+        has been the first of this kind using simulated annealing to
+        find modulated sub-networks~\cite{jActiveModules}. The benefit
+        of omitting pathways is bought by reduced interpretability of
+        the results due to the lack of functional labels on the
+        networks.
 
@@
 
-	As input, information on functional links between genes provided by
-	e.g.~KEGG or Reactome and information on the differential expression status
-	of single genes resulting from a DGEA, are required. As a result the
-	analysis returns sub-graphs and their joint confidence scores, reflecting
-	how the perturbation is migrated through the network. Furthermore, the
-	entry points of perturbation in the networks and overlap with conventional
-	pathway categories are returned.
-These entry points are given impact scores showing the degree of perturbations introduced by them.
-The reliability of the entry points is checked using a t-test on the \textit{p}-values by investigating the significant difference between the downstream and upstream genes of each entry point.
-The output is prepared in a directed
-	adjacency file, convenient for visualization, e.g., with
-	StringApp~\cite{StringApp}, available as a Cytoscape plug-in~\cite{Cyto}.
+	As input, information on functional links between genes
+        provided by e.g.~KEGG or Reactome and information on the
+        differential expression status of single genes resulting from
+        a DGEA, are required. As a result the analysis returns
+        sub-graphs and their joint confidence scores, reflecting how
+        the perturbation migrates through the network. Furthermore,
+        the entry points of perturbation in the networks and overlap
+        with conventional pathway categories are returned. To
+        facilitate prioritization of the perturbation entry points, to
+        each a impact score and a measure of its reliability is
+        assigned. The impact score expresses the fraction of the
+        sub-module being downstream of the entry point. The
+        reliability is measured using a t-test on the
+        \textit{p}-values of the immediate upstream and downstream
+        genes. The output is prepared in a directed adjacency file,
+        convenient for visualization, e.g., with
+        StringApp~\cite{StringApp}, available as a Cytoscape
+        plug-in~\cite{Cyto}.
 
-	The proposed algorithm is named
-	\texttt{Modulated Sub-graph Finder} (abbreviated
-	\texttt{MSF}). \texttt{MSF} can help transform the information obtained
-	from DGEA into comprehensible knowledge of signal transduction of genes and
-	thereby being a valuable complement to existing pathway based
-	methods. \texttt{MSF} is freely accessible on git hub under the terms of the
-	Creative Commons Attribution 4.0 International License.
+	The proposed algorithm is named \texttt{Modulated Sub-graph
+          Finder} (abbreviated \texttt{MSF}). \texttt{MSF} can help
+        transform the information obtained from DGEA into
+        comprehensible knowledge of signal transduction of genes,
+        hence being a valuable complement to existing pathway based
+        methods. \texttt{MSF} is freely accessible on git hub under
+        the terms of the Creative Commons Attribution 4.0
+        International License.
 
@@
 
-	\texttt{MSF} was implemented as a java program. It is developed as a novel heuristic approach to find concertedly
-	modulated sub-graphs in networks of biological interactions. \texttt{MSF}
-	does not use predefined gene sets grouped into functional units, but rather
-	relies purely on the network of interacting genes. The input network
-	consists of nodes corresponding to genes and edges representing
-	interactions. Furthermore it utilizes comprehensive results from a
-	differential gene expression analysis to discover the sub-graphs, or
-	modules, which are as a whole modulated.
+	\texttt{MSF} was implemented as a Java program. It is
+        developed as a novel heuristic approach to find concertedly
+        modulated sub-graphs in networks of biological
+        interactions. \texttt{MSF} does not use predefined gene sets
+        grouped into functional units, but rather relies purely on the
+        network of interacting genes. The input network consists of
+        nodes corresponding to genes and edges representing
+        interactions. Furthermore it utilizes comprehensive results
+        from a differential gene expression analysis to discover the
+        sub-graphs, or modules, which are as a whole modulated.
 
-	\texttt{MSF} uses the individual gene's \textit{p}-values generated from
-	the DGEA. The \textit{p}-value expresses the probability that the null
-	hypothesis of unmodified gene expression can't be rejected for a given
-	statistical model. To find significantly modulated sub-graphs individual
-	\textit{p}-values of the vicinal genes in the global network are combined
-	into a single combined \textit{p}-value, using a statistical method for
-	combining dependent \textit{p}-values described by
-	Hartung~\cite{Hartung}. Hartung's method uses the inverse of standard
-	normal distribution function. Using the inverse normal cumulative distribution function $\Phi^{-1}$, individual gene \textit{p}-values $T_{i}$ are 
-	transformed to their corresponding normal score $t_{i}=\Phi^{-1}(1-T_{i})$ that is uniformly distributed on (0,1). Then using these normal
-	scores, the correlation between genes is calculated $Cov(t_{i},t_{j})=\rho$. The normal scores and
-	correlation are applied to the weighted inverse normal function to calculate the
-	combined \textit{p}-value $t(\rho)$ for all genes examined, namely the examined
-	sub-graph
+	\texttt{MSF} uses the individual gene's \textit{p}-values
+        generated from the DGEA. The \textit{p}-value expresses the
+        probability that the null hypothesis of unmodified gene
+        expression can't be rejected for a given statistical model. To
+        find significantly modulated sub-graphs individual
+        \textit{p}-values of the vicinal genes in the global network
+        are combined into a single combined \textit{p}-value, using a
+        statistical method for combining dependent \textit{p}-values
+        described by Hartung~\cite{Hartung}. Hartung's method uses the
+        inverse of standard normal distribution function. Using the
+        inverse normal cumulative distribution function $\Phi^{-1}$,
+        individual gene \textit{p}-values $T_{i}$ are transformed to
+        their corresponding normal score $t_{i}=\Phi^{-1}(1-T_{i})$
+        that is uniformly distributed on (0,1). Then using these
+        normal scores, the correlation between genes is calculated
+        $Cov(t_{i},t_{j})=\rho$. The normal scores and correlation are
+        applied to the weighted inverse normal function to calculate
+        the combined \textit{p}-value $t(\rho)$ for all genes
+        examined, namely the examined sub-graph
 	\newline
@@
 
-	Lambda $\lambda$ be the weights for each gene, currently all genes have equal weight i.e. 1. The combined \textit{p}-value  $t(\rho)$ of a sub-graph will express the
-	significance of all genes in the sub-graph being modulated
-	together. Thereby, the information from the different genes are used as,
-	although not independent, replicated measurement of the behavior of the
-	whole sub-graph. This potentially increases the power to detect also
-	significant sub-graphs consisting of genes which are not significant on
-	there own.
+	Lambda $\lambda$ be the weights for each gene, currently all
+        genes have equal weight, i.e.~1. The combined \textit{p}-value
+        $t(\rho)$ of a sub-graph will express the significance of all
+        genes in the sub-graph being modulated together. The
+        information from the different genes are used as, although not
+        independent, replicated measurement of the behavior of the
+        whole sub-graph. This potentially increases the power to
+        detect also significant sub-graphs consisting of genes which
+        are not significant on there own.
 
@@
         limit of how much of the sub-graph's perturbation could have
-          been introduced by the respective source, and thereby could
-          be helpful to prioritize different identified sources for
-          larger sub-graphs. In the same
-        spirit as sources, the most downstream genes of the
-        modulated sub-graphs are identified and defined as sinks. Due
-        to loops not all sub-graphs are guaranteed to have sources or
-        sinks.
-The reliability of each source is inspected using a t-test on the \textit{p}-values .
-The significant difference between the two groups, genes downstream the source to the genes upstream of the source is determined.
-This would help to assess if the source identified is reliable and marks the border between two different regulation regimes.
+        been introduced by the respective source, and thereby could be
+        helpful to prioritize different identified sources for larger
+        sub-graphs. In the same spirit as sources, the most downstream
+        genes of the modulated sub-graphs are identified and defined
+        as sinks. Due to loops not all sub-graphs are guaranteed to
+        have sources or sinks. The reliability of each source is
+        inspected using a t-test on the \textit{p}-values . The
+        significant difference between the two groups, genes
+        downstream the source to the genes upstream of the source is
+        determined. This would help to assess if the source identified
+        is reliable and indeed marks the border between two different
+        regulation regimes.
 
@@
         sub-graphs identified. This file could be imported into
-        Cytoscape~\cite{Cyto} for visualization. Additionally, a
-          file containing details on all sources and sinks for all
-          modulated sub-graphs is reported. Furthermore, for
-          facilitated visualization in Cytoscape, a node attribute
-          file is provided, containing the source weight-age and the
-          log-fold chances of all considered genes.
+        Cytoscape~\cite{Cyto} for visualization. Additionally, a file
+        containing details on all sources and sinks for all modulated
+        sub-graphs is reported. Furthermore, for facilitated
+        visualization in Cytoscape, a node attribute file is provided,
+        containing the source weightage and the log-fold chances of
+        all considered genes.
 
@@
         vast literature is available for the pathogenesis of Ebola,
-        thereby, facilitating the verification of the results of
+        hence facilitating the verification of the results of
         \texttt{MSF} with the vast literature present on Ebola
@@
         joining neighbors to modules. The numbers indicate that
-        \texttt{MSF} detects less but larger and interpretable
+        \texttt{MSF} detects less but larger and easier interpretable
         sub-modules, applying its statistical test. Furthermore, the
@@
 
-	\subsection*{Robustness}Gene set enrichment analysis
+	\subsection*{Robustness}\sout{Gene set enrichment analysis}
 
@@
         range of possible values. The observed median recall rates lay
-        between 71~\% (6~hpi) and 84~\% (48~hpi).Gene set enrichment analysis
+        between 71~\% (6~hpi) and 84~\% (48~hpi).\sout{Gene set enrichment analysis}
 
@@
 
-	All enriched pathways with a cut-off of \textit{p}-value
-        < 0.05 for \texttt{MSF} and DEG list for the
-        three time-points were selected. The comparison shows most of
-        the pathways known from literature to be dis-regulated by
-        Ebola infection are enriched in both the enrichment
-        analysis. Toll-Like receptor signaling pathway when interacts
-        with EBOV glycoprotein (GP), it triggers the activation of
+	All enriched pathways with a cut-off of \textit{p}-value <
+        0.05 for \texttt{MSF} and DEG list for the three time-points
+        were selected. The comparison shows most of the pathways known
+        from literature to be dis-regulated by Ebola infection are
+        enriched in both the enrichment analysis. Toll-like receptor
+        signaling pathway when interacts with EBOV glycoprotein (GP)\TODO{???},
+        it triggers the activation of
         cytokines~\cite{Olejnik}. Toll-like receptor pathway is
@@
         highest ranked pathways. Although GSEA identified the
-        important pathways for Ebola infection, it did not show how
+        important pathways for Ebola infection, it did not show how \TODO{???}
         the cross-talk and topology of genes between the pathways
         identified. \texttt{MSF} and GSEA uses complete DEG list
-        without any cut-off, is why pathways important for Ebola
+        without any cut-off, \TODO{????} is why pathways important for Ebola
         infection showed up even with weak signal genes in it.
@@
         concertedly modulated sub-graphs in a given network. Its
-        implementation in java enables its use across many operating
+        implementation in Java enables its use across many operating
         systems e.g. linux and windows. So far the raw output from
